### 2017-09-18 03Writeup/paper.tex
Bit of re-organizing; put the GET data in .py scripts rather than notebooks

%-------------------------------------------------------------------------------
% Preamble
%-------------------------------------------------------------------------------

\documentclass[twocol]{ametsoc}
\journal{mwr}

%Citations should be of the form ``author year''  not ``author, year''
\bibpunct{(}{)}{;}{a}{}{,}

%Packages
\usepackage{
	siunitx, % units
	cleveref,% cross-references
	physics, % better physics notation
	url, % for URLs
	amsmath, % for math
}

% TODO Notes -- DO NOT KEEP FOR FINAL

%-------------------------------------------------------------------------------
% Title and Authors
%-------------------------------------------------------------------------------

%\title{Causes and subseasonal-to-seasonal model skill of the persistent intense rainfall and flooding in Paraguay during the austral summer 2015-2016}
\title{Extreme rainfall in Paraguay during the 2015-2016 austral summer: causes and subseasonal-to-seasonal predictive skill}
\authors{James Doss-Gollin\correspondingauthor{James Doss-Gollin. Dept. of Earth and Environmental Engineering, Columbia University, 500 W. 120th St., New York, NY. USA.}}
\affiliation{Dept. of Earth and Environmental Engineering, Columbia University, 500 W. 120th St., New York, NY. USA \\ Columbia Water Center, Columbia University, 500 W. 120th St., New York, NY. USA}
\email{james.doss-gollin@columbia.edu}
\extraauthor{\'Angel G. Mu\~noz}
\extraaffil{Atmospheric and Oceanic Sciences (AOS). Princeton University. Princeton, NJ 08540-6649. USA  \\
	International Research Institute for Climate and Society (IRI). The Earth Institute. Columbia University. Palisades, NY. 10964-1000. USA}
    \extraauthor{Simon J. Mason}
\extraaffil{International Research Institute for Climate and Society (IRI). The Earth Institute. Columbia University. Palisades, NY. 10964-1000. USA}
\extraauthor{Max Past\'{e}n}
\extraaffil{Direcci\'{o}n de Meteorolog\'{i}a e Hidrolog\'{i}a. Paraguay \\
	Facultad Polit\'ecnica. Universidad Nacional de Asunci\'{o}n. Paraguay}

%-------------------------------------------------------------------------------
% Abstract
% Up to 250 words in length
%-------------------------------------------------------------------------------

\abstract{
	During the austral summer 2015-16 severe flooding displaced over 150,000 people on the Paraguay River system in Paraguay, Argentina, and Southern Brazil. This flooding was out of phase with the typical seasonal cycle of the Paraguay River, and was driven by repeated intense rainfall events in the Lower Paraguay River basin. Using a weather typing approach within a diagnostic framework, we show that enhanced moisture inflow from the South American Low-Level Jet, and local convergence associated with baroclinic systems favored the development of mesoscale convective activity and enhanced precipitation. The observed circulation patterns were made more likely by the cross-timescale interactions of multiple climate modes, including a very strong, mature El Ni\~no event and an unusually persistent Madden-Julian Oscillation in phases four and five.
	The rainfall skill is assessed using seasonal forecasts produced by the Latin American Observatory of Climate Events (OLE2), and sub-seasonal forecasts produced by the European Centre for Medium-Range Weather Forecasts. The probabilistic seasonal forecasts accurately predicted above-normal rainfall over the region. Although raw sub-seasonal forecasts exhibited limited skill at lead times beyond the first predicted week (i.e., week 1), it is found that a Model Output Statistics approach involving principal component regression substantially improves skill for week 3, being also better than other bias-correction methods based on canonical correlation analysis and extended logistic regression. Possible implications for flood preparedness are briefly discussed.
}

\begin{document}

%% Necessary!
\maketitle

%-------------------------------------------------------------------------------
% Introduction
%-------------------------------------------------------------------------------

\section{Introduction}

% Why this is worthy of study -- needs some work!
During the austral summer of 2015-16, repeated intense rainfall events led to severe flooding on the Paraguay River, displacing approximately \num{170000} people \citep{BrackenridgeDFO} and generating tremendous property damage.
The Paraguay River (\cref{fig:study-area}) is the primary tributary of the Paran\'{a} River and an important contributor to the La Plata basin, and its flooding directly impacts the Paraguayan cities of Asunci\'{o}n, Concepci\'{o}n, and Pilar, and the Argentine cities of Formosa and Corrientes.
Intense rainfall and flooding along the Paraguay River also has important implications for hydropower generation along the Paran\'{a} River, for agriculture, and for regional water resource management.
Until now, the drivers of the November-February (NDJF) 2015-16 rainfall events and the extent to which they could have been skilfully predicted has not been analyzed in the scientific literature.
The goal of this paper is to address these issues.

% Climatology and past floods
The rainfall climatology of the Paraguay River basin varies strongly by season, with extratropical characteristics in the winter and monsoonal characteristics in the summer.
In the summer season (NDJF), the focus of this study, circulation is dominated by the upper-tropospheric Bolivian High, the lower-level subtropical highs, the Chaco Low over northern Argentina, the South Atlantic Convergence Zone (SACZ), and the South American Low-Level Jet, or SALLJ \citep{Grimm2009,Marengo:2012cm}.
Rainfall in the Lower Paraguay River Basin peaks around \SI{5}{\milli\meter\per\day} during the warm months (October-May) and reaches a minimum near \SI{2}{\milli\meter\per\day} in July and August.
However, the flat topography limits the ability to carry the summer runoff, causing seasonal innundation of the Pantanal and distributing the river discharge in time \citep{Bravo:2011et,Barros:2004bn}.
Thus, the streamflow maxima typically occurs in phase with precipitation upstream of the Pantanal, while downstream of the Pantanal -- over the Lower Paraguay River -- the annual peak, including for extreme flood events, typically occurs between April and July \citep{Barros:2004bn}.

% SALLJ and MCS
A large fraction of rainfall, and nearly all intense rainfall, in the Paraguay River Basin during the summer is associated with mesoscale convection \citep{Velasco1987}.
Previous studies of organized convection and precipitation across subtropical continental South America have found close correspondence to the exit region of the low-level jets \citep{Saulo:2007km,Salio:2007gd,Marengo2004,Velasco1987}, which is influenced in both summer and winter by midlatitude baroclinic wave trains that interact with the Andes topography to generate orographically bound cyclones and northerly low-level flow \citep{Campetella:2002hx,Seluchi:2006bi,Boers:2014gt}.
The strength and direction of this moisture transport varies substantially between events; SALLJ exit regions range from central Argentina \citep[``Chaco Jet Events'';][]{Salio:2002ev} to southeastern Brazil. In addition, the relationship between the low-level jet and convection involves two-way feedback \citep{Saulo:2007km}.

% S2S Drivers
At sub-seasonal timescales, convection in the Paraguay River Basin is modulated by a variety of drivers, notably including the SACZ and the Madden-Julien Oscillation (MJO).
During SACZ (NoSACZ) conditions, strong (weak) convergence is observed over the Amazon basin with divergence (convergence) over southwestern Brazil, northern Argentina and Paraguay \citep{Herdies:2002jy,Carvalho2010}.
SACZ occurrence is related to westerly wind regimes, as well as ``active'' and ``break'' periods of the South American Monsoon \citep{Marengo2004}.
\citet{Carvalho2004} and \citet{Jones2002} demonstrate that mid-latitude dynamics influence low-level wind anomalies on many time scales, but also find that tropical-extratropical interactions complicate attempts to find one-way casual relationships.

% Seasonal Drivers
At seasonal timescales, El Ni\~no-Southern Oscillation (ENSO) is the dominant driver of convection variability in the Paraguay River Basin.
During El Ni\~no years, a low-level anticyclonic anomaly over central Brazil enhances occurrence of the low-level jet, favoring the development of mesoscale convective systems \citep{Velasco1987}.
The impact of ENSO events is felt partly by changes in the SACZ, which change due to ENSO-induced extratropical Rossby wave trains \citep{Grimm:2011vp}.
There is also intraseasonal variability of rainfall during El Ni\~no years, typically including a reversal between November and January of the following year, which are likely dominated by land-surface interactions \citep[see][]{Grimm2009,Grimm2003}.
These findings suggest that while interannual variability during the spring may be driven by remote processes,  during the summer local processes dominate.
In particular, regional land-surface feedbacks can cause regions that exhibit wet anomalies in the spring to experience, on average, more summer precipitation \citep{Grimm:2009bq}.

The paper continues as follows.
In \cref{sec:data} we describe our data sources, while in \cref{sec:methods} we provide an overview of technical methods carried out.
In \cref{sec:diag} we diagnose the circulation patterns associated with intense rainfall during NDJF 2015-16, using composite and weather typing approaches.
In \cref{sec:fcsts} we examine forecasts of rainfall in the Lower Paraguay River Basin and assess forecast skill, with special attention to the first week of December, when the most important flooding events started.
In \cref{sec:discussion} we discuss potential implications of our findings and potential future work.
Finally, in \cref{sec:summary} we summarize our results.

%-------------------------------------------------------------------------------
% DATA
%-------------------------------------------------------------------------------

\section{Data} \label{sec:data}

The analysis involves observations and model forecasts. The period analyzed for diagnostics purposes is from Nov 1979 through Feb 2016. Forecasts consider the period starting in Dec 2015 until Mar 2016, paying special attention to the week of Dec 1-7, 2015, as indicated in the Introduction. Hindcasts to assess the real-time predictive skill consider the period Dec 1-7, 1995-2014. All codes used to access the data and to create the plots and results described in this paper are available at \url{https://github.com/jdossgollin}.

\subsection{Observations}
Rainfall data is taken from the CPC Unified Gauge-based Analysis of Global Daily Precipitation dataset, available online at \url{http://iridl.ldeo.columbia.edu/SOURCES/.NOAA/.NCEP/.CPC/.UNIFIED_PRCP/.GAUGE_BASED/.GLOBAL/.v1p0/}.
%Extreme rainfall events are computed as those exceeding the 90th percentile in the observations.
Spatial resolution is \SI{0.5}{\degree}, and temporal resolution is daily.

Atmospheric circulations are diagnosed using daily data from the NCAR-NCEP Reanalysis I data set \citep{Kalnay1996}, and are referred to as observations, available online at \url{https://www.esrl.noaa.gov/psd/thredds/dodsC/Datasets/ncep.reanalysis/}.
Spatial resolution is \SI{2.5}{\degree}, which is adequate for synoptic-scale analyses.
Because the end-of-day time for the rainfall data is 12:00GMT over most of South America, we use six-hour reanalysis data, and shift by twelve hours before re-sampling to the daily time step.
This ensures that the time steps in the reanalysis and rainfall data sets are the same, but means that a day is defined as beginning at 12:00 GMT.
Since most summer rainfall in this region occurs overnight, this end-of-day time (which translates to approximately 8:00 AM locally depending on the exact time zone) tends to separate distinct events.

Streamflow station data was collected by the Paraguayan Navy and National Administration of Navigation and Ports of Paraguay, and was processed and distributed by the Paraguayan Directorate of Meteorology and Hydrology (DINAC-DMH).
Locations of streamflow gauges are shown in \cref{fig:study-area}.

\subsection{Model forecasts}
This study analyzes probabilistic seasonal and sub-seasonal forecasts of extremely wet rainfall events.
The seasonal predictions correspond to the ``flexible format" forecasts provided by the International Research Institute for Climate and Society (IRI).
This product provides the probability of exceedance (or non-exceedance) of a user-specified percentile of the climatological distribution.
The DJF 2015-2016 forecasts analyzed were produced in November 2015.
Due to the short record of flexible format forecasts (i.e., 2012-2016), no verification was performed for seasonal predictions.
These forecasts are provided at a horizontal resolution of \SI{2.5}{\degree}.

Sub-seasonal forecasts used in this research correspond to those of the European Centre for Medium-Range Weather Forecasts (ECMWF) coupled model, available via the Subseasonal-to-Seasonal Prediction Project Database \citep{Vitart2016}.
The horizontal resolution is \SI{1.5}{\degree} deg.
The rainfall forecasts and hindcasts (or re-forecasts) considered were initialized on November 12th and 16th, 2015, targeting the first week of December of 2015 (defined as Dec 1-7).
There is a total of 51 ensemble members for each forecast, and 11 ensemble members for each one of the 20 hindcasts (Dec 1-7, 1995-2014).
The hindcasts were used to define the extreme event threshold, and for probabilistic forecast verification.
% NEED TO STATE HERE THAT ACTUALLY ALL FORECASTS WERE USED FOR THE `CHICLET' DIAGRAM

%-------------------------------------------------------------------------------
% METHODS
%-------------------------------------------------------------------------------

\section{Methods} \label{sec:methods}

Different types of analyses are used to diagnose the causes of the extreme rainfall events, and to bias-correct and verify the forecasts.
These are summarized in the following sub-sections.

\subsection{Cluster analysis}

A cluster algorithm is used on daily data to help diagnose the mechanisms associated with the rainfall events of interest in this research.
The $k$-means technique is a partitioning method that classifies all days in study into a predefined number of clusters.
It minimizes the intracluster sum of variance, $\mathcal{W}$, for a partition $\mathcal{P}$ defined as the Voronoi diagram generated by the classification process:
\begin{equation}
	\mathcal{W} \qty( \mathcal{P} ) = \sum_{i}^{k}\sum_{X\in p_{i}} \dd s^{2} \qty(X,\bar{p}_{i})
\end{equation}
with $i=1...k$,  $k$ the total number of classes in the solution, $p_i$ the different clusters in the partition and $\dd s^2(X,\bar{p}_i)$ the Euclidean distance between the cluster's maps $X$ and centroid $\bar{p}_i$.
The algorithm is guaranteed to converge to at least a local minimum; to select the best partition, 500 simulations were created using the implementation in Python's \texttt{scikit-learn} package \citep{Pedregosa:2012tv}, and the classifiability index of \citet{Michelangeli1995} was used to select the optimal partition.
Calculation of the classifiability index for several values of $k$ suggests that states with $k=2,4,6$ are all reasonable; we chose the solution $k=6$ because the weather types identified are qualitatively similar to those determined over southeastern South America \citep{Munoz2015,Munoz2016} and have an intuitive physical meaning, discussed further in section \ref{sec:diag}.

This procedure assigns only one cluster to each day on record.
Daily NDJF geopotential data over the box spanning \SIrange{15}{30}{\degree S} and \SIrange{65}{45}{\degree W} (\cref{fig:study-area}) was first projected onto its four leading empirical orthogonal functions, accounting for 95\% of the total observed variance.
No additional time filtering was applied to the data before clustering, thus retaining the annual seasonal cycle, inter-annual, sub-seasonal and synoptic weather time scales for diagnosis.
No latitudinal weighting was necessary as the selected domain is relatively small and it does not extevernd into high latitudes.
The time series were scaled to unit variance prior to running the clustering algorithm.

\subsection{Forecasts and Model Output Statistics (MOS)}

Since no model is perfect, several methods --generically known as Model Output Statistics \citep[MOS;][]{Glahn:1972vt} -- have been developed to correct for different kind of biases.
In this work, we analyze how well the rainfall events could have been predicted, both using the raw sub-seasonal forecasts and bias-corrected seasonal and sub-seasonal forecasts.

The IRI's seasonal predictions are already provided with bias-correction \citep{Barnston:2010ge}, and hence we did not perform any further MOS on seasonal rainfall fields.

We use three different Model Output Statistics strategies in this study to correct for biases in sub-seasonal forecasts.
The approach to perform the corrections is based on both the homoscedastic and heteroscedastic versions of the extended logistic regression (XLR and HXLR, respectively), principal component regression (PCR), and canonical correlation analysis (CCA).
These different regression models are deemed here to be of increasing complexity, with the logistic regression models being relatively simple and the CCA one the more complex approach. More details and references on each model are given below.

The original logistic regression models the conditional probability of binary events, and it has been used in MOS approaches for binary predictands \citep{Hamill:2004hk}.
Nonetheless, when using the logistic regression to address multiple thresholds via independent fits, the predicted probabilities are in general not mutually consistent \citep{Messner:2014gp}.
The (homoscedastic) XLR was designed to solve this problem \citep{Wilks:2009bk}, via the consideration of a transformation of the thresholds of interest as an additional predictor variable.
The HXLR, a heteroscedastic generalization of the XLR, was proposed to appropriately use the ensemble spread as predictor for the dispersion of the predictive distribution \citep{Messner:2014gp}.
In this study we explore the potential benefits and disadvantages of both versions of the extended logistic regression.

In PCR, a special case of CCA, each grid box in the predictand field is regressed using a linear combination of the predictor's EOFs \citep{Mason:2008da,Wilks:2006fx}.

CCA is a common statistical method frequently used to forecast rainfall using a purely empirical approach \citep{Mason:2008da,Barnston:2012ce}.
In brief, CCA \citep{Barnston:1992gd,Wilks:2006fx} identifies modes of co-variability, called canonical variates or canonical modes, by maximizing the correlation between linear combinations of the predictor and predictand's empirical orthogonal functions (EOFs).
The method forecasts spatial patterns of variability spanning across the region of interest rather than making forecasts for individual locations.
Since the canonical modes describe the preferred coupled spatial patterns relating predictors and predictands, CCA has the potential to correct the model rainfall spatial patterns using statistical relationships involving the observed patterns.

Because CCA and PCR involve regressions of spatial modes, their use addresses bias in both the magnitude and the spatial distribution of the modeled precipitation patterns. Conversely, because XLR and HXLR are applied uniquely to each grid cell, these procedures correct only for magnitude biases in the probability distribution.

For the purposes of MOS corrections, the predictand (or variable to forecast) is the observed rainfall for the target period of interest, and the predictor (or variable to be corrected) is the raw model rainfall for the same period.
We use the same spatial domain [\SI{39}{\degree S}-\SI{17}{\degree S}; \SI{66}{\degree W}-\SI{49}{\degree W}] for both the predictor and the predictand, except for the PCR and CCA cases, in which a larger domain [\SI{0}{\degree S}-\SI{60}{\degree S}; \SI{80}{\degree W}-\SI{30}{\degree W}] was used to better capture the spatial patterns in the raw model field.

To avoid artificial skill when building the statistical models, we employ a cross-validation approach that uses a 5-year-out window: five continuous years are left out of the record, the regression coefficients are computed with the remaining of the time series, and the resulting model is validated comparing the prediction for the third year left out (middle of the window) against observations.
The 5-year-long window is redefined a year at a time, moving from the beginning of the record to its end.

Due to the nature of each model used for the MOS-correction, different analyses were explored to define the best predictor(s) for each case.
For example, for the PCR and CCA models, predictors were defined using different geographical domains and using (a) an ensemble of the members available in both initialization times (Nov 12th and 16th, 2015), and (b) computing the ensemble mean for each one of these same initializations and treating them as independent predictors.
Since the skill was shown to be higher in option (b), this was selected as the final predictor (for the spatial domain indicated in the previous paragraph).
A summary of the final selection for each model predictors is presented in Table \ref{t1}.

Probabilistic forecasts are normally used to convey information about the uncertainties associated to the prediction of interest.
It is a common practice to present these forecasts as probabilities for occurrence of below-normal, normal and above-normal precipitation, where the normal is defined as the long-term average of at least 20 years.
Here we use the period 1981-2010 for seasonal forecasts, and 1995-2014 for sub-seasonal ones.
Nonetheless, since we analyze extreme event predictions (90th percentile exceedances), rather than terciles.
Furthermore, we present all predictions in terms of odds relative to the climatological odds:
\begin{equation}
odds_{r} \equiv \frac{\frac{p}{1-p}}{\frac{p_c}{1-p_c}}
\end{equation}
where $p$ and $p_c$ represent the forecast probability and the climatological probability, respectively.
This choice helps to easily visualize in the forecast maps locations where the odds of occurring an extreme rainfall event are considerably higher than expected.
In contrast, using probabilities in these forecasts requires that the reader compares values against the climatological probability.
Our choice thus simplifies the comparison in a way that  also tends to be better understood by decision makers.

\subsection{Probabilistic forecast verification}

In addition to visually comparing predictions and observations to verify how well the extreme events could have been predicted, we also provide some context to the analysis of the sub-seasonal skill conducting a formal cross-validated verification.
This is not performed for the seasonal predictions due to the prohibitively short record available in that case (see section \ref{sec:data}).

The Generalized Relative Operating Characteristics (GROC), also called 2AFC score \citep{Mason:2009kr}, is used to evaluate skill of probabilistic rainfall forecasts at a local scale, via spatial maps.
This score measures the ``proportion of all available pairs of observations of differing category whose probability forecasts are discriminated in the correct direction'' \citep{Mason:2009kr}.
It has an intuitive interpretation as an indication of how often the forecasts are correct.
To conduct the verification in a consistent manner, we use the Climate Predictibility Tool (CPT), developed by the International Research Institute for Climate and Society \citep{Mason:2017gg}.

%-------------------------------------------------------------------------------
% DIAGNOSTICS
%-------------------------------------------------------------------------------

\section{Atmospheric circulation and intense rainfall} \label{sec:diag}

This section discusses the synoptic circulation patterns related to the intense rainfall events observed during the austral summer of 2015-2016. To provide context to the analysis of this particular season, we first characterize recurrent daily circulation types (or weather types) over the region under study, and discuss the average synoptic patterns conducive to intense rainfall events there.

\subsection{Weather Type Analysis: Daily Circulation Patterns} \label{sec:weather-types}

Visual inspection of the rainfall and circulation anomalies composited on each weather type (\cref{fig:wt-composite}) shows that although they were constructed from the leading EOFs of \SI{850}{\hecto\pascal} geopotential height over a relatively small region (see \cref{fig:study-area}), the circulation patterns are associated with regional-scale wave trains.

The \SI{850}{\hecto\pascal} geopotential height anomalies associated with the weather types resemble the first three EOFS of the field, and weather types 2, 4, and 6 are nearly the inverse of weather types 1, 3, and 5, respectively.
The weather types also resemble those found by \citet{Munoz2015} for DJF using a similar approach over an area further to the southeast, though the relative frequencies are quite different.
This is likely due to the fact that the in the present research we use NDJF instead of DJF, and the periods considered are not the same.
The rainfall anomalies associated with these weather types are also consistent with the leading modes of spring/summer rainfall variability identified by other authors \citep{Grimm:2009bq}.

Weather types, 1, 3, and 6 all show enhanced low-level jet activity, while weather types 2, 4, and 5 all show a weakening or reversal of the climatological low-level wind.
As discussed by previous studies \citep{Saulo:2007km,Salio:2007gd,Marengo2004,Velasco1987}, there is a close correspondence between jet exit regions and the regions of highest rainfall.
For example, the low-level jet for weather type 3 has a very small zonal component, while for weather type 6 the zonal component is much greater.
Correspondingly, the maximum rainfall intensities for weather type 3 are farther to the west than for weather types 1, which is still further to the west than for weather type 6.

Analysis of the relative frequency of different weather types across years (not shown) reveals strong inter-annual variability.
This is consistent with the findings of \citet{Munoz2015,Munoz2016}, which identify strong modulation of weather type occurrence at sub-seasonal to seasonal time scales.

\subsection{Mean Circulations Associated with Paraguay River Basin Intense Rainfall} \label{sec:rainfall-circulation}

To understand how circulation anomalies observed during NDJF 2015-2016 led to the observed floods it is helpful to first explore the atmospheric circulations which climatologically are associated with intense rainfall in the lower Paraguay River.

\Cref{fig:lagged-rain} shows time-lagged anomalies up to and after days exceeding the 90th percentile of area-averaged daily rainfall in the lower Paraguay River basin (see \cref{fig:study-area}), and is consistent with the findings of \citet{Marengo2004,Salio:2007gd}.
\Cref{fig:lagged-rain} suggests that the key mechanism involves two components.
The first, at $t=-2$ and $t=-1$, is a northerly inflow of heat and moisture via a low-level jet, which may be favored by a polar trough extending slightly to the east of Paraguay, consistent with \SI{850}{\hecto\pascal} circulation shown in weather types 1, 3, or 6.
Previous studies have demonstrated that this jet is also associated with strong vertical wind shear \citep{Marengo2004,Silva2009}, favoring instability and intense convection.
The second factor is low-level convergence (\cref{fig:lagged-rain} top row) and upper-level divergence (\cref{fig:lagged-rain} middle row) associated with the eastward progression of the frontal system.
While \cref{fig:lagged-rain} shows composites of winds anomalies, analysis of particular events reveals that the anomalies are often, though not always, sufficient to fully reverse the direction of climatological wind.
Similar analysis for the 85th or 95th percentiles of daily rainfall (not shown) yield essentially identical results, implying that the mechanism for the most extreme events is not fundamentally distinct from the mechanism for moderate-intensity events.

\texttt{RE: need to identify what makes it rain over Paraguay rather than farther South: agreed but we should consider what analysis is required for this step.}

\subsection{Circulation Patterns Observed during NDJF 2015-16}

An important question is whether the circulations that drove intense rainfall of NDJF 2015-16 differed from the circulations described above which typically drive intense rainfall.

While the weather typing approach requires simplifying the dynamics of daily circulation patterns, its advantage is that it greatly facilitates the analysis of sequences of precipitation.
A Klee diagram \citep{Munoz2015, Munoz2016} is used to visualize these sequences, which are presented in \cref{fig:klee} along with rainfall time series for the season of interest which shows that intense rainfall concentrated over a period spanning from mid-November 2015 through early January 2016, with smaller peaks in late January and mid-February.
As discussed in \cref{sec:rainfall-circulation}, intense rainfall is typically preceded by a northerly, low-level influx of energy and moisture.
Examination of \cref{fig:klee} reveals that while the most intense rainfall occurred during many different weather types, weather type 1 immediately precedes nearly almost all of the most intense rainfall peaks.
Thus, while the intensity and persistence of intense rainfall was atypical, the causal mechanism of the intense rainfall observed during this season was consistent with climatology.

At the monthly scale (\cref{fig:anomalies}), the largest observed rainfall anomalies overlap with the direction, intensity, and exit region of the low-level jet apparent in the \SI{850}{\hecto\pascal} field.
During all months, except January 2016, the \SI{850}{\hecto\pascal} wind field indicates enhancement of the SALLJ above approximately \SI{20}{\degree S} but does not penetrate further south.
In the upper troposphere, enhanced convergence over the tropical Atlantic just off the coast of Northeast Brazil and enhanced divergence over Paraguay, Uruguay, and Northern Argentina coincides with the enhanced SALLJ activity.
The latter feature in particular is enhanced during November and December 2015 and February 2016, but weakened during January 2016.
\texttt{Add more here James}

\section{Forecasts} \label{sec:fcsts}

In this section we analyze the extent to which forecasts were able to predict the persistent rainfall during summer of 2015-16. There are advantages in simultaneously considering useful climate information at multiple timescales, rather than just focusing on one of them. This multistage prediction system is known as the IRI's ``Ready-Set-Go" approach \citep{Goddard:2014kf,Munoz2016}.
For the sake of concision, in this section we discuss only probabilistic seasonal (DJF 2015-2016) and sub-seasonal (Dec 1-7, 2015) forecasts, and also an integrated subseasonal-to-seasonal outlook.

\subsection{Seasonal Forecast}

Extreme rainfall over the region was accurately predicted by the IRI for the DJF 2015-2016 season since at least November 2015 (see \Cref{fig:probfcst}). Probabilities of 30\%-50\% for the 90th percentile, three to five times the expected climatological values, are visible over southern Paraguay and Brazil, and northern Uruguay and Argentina, mostly in agreement with observations (hashed region in \cref{fig:probfcst}).
Although this seasonal prediction was not skillful over eastern Paraguay, the forecast  showed a clear signal in the neighborhood, that could have been used for disaster-preparedness weeks in advance of the events.

\subsection{Sub-seasonal Forecasts}
Sub-seasonal predictions are still too new to be used as operational tools, and their skill is normally not high enough as to be useful for decision-making [REF recent paper by Andy on s2s]. Nonetheless, the international S2S Prediction Project \citep{Vitart2016} has been providing since 2015 free access to almost-real-time sub-seasonal forecasts from multiple models, an opportunity to explore how well the extreme rainfall events of the first week of December 2015 could have been predicted.

As seen in \Cref{fig:subsprobfcst}a, the raw (uncorrected) sub-seasonal forecast of the ECMF model for Dec 1-7 2015, indicated very high odds for occurrence of extreme rainfall but with important biases in the actual location and spatial pattern; for Paraguay, it confidently suggests occurrence of extremes to the south-southeast of the country, which was mostly not observed. Overall, the 20-year based skill of probabilistic forecasts for the first week of December --even when using multiple initializations-- is worse than using climatology for most of Paraguay (\Cref{fig:subsprobfcst}b), being best for the zone between that country and northern Uruguay, and for most of southern Brazil.

Nonetheless, despite the biases, the model is capturing a signal, which suggests the use of Model Output Statistics to explore the extent to which corrections in the magnitudes and spatial patterns can improve the forecast. To the best of our knowledge, this is the first time this kind of corrections is performed for sub-seasonal forecasts in Latin America and the Caribbean, and probably in other places of the world.

In general, the use of extended logistic regression models does not improve the forecast for the week. For example, with respect to the raw prediction, XLR tends to amplify the relative odds, and to cluster and shift the forecast location of the extremes (\Cref{fig:subsprobfcst}c); the forecast tend to be better for Uruguay, but suggests extreme rainfall in the Paraguayan Chaco, which was not present in the raw prediction. On the other hand, the use of the ensemble spread in the HLXR model does not help: this forecast tends to be very confident (always compared with the uncorrected case) on the extremes occurring in almost all the region of interest (\Cref{fig:subsprobfcst}e).

The probabilistic skill verification for both extended logistic regression models is very similar to the one of the raw forecast ( \Cref{fig:subsprobfcst}b,d,f), varying just in magnitude in a few places. This is attributed to the fact that these models only correct for magnitude, and they do it on a grid-box by grid-box basis, so the spatial distribution of skill is basically the same. I WROTE AN EMAIL TO SIMON TO ASK ABOUT THIS, AS I'M NOT CONVINCED.

Better forecasts are obtained when both magnitude and spatial corrections are performed, although with relative odds considerably less confident than the ones in the raw forecast. The Principal Component Regression model correctly shows high relative odds in most of the places where extreme rainfall was observed (\Cref{fig:subsprobfcst}g), although it also erroneously forecasts extremes in other locations, like zones of western Paraguay and northeastern Argentina. On the other hand, the main problem with the Canonical Correlation Analysis model is its lack of discrimination between occurrence or non-occurrence of extremes in the region: the spatial distribution of odds is too homogeneous (\Cref{fig:subsprobfcst}i). Still, it tends to exhibit slightly higher odds than the spatial average in regions where extremes were observed, and it forecast similar false alarms than the PCR model.

The 20-year based skill maps of probabilistic forecasts (see \Cref{fig:subsprobfcst}h,j) computed with these models are very similar to each other, although the one for PCR is definitively better and shows high skill for almost all the region of interest, although there are still zones along the eastern part of Paraguay exhibiting equal or lower skill than climatological values. Hence, despite of the particular errors in the Dec 1-7 2015 forecasts, on the long term both PCR and CCA verify considerably better than the raw, XLR and HXLR predictions.

\subsection{Subseasonal-to-seasonal outlook}
Under certain conditions, like in particular combinations or entanglement of ENSO and MJO phases, the skill of climate models tend to be higher than in other occasions \citep{Vitart2016,Moron2015,Munoz2016} These ``windows of opportunity'' can be understood as the result of pre-conditioning of the space state of the system, making certain trajectories in that space more frequently visited than when the pre-conditioning is not present, i.e., when the climate modes are not phase-locked.

%Following these ideas, in a recent work for southeastern South America \citet{Munoz2016} discussed a new probabilistic forecast approach for subseasonal-to-seasonal scenarios based on classification of sequences of weather types.  that are very similar to the ones obtained in the present research.

I'M NOT SURE WHAT NEW MESSAGE WE WANT TO CONVEY WITH THIS FIGURE AND ANALYSIS. MAYBE THERE'S SOMETHIG WE CAN RESCUE FROM THE ORIGINAL CHICLET DIAGRAM OF MIKE TIPPET?
\Cref{fig:chiclet} presents the ensemble spread and mean, as compared to observations, at several lead times. Because the models are initialized twice-weekly, we smooth the rainfall field with a 3.5-day running mean (past three days weighted equally, fourth day before weighted one-half) in order to more easily compare observation with model outputs.
\Cref{fig:chiclet} indicates that on weather timescales (\SI{2}{\day}) and even on synoptic timescales (\SI{10}{\day}), the ensemble mean skillfully predicts the timing and the amplitude of the area-averaged rainfall.
However, at longer sub-seasonal (\SIlist{25;40}{\day}) timescales, the ensemble mean shows little deviation from climatology, and so did not provide meaningful flood risk guidance for this region.

MAYBE I NEED TO USE THE METHOD DEVELOPED IN MUNOZ ET AL 2016. OR WE SHOULD EXCISE THIS SUBSECTION.
%-------------------------------------------------------------------------------
% DISCUSSION
%-------------------------------------------------------------------------------

\section{Discussion}
\label{sec:discussion}

I SUGGEST TO START WITH (OR TO HAVE IT CLOSE TO THE START) RAINFALL STREAMFLOW ANALYSIS AND THE CORRESPONDING FIGURE.

CLEARLY ADDRESS THE QUESTIONS:
1. WHAT CAUSED THE RAINFALL? SPECIFIC CIRCULATION PATTERNS.
2. WHY THESE CIRCULATION PATTERNS? ENTANGLEMENT OF ENSO AND MJO
3. COULD WE HAVE PREDICTED THAT? AT DIFFERENT TIMESCALES.

Up to this point I think that we do a good job presenting what happened, and how the circulation anomalies led to rainfall.
I think the first piece of the discussion should focus on \emph{why} the observed circulation anomalies occurred.
\begin{itemize}
	\item I think this is the appropriate place to discuss ENSO and MJO teleconnections --the interaction of the two is very consistent with observed rainfall and circulation patterns
	\item Compared to other El Ni\~{n}o years, there were a lot of similarities, but also some very important differences including increased upper-level divergence over Paraguay, and
	\item \citet{Grimm2009} argues that land-surface interactions are important for the lifecycle of ENSO's impact on the region.
	Since the 2015-16 El Ni\~no was super strong, and 2014-15 was a weak EN year, it seems likely that these land-surface interactions and land-ocean temperature gradients were different amongst EN years historically; we don't know that this *is* the mechanism, but it would be interesting to explore further. MENTION THIS BUT I SUGGEST NOT TO FURTHER EXPLORE IT HERE. WE HAVE ENOUGH.
	\item By examining \cref{fig:klee}, we can see that some weather types tend be more likely to occur during specific parts of the 120-day season; this is consistent with \citet{Munoz2016} and with the notion that there is some structure to the nonlinear dynamical system, which could support the land-surface feedbacks that are proposed by i.e. \citep{Grimm:2009bq}.
\end{itemize}

The second piece we need to discuss a bit more is the lack of model skill at S2S timescales.
I think we would benefit from a comparison of other case studies with this dataset, and with literature on model skill in this region.
%\improvement[inline]{\'{A}ngel to take lead}
NO PREVIOUS SKILL ANALYSISI IN LITERATURE FOR THE REGION OR SOUTH AMERICA. I THINK THAT WITH THE DETAILED WORK ON MOS FOR S2S THAT WE CONDUCTED THIS IS IN BETTER SHAPE.

Finally, I think we would also benefit from discussing more clearly the relationship between rainfall and streamflow, and potential implications for management and planning.
There are a bunch of relevant citations in \texttt{delted.tex}.
\begin{itemize}
	\item The relationship between streamflow and rainfall is complicated in this region. A couple of things that make it particularly interesting and complex are: (a) the wetlands of the Pantanal, which act as water storage units \citep{Bravo:2011et,Barros:2004bn}, (b) The region is very flat, so that water moves slowly and thus local flood waves move relatively slowly; (c) There's significant human modification of the Paran\'a River via the hydroelectric sites at Itaipu and Yacyreta.
  \item This is probably the place to mention that \cref{fig:streamflow} suggests multi-decadal oscillation, possibly associated with PDO, differs from changepoint perspective of \citet{Collischonn:2001bi}.
	This is definitely beyond scope of our paper but worth considering in the future, particularly for flood management planning.
	\item If we're going to discuss implications (and I think we should), we should definitely comment on the lack of data on flood impacts -- it is very hard to engage in cost-benefit planning of any sort without this information, and given recent flooding, as well as the possibility of being in a multi-year ``active'' phase of the river, Paraguay can't run from this problem
\end{itemize}

%-------------------------------------------------------------------------------
% Summary
%-------------------------------------------------------------------------------

\section{Summary} \label{sec:summary}

The flooding was made more likley by antecedent soil moisture left by the preceding rainy season, and by several successive wet years; \citet{Santos:2016td} posit that groundwater may be important in this region and this would be consistent. These antecedent conditions combined with the persistent intense rainfall occurring in December to produce heavy flooding on the Lower Paraguay river.

The intense rainfall observed was driven by \emph{persistent} rainfall, associated with enhanced moisture inflow from the low-level jet and convergence associated with baroclinic systems.

This circulation pattern was made more likely by sub-seasonal (MJO), seasonal (ENSO) and possibly even by multi-decadal variability, but the anomalies were unusually strong

%-------------------------------------------------------------------------------
% Acknowledgments
%-------------------------------------------------------------------------------

\acknowledgments
The authors thank funding sources, data providers (DINAC?), etc. S2S Database.
Streamflow data was collected by the Paraguayan Navy and National Administration of Navigation and Ports of Paraguay, and was processed and distributed by the Paraguayan Directorate of Meteorology and Hydrology (DINAC-DMH).

%-------------------------------------------------------------------------------
% REFERENCES
%-------------------------------------------------------------------------------

\bibliographystyle{ametsoc2014}
\bibliography{library}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% TABLES
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table*}[h]
%
\caption{Model Output Statistics (MOS) methods used to correct the ECMF sub-seasonal forecasts. Spatial domain for predictand is always the same (\SI{39}{\degree S}-\SI{17}{\degree S}; \SI{66}{\degree W}-\SI{49}{\degree W}). Two initializations are used: Nov 12th and 16th, 2015.} \label{t1}
\begin{center}
%
\scalebox{0.8}{
\begin{tabular}{ccccccccc}
%
\topline
%
Model & Region (Predictor) & Final predictor(s) selected \\
%
\midline
%
 \emph{Raw} & (\SIrange{39}{17}{\degree S}; \SIrange{66}{49}{\degree W}) & \begin{tabular}{@{}c@{}}Ensemble mean, computed using members from the  \\ two initializations. No correction performed. \end{tabular} \\
%
%
 \emph{XLR} & (\SIrange{39}{17}{\degree S}; \SIrange{66}{49}{\degree W}) & \begin{tabular}{@{}c@{}}Ensemble mean, computed using members from the  \\ two initializations. \end{tabular}  \\
%
 \emph{HLXR} & (\SIrange{39}{17}{\degree S}; \SIrange{66}{49}{\degree W}) & \begin{tabular}{@{}c@{}} Ensemble mean and spread, computed using  \\ members from the two initializations.\end{tabular}  \\
%
  \emph{PCR} & (\SIrange{60}{0}{\degree S}; \SIrange{80}{30}{\degree W}) & \begin{tabular}{@{}c@{}} Linear combination of model's EOFs, computed using \\ both initializations as independent predictors (10 EOFs). \end{tabular} \\
%
  \emph{CCA} & (\SIrange{60}{0}{\degree S}; \SIrange{80}{30}{\degree W}) & \begin{tabular}{@{}c@{}} Canonical modes computed using \\ both initializations as independent predictors. (10 predictor EOFs, \\ 4 predictand EOFs, 4 canonical modes)\end{tabular}  \\
%
\botline
%
\end{tabular}
}
%
\end{center}
%
\end{table*}
%
%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% FIGURES
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{figure*}
	\noindent\includegraphics[width=6.5in]{../_figs/StudyArea.pdf}
	\caption{
		The study area.
		(L): all of South America.
		The weather typing described in \cref{sec:methods,sec:weather-types} was calculated using geopotential height within the area indicated by the red box.
		The ``Lower Paraguay River Basin'', over which rainfall is averaged to create plots in \cref{fig:klee,fig:chiclet} and define intense rainfall days in \cref{fig:lagged-rain}, is shown by the black box.
		(R): The Paraguay River.
		As for (L), the ``Lower Paraguay River Basin'' over which rainfall is averaged is indicated.
		Streamflows shown in \cref{fig:streamflow} were taken from Asunci\'on, Paraguay, shown by the black dot.
		The Paraguay River and its tributaries, from the Natural Earth database (\texttt{www.naturalearthdata.com}), are also shown.
	}
  \label{fig:study-area}
\end{figure*}

\begin{figure*}
	\noindent\includegraphics[width=6.5in,height=9in,keepaspectratio=true]{../_figs/WTComposite.pdf}
	\caption{
		Circulation anomalies associated with each weather type.
    Weather types are separated by column.
		Top row gives anomalies of \SI{200}{\hecto\pascal} wind; arrows indicate the direction, and both shaded contours and arrow length indicate the magnitude of the wind anomaly.
    Second row shows circulation at \SI{850}{\hecto\pascal}; shaded contours give anomalies of geopotential height and arrows indicate the wind speed and direction.
    Bottom row shows monthly rainfall anomalies in units of \si{\milli\meter\per\day}.
	}
	\label{fig:wt-composite}
\end{figure*}

\begin{figure*}	\noindent\includegraphics[width=6.5in]{../_figs/LaggedRain.pdf}
	\caption{
		Circulation anomalies associated with intense rainfall (99th percentile exceedance of area-averaged rainfall) in the Lower Paraguay River region. Lagged composites are shown, by column, for $t = $ \SIlist{-2;-1;0;1}{\day} relative to the date of intense rainfall.
		Top row gives anomalies of \SI{200}{\hecto\pascal} wind; arrows indicate the direction, and both shaded contours and arrow length indicate the magnitude of the wind anomaly.
    Second row shows circulation at \SI{850}{\hecto\pascal}; shaded contours give anomalies of geopotential height and arrows indicate the wind speed and direction.
    Bottom row shows rainfall anomalies in units of \si{\milli\meter\per\day}.
	}
  \label{fig:lagged-rain}
\end{figure*}

\begin{figure*}
	\noindent\includegraphics[width=6.5in]{../_figs/RainfallWeatherType}
	\caption{
		Weather type occurrence during two years.
		Time series of area-averaged rainfall (shown by the box in \cref{fig:study-area}) and weather type for each day of NDJF  2015-16.
		Dashed lines indicate the climatological 50th, 90th, and 99th percentiles of NDJF area-averaged rain.
	}
  \label{fig:klee}
\end{figure*}

\begin{figure*}
	\noindent\includegraphics[width=6.5in]{../_figs/NDJF201516Anomaly.pdf}
	\caption{
    	Monthly circulation anomalies during NDJF 2015-16.
      Top row gives anomalies of \SI{200}{\hecto\pascal} wind; arrows indicate the direction, and both shaded contours and arrow length indicate the magnitude of the wind anomaly.
      Second row shows circulation at \SI{850}{\hecto\pascal}; shaded contours give anomalies of geopotential height and arrows indicate the wind speed and direction.
      Bottom row shows monthly rainfall anomalies, in unites of \si{\milli\meter\per\day}.
	}
  \label{fig:anomalies}
\end{figure*}

\begin{figure*}
	\noindent\includegraphics[width=6.4in]{../_figs/ForecastSkill.pdf}
	\caption{
		caption
	}
  \label{fig:subsprobfcst}
\end{figure*}

\begin{figure*}
	\noindent\includegraphics[width=6.5in]{../_figs/Chiclet.pdf}
	\caption{
    Assessment of S2S model forecasts.
    Forecast time series from the ECMWF S2S model are shown for several lead times (\SIlist{2;10;25;40}{\day}) of each individual ensemble member (gray), the ensemble mean (black, dashed), and the observed (blue) rainfall.
		The observed rainfall is smoothed with a \SI{3.5}{\day} running mean in order to facilitate comparison between daily observations and the models, which are initialized every 3-4 days.
	}
  \label{fig:chiclet}
\end{figure*}

\begin{figure*}
	\noindent\includegraphics[width=6.5in]{Streamflow.pdf}
	\caption{
		River stage (height; in \si{\meter}) for the Paraguay River at four gauges along the Paraguay River.
		(a) Seasonality (black) and time series of 2015-16 observations (orange) at each stream gauge.
			The seasonality was fit using local polynomial regression as implemented in the \texttt{locfit} package in
			the \textbf{R} statistical programming environment \citep{Loader1999}.
		(b) Time series of daily stage measurements from 1929 to 2016.
	}
  \label{fig:streamflow}
\end{figure*}
\end{document}

### 2017-09-18 01Get/01-Reanalysis.py

@@ -1,0 +1,60 @@
+'''
+Read in reanalysis data from NCEP-NCAR Reanalysis I data set.
+
+The approach to data management here is to try to minimize the number of times data needs to be re-downloaded (trial and error is a tough teacher...).
+Consequently, we first download a relatively complete data set (all years in our range, and all levels in our range).
+That is implemented in this notebook.
+Later, we calculate anomalies and climatologies for a subset of the data -- this is very quick and if we change our spatial or temporal domain, we don't need to re-download anything.
+
+Now, for each level, we will run through each year from `syear` to `eyear`.
+Then, for each year, we will:
+
+ 1. Download the six-hour zonal ($u$) wind for that level and year and save to file
+ 2. Download the six-hour meridional ($v$) wind for that level ande year and save to file
+ 3. Calculate the six-hour streamfunction ($\psi$) for that year and save to file
+
+The `overwrite` parameter specified above determines whether the data is downloaded even if the file is already on disk -- this can be useful for downloading a bit at a time.
+'''
+
+import numpy as np
+import xarray as xr
+import os
+from paraguayfloodspy.pars  import GetPars
+from windspharm.xarray import VectorWind
+
+def main():
+    print("\nWelcome to Reanalysis.py\n")
+    # Define the Parameters
+    time_pars = GetPars('time')
+    syear,eyear = time_pars['syear'], time_pars['eyear']
+    levels = [850]
+    years = np.arange(syear, eyear+1)
+    overwrite = False # overwrite existing data?
+
+    # Loop through and download each required file
+    for level_i in levels:
+        print("Now working on {}hPa data".format(level_i))
+        for year_i in years:
+            # Define URL and Output File Names
+            url_u = 'https://www.esrl.noaa.gov/psd/thredds/dodsC/Datasets/ncep.reanalysis/pressure/uwnd.{}.nc'.format(year_i)
+            url_v = 'https://www.esrl.noaa.gov/psd/thredds/dodsC/Datasets/ncep.reanalysis/pressure/vwnd.{}.nc'.format(year_i)
+            fname_u = '_data/reanalysis/raw/uwnd_{}_{}.nc'.format(level_i, year_i)
+            fname_v = '_data/reanalysis/raw/vwnd_{}_{}.nc'.format(level_i, year_i)
+            fname_streamfunc = '_data/reanalysis/raw/streamfunc_{}_{}.nc'.format(level_i, year_i)
[…]

### 2017-09-18 01Get/02-Rainfall.py

@@ -1,0 +1,67 @@
+'''
+Here we download Get the CPC data from the IRI Data Library.
+We only get rainfall estimate -- neglects information on number of stations used in analysis.
+Please see caveats and documentation on IRI Server at http://iridl.ldeo.columbia.edu/SOURCES/.NOAA/.NCEP/.CPC/.UNIFIED_PRCP/.GAUGE_BASED/.GLOBAL/.v1p0/
+'''
+
+import pandas as pd
+import numpy as np
+import xarray as xr
+import datetime
+import os
+from paraguayfloodspy.pars  import GetPars
+
+# Work with Times from the IRIDL
+def convert_t_to_time(Tvec):
+    times = np.array([datetime.date(1960, 1, 1) + datetime.timedelta(np.int(ti)) for ti in Tvec])
+    return(np.array(times))
+def convert_time_to_t(date):
+    date_diff = date - datetime.date(1960,1,1)
+    T = date_diff.days
+    return(T)
+
+def IRICPCYear(year, verbose=True):
+        if verbose:
+            print('Downloading data for {}...'.format(year))
+        # realtime or retro?
+        if year >= 1979 and year <= 2005:
+            url = 'http://iridl.ldeo.columbia.edu/SOURCES/.NOAA/.NCEP/.CPC/.UNIFIED_PRCP/.GAUGE_BASED/.GLOBAL/.v1p0/.RETRO/.rain/dods'
+        elif year >= 2006 and year <= 2016:
+            url = 'http://iridl.ldeo.columbia.edu/SOURCES/.NOAA/.NCEP/.CPC/.UNIFIED_PRCP/.GAUGE_BASED/.GLOBAL/.v1p0/.REALTIME/.rain/dods'
+        else:
+            raise ValueError('You have entered an invalid year. {} is outside range [1979, 2016]'.format(year))
+
+        # get the data
+        Tstart = convert_time_to_t(datetime.date(year, 1, 1))
+        Tend = convert_time_to_t(datetime.date(year, 12, 31))
+        ds = xr.open_dataarray(url, decode_times=False)
+        ds = ds.sel(T = slice(Tstart, Tend))
+        ds.load() # force it to download
+
+        # convert to more standard format
+        ds = ds.rename({'X': 'lon', 'Y': 'lat', 'T': 'time'})
+        ds['time'] = convert_t_to_time(ds['time'])
+        ds['time'] = ds['time'].astype('datetime64')
+        return(ds)
+
+def GetFileName(year):
[…]

### 2017-09-18 01Get/04-DailyIndices.py

@@ -1,0 +1,34 @@
+'''
+Read in daily indices (here just MJO) and save to file
+'''
+
+import pandas as pd
+from paraguayfloodspy.pars  import GetPars
+
+def main():
+    print("\nWelcome to Daily-Indices.py")
+    time_pars = GetPars('time')
+    syear,eyear = time_pars['syear'], time_pars['eyear']
+    outfile = "_data/indices/daily_indices.csv"
+
+    raw_url = 'http://www.bom.gov.au/climate/mjo/graphics/rmm.74toRealtime.txt'
+    col_names = ['year', 'month', 'day', 'RMM1', 'RMM2', 'phase', 'amplitude', 'source']
+
+
+    # Now read in the data
+    df = pd.read_table(raw_url, delim_whitespace=True,index_col=None, skiprows=2, names=col_names)
+
+    # Convert the year-month-day structure to a standard time object and select only relevant columns
+    df['time'] = pd.to_datetime(df[['year', 'month', 'day']])
+    df.index = df['time']
+    df = df[['RMM1', 'RMM2', 'phase', 'amplitude']]
+    df = df['{}'.format(syear):'{}'.format(eyear)]
+
+    # Check
+    df.head(10)
+
+    # Save to file
+    df.to_csv(outfile)
+
+if __name__ == '__main__':
+    main()

### 2017-09-18 01Get/06-Anomalies.py

@@ -1,0 +1,98 @@
+'''
+Subsetting, Anomalies, and Climatologies
+
+In a previous notebook (`Reanalysis`) we created a relatively large database of raw $U,V,\psi$.
+This is helpful, but for our visualizations it's nice to sub-set the data in advance so that all we have to do is load the relevant bits.
+We'll also pre-compute anomalies and climatologies, which saves substantial computational time down the road at the expense of a bit of file storage.
+Because we don't need the fine time resolution, we'll also re-sample everything to a daily time step (note that we didn't just download daily reanalysis data because we need to shift the end-of-day time!)
+'''
+
+import xarray as xr
+import glob
+import os
+from paraguayfloodspy.xrutil import *
+from paraguayfloodspy.pars import *
+
+def main():
+    print("\nWelcome to Anomalies.py")
+    pars = GetPars('all')
+    syear, eyear = pars['time']['syear'], pars['time']['eyear']
+    overwrite = False
+
+    # 1. REANALYSIS DATA
+    def transfer_fn(var, lonmin, lonmax, latmin, latmax, shift_time_h, return_daily, months):
+        var = var.sel(lon = slice(lonmin, lonmax), lat = slice(latmax, latmin))
+        # shift time by hours, if desired
+        if shift_time_h != 0:
+            time_old = var.time.values
+            time_new = time_old + np.timedelta64(12, 'h')
+            var['time'] = time_new
+        # resample to daily time step, if desired
+        if return_daily:
+            var = var.resample('1D', dim = 'time')
+        # subset months if required
+        if months is not None:
+            var = var.sel(time = np.in1d(var['time.month'], months))
+        return(var)
+
+    trans_lam = lambda ds: transfer_fn(
+        var=ds,
+        lonmin=pars['clim_plot']['lonmin'],
+        lonmax=pars['clim_plot']['lonmax'],
+        latmin=pars['clim_plot']['latmin'],
+        latmax=pars['clim_plot']['latmax'],
+        months=pars['months'],
+        shift_time_h=12, return_daily=True)
+
+    for var in ['uwnd', 'vwnd', 'streamfunc']:
+        for level_i in [850]:
+            fnames = glob.glob('_data/reanalysis/raw/{}_{}_*.nc'.format(var, level_i))
+            raw_fname = "_data/reanalysis/subset/{}_{}_raw.nc".format(var, level_i)
[…]

### 2017-09-18 run.sh

@@ -1,0 +1,7 @@
+# GET RAW DATA
+python 01Get/01-Reanalysis.py
+python 01Get/02-Rainfall.py
+python 01Get/03-S2S-Data.py
+python 01Get/04-DailyIndices.py
+python 01Get/05-MonthlyIndices.py
+python 01Get/06-Anomalies.py

### 2017-09-18 03Writeup/Supplement.tex
Work on writeup

%-------------------------------------------------------------------------------
% Preamble
%-------------------------------------------------------------------------------

\documentclass{article}

%Packages
\usepackage{
	siunitx, % units
	cleveref,% cross-references
	physics, % better physics notation
	url, % for URLs
	amsmath, % for math
	array, % for fixed-width tables
  graphicx, % for images
}

\usepackage[letterpaper]{geometry}
\renewcommand{\thefigure}{S\arabic{figure}}

%-------------------------------------------------------------------------------
% Title and Authors
%-------------------------------------------------------------------------------

\title{Supplemental Information for\\``Extreme rainfall in Paraguay during the 2015-2016 austral summer: causes and subseasonal-to-seasonal predictive skill''}
\author{James Doss-Gollin\and \'{A}ngel G. Mu\~{n}oz \and Simon J. Mason \and Max Past\'{e}n}
\date{\today}

\begin{document}

%% Necessary!
\maketitle

This document contains several supplemental figures described in the main text.
Further supporting information is available in the form of source code and \texttt{jupyter} notebooks at \url{github.com/jdossgollin/PYFloods}.
\listoffigures
\clearpage

\begin{figure}
	\includegraphics[width=\textwidth]{../_figs/Climatology.pdf}
	\caption{
		NDJF monthly circulation climatology over the study domain and from 1980 to 2010.
    First row shows the \SI{850}{\hecto\pascal} streamfunction $\psi$ climatology, and second row shows rainfall climatology.
	}
\end{figure}

\begin{figure}
  \includegraphics[width=\textwidth]{../_figs/PSI_Var_Explained.pdf}
	\caption{
		Variance explained for each EOF loading of the \SI{850}{\hecto\pascal} streamfunction $\psi$.
    Left figure shows the individual variance explained for each EOF and right figure shows cumulative variance explained.
	}
\end{figure}

\begin{figure}
  \includegraphics[width=\textwidth]{../_figs/WT_Classifiability.pdf}
	\caption{
		Classifiability index (see Methods) calculated for several chosen values of $k$, representing the number of weather types used for clustering.
	}
\end{figure}

\begin{figure}
  \includegraphics[width=\textwidth]{../_figs/WT_Occurrence_Proportion.pdf}
	\caption{
		Occurrence fraction of NDJF weather types for all years (blue) and 2015-16 (orange).
	}
\end{figure}

\begin{figure}
  \includegraphics[width=\textwidth]{../_figs/ENSO_EOF_Scatter.pdf}
	\caption{
		Pair plots for three time series: NINO3.4, and EOFs 2 and 3 of the \SI{850}{\hecto\pascal} field over the chosen domain.
    All time series are degraded to monthly time steps.
    Across the diagonal, marginal distributions are shown.
    Straight lines represent linear regression estimates.
	}
\end{figure}

\begin{figure}
  \includegraphics[width=\textwidth]{../_figs/MJO_EOF_Scatter}
	\caption{
		Pair plots for four time series: MJO RMM1 and RMM2, and EOFs 2 and 3 of the \SI{850}{\hecto\pascal} field over the chosen domain.
    All time series are at daily time steps.
    Across the diagonal, marginal distributions are shown.
    Straight lines represent linear regression estimates.
	}
\end{figure}

\begin{figure}
  \includegraphics[width=\textwidth]{../_figs/knn_rainfall_EOF.pdf}
	\caption{
    Expected rainfall over two regions conditional on the (simultaneous) values of EOF2 and EOF3 of the \SI{850}{\hecto\pascal} streamfunction $\psi$ shown in the main text.
    The ``Chaco Jet Extension Region'' is defined as extending from \SIrange{299.25}{302.75}{\degree E} and \SIrange{34.75}{30.75}{\degree S} and the ``Lower Paraguay River Basin'' is defined in the main text.
    Both values represent area-averaged rainfall time series.
    Expected rainfall is calculated by using a nearest-neighbors regression with 100 neighbors.
    Because this methodology performs poorly under extrapolation (where there are few observations), the predicted domain is limited to include only regions with high density of observations.
    }
\end{figure}

\begin{figure}
  \includegraphics[width=\textwidth]{../_figs/ENSO_MJO_EOF_PRCP_Time_Series.pdf}
	\caption{
		Time series of S2S predictors and their influence on evolution of circulation and rainfall.
    Top row shows two predictors shown to influence the circulation: monthly NINO3.4 index and daily MJO RMM1 index.
    Blue dots show daily observations of MJO RMM1, while the blue line shows the MJO RMM1 time series with a 7-day rolling mean applied.
    Bottom row shows the evolution of EOFS 2 and 3 of the \SI{850}{\hecto\pascal} field.
    As for the MJO RMM1, dots show the daily fields and lines show the fields with a 7-day running mean applied.
    Bottom rain shows time series of daily rainfall over the ``Lower Paraguay River Basin'' and ``Chaco Jet Extension Region'' defined above.
	}
\end{figure}

\end{document}

### 2017-09-18 03Writeup/paper.tex
Work on writeup

@@
 
-\documentclass[twocol]{ametsoc}
+\documentclass{ametsoc}
 \journal{mwr}
@@
 	amsmath, % for math
+	array, % for fixed-width tables
 }
-
-% TODO Notes -- DO NOT KEEP FOR FINAL
+\newcolumntype{L}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
 
@@
 
-%\title{Causes and subseasonal-to-seasonal model skill of the persistent intense rainfall and flooding in Paraguay during the austral summer 2015-2016}
 \title{Extreme rainfall in Paraguay during the 2015-2016 austral summer: causes and subseasonal-to-seasonal predictive skill}
@@
 
-% Why this is worthy of study -- needs some work!
+% THIS PARAGRAPH NEEDS SOME WORK
+% COME BACK TO IT LATER
 During the austral summer of 2015-16, repeated intense rainfall events led to severe flooding on the Paraguay River, displacing approximately \num{170000} people \citep{BrackenridgeDFO} and generating tremendous property damage.
@@
 Furthermore, we present all predictions in terms of odds relative to the climatological odds:
-\begin{equation}
+\begin{equation} \label{eq:odds-ratio}
 odds_{r} \equiv \frac{\frac{p}{1-p}}{\frac{p_c}{1-p_c}}
@@
 %
-\caption{Model Output Statistics (MOS) methods used to correct the ECMF sub-seasonal forecasts.
-Spatial domain for predictand is always the same (\SI{39}{\degree S}-\SI{17}{\degree S}; \SI{66}{\degree W}-\SI{49}{\degree W}).
-Two initializations are used: Nov 12th and 16th, 2015.} \label{t1}
+\caption{
+	Model Output Statistics (MOS) methods used to correct the ECMF sub-seasonal forecasts.
+	Spatial domain for predictand is always the same (\SIrange{39}{17}{\degree S}; \SIrange{66}{49}{\degree W}).
+	Two initializations are used: Nov 12th and 16th, 2015.} \label{tab:mos-methods}
 \begin{center}
-%
-\scalebox{0.8}{
-\begin{tabular}{ccccccccc}
-%
-\topline
-%
+\begin{tabular}{L{0.75in} L{1.15in} L{3.5in}}
+\hline\hline
 Model & Region (Predictor) & Final predictor(s) selected \\
 %
-\midline
+\hline
 %
- \emph{Raw} & (\SIrange{39}{17}{\degree S}; \SIrange{66}{49}{\degree W}) & \begin{tabular}{@{}c@{}}Ensemble mean, computed using members from the  \\ two initializations.
-No correction performed.
-\end{tabular} \\
+\emph{Raw} & \SIrange{39}{17}{\degree S}; \SIrange{66}{49}{\degree W} & Ensemble mean, computed using members from the  two initializations. No correction performed. \\
 %
+\emph{XLR} & \SIrange{39}{17}{\degree S}; \SIrange{66}{49}{\degree W} & Ensemble mean, computed using members from the  two initializations.  \\
 %
- \emph{XLR} & (\SIrange{39}{17}{\degree S}; \SIrange{66}{49}{\degree W}) & \begin{tabular}{@{}c@{}}Ensemble mean, computed using members from the  \\ two initializations. \end{tabular}  \\
+\emph{HLXR} & \SIrange{39}{17}{\degree S}; \SIrange{66}{49}{\degree W} & Ensemble mean and spread, computed using  members from the two initializations.\\
 %
- \emph{HLXR} & (\SIrange{39}{17}{\degree S}; \SIrange{66}{49}{\degree W}) & \begin{tabular}{@{}c@{}} Ensemble mean and spread, computed using  \\ members from the two initializations.\end{tabular}  \\
+\emph{PCR} & \SIrange{60}{0}{\degree S}; \SIrange{80}{30}{\degree W} & Linear combination of model's EOFs, computed using both initializations as independent predictors (10 EOFs).\\
 %
-  \emph{PCR} & (\SIrange{60}{0}{\degree S}; \SIrange{80}{30}{\degree W}) & \begin{tabular}{@{}c@{}} Linear combination of model's EOFs, computed using \\ both initializations as independent predictors (10 EOFs).
-\end{tabular} \\
+\emph{CCA} & \SIrange{60}{0}{\degree S}; \SIrange{80}{30}{\degree W} & Canonical modes computed using both initializations as independent predictors.
+(10 predictor EOFs, 4 predictand EOFs, 4 canonical modes) \\
 %
-  \emph{CCA} & (\SIrange{60}{0}{\degree S}; \SIrange{80}{30}{\degree W}) & \begin{tabular}{@{}c@{}} Canonical modes computed using \\ both initializations as independent predictors.
-(10 predictor EOFs, \\ 4 predictand EOFs, 4 canonical modes)\end{tabular}  \\
-%
-\botline
+\hline\hline
 %
 \end{tabular}
-}
-%
 \end{center}
-%
 \end{table*}
-%
-%
 
@@
 	\caption{
-		The study area.
+		Map of the study area.
 		(L): all of South America.
-		The weather typing described in \cref{sec:methods,sec:weather-types} was calculated using geopotential height within the area indicated by the red box.
-		The ``Lower Paraguay River Basin'', over which rainfall is averaged to create plots in \cref{fig:klee,fig:chiclet} and define intense rainfall days in \cref{fig:lagged-rain}, is shown by the black box.
+		The domains of the ``Lower Paraguay River Basin'' and the domain used for weather typing are indicated in black and red, respectively.
 		(R): The Paraguay River.
-		As for (L), the ``Lower Paraguay River Basin'' over which rainfall is averaged is indicated.
-		Streamflows shown in \cref{fig:streamflow} were taken from Asunci\'on, Paraguay, shown by the black dot.
+		As for (L), the ``Lower Paraguay River Basin'' is indicated.
+		Streamflows shown in \cref{fig:streamflow} were taken from the fours stations indicated in green.
 		The Paraguay River and its tributaries, from the Natural Earth database (\texttt{www.naturalearthdata.com}), are also shown.
@@
 \begin{figure*}
+	\noindent\includegraphics[width=6.5in]{./Streamflow.pdf}
+	\caption{
+		River stage (height; in \si{\meter}) for the Paraguay River at four gauges along the Paraguay River.
+		(a) Seasonality (orange) and time series of 2015-16 observations (black) at each stream gauge.
+			The seasonality was fit using local polynomial regression as implemented in the \texttt{locfit} package in
+			the \textbf{R} statistical programming environment \citep{Loader1999}.
+		(b) Time series of daily stage measurements from 1929 to 2016 at each station.
+	}
+  \label{fig:streamflow}
+\end{figure*}
+
+\begin{figure*}
+	\noindent\includegraphics[width=6.5in]{../_figs/NDJF201516Anomaly.pdf}
+	\caption{
+    	Monthly composite anomalies observed during NDJF 2015-16.
+      Top row shows anomalies of streamfunction $\psi$ at \SI{850}{\hecto\pascal}.
+      Bottom row shows monthly rainfall anomalies, in units of \si{\milli\meter\per\day}.
+	}
+  \label{fig:anomalies}
+\end{figure*}
+
+\begin{figure*}
+	\noindent\includegraphics[width=6.5in]{../_figs/PSI_EOF_Loadings.pdf}
+	\caption{
+    	Leading four EOF loadings of daily NDJF \SI{850}{\hecto\pascal} streamfunction $\psi$ over chosen domain.
+	}
+  \label{fig:eof-loading}
+\end{figure*}
+
+\begin{figure*}
 	\noindent\includegraphics[width=6.5in,height=9in,keepaspectratio=true]{../_figs/WTComposite.pdf}
 	\caption{
-		Circulation anomalies associated with each weather type.
-    Weather types are separated by column.
-		Top row gives anomalies of \SI{200}{\hecto\pascal} wind; arrows indicate the direction, and both shaded contours and arrow length indicate the magnitude of the wind anomaly.
-    Second row shows circulation at \SI{850}{\hecto\pascal}; shaded contours give anomalies of geopotential height and arrows indicate the wind speed and direction.
-    Bottom row shows monthly rainfall anomalies in units of \si{\milli\meter\per\day}.
+		Composite anomalies associated with each weather type (column).
+    Top row shows anomalies of streamfunction $\psi$ at \SI{850}{\hecto\pascal}.
+    Bottom row shows precipitation anomalies in units of \si{\milli\meter\per\day}.
 	}
@@
 
+\begin{figure*}
+	\noindent\includegraphics[width=6.5in]{../_figs/RainfallWeatherType}
+	\caption{
+		Time series of area-averaged rainfall in the Lower Paraguay River Basin (\cref{fig:study-area}) for each day of NDJF  2015-16 (lines).
+		The weather types corresponding to each day are indicated in both color and adjacent label.
+		Dashed lines indicate the climatological 50th, 90th, and 99th percentiles of NDJF area-averaged rain over the Lower Paragauy River basin.
+	}
+  \label{fig:rain-wt}
+\end{figure*}
+
 \begin{figure*}	\noindent\includegraphics[width=6.5in]{../_figs/LaggedRain.pdf}
 	\caption{
-		Circulation anomalies associated with intense rainfall (99th percentile exceedance of area-averaged rainfall) in the Lower Paraguay River region.
+		Circulation anomalies associated with intense rainfall (99th percentile exceedance of area-averaged rainfall in the Lower Paraguay River Basin shown in \cref{fig:study-area}).
 		Lagged composites are shown, by column, for $t = $ \SIlist{-2;-1;0;1}{\day} relative to the date of intense rainfall.
-		Top row gives anomalies of \SI{200}{\hecto\pascal} wind; arrows indicate the direction, and both shaded contours and arrow length indicate the magnitude of the wind anomaly.
-    Second row shows circulation at \SI{850}{\hecto\pascal}; shaded contours give anomalies of geopotential height and arrows indicate the wind speed and direction.
-    Bottom row shows rainfall anomalies in units of \si{\milli\meter\per\day}.
+		Top row shows anomalies of streamfunction $\psi$ at \SI{850}{\hecto\pascal}.
+    Bottom row shows precipitation anomalies in units of \si{\milli\meter\per\day}.
 	}
@@
 \begin{figure*}
-	\noindent\includegraphics[width=6.5in]{../_figs/RainfallWeatherType}
+	\noindent\includegraphics[width=6.5in]{../_figs/Chiclet.pdf}
 	\caption{
-		Weather type occurrence during two years.
-		Time series of area-averaged rainfall (shown by the box in \cref{fig:study-area}) and weather type for each day of NDJF  2015-16.
-		Dashed lines indicate the climatological 50th, 90th, and 99th percentiles of NDJF area-averaged rain.
+    S2S model forecasts of area-averaged rainfall over the Lower Paraguay River Basin (\cref{fig:study-area}).
+    Forecast time series from the ECMWF S2S model are shown for several lead times (\SIlist{5;14;19;30}{\day}).
+		Each gray dot represents the forecast of a single ensemble member issued for the target date ($x$-axis) of area-averaged rainfall, in \si{\milli\meter\per\day} ($y$-axis).
+		The blue line connects the ensemble means, and the black line shows the observed rainfall.
+		Because the model is initialized twice-weekly, for a given lead time forecasts will only exist for two in seven possible target dates;  consequently the rainfall is smoothed with a 3-day running mean to facilitate visual comparison.
 	}
-  \label{fig:klee}
-\end{figure*}
-
-\begin{figure*}
-	\noindent\includegraphics[width=6.5in]{../_figs/NDJF201516Anomaly.pdf}
-	\caption{
-    	Monthly circulation anomalies during NDJF 2015-16.
-      Top row gives anomalies of \SI{200}{\hecto\pascal} wind; arrows indicate the direction, and both shaded contours and arrow length indicate the magnitude of the wind anomaly.
-      Second row shows circulation at \SI{850}{\hecto\pascal}; shaded contours give anomalies of geopotential height and arrows indicate the wind speed and direction.
-      Bottom row shows monthly rainfall anomalies, in unites of \si{\milli\meter\per\day}.
-	}
-  \label{fig:anomalies}
+  \label{fig:chiclet}
 \end{figure*}
@@
 	\caption{
-		caption
+		Model forecasts and skill scores for the methods indicated in \cref{tab:mos-methods}.
+		Top row shows the intense rainfall forecast for 1-7 December 2015 as the odds ratio of \cref{eq:odds-ratio} over the target domain.
+		A value greater than 1 indicates that the model forecast greater-than-average odds of rainfall exceeding the 90th percentile.
+		The second shows the 2AFC skill score for each grid cell; a value greater than 50 indicates that the model outperforms climatology.
 	}
@@
 
-\begin{figure*}
-	\noindent\includegraphics[width=6.5in]{../_figs/Chiclet.pdf}
-	\caption{
-    Assessment of S2S model forecasts.
-    Forecast time series from the ECMWF S2S model are shown for several lead times (\SIlist{2;10;25;40}{\day}) of each individual ensemble member (gray), the ensemble mean (black, dashed), and the observed (blue) rainfall.
-		The observed rainfall is smoothed with a \SI{3.5}{\day} running mean in order to facilitate comparison between daily observations and the models, which are initialized every 3-4 days.
-	}
-  \label{fig:chiclet}
-\end{figure*}
-
-\begin{figure*}
-	\noindent\includegraphics[width=6.5in]{Streamflow.pdf}
-	\caption{
-		River stage (height; in \si{\meter}) for the Paraguay River at four gauges along the Paraguay River.
-		(a) Seasonality (black) and time series of 2015-16 observations (orange) at each stream gauge.
-			The seasonality was fit using local polynomial regression as implemented in the \texttt{locfit} package in
-			the \textbf{R} statistical programming environment \citep{Loader1999}.
-		(b) Time series of daily stage measurements from 1929 to 2016.
-	}
-  \label{fig:streamflow}
-\end{figure*}
 \end{document}

### 2017-09-18 run.sh

@@ -1,7 +1,24 @@
-# GET RAW DATA
+# You must run `setupenv.sh` separately
+
+# 1. GET RAW DATA
 python 01Get/01-Reanalysis.py
 python 01Get/02-Rainfall.py
 python 01Get/03-S2S-Data.py
 python 01Get/04-DailyIndices.py
 python 01Get/05-MonthlyIndices.py
 python 01Get/06-Anomalies.py
+
+# 2. RUN ANALYSIS
+function JUPCONV(){
+  jupyter nbconvert --execute $1
+}
+JUPCONV 02Analysis/01-StudyArea.ipynb
+JUPCONV 02Analysis/02-Climatology-Anomalies.ipynb
+JUPCONV 02Analysis/03-Get-WTs.ipynb
+JUPCONV 02Analysis/04-WT-Plots.ipynb
+JUPCONV 02Analysis/05-Lagged-Rain.ipynb
+JUPCONV 02Analysis/06-S2S-Skill.ipynb
+JUPCONV 02Analysis/07-Jet-Extension.ipynb
+
+# 3. WRITEUP
+latexmk -cd -e -f -pdf -interaction=nonstopmode "03Writeup/paper.tex"

### 2017-09-18 03Writeup/paper.tex
re-wrote methods and data

@@
 
-\documentclass{ametsoc}
+\documentclass[twocol]{ametsoc}
 \journal{mwr}
@@
 The analysis involves observations and model forecasts.
-The period analyzed for diagnostics purposes is from Nov 1979 through Feb 2016.
-Forecasts consider the period starting in Dec 2015 until Mar 2016, paying special attention to the week of Dec 1-7, 2015, as indicated in the Introduction.
-Hindcasts to assess the real-time predictive skill consider the period Dec 1-7, 1995-2014.
-All codes used to access the data and to create the plots and results described in this paper are available at \url{https://github.com/jdossgollin}.
 
 \subsection{Observations}
-Rainfall data is taken from the CPC Unified Gauge-based Analysis of Global Daily Precipitation dataset, available online at \url{http://iridl.ldeo.columbia.edu/SOURCES/.NOAA/.NCEP/.CPC/.UNIFIED_PRCP/.GAUGE_BASED/.GLOBAL/.v1p0/}.
-%Extreme rainfall events are computed as those exceeding the 90th percentile in the observations.
+
+The period analyzed for diagnostics purposes is from 01 Nov 1979 through 29 Feb 2016.
+
+Rainfall data is taken from the CPC Unified Gauge-based Analysis of Global Daily Precipitation dataset \citep{xie2010cpc}, available online at \url{http://iridl.ldeo.columbia.edu/SOURCES/.NOAA/.NCEP/.CPC/.UNIFIED_PRCP/.GAUGE_BASED/.GLOBAL/.v1p0/}.
 Spatial resolution is \SI{0.5}{\degree}, and temporal resolution is daily.
+Both the retro and real-time datasets are used.
 
-Atmospheric circulations are diagnosed using daily data from the NCAR-NCEP Reanalysis I data set \citep{Kalnay1996}, and are referred to as observations, available online at \url{https://www.esrl.noaa.gov/psd/thredds/dodsC/Datasets/ncep.reanalysis/}.
+Atmospheric circulations are diagnosed using daily data from the NCAR-NCEP Reanalysis I data set \citep{Kalnay1996}, available online at \url{https://www.esrl.noaa.gov/psd/thredds/dodsC/Datasets/ncep.reanalysis/}.
 Spatial resolution is \SI{2.5}{\degree}, which is adequate for synoptic-scale analyses.
+
 Because the end-of-day time for the rainfall data is 12:00GMT over most of South America, we use six-hour reanalysis data, and shift by twelve hours before re-sampling to the daily time step.
@@
 
-Streamflow station data was collected by the Paraguayan Navy and National Administration of Navigation and Ports of Paraguay, and was processed and distributed by the Paraguayan Directorate of Meteorology and Hydrology (DINAC-DMH).
+Streamflow data was collected by the Paraguayan Navy and National Administration of Navigation and Ports of Paraguay, and was processed and distributed by the Paraguayan Directorate of Meteorology and Hydrology (DINAC-DMH).
 Locations of streamflow gauges are shown in \cref{fig:study-area}.
@@
 \subsection{Model forecasts}
-This study analyzes probabilistic seasonal and sub-seasonal forecasts of extremely wet rainfall events.
-The seasonal predictions correspond to the ``flexible format" forecasts provided by the International Research Institute for Climate and Society (IRI).
-This product provides the probability of exceedance (or non-exceedance) of a user-specified percentile of the climatological distribution.
-The DJF 2015-2016 forecasts analyzed were produced in November 2015.
-Due to the short record of flexible format forecasts (i.e., 2012-2016), no verification was performed for seasonal predictions.
-These forecasts are provided at a horizontal resolution of \SI{2.5}{\degree}.
 
-Sub-seasonal forecasts used in this research correspond to those of the European Centre for Medium-Range Weather Forecasts (ECMWF) coupled model, available via the Subseasonal-to-Seasonal Prediction Project Database \citep{Vitart2016}.
+%The seasonal predictions correspond to the ``flexible format" forecasts provided by the International Research Institute for Climate and Society (IRI).
+%This product provides the probability of exceedance (or non-exceedance) of a user-specified percentile of the climatological distribution.
+%The DJF 2015-2016 forecasts analyzed were produced in November 2015.
+%Due to the short record of flexible format forecasts (i.e., 2012-2016), no verification was performed for seasonal predictions.
+%These forecasts are provided at a horizontal resolution of \SI{2.5}{\degree}.
+
+This study analyzes probabilistic sub-seasonal forecasts of extremely wet rainfall events.
+The sub-seasonal forecasts used were isseud by the European Centre for Medium-Range Weather Forecasts (ECMWF) using a coupled model, available via the Subseasonal-to-Seasonal Prediction Project Database \citep{Vitart2016}.
 The horizontal resolution is \SI{1.5}{\degree} deg.
-The rainfall forecasts and hindcasts (or re-forecasts) considered were initialized on November 12th and 16th, 2015, targeting the first week of December of 2015 (defined as Dec 1-7).
+Forecasts consider the period starting in Dec 2015 until Mar 2016 and hindcasts to assess the real-time predictive skill consider the period Dec 1-7, 1995-2014.
 There is a total of 51 ensemble members for each forecast, and 11 ensemble members for each one of the 20 hindcasts (Dec 1-7, 1995-2014).
-The hindcasts were used to define the extreme event threshold, and for probabilistic forecast verification.
-% NEED TO STATE HERE THAT ACTUALLY ALL FORECASTS WERE USED FOR THE `CHICLET' DIAGRAM
+
+Hindcasts were used to define the extreme event threshold, and for probabilistic forecast verification; forecasts were used to analyze modeled rainfall during the entire NDJF 2015-16 season and in particular the week of Dec 1-7, 2015.
+For probabilistic analysis of the rainfall during the week 1-7 December 2015, rainfall forecasts and hindcasts considered were initialized on November 12th and 16th, 2015.
+For plotting, all forecast initializations were considered.
+These steps are discussed further in subsequent sections.
 
@@
 Different types of analyses are used to diagnose the causes of the extreme rainfall events, and to bias-correct and verify the forecasts.
-These are summarized in the following sub-sections.
+All codes for data and analyis are available at \url{https://github.com/jdossgollin/PYFloods}.
 
-\subsection{Cluster analysis}
+\subsection{Climatology and Anomalies} \label{sec:climatology-anomalies}
+
+The climatological value of a field (rainfall or a circulation variable from reanalysis) is defined as the mean, at each grid cell, during the period $\qty[1980, 2010]$.
+The anomaly, for a particular date, is defined as the departure from this climatology for the corresponding month; the anomaly can be defined outside the range $\qty[1980, 2010]$.
+
+\subsection{Weather Typing}
 
 A cluster algorithm is used on daily data to help diagnose the mechanisms associated with the rainfall events of interest in this research.
+
+The clustering was performed on the daily NDJF $\SI{850}{\hecto\pascal}$ streamfunction field ($\psi$), calculated from the meridional and zonal wind fields using spherical harmonics as implemented in the \texttt{windspharm} python package \citep{Dawson:2016ws}, over the domain spanning \SIrange{15}{30}{\degree S} and \SIrange{65}{45}{\degree W} (\cref{fig:study-area}).
+To facilitate clustering (which tends to perform poorly in high-dimensional spaces), the anomaly field of $\psi$ (see \cref{sec:climatology-anomalies}) was projected  onto its four leading empirical orthogonal functions (EOFs), accounting for 95\% of the total observed variance (supplemental figure S1).
+The \texttt{eofs} python package was used to perform this dimension reduction \citep{Dawson:2016ge}.
+No additional time filtering was applied to the data before clustering, thus retaining the annual seasonal cycle, inter-annual, sub-seasonal and synoptic weather time scales for diagnosis.
+No latitudinal weighting was necessary as the selected domain is relatively small and it does not extend into high latitudes.
+Once the EOFs were calculated, the principal component time series were computed for each day and scaled to unit variance.
+
+Next, the $k$-means algorithm was used to assign a single cluster value to each day on record.
 The $k$-means technique is a partitioning method that classifies all days in study into a predefined number of clusters.
@@
 \begin{equation}
-	\mathcal{W} \qty( \mathcal{P} ) = \sum_{i}^{k}\sum_{X\in p_{i}} \dd s^{2} \qty(X,\bar{p}_{i})
+	\mathcal{W} \qty( \mathcal{P} ) = \sum_{i=1}^{k}\sum_{\vb{x} \in \vb{p}_{i}} \abs{\abs{\vb{x} - \vb{p}_i}}
 \end{equation}
-with $i=1...k$,  $k$ the total number of classes in the solution, $p_i$ the different clusters in the partition and $\dd s^2(X,\bar{p}_i)$ the Euclidean distance between the cluster's maps $X$ and centroid $\bar{p}_i$.
-The algorithm is guaranteed to converge to at least a local minimum; to select the best partition, 500 simulations were created using the implementation in Python's \texttt{scikit-learn} package \citep{Pedregosa:2012tv}, and the classifiability index of \citet{Michelangeli1995} was used to select the optimal partition.
-Calculation of the classifiability index for several values of $k$ suggests that states with $k=2,4,6$ are all reasonable; we chose the solution $k=6$ because the weather types identified are qualitatively similar to those determined over southeastern South America \citep{Munoz2015,Munoz2016} and have an intuitive physical meaning, discussed further in section \ref{sec:diag}.
+where $k$ is the total number of classes in the solution, $\abs{\abs{\cdot}}$ is the Euclidan norm, $\vb{p}_i$ the different clusters in the partition and $\vb{x}$ the maps corresponding to cluster $i$.
 
-This procedure assigns only one cluster to each day on record.
-Daily NDJF geopotential data over the box spanning \SIrange{15}{30}{\degree S} and \SIrange{65}{45}{\degree W} (\cref{fig:study-area}) was first projected onto its four leading empirical orthogonal functions, accounting for 95\% of the total observed variance.
-No additional time filtering was applied to the data before clustering, thus retaining the annual seasonal cycle, inter-annual, sub-seasonal and synoptic weather time scales for diagnosis.
-No latitudinal weighting was necessary as the selected domain is relatively small and it does not extevernd into high latitudes.
-The time series were scaled to unit variance prior to running the clustering algorithm.
+The $k$-means algorithm is guaranteed to converge to at least a local minimum; to select the best partition, 500 simulations were created using the implementation in Python's \texttt{scikit-learn} package \citep{Pedregosa:2012tv}.
+The classifiability index of \citet{Michelangeli1995} was then used to select the optimal partition.
+Calculation of the classifiability index for several values of $k$ (supplemental figure S2) suggests that states with $k=4,5,6,7,8$ are all reasonable; we chose the solution $k=6$ because the weather types identified are qualitatively similar to those determined over southeastern South America \citep{Munoz2015,Munoz2016} and have an intuitive physical meaning, discussed further in the following sections.
 
@@
 
-Since no model is perfect, several methods --generically known as Model Output Statistics \citep[MOS;][]{Glahn:1972vt} -- have been developed to correct for different kind of biases.
-In this work, we analyze how well the rainfall events could have been predicted, both using the raw sub-seasonal forecasts and bias-corrected seasonal and sub-seasonal forecasts.
+A wide variety of methods, generically known as Model Output Statistics \citep[MOS;][]{Glahn:1972vt}, have been proposed to correct for different types of bias in model outputs.
+In this work, we analyze how well the rainfall events could have been predicted, both using the raw sub-seasonal forecasts and  MOS-adjusted sub-seasonal forecasts.
+We use apply three types of MOS models: the homoscedastic and heteroscedastic versions of the extended logistic regression (XLR and HXLR, respectively); principal component regression (PCR); and canonical correlation analysis (CCA).
 
-The IRI's seasonal predictions are already provided with bias-correction \citep{Barnston:2010ge}, and hence we did not perform any further MOS on seasonal rainfall fields.
-
-We use three different Model Output Statistics strategies in this study to correct for biases in sub-seasonal forecasts.
-The approach to perform the corrections is based on both the homoscedastic and heteroscedastic versions of the extended logistic regression (XLR and HXLR, respectively), principal component regression (PCR), and canonical correlation analysis (CCA).
-These different regression models are deemed here to be of increasing complexity, with the logistic regression models being relatively simple and the CCA one the more complex approach.
-More details and references on each model are given below.
-
-The original logistic regression models the conditional probability of binary events, and it has been used in MOS approaches for binary predictands \citep{Hamill:2004hk}.
+Logistic regression models the conditional probability of binary events, and it has been used in MOS approaches for binary predictands \citep{Hamill:2004hk}.
 Nonetheless, when using the logistic regression to address multiple thresholds via independent fits, the predicted probabilities are in general not mutually consistent \citep{Messner:2014gp}.
-The (homoscedastic) XLR was designed to solve this problem \citep{Wilks:2009bk}, via the consideration of a transformation of the thresholds of interest as an additional predictor variable.
+The (homoscedastic) XLR was designed to address this shortcoming \citep{Wilks:2009bk}, via the consideration of a transformation of the thresholds of interest as an additional predictor variable.
 The HXLR, a heteroscedastic generalization of the XLR, was proposed to appropriately use the ensemble spread as predictor for the dispersion of the predictive distribution \citep{Messner:2014gp}.
-In this study we explore the potential benefits and disadvantages of both versions of the extended logistic regression.
-
-In PCR, a special case of CCA, each grid box in the predictand field is regressed using a linear combination of the predictor's EOFs \citep{Mason:2008da,Wilks:2006fx}.
 
@@
 The method forecasts spatial patterns of variability spanning across the region of interest rather than making forecasts for individual locations.
-Since the canonical modes describe the preferred coupled spatial patterns relating predictors and predictands, CCA has the potential to correct the model rainfall spatial patterns using statistical relationships involving the observed patterns.
-
-Because CCA and PCR involve regressions of spatial modes, their use addresses bias in both the magnitude and the spatial distribution of the modeled precipitation patterns.
-Conversely, because XLR and HXLR are applied uniquely to each grid cell, these procedures correct only for magnitude biases in the probability distribution.
+In PCR, a special case of CCA, each grid cell in the predictand field estimated by regression using a linear combination of the predictor's EOFs \citep{Mason:2008da,Wilks:2006fx} rather than by identifying canonical modes.
+Unlike the XLR and HXLR models, which perform bias correction independently for each grid cell, the CCA and PCR models can address both biases in the magnitude and the spatial distribution of the modeled precipitation patterns.
 
@@
 We use the same spatial domain [\SI{39}{\degree S}-\SI{17}{\degree S}; \SI{66}{\degree W}-\SI{49}{\degree W}] for both the predictor and the predictand, except for the PCR and CCA cases, in which a larger domain [\SI{0}{\degree S}-\SI{60}{\degree S}; \SI{80}{\degree W}-\SI{30}{\degree W}] was used to better capture the spatial patterns in the raw model field.
+A variety of approaches for modeling the output of many ensemble members was explored; a summary of the predictors found to be most skillful for each MOS model is presented in \cref{tab:mos-methods}.
 
-To avoid artificial skill when building the statistical models, we employ a cross-validation approach that uses a 5-year-out window: five continuous years are left out of the record, the regression coefficients are computed with the remaining of the time series, and the resulting model is validated comparing the prediction for the third year left out (middle of the window) against observations.
+To avoid artificial skill when building the statistical models, we employ a cross-validation approach with a 5-year-out window.
+In this framwork, five continuous years are left out of the record, the regression coefficients are computed with the remaining of the time series, and the resulting model is validated comparing the prediction for the third year left out (middle of the window) against observations.
 The 5-year-long window is redefined a year at a time, moving from the beginning of the record to its end.
 
-Due to the nature of each model used for the MOS-correction, different analyses were explored to define the best predictor(s) for each case.
-For example, for the PCR and CCA models, predictors were defined using different geographical domains and using (a) an ensemble of the members available in both initialization times (Nov 12th and 16th, 2015), and (b) computing the ensemble mean for each one of these same initializations and treating them as independent predictors.
-Since the skill was shown to be higher in option (b), this was selected as the final predictor (for the spatial domain indicated in the previous paragraph).
-A summary of the final selection for each model predictors is presented in Table \ref{t1}.
-
-Probabilistic forecasts are normally used to convey information about the uncertainties associated to the prediction of interest.
-It is a common practice to present these forecasts as probabilities for occurrence of below-normal, normal and above-normal precipitation, where the normal is defined as the long-term average of at least 20 years.
-Here we use the period 1981-2010 for seasonal forecasts, and 1995-2014 for sub-seasonal ones.
-Nonetheless, since we analyze extreme event predictions (90th percentile exceedances), rather than terciles.
-Furthermore, we present all predictions in terms of odds relative to the climatological odds:
+To visualize the probability of intense rainfall (exceedance of the 90th percentile of modeled rainfall during the 1995-2014 period) at each grid cell, we present all predictions in terms of odds relative to the climatological odds:
 \begin{equation} \label{eq:odds-ratio}
@@
 where $p$ and $p_c$ represent the forecast probability and the climatological probability, respectively.
-This choice helps to easily visualize in the forecast maps locations where the odds of occurring an extreme rainfall event are considerably higher than expected.
-In contrast, using probabilities in these forecasts requires that the reader compares values against the climatological probability.
-Our choice thus simplifies the comparison in a way that  also tends to be better understood by decision makers.
+This is analagous to the commonly used practice of showing probability of below-normal, normal, and above-normal precipitationand simplifies the comparison in a way that tends to be better understood by decision-makers.
 
-\subsection{Probabilistic forecast verification}
+\subsection{Probabilistic Forecast Verification}
 
-In addition to visually comparing predictions and observations to verify how well the extreme events could have been predicted, we also provide some context to the analysis of the sub-seasonal skill conducting a formal cross-validated verification.
-This is not performed for the seasonal predictions due to the prohibitively short record available in that case (see section \ref{sec:data}).
-
-The Generalized Relative Operating Characteristics (GROC), also called 2AFC score \citep{Mason:2009kr}, is used to evaluate skill of probabilistic rainfall forecasts at a local scale, via spatial maps.
+In addition to visually comparing predictions and observations to verify how well the extreme events could have been predicted, we also compute the Generalized Relative Operating Characteristics (GROC), also called 2AFC score \citep{Mason:2009kr}, to evaluate skill of probabilistic rainfall forecasts.
 This score measures the ``proportion of all available pairs of observations of differing category whose probability forecasts are discriminated in the correct direction'' \citep{Mason:2009kr}.
@@
 
-\section{Atmospheric circulation and intense rainfall} \label{sec:diag}
+\section{Atmospheric Circulations and Intense Rainfall} \label{sec:diag}
 

### 2017-09-19 04-Writeup/paper.tex
Get up and running with a working Makefile

--- 2017-09-18 03Writeup/paper.tex
@@
 
+\subsection{Climate Indices}
+
+This study also makes use of some climate indices.
+Data on the El Ni\~{n}o-Southern Oscillation (ENSO) came from a statistical-dynamical interpolation \citep{Kaplan:1998df}, which is constrained by relatively high-quality observations during the study period and is available online at \url{http://iridl.ldeo.columbia.edu/SOURCES/.Indices/.nino/.EXTENDED/.NINO34/}.
+Data on the Madden-Julien Oscillation (MJO) came from the Australian Bureau of Meteorology and is available online at \url{http://www.bom.gov.au/climate/mjo/graphics/rmm.74toRealtime.txt}.
+
 \subsection{Model forecasts}
 
-%The seasonal predictions correspond to the ``flexible format" forecasts provided by the International Research Institute for Climate and Society (IRI).
-%This product provides the probability of exceedance (or non-exceedance) of a user-specified percentile of the climatological distribution.
-%The DJF 2015-2016 forecasts analyzed were produced in November 2015.
-%Due to the short record of flexible format forecasts (i.e., 2012-2016), no verification was performed for seasonal predictions.
-%These forecasts are provided at a horizontal resolution of \SI{2.5}{\degree}.
+This study analyzes probabilistic seasonal and sub-seasonal forecasts of extremely wet rainfall events, defined in the same way as in observations (90th percentile).
 
-This study analyzes probabilistic sub-seasonal forecasts of extremely wet rainfall events.
 The sub-seasonal forecasts used were isseud by the European Centre for Medium-Range Weather Forecasts (ECMWF) using a coupled model, available via the Subseasonal-to-Seasonal Prediction Project Database \citep{Vitart2016}.
@@
 
+The seasonal predictions used were ``flexible format'' forecasts provided by the International Research Institute for Climate and Society (IRI).
+This product provides the probability of exceedance (or non-exceedance) of a user-specified percentile of the climatological distribution.
+The DJF 2015-2016 forecasts analyzed were produced in November 2015.
+Due to the short record of flexible format forecasts (2012-2016), no verification was performed for seasonal predictions.
+These forecasts are provided at a horizontal resolution of \SI{2.5}{\degree}.
+
 %-------------------------------------------------------------------------------
@@
 Different types of analyses are used to diagnose the causes of the extreme rainfall events, and to bias-correct and verify the forecasts.
-All codes for data and analyis are available at \url{https://github.com/jdossgollin/PYFloods}.
+All codes used for data and analyis are available at \url{https://github.com/jdossgollin/PYFloods}.
+Computation with multidimensional data was performed using the \texttt{xarray} \citep{hoyer2017xarray} and \texttt{numpy} \citep{vanderWalt:2011dp} packages of the python computing environment; computations on tabular data utilized the \texttt{pandas} \citep{McKinney:2010un} package; and plotting was performed with the \texttt{matplotlib} \citep{Hunter:2007ux} and \texttt{cartopy} \citep{Cartopy} packages.
 
@@
 
+The IRI's seasonal predictions are already provided with bias correction \citep{Barnston:2010ge}, and hence we did not perform any further MOS on seasonal rainfall fields.
+
 \subsection{Probabilistic Forecast Verification}
@@
 %-------------------------------------------------------------------------------
+% OBSERVED FLOODS
+%-------------------------------------------------------------------------------
+
+\section{Observed Flooding} \label{sec:diag}
+
+\Cref{fig:streamflow} shows the streamflow time series at several gauges on the Paraguay River during NDJF 2015-16 in the context of their seasonality and decadal variability.
+Because no stage-discharge curves are available, we present only the river stage.
+While this is relevant from the perspective of flood damage, flow rates cannot be estimated without these curves (which are also highly variable in time and consequently must be updated frequently).
+In a very flat river basin such as the Lower Paraguay River Basin, such one might expect to expect that discharge scales approximately with $\text{stage}^3$.
+
+The most notable feature of \cref{fig:streamflow}(a) is that during November and December 2015, the river rose rapidly at Concepci\'on, Asunci\'on,  and Pilar but not at Bah\'ia Negra.
+As discussed in \citet{Bravo:2011et,Barros:2004bn}, the location of the Bah\'ia Negra gauge (see \cref{fig:study-area}) in the Pantanal region means that it responds very slowly to rainfall input.
+However, the three downstream gauges respond to the rainfall forcing with a slow but steady rise.
+Despite several very intense storms (see following sections), the streamflow record at Asunci\'on and Pilar (which are downstream of Concepci\'on) indicates relatively little response to individual storms.
+
+Interestingly, examination of \cref{fig:streamflow}(b) suggests multidecadal oscillation in the streamflow record.
+This is in agreement with \citet{Collischonn:2001bi,Carvalho2011} who find a changepoint in the 1970s possibly associated with changes in the Pacific Decadal Oscillation.
+However, because only river stage data (and not discharge) data are available, it is not possible to discern whether the observed changes in river stage are driven by sediment loading and local measurement characteristics or by large-scale climate fluctuations.
+Further treatment of this question is beyond the scope of this paper.
+
+%-------------------------------------------------------------------------------
 % DIAGNOSTICS
@@
 
-\section{Atmospheric Circulations and Intense Rainfall} \label{sec:diag}
+\section{Atmospheric circulation and intense rainfall} \label{sec:diagnostics}
 
-This section discusses the synoptic circulation patterns related to the intense rainfall events observed during the austral summer of 2015-2016.
-To provide context to the analysis of this particular season, we first characterize recurrent daily circulation types (or weather types) over the region under study, and discuss the average synoptic patterns conducive to intense rainfall events there.
+\subsection{Intense Rainfall: Climatological Drivers} \label{sec:rainfall-circulation}
+
+To understand how circulation anomalies observed during NDJF 2015-2016 led to the observed floods it is helpful to first explore the atmospheric circulations which are typically associated with intense rainfall in the lower Paraguay River.
+
+\Cref{fig:lagged-rain} shows time-lagged anomalies up to and after days exceeding the 90th percentile of area-averaged daily rainfall in the lower Paraguay River basin (see \cref{fig:study-area}), and is consistent with the findings of \citet{Marengo2004,Salio:2007gd}.
+At $t=-2$ and $t=-1$ a low-level northerry jet injects inflow of heat and moisture to the region, favored by a polar trough extending slightly to the east of Paraguay.
+This is in agreement with previous studies which find a close correspondence between the exit regions of the low-level jet and the observed rainfall \citep{Saulo:2007km,Salio:2007gd,Marengo2004,Velasco1987} as the moisture influx, energy influx, and strong vertical wind shear favor instability and intense convection \citep{Marengo2004,Silva2009}.
+Next, low-level convergence and upper-level divergence (not shown) associated with the eastward progression of the frontal system further favor rainfall locally.
+Similar analysis for the 85th or 95th percentiles of daily rainfall (not shown) yield highly similar results, implying that the mechanism for the most extreme events is not fundamentally distinct from the mechanism for moderate-intensity events.
 
@@
 
-Visual inspection of the rainfall and circulation anomalies composited on each weather type (\cref{fig:wt-composite}) shows that although they were constructed from the leading EOFs of \SI{850}{\hecto\pascal} geopotential height over a relatively small region (see \cref{fig:study-area}), the circulation patterns are associated with regional-scale wave trains.
+We next use the weather typing algorithm to understand circulations and sequences of circulations associated with intense rainfall in the Lower Paraguay River Basin.
 
-The \SI{850}{\hecto\pascal} geopotential height anomalies associated with the weather types resemble the first three EOFS of the field, and weather types 2, 4, and 6 are nearly the inverse of weather types 1, 3, and 5, respectively.
-The weather types also resemble those found by \citet{Munoz2015} for DJF using a similar approach over an area further to the southeast, though the relative frequencies are quite different.
-This is likely due to the fact that the in the present research we use NDJF instead of DJF, and the periods considered are not the same.
-The rainfall anomalies associated with these weather types are also consistent with the leading modes of spring/summer rainfall variability identified by other authors \citep{Grimm:2009bq}.
+The first step of the weather typing algorithm is to identify leading EOFs of the \SI{850}{\hecto\pascal} streamfunction $\psi$.
+The EOF laodings are shown in \cref{fig:eof-loading}.
+Of these, EOF 1 explains by far the most variance ($\approx 67\%$) while EOFs 2, 3, and 4 collectively explain approximately $28\%$ of total variance (supplemental figure S2).
+To a first order of approximation, EOF 1 explains the large-scale anomalies induced by passing baroclinic waves (weaker in the West, likely due to the Andes Mountains).
+EOF 2 indicates the strength of the low-level jet circulation, while EOFs 3 and 4 control the exit region of this jet.
+Given these features in the EOF loadings, it is no surprise that composites of the composite \SI{850}{\hecto\pascal} circulations and rainfall associated with each weather type (\cref{fig:wt-composite}) reveal circulation patterns are associated with synoptic- and regional-scale circulation patterns.
 
-Weather types, 1, 3, and 6 all show enhanced low-level jet activity, while weather types 2, 4, and 5 all show a weakening or reversal of the climatological low-level wind.
-As discussed by previous studies \citep{Saulo:2007km,Salio:2007gd,Marengo2004,Velasco1987}, there is a close correspondence between jet exit regions and the regions of highest rainfall.
-For example, the low-level jet for weather type 3 has a very small zonal component, while for weather type 6 the zonal component is much greater.
-Correspondingly, the maximum rainfall intensities for weather type 3 are farther to the west than for weather types 1, which is still further to the west than for weather type 6.
+As predicted, there is a close correspondence between the exit regions of the low-level jet and the observed rainfall.
+Weather type 1 represnts a ``Chaco Jet'' event \citep{Salio:2002ev} in which the maximum wind speed penetrates past \SI{25}{\degree S}, leading to intense rainfall NE Argentina and Uruguay.
+The other weather type associated with a SALLJ is weather type 4, the ``No-Chaco Jet'' event in which the wind turns to the East before \SI{25}{\degree S}, leading to heavy rainfall over Eastern Paraguay and SW Brazil.
+Weather types 2 and 3 resemble the weak and intense phases, respectively, of the South Atlantic Convergence Zone \citep[SACZ;][]{Carvalho2004} while weather types 5 and 6 represent southerly and easterly wind, respectively, over the Lower Paraguay River Basin.
+\texttt{ANY THOUGHTS ON WT5,6?}
 
-Analysis of the relative frequency of different weather types across years (not shown) reveals strong inter-annual variability.
-This is consistent with the findings of \citet{Munoz2015,Munoz2016}, which identify strong modulation of weather type occurrence at sub-seasonal to seasonal time scales.
+\subsection{NDJF 2015-16: Circulation Sequences}
 
-\subsection{Mean Circulations Associated with Paraguay River Basin Intense Rainfall} \label{sec:rainfall-circulation}
-
-To understand how circulation anomalies observed during NDJF 2015-2016 led to the observed floods it is helpful to first explore the atmospheric circulations which climatologically are associated with intense rainfall in the lower Paraguay River.
-
-\Cref{fig:lagged-rain} shows time-lagged anomalies up to and after days exceeding the 90th percentile of area-averaged daily rainfall in the lower Paraguay River basin (see \cref{fig:study-area}), and is consistent with the findings of \citet{Marengo2004,Salio:2007gd}.
-\Cref{fig:lagged-rain} suggests that the key mechanism involves two components.
-The first, at $t=-2$ and $t=-1$, is a northerly inflow of heat and moisture via a low-level jet, which may be favored by a polar trough extending slightly to the east of Paraguay, consistent with \SI{850}{\hecto\pascal} circulation shown in weather types 1, 3, or 6.
-Previous studies have demonstrated that this jet is also associated with strong vertical wind shear \citep{Marengo2004,Silva2009}, favoring instability and intense convection.
-The second factor is low-level convergence (\cref{fig:lagged-rain} top row) and upper-level divergence (\cref{fig:lagged-rain} middle row) associated with the eastward progression of the frontal system.
-While \cref{fig:lagged-rain} shows composites of winds anomalies, analysis of particular events reveals that the anomalies are often, though not always, sufficient to fully reverse the direction of climatological wind.
-Similar analysis for the 85th or 95th percentiles of daily rainfall (not shown) yield essentially identical results, implying that the mechanism for the most extreme events is not fundamentally distinct from the mechanism for moderate-intensity events.
-
-\texttt{RE: need to identify what makes it rain over Paraguay rather than farther South: agreed but we should consider what analysis is required for this step.}
-
-\subsection{Circulation Patterns Observed during NDJF 2015-16}
-
-An important question is whether the circulations that drove intense rainfall of NDJF 2015-16 differed from the circulations described above which typically drive intense rainfall.
+To understand what happened specifically during NDJF 2015-16, we show both sequences of daily weather types and monthly-mean circulation anomalies.
 
 While the weather typing approach requires simplifying the dynamics of daily circulation patterns, its advantage is that it greatly facilitates the analysis of sequences of precipitation.
-A Klee diagram \citep{Munoz2015, Munoz2016} is used to visualize these sequences, which are presented in \cref{fig:klee} along with rainfall time series for the season of interest which shows that intense rainfall concentrated over a period spanning from mid-November 2015 through early January 2016, with smaller peaks in late January and mid-February.
-As discussed in \cref{sec:rainfall-circulation}, intense rainfall is typically preceded by a northerly, low-level influx of energy and moisture.
-Examination of \cref{fig:klee} reveals that while the most intense rainfall occurred during many different weather types, weather type 1 immediately precedes nearly almost all of the most intense rainfall peaks.
-Thus, while the intensity and persistence of intense rainfall was atypical, the causal mechanism of the intense rainfall observed during this season was consistent with climatology.
+\Cref{fig:rain-wt} shows a time series of area-averaged rainfall over the Lower Paraguay River Basin for NDJF 2015-16; this plot shows that intense rainfall concentrated over a period spanning from mid-November 2015 through early January 2016, with smaller peaks in late January and mid-February.
+As predicted by \cref{fig:wt-composite}, the most intense rainfall occurred during weather types 1 and 4.
+In particular, many (though not all) such events occured during a transition of WT1-WT4-WT5, which is consistent with the westerlay passage of a mid-latitude system as shown in \cref{fig:lagged-rain}.
+During NDJF 2015-16 weather type 1 occurred nearly twice as often (33\% of days) as its climatology (19\% of days), weather type 4 occurred slightly more often (18\% of days) than its climatology (16\% of days), and other weather types occurred less frequently than their climatology (see Figure S3).
+During mid-January, during a sequence of persistent low rainfall, weather type 3 featured persistently, leading to intense rainfall over central Brazil (not shown) and negligible rainfall over the Lower Paraguay River Basin.
+Thus while the intensity and persistence of intense rainfall was atypical, the causal mechanism of the intense rainfall observed during this season was consistent with climatology.
 
-At the monthly scale (\cref{fig:anomalies}), the largest observed rainfall anomalies overlap with the direction, intensity, and exit region of the low-level jet apparent in the \SI{850}{\hecto\pascal} field.
-During all months, except January 2016, the \SI{850}{\hecto\pascal} wind field indicates enhancement of the SALLJ above approximately \SI{20}{\degree S} but does not penetrate further south.
-In the upper troposphere, enhanced convergence over the tropical Atlantic just off the coast of Northeast Brazil and enhanced divergence over Paraguay, Uruguay, and Northern Argentina coincides with the enhanced SALLJ activity.
-The latter feature in particular is enhanced during November and December 2015 and February 2016, but weakened during January 2016.
-\texttt{Add more here James}
+At monthly scale (\cref{fig:anomalies}), a weak anticyclonic circulation set up over central Brazil during November 2015 and strengthened into the following month.
+In January 2016 it weakened before returning in February 2016.
+When analyzing these anomalies it is important to consider that these anomalies are all relative to the climatology (supplemental figure S4) which exhibits substantial seasonality.
+The observed rainfall and circulation anomalies are consistent with the aggregation of the observed weather types shown in \cref{fig:rain-wt}; November 2015 most closely resembles weather type 1, December 2015 weather types 1 and 4, January 2016 weather type
+3, and February 2016 weather types 2 and 4.
+
+%-------------------------------------------------------------------------------
+% FORECASTS
+%-------------------------------------------------------------------------------
 
@@
 In this section we analyze the extent to which forecasts were able to predict the persistent rainfall during summer of 2015-16.
+
 There are advantages in simultaneously considering useful climate information at multiple timescales, rather than just focusing on one of them.
@@
 
-Extreme rainfall over the region was accurately predicted by the IRI for the DJF 2015-2016 season since at least November 2015 (see \Cref{fig:probfcst}).
+Extreme rainfall over the region was accurately predicted by the IRI for the DJF 2015-2016 season since at least November 2015 (see \cref{fig:subs-prob-fcst}).
 Probabilities of 30\%-50\% for the 90th percentile, three to five times the expected climatological values, are visible over southern Paraguay and Brazil, and northern Uruguay and Argentina, mostly in agreement with observations (hashed region in \cref{fig:probfcst}).
@@
 
-As seen in \Cref{fig:subsprobfcst}a, the raw (uncorrected) sub-seasonal forecast of the ECMF model for Dec 1-7 2015, indicated very high odds for occurrence of extreme rainfall but with important biases in the actual location and spatial pattern; for Paraguay, it confidently suggests occurrence of extremes to the south-southeast of the country, which was mostly not observed.
-Overall, the 20-year based skill of probabilistic forecasts for the first week of December --even when using multiple initializations-- is worse than using climatology for most of Paraguay (\Cref{fig:subsprobfcst}b), being best for the zone between that country and northern Uruguay, and for most of southern Brazil.
+As seen in \Cref{fig:subs-prob-fcst}a, the raw (uncorrected) sub-seasonal forecast of the ECMF model for Dec 1-7 2015, indicated very high odds for occurrence of extreme rainfall but with important biases in the actual location and spatial pattern; for Paraguay, it confidently suggests occurrence of extremes to the south-southeast of the country, which was mostly not observed.
+Overall, the 20-year based skill of probabilistic forecasts for the first week of December --even when using multiple initializations-- is worse than using climatology for most of Paraguay (\Cref{fig:subs-prob-fcst}b), being best for the zone between that country and northern Uruguay, and for most of southern Brazil.
 
@@
 In general, the use of extended logistic regression models does not improve the forecast for the week.
-For example, with respect to the raw prediction, XLR tends to amplify the relative odds, and to cluster and shift the forecast location of the extremes (\Cref{fig:subsprobfcst}c); the forecast tend to be better for Uruguay, but suggests extreme rainfall in the Paraguayan Chaco, which was not present in the raw prediction.
-On the other hand, the use of the ensemble spread in the HLXR model does not help: this forecast tends to be very confident (always compared with the uncorrected case) on the extremes occurring in almost all the region of interest (\Cref{fig:subsprobfcst}e).
+For example, with respect to the raw prediction, XLR tends to amplify the relative odds, and to cluster and shift the forecast location of the extremes (\Cref{fig:subs-prob-fcst}c); the forecast tend to be better for Uruguay, but suggests extreme rainfall in the Paraguayan Chaco, which was not present in the raw prediction.
+On the other hand, the use of the ensemble spread in the HLXR model does not help: this forecast tends to be very confident (always compared with the uncorrected case) on the extremes occurring in almost all the region of interest (\Cref{fig:subs-prob-fcst}e).
 
-The probabilistic skill verification for both extended logistic regression models is very similar to the one of the raw forecast ( \Cref{fig:subsprobfcst}b,d,f), varying just in magnitude in a few places.
+The probabilistic skill verification for both extended logistic regression models is very similar to the one of the raw forecast ( \Cref{fig:subs-prob-fcst}b,d,f), varying just in magnitude in a few places.
 This is attributed to the fact that these models only correct for magnitude, and they do it on a grid-box by grid-box basis, so the spatial distribution of skill is basically the same.
@@
 Better forecasts are obtained when both magnitude and spatial corrections are performed, although with relative odds considerably less confident than the ones in the raw forecast.
-The Principal Component Regression model correctly shows high relative odds in most of the places where extreme rainfall was observed (\Cref{fig:subsprobfcst}g), although it also erroneously forecasts extremes in other locations, like zones of western Paraguay and northeastern Argentina.
-On the other hand, the main problem with the Canonical Correlation Analysis model is its lack of discrimination between occurrence or non-occurrence of extremes in the region: the spatial distribution of odds is too homogeneous (\Cref{fig:subsprobfcst}i).
+The Principal Component Regression model correctly shows high relative odds in most of the places where extreme rainfall was observed (\Cref{fig:subs-prob-fcst}g), although it also erroneously forecasts extremes in other locations, like zones of western Paraguay and northeastern Argentina.
+On the other hand, the main problem with the Canonical Correlation Analysis model is its lack of discrimination between occurrence or non-occurrence of extremes in the region: the spatial distribution of odds is too homogeneous (\Cref{fig:subs-prob-fcst}i).
 Still, it tends to exhibit slightly higher odds than the spatial average in regions where extremes were observed, and it forecast similar false alarms than the PCR model.
 
-The 20-year based skill maps of probabilistic forecasts (see \Cref{fig:subsprobfcst}h,j) computed with these models are very similar to each other, although the one for PCR is definitively better and shows high skill for almost all the region of interest, although there are still zones along the eastern part of Paraguay exhibiting equal or lower skill than climatological values.
+The 20-year based skill maps of probabilistic forecasts (see \Cref{fig:subs-prob-fcst}h,j) computed with these models are very similar to each other, although the one for PCR is definitively better and shows high skill for almost all the region of interest, although there are still zones along the eastern part of Paraguay exhibiting equal or lower skill than climatological values.
 Hence, despite of the particular errors in the Dec 1-7 2015 forecasts, on the long term both PCR and CCA verify considerably better than the raw, XLR and HXLR predictions.
@@
 	\caption{
-		Model forecasts and skill scores for the methods indicated in \cref{tab:mos-methods}.
+		MOS-adjusted S2S model forecasts and skill scores for the methods indicated in \cref{tab:mos-methods}.
 		Top row shows the intense rainfall forecast for 1-7 December 2015 as the odds ratio of \cref{eq:odds-ratio} over the target domain.
@@
 	}
-  \label{fig:subsprobfcst}
+  \label{fig:subs-prob-fcst}
+\end{figure*}
+
+\begin{figure*}
+	\noindent\includegraphics[width=6.5in]{../_figs/SeasonalForecast.pdf}
+	\caption{
+	Seasonal model forecast.
+	Shading indicates the forecast probability of exceeding the 90th percentile of climatological rainfall during DJF 2015-16.
+	As \cref{fig:subs-prob-fcst}, the variable shown is the odds ratio of \cref{eq:odds-ratio} over the target domain.
+	A value greater than 1 indicates that the model forecast greater-than-average odds of rainfall exceeding the 90th percentile.
+	}
+  \label{fig:seas-prob-fcst}
 \end{figure*}

### 2017-09-19 01-Get/DailyIndices.py

@@ -6,7 +6,6 @@ import pandas as pd
 from paraguayfloodspy.pars  import GetPars
 
 def main():
-    print("\nWelcome to Daily-Indices.py")
     time_pars = GetPars('time')
     syear,eyear = time_pars['syear'], time_pars['eyear']
     outfile = "_data/indices/daily_indices.csv"

### 2017-09-19 01-Get/MonthlyIndices.py

@@ -19,23 +19,15 @@ def read_indices(raw_url, col_name):
     return(df)
 
 def main():
-    print("\nWelcome to Monthly-Indices.py")
     time_pars = GetPars('time')
     syear,eyear = time_pars['syear'], time_pars['eyear']
     outfile = "_data/indices/monthly_indices.csv"
 
-    pdo = read_indices(
-        raw_url = 'http://iridl.ldeo.columbia.edu/SOURCES/.Indices/.PDO/.pdo/gridtable.tsv',
-        col_name = 'pdo'
-    )
-    enso = read_indices(
+    df = read_indices(
         raw_url = 'http://iridl.ldeo.columbia.edu/SOURCES/.Indices/.nino/.EXTENDED/.NINO34/gridtable.tsv',
         col_name = 'nino_34'
     )
-
-    df = pdo.join(enso,  how='outer')
     df = df['{}'.format(syear):'{}'.format(eyear)]
-    df.head(10)
 
     # Save to file
     df.to_csv(outfile)

### 2017-09-19 01-Get/Rainfall.py

@@ -49,7 +49,6 @@ def GetFileName(year):
     return(fn)
 
 def main():
-    print("\nWelcome to Rainfall.py\n")
     pars = GetPars('time')
     syear, eyear = pars['syear'], pars['eyear']
     overwrite = False

### 2017-09-19 02-Process/Anomalies.py

@@ -14,7 +14,6 @@ from paraguayfloodspy.xrutil import *
 from paraguayfloodspy.pars import *
 
 def main():
-    print("\nWelcome to Anomalies.py")
     pars = GetPars('all')
     syear, eyear = pars['time']['syear'], pars['time']['eyear']
     overwrite = False
@@ -44,7 +43,7 @@ def main():
         months=pars['months'],
         shift_time_h=12, return_daily=True)
 
-    for var in ['uwnd', 'vwnd', 'streamfunc']:
+    for var in ['streamfunc']:
         for level_i in [850]:
             fnames = glob.glob('_data/reanalysis/raw/{}_{}_*.nc'.format(var, level_i))
             raw_fname = "_data/reanalysis/subset/{}_{}_raw.nc".format(var, level_i)

### 2017-09-19 run.sh

@@ -1,24 +1,0 @@
-# You must run `setupenv.sh` separately
-
-# 1. GET RAW DATA
-python 01Get/01-Reanalysis.py
-python 01Get/02-Rainfall.py
-python 01Get/03-S2S-Data.py
-python 01Get/04-DailyIndices.py
-python 01Get/05-MonthlyIndices.py
-python 01Get/06-Anomalies.py
-
-# 2. RUN ANALYSIS
-function JUPCONV(){
-  jupyter nbconvert --execute $1
-}
-JUPCONV 02Analysis/01-StudyArea.ipynb
-JUPCONV 02Analysis/02-Climatology-Anomalies.ipynb
-JUPCONV 02Analysis/03-Get-WTs.ipynb
-JUPCONV 02Analysis/04-WT-Plots.ipynb
-JUPCONV 02Analysis/05-Lagged-Rain.ipynb
-JUPCONV 02Analysis/06-S2S-Skill.ipynb
-JUPCONV 02Analysis/07-Jet-Extension.ipynb
-
-# 3. WRITEUP
-latexmk -cd -e -f -pdf -interaction=nonstopmode "03Writeup/paper.tex"

### 2017-09-20 writeup/paper.tex
Use Make much more

--- 2017-09-19 04-Writeup/paper.tex
@@
 
-The climatological value of a field (rainfall or a circulation variable from reanalysis) is defined as the mean, at each grid cell, during the period $\qty[1980, 2010]$.
-The anomaly, for a particular date, is defined as the departure from this climatology for the corresponding month; the anomaly can be defined outside the range $\qty[1980, 2010]$.
+The climatological value of a field (rainfall or a circulation variable from reanalysis) is defined as the mean, at each grid cell, during all NDJF months in the period $\qty[1980, 2010]$.
+The anomaly, for a particular date, is defined as the departure from this climatology.
+Taking climatologies against the NDJF combined climatology, rather than against monthly climatologies, is necessary for weather typing (see below) as discontinuities in the climatological field lead to discontinuities in the anomaly field which leads to inconsistent and results which are difficult to interpret.
 
@@
 
-\Cref{fig:lagged-rain} shows time-lagged anomalies up to and after days exceeding the 90th percentile of area-averaged daily rainfall in the lower Paraguay River basin (see \cref{fig:study-area}), and is consistent with the findings of \citet{Marengo2004,Salio:2007gd}.
+\Cref{fig:lagged-rain} shows time-lagged anomalies up to and after days exceeding the 95th percentile of area-averaged daily rainfall in the lower Paraguay River basin (see \cref{fig:study-area}), and is consistent with the findings of \citet{Marengo2004,Salio:2007gd}.
 At $t=-2$ and $t=-1$ a low-level northerry jet injects inflow of heat and moisture to the region, favored by a polar trough extending slightly to the east of Paraguay.
@@
 Next, low-level convergence and upper-level divergence (not shown) associated with the eastward progression of the frontal system further favor rainfall locally.
-Similar analysis for the 85th or 95th percentiles of daily rainfall (not shown) yield highly similar results, implying that the mechanism for the most extreme events is not fundamentally distinct from the mechanism for moderate-intensity events.
+Similar analysis for the 90th or 99th percentiles of daily rainfall (not shown) yield highly similar results, implying that the mechanism for the most extreme events is not fundamentally distinct from the mechanism for moderate-intensity events.
 
@@
 	\caption{
-		Circulation anomalies associated with intense rainfall (99th percentile exceedance of area-averaged rainfall in the Lower Paraguay River Basin shown in \cref{fig:study-area}).
+		Circulation anomalies associated with intense rainfall (95th percentile exceedance of area-averaged rainfall in the Lower Paraguay River Basin shown in \cref{fig:study-area}).
 		Lagged composites are shown, by column, for $t = $ \SIlist{-2;-1;0;1}{\day} relative to the date of intense rainfall.

### 2017-09-20 01-Get/Rainfall.py

@@ -1,66 +1,0 @@
-'''
-Here we download Get the CPC data from the IRI Data Library.
-We only get rainfall estimate -- neglects information on number of stations used in analysis.
-Please see caveats and documentation on IRI Server at http://iridl.ldeo.columbia.edu/SOURCES/.NOAA/.NCEP/.CPC/.UNIFIED_PRCP/.GAUGE_BASED/.GLOBAL/.v1p0/
-'''
-
-import pandas as pd
-import numpy as np
-import xarray as xr
-import datetime
-import os
-from paraguayfloodspy.pars  import GetPars
-
-# Work with Times from the IRIDL
-def convert_t_to_time(Tvec):
-    times = np.array([datetime.date(1960, 1, 1) + datetime.timedelta(np.int(ti)) for ti in Tvec])
-    return(np.array(times))
-def convert_time_to_t(date):
-    date_diff = date - datetime.date(1960,1,1)
-    T = date_diff.days
-    return(T)
-
-def IRICPCYear(year, verbose=True):
-        if verbose:
-            print('Downloading data for {}...'.format(year))
-        # realtime or retro?
-        if year >= 1979 and year <= 2005:
-            url = 'http://iridl.ldeo.columbia.edu/SOURCES/.NOAA/.NCEP/.CPC/.UNIFIED_PRCP/.GAUGE_BASED/.GLOBAL/.v1p0/.RETRO/.rain/dods'
-        elif year >= 2006 and year <= 2016:
-            url = 'http://iridl.ldeo.columbia.edu/SOURCES/.NOAA/.NCEP/.CPC/.UNIFIED_PRCP/.GAUGE_BASED/.GLOBAL/.v1p0/.REALTIME/.rain/dods'
-        else:
-            raise ValueError('You have entered an invalid year. {} is outside range [1979, 2016]'.format(year))
-
-        # get the data
-        Tstart = convert_time_to_t(datetime.date(year, 1, 1))
-        Tend = convert_time_to_t(datetime.date(year, 12, 31))
-        ds = xr.open_dataarray(url, decode_times=False)
-        ds = ds.sel(T = slice(Tstart, Tend))
-        ds.load() # force it to download
-
-        # convert to more standard format
-        ds = ds.rename({'X': 'lon', 'Y': 'lat', 'T': 'time'})
-        ds['time'] = convert_t_to_time(ds['time'])
-        ds['time'] = ds['time'].astype('datetime64')
-        return(ds)
-
-def GetFileName(year):
[…]

### 2017-09-20 01-Get/Reanalysis.py

@@ -1,57 +1,0 @@
-'''
-Read in reanalysis data from NCEP-NCAR Reanalysis I data set.
-
-The approach to data management here is to try to minimize the number of times data needs to be re-downloaded (trial and error is a tough teacher...).
-Consequently, we first download a relatively complete data set (all years in our range, and all levels in our range).
-That is implemented in this notebook.
-Later, we calculate anomalies and climatologies for a subset of the data -- this is very quick and if we change our spatial or temporal domain, we don't need to re-download anything.
-
-Now, for each level, we will run through each year from `syear` to `eyear`.
-Then, for each year, we will:
-
- 1. Download the six-hour zonal ($u$) wind for that level and year and save to file
- 2. Download the six-hour meridional ($v$) wind for that level ande year and save to file
- 3. Calculate the six-hour streamfunction ($\psi$) for that year and save to file
-
-The `overwrite` parameter specified above determines whether the data is downloaded even if the file is already on disk -- this can be useful for downloading a bit at a time.
-'''
-
-import numpy as np
-import xarray as xr
-import os
-from paraguayfloodspy.pars  import GetPars
-from windspharm.xarray import VectorWind
-
-def main():
-    # Define the Parameters
-    time_pars = GetPars('time')
-    syear,eyear = time_pars['syear'], time_pars['eyear']
-    levels = [850]
-    years = np.arange(syear, eyear+1)
-    overwrite = False # overwrite existing data?
-
-    # Loop through and download each required file
-    for level_i in levels:
-        print("Now working on {}hPa data".format(level_i))
-        for year_i in years:
-            # Define URL and Output File Names
-            url_u = 'https://www.esrl.noaa.gov/psd/thredds/dodsC/Datasets/ncep.reanalysis/pressure/uwnd.{}.nc'.format(year_i)
-            url_v = 'https://www.esrl.noaa.gov/psd/thredds/dodsC/Datasets/ncep.reanalysis/pressure/vwnd.{}.nc'.format(year_i)
-            fname_u = '_data/reanalysis/raw/uwnd_{}_{}.nc'.format(level_i, year_i)
-            fname_v = '_data/reanalysis/raw/vwnd_{}_{}.nc'.format(level_i, year_i)
-            fname_streamfunc = '_data/reanalysis/raw/streamfunc_{}_{}.nc'.format(level_i, year_i)
-            if (os.path.isfile(fname_u) and os.path.isfile(fname_v) and os.path.isfile(fname_streamfunc)) and not overwrite:
[…]

### 2017-09-20 02-Process/Anomalies.py

@@ -1,97 +1,0 @@
-'''
-Subsetting, Anomalies, and Climatologies
-
-In a previous notebook (`Reanalysis`) we created a relatively large database of raw $U,V,\psi$.
-This is helpful, but for our visualizations it's nice to sub-set the data in advance so that all we have to do is load the relevant bits.
-We'll also pre-compute anomalies and climatologies, which saves substantial computational time down the road at the expense of a bit of file storage.
-Because we don't need the fine time resolution, we'll also re-sample everything to a daily time step (note that we didn't just download daily reanalysis data because we need to shift the end-of-day time!)
-'''
-
-import xarray as xr
-import glob
-import os
-from paraguayfloodspy.xrutil import *
-from paraguayfloodspy.pars import *
-
-def main():
-    pars = GetPars('all')
-    syear, eyear = pars['time']['syear'], pars['time']['eyear']
-    overwrite = False
-
-    # 1. REANALYSIS DATA
-    def transfer_fn(var, lonmin, lonmax, latmin, latmax, shift_time_h, return_daily, months):
-        var = var.sel(lon = slice(lonmin, lonmax), lat = slice(latmax, latmin))
-        # shift time by hours, if desired
-        if shift_time_h != 0:
-            time_old = var.time.values
-            time_new = time_old + np.timedelta64(12, 'h')
-            var['time'] = time_new
-        # resample to daily time step, if desired
-        if return_daily:
-            var = var.resample('1D', dim = 'time')
-        # subset months if required
-        if months is not None:
-            var = var.sel(time = np.in1d(var['time.month'], months))
-        return(var)
-
-    trans_lam = lambda ds: transfer_fn(
-        var=ds,
-        lonmin=pars['clim_plot']['lonmin'],
-        lonmax=pars['clim_plot']['lonmax'],
-        latmin=pars['clim_plot']['latmin'],
-        latmax=pars['clim_plot']['latmax'],
-        months=pars['months'],
-        shift_time_h=12, return_daily=True)
-
-    for var in ['streamfunc']:
-        for level_i in [850]:
-            fnames = glob.glob('_data/reanalysis/raw/{}_{}_*.nc'.format(var, level_i))
-            raw_fname = "_data/reanalysis/subset/{}_{}_raw.nc".format(var, level_i)
[…]

### 2017-09-20 scripts/access/CPCRainYear.py

@@ -1,0 +1,56 @@
+'''
+Download one year of CPC Rainfall Data from the IRI Data Library.
+We only get rainfall estimate -- neglects information on number of stations used in analysis.
+Please see caveats and documentation on IRI Server at http://iridl.ldeo.columbia.edu/SOURCES/.NOAA/.NCEP/.CPC/.UNIFIED_PRCP/.GAUGE_BASED/.GLOBAL/.v1p0/
+'''
+
+import pandas as pd
+import numpy as np
+import xarray as xr
+import datetime
+import argparse
+
+parser = argparse.ArgumentParser()
+parser.add_argument('--year', nargs=1, type=int, help='The year of data to download')
+parser.add_argument('--outfile', nargs=1, help='The output file to save')
+
+# Work with Times from the IRI DL
+def convert_t_to_time(Tvec):
+    times = np.array([datetime.date(1960, 1, 1) + datetime.timedelta(np.int(ti)) for ti in Tvec])
+    return(np.array(times))
+def convert_time_to_t(date):
+    date_diff = date - datetime.date(1960,1,1)
+    T = date_diff.days
+    return(T)
+
+def main():
+    # parse the arguments
+    args = parser.parse_args()
+    year = args.year[0]
+    outfile = args.outfile[0]
+
+    # Get the URL
+    if year >= 1979 and year <= 2005:
+        url = 'http://iridl.ldeo.columbia.edu/SOURCES/.NOAA/.NCEP/.CPC/.UNIFIED_PRCP/.GAUGE_BASED/.GLOBAL/.v1p0/.RETRO/.rain/dods'
+    elif year >= 2006 and year <= 2016:
+        url = 'http://iridl.ldeo.columbia.edu/SOURCES/.NOAA/.NCEP/.CPC/.UNIFIED_PRCP/.GAUGE_BASED/.GLOBAL/.v1p0/.REALTIME/.rain/dods'
+    else:
+        raise ValueError('You have entered an invalid year. {} is outside range [1979, 2016]'.format(year))
+
+    # get the data
+    Tstart = convert_time_to_t(datetime.date(year, 1, 1))
+    Tend = convert_time_to_t(datetime.date(year, 12, 31))
+    ds = xr.open_dataarray(url, decode_times=False)
+    ds = ds.sel(T = slice(Tstart, Tend))
+    ds.load() # force it to download
+
+    # convert to more standard format
[…]

### 2017-09-20 scripts/access/DailyIndices.py

@@ -3,12 +3,18 @@ Read in daily indices (here just MJO) and save to file
 '''
 
 import pandas as pd
-from paraguayfloodspy.pars  import GetPars
+import numpy as np
+import argparse
+
+parser = argparse.ArgumentParser()
+parser.add_argument('--years', type=int, nargs=2, help = 'LONMIN, LATMAX')
+parser.add_argument('--outfile', nargs=1, help='The output file to save')
 
 def main():
-    time_pars = GetPars('time')
-    syear,eyear = time_pars['syear'], time_pars['eyear']
-    outfile = "_data/indices/daily_indices.csv"
+
+    args = parser.parse_args()
+    syear, eyear = np.min(args.years), np.max(args.years)
+    outfile=args.outfile[0]
 
     raw_url = 'http://www.bom.gov.au/climate/mjo/graphics/rmm.74toRealtime.txt'
     col_names = ['year', 'month', 'day', 'RMM1', 'RMM2', 'phase', 'amplitude', 'source']

### 2017-09-20 scripts/access/MonthlyIndices.py

@@ -5,7 +5,11 @@ Read in monthly indices (ENSO and PDO) and save to file
 import pandas as pd
 import datetime
 import numpy as np
-from paraguayfloodspy.pars  import GetPars
+import argparse
+
+parser = argparse.ArgumentParser()
+parser.add_argument('--years', type=int, nargs=2, help = 'LONMIN, LATMAX')
+parser.add_argument('--outfile', nargs=1, help='The output file to save')
 
 def read_indices(raw_url, col_name):
     df = pd.read_table(raw_url, delim_whitespace=True,index_col=None, skiprows=2, names=['time', '{}'.format(col_name)])
@@ -19,9 +23,9 @@ def read_indices(raw_url, col_name):
     return(df)
 
 def main():
-    time_pars = GetPars('time')
-    syear,eyear = time_pars['syear'], time_pars['eyear']
-    outfile = "_data/indices/monthly_indices.csv"
+    args = parser.parse_args()
+    syear, eyear = np.min(args.years), np.max(args.years)
+    outfile=args.outfile[0]
 
     df = read_indices(
         raw_url = 'http://iridl.ldeo.columbia.edu/SOURCES/.Indices/.nino/.EXTENDED/.NINO34/gridtable.tsv',

### 2017-09-20 scripts/derive/Anomalies.py

@@ -1,0 +1,76 @@
+'''
+Subsetting, Anomalies, and Climatologies
+'''
+
+import xarray as xr
+import glob
+import numpy as np
+from paraguayfloodspy.xrutil import *
+import argparse
+
+parser = argparse.ArgumentParser()
+parser.add_argument('--mode', nargs=1, help='Either [rain] or [reanalysis]')
+parser.add_argument('--pattern', nargs=1, help='The pattern of files to read in')
+parser.add_argument('--latlim', type=float, nargs=2, help = 'LATMIN, LATMAX')
+parser.add_argument('--lonlim', type=float, nargs=2, help = 'LONMIN, LATMAX')
+parser.add_argument('--years', type=int, nargs=2, help = 'LONMIN, LATMAX')
+parser.add_argument('--outfile', nargs=1, help='The output file to save')
+
+def main():
+    args = parser.parse_args()
+    latmin, latmax = np.min(args.latlim), np.max(args.latlim)
+    lonmin, lonmax = np.min(args.lonlim), np.max(args.lonlim)
+    syear, eyear = np.min(args.years), np.max(args.years)
+
+    def transfer_fn(ds, mode, lonmin, lonmax, latmin, latmax):
+        '''
+        This function sub-sets data as we read it
+        '''
+        assert lonmax > lonmin, "lonmax must be greater than lonmin"
+        assert latmax > latmin, "latnmax must be greater than latmin"
+        if mode == 'rain':
+            ds = ds.sel(lon = slice(lonmin, lonmax), lat = slice(latmin, latmax))
+        elif mode == 'reanalysis':
+            ds = ds.sel(lon = slice(lonmin, lonmax), lat = slice(latmax, latmin))
+        else:
+            raise ValueError('mode must be rain or reanalysis')
+        return(ds)
+
+    # Read in raw data
+    fnames = glob.glob(args.pattern[0])
+    if len(fnames) < 1:
+        raise ValueError("Pattern {} doesn't match any files".format(args.pattern[0]))
+
+    # subset data when we read it in
+    trans_lam = lambda ds: transfer_fn(ds, mode=args.mode[0], lonmin=lonmin, lonmax=lonmax, latmin=latmin, latmax=latmax)
+    raw = read_netcdfs(fnames, dim='time', transform_func=trans_lam)
+    if args.mode[0] =='rain':
+        raw = raw['rain']
+    elif args.mode[0] == 'reanalysis':
[…]

### 2017-09-21 writeup/paper.tex
Re-vamp figures

@@
 Different types of analyses are used to diagnose the causes of the extreme rainfall events, and to bias-correct and verify the forecasts.
+
 All codes used for data and analyis are available at \url{https://github.com/jdossgollin/PYFloods}.
@@
 
-The $k$-means algorithm is guaranteed to converge to at least a local minimum; to select the best partition, 500 simulations were created using the implementation in Python's \texttt{scikit-learn} package \citep{Pedregosa:2012tv}.
+The $k$-means algorithm is guaranteed to converge to at least a local minimum; to select the best partition, 250 simulations were created using the implementation in Python's \texttt{scikit-learn} package \citep{Pedregosa:2012tv}.
 The classifiability index of \citet{Michelangeli1995} was then used to select the optimal partition.
@@
 
-\Cref{fig:lagged-rain} shows time-lagged anomalies up to and after days exceeding the 95th percentile of area-averaged daily rainfall in the lower Paraguay River basin (see \cref{fig:study-area}), and is consistent with the findings of \citet{Marengo2004,Salio:2007gd}.
-At $t=-2$ and $t=-1$ a low-level northerry jet injects inflow of heat and moisture to the region, favored by a polar trough extending slightly to the east of Paraguay.
+\Cref{fig:lagged-rain} shows time-lagged anomalies up to and after days exceeding the 95th percentile of area-averaged daily rainfall in the lower Paraguay River basin (see \cref{fig:study-area}), and is consistent with previous findings \citep{Marengo2004,Salio:2007gd}.
+At $t=-2$ a mid-latitude baroclinic system approaches the South American continent, intensifying and moving to the East at $t=-1,0,1$.
+Along the cold front associated with this system, a low-level northerly jet advects of heat and moisture to the region, favored by a polar trough extending slightly to the east of Paraguay.
 This is in agreement with previous studies which find a close correspondence between the exit regions of the low-level jet and the observed rainfall \citep{Saulo:2007km,Salio:2007gd,Marengo2004,Velasco1987} as the moisture influx, energy influx, and strong vertical wind shear favor instability and intense convection \citep{Marengo2004,Silva2009}.
-Next, low-level convergence and upper-level divergence (not shown) associated with the eastward progression of the frontal system further favor rainfall locally.
-Similar analysis for the 90th or 99th percentiles of daily rainfall (not shown) yield highly similar results, implying that the mechanism for the most extreme events is not fundamentally distinct from the mechanism for moderate-intensity events.
+Next, low-level convergence and upper-level divergence (not shown) associated with the eastward progression of the frontal system further favor rainfall over the Lower Paraguay River Basin.
+As the system progresses, the rainfall anomaly approximately follows the anomaly of the low-level wind.
+
+Similar analysis for the 90th or 99th percentiles of daily rainfall (not shown) yield similar results, implying that the synoptic mechanism for the most extreme events is not fundamentally distinct from the mechanism for moderate-intensity events.
 
@@
 
-We next use the weather typing algorithm to understand circulations and sequences of circulations associated with intense rainfall in the Lower Paraguay River Basin.
+We next use the weather typing algorithm to understand particular circulations and sequences of circulations associated with intense rainfall in the Lower Paraguay River Basin.
 
@@
 
-Extreme rainfall over the region was accurately predicted by the IRI for the DJF 2015-2016 season since at least November 2015 (see \cref{fig:subs-prob-fcst}).
-Probabilities of 30\%-50\% for the 90th percentile, three to five times the expected climatological values, are visible over southern Paraguay and Brazil, and northern Uruguay and Argentina, mostly in agreement with observations (hashed region in \cref{fig:probfcst}).
+Extreme rainfall over the region was accurately predicted by the IRI for the DJF 2015-2016 season since at least November 2015 (see \cref{fig:seas-prob-fcst}).
+Probabilities of 30\%-50\% for the 90th percentile, three to five times the expected climatological values, are visible over southern Paraguay and Brazil, and northern Uruguay and Argentina, mostly in agreement with observations (hashed region in \cref{fig:seas-prob-fcst}).
 Although this seasonal prediction was not skillful over eastern Paraguay, the forecast  showed a clear signal in the neighborhood, that could have been used for disaster-preparedness weeks in advance of the events.
@@
 As seen in \Cref{fig:subs-prob-fcst}a, the raw (uncorrected) sub-seasonal forecast of the ECMF model for Dec 1-7 2015, indicated very high odds for occurrence of extreme rainfall but with important biases in the actual location and spatial pattern; for Paraguay, it confidently suggests occurrence of extremes to the south-southeast of the country, which was mostly not observed.
-Overall, the 20-year based skill of probabilistic forecasts for the first week of December --even when using multiple initializations-- is worse than using climatology for most of Paraguay (\Cref{fig:subs-prob-fcst}b), being best for the zone between that country and northern Uruguay, and for most of southern Brazil.
+Overall, the 20-year based skill of probabilistic forecasts for the first week of December --even when using multiple initializations-- is worse than using climatology for most of Paraguay (\cref{fig:subs-prob-fcst}b), being best for the zone between that country and northern Uruguay, and for most of southern Brazil.
 
@@
 WE HAVE ENOUGH.
-	\item By examining \cref{fig:klee}, we can see that some weather types tend be more likely to occur during specific parts of the 120-day season; this is consistent with \citet{Munoz2016} and with the notion that there is some structure to the nonlinear dynamical system, which could support the land-surface feedbacks that are proposed by i.e.
+	\item By examining \cref{fig:rain-wt}, we can see that some weather types tend be more likely to occur during specific parts of the 120-day season; this is consistent with \citet{Munoz2016} and with the notion that there is some structure to the nonlinear dynamical system, which could support the land-surface feedbacks that are proposed by i.e.
 \citep{Grimm:2009bq}.
@@
 \begin{figure*}
-	\noindent\includegraphics[width=6.5in]{../_figs/StudyArea.pdf}
+	\noindent\includegraphics[width=6.5in]{../figs/StudyArea.pdf}
 	\caption{
@@
 \begin{figure*}
-	\noindent\includegraphics[width=6.5in]{../_figs/NDJF201516Anomaly.pdf}
+	\noindent\includegraphics[width=6.5in]{../figs/NDJF201516Anomaly.pdf}
 	\caption{
@@
 \begin{figure*}
-	\noindent\includegraphics[width=6.5in]{../_figs/PSI_EOF_Loadings.pdf}
+	\noindent\includegraphics[width=6.5in]{../figs/PSI_EOF_Loadings.pdf}
 	\caption{
@@
 \begin{figure*}
-	\noindent\includegraphics[width=6.5in,height=9in,keepaspectratio=true]{../_figs/WTComposite.pdf}
+	\noindent\includegraphics[width=6.5in,height=9in,keepaspectratio=true]{../figs/WTComposite.pdf}
 	\caption{
@@
 \begin{figure*}
-	\noindent\includegraphics[width=6.5in]{../_figs/RainfallWeatherType}
+	\noindent\includegraphics[width=6.5in]{../figs/RainfallWeatherType}
 	\caption{
@@
 
-\begin{figure*}	\noindent\includegraphics[width=6.5in]{../_figs/LaggedRain.pdf}
+\begin{figure*}	\noindent\includegraphics[width=6.5in]{../figs/LaggedRain.pdf}
 	\caption{
@@
 \begin{figure*}
-	\noindent\includegraphics[width=6.5in]{../_figs/Chiclet.pdf}
+	\noindent\includegraphics[width=6.5in]{../figs/Chiclet.pdf}
 	\caption{
@@
 \begin{figure*}
-	\noindent\includegraphics[width=6.4in]{../_figs/ForecastSkill.pdf}
+	\noindent\includegraphics[width=6.4in]{../figs/ForecastSkill.pdf}
 	\caption{
@@
 \begin{figure*}
-	\noindent\includegraphics[width=6.5in]{../_figs/SeasonalForecast.pdf}
+	\noindent\includegraphics[width=6.5in]{../figs/SeasonalForecast.pdf}
 	\caption{

### 2017-09-21 config/PlotParameters.py

@@ -1,0 +1,12 @@
+import numpy as np
+# Specify some parameters for plotting
+# this will keep the layout and look the same
+
+# Define color maps
+cmap = {'psi': 'PuOr', 'psi_a': 'PuOr', 'rain': 'Greens', 'rain_a': 'BrBG'}
+
+# Some extents for plotting
+extent = {'SAm': [275, 330, 5, -45], 'SH': [-140,10,-60,10]}
+
+# Some levels
+levels = {'psi': np.linspace(-2.4e7, 2.4e7, 13), 'psi_a': np.linspace(-3.5e6, 3.5e6, 8)}

### 2017-09-22 writeup/Supplement.tex
Use Make much more
Final weather types calculated w/ 500 sims

--- 2017-09-18 03Writeup/Supplement.tex
@@
 \begin{figure}
-	\includegraphics[width=\textwidth]{../_figs/Climatology.pdf}
-	\caption{
-		NDJF monthly circulation climatology over the study domain and from 1980 to 2010.
-    First row shows the \SI{850}{\hecto\pascal} streamfunction $\psi$ climatology, and second row shows rainfall climatology.
-	}
-\end{figure}
-
-\begin{figure}
-  \includegraphics[width=\textwidth]{../_figs/PSI_Var_Explained.pdf}
+  \includegraphics[width=\textwidth]{../figs/PSI_Var_Explained.pdf}
 	\caption{
@@
 \begin{figure}
-  \includegraphics[width=\textwidth]{../_figs/WT_Classifiability.pdf}
+  \includegraphics[width=\textwidth]{../figs/WT_Classifiability.pdf}
 	\caption{
 		Classifiability index (see Methods) calculated for several chosen values of $k$, representing the number of weather types used for clustering.
+		Only $n=50$ simulations were used for each value of $k$.
 	}
@@
 \begin{figure}
-  \includegraphics[width=\textwidth]{../_figs/WT_Occurrence_Proportion.pdf}
+  \includegraphics[width=\textwidth]{../figs/WT_Occurrence_Proportion.pdf}
 	\caption{
@@
 \begin{figure}
-  \includegraphics[width=\textwidth]{../_figs/ENSO_EOF_Scatter.pdf}
-	\caption{
-		Pair plots for three time series: NINO3.4, and EOFs 2 and 3 of the \SI{850}{\hecto\pascal} field over the chosen domain.
-    All time series are degraded to monthly time steps.
-    Across the diagonal, marginal distributions are shown.
-    Straight lines represent linear regression estimates.
-	}
-\end{figure}
-
-\begin{figure}
-  \includegraphics[width=\textwidth]{../_figs/MJO_EOF_Scatter}
-	\caption{
-		Pair plots for four time series: MJO RMM1 and RMM2, and EOFs 2 and 3 of the \SI{850}{\hecto\pascal} field over the chosen domain.
-    All time series are at daily time steps.
-    Across the diagonal, marginal distributions are shown.
-    Straight lines represent linear regression estimates.
-	}
-\end{figure}
-
-\begin{figure}
-  \includegraphics[width=\textwidth]{../_figs/knn_rainfall_EOF.pdf}
+  \includegraphics[width=\textwidth]{../figs/knn_rainfall_EOF.pdf}
 	\caption{
@@
 \begin{figure}
-  \includegraphics[width=\textwidth]{../_figs/ENSO_MJO_EOF_PRCP_Time_Series.pdf}
+  \includegraphics[width=\textwidth]{../figs/ENSO_MJO_EOF_PRCP_Time_Series.pdf}
 	\caption{

### 2017-09-22 writeup/paper.tex
Final weather types calculated w/ 500 sims

@@
 Locations of streamflow gauges are shown in \cref{fig:study-area}.
+Because no stage-discharge curves are available, we present only the river stage values.
+While this is relevant from the perspective of flood damage, flow rates cannot be estimated without these curves (which vary in time and consequently must be updated frequently).
+In a very flat river basin such as the Lower Paraguay River Basin, such one might expect that discharge scales approximately as $\sim O\qty(\text{stage}^3)$.
 
@@
 
-The climatological value of a field (rainfall or a circulation variable from reanalysis) is defined as the mean, at each grid cell, during all NDJF months in the period $\qty[1980, 2010]$.
-The anomaly, for a particular date, is defined as the departure from this climatology.
-Taking climatologies against the NDJF combined climatology, rather than against monthly climatologies, is necessary for weather typing (see below) as discontinuities in the climatological field lead to discontinuities in the anomaly field which leads to inconsistent and results which are difficult to interpret.
+We define the climatology of a circulation field $X$ (such as \SI{850}{\hecto\pascal} streamfunction) for each julian day using a moving-window approach.
+For each julian day $d = 1, \ldots, 366$, the climatological value of the field $\overline{X}(d)$ is defined as the mean value of all observations $X_t, t=1,\ldots,T$ whose julian day $d_t$ satisfies
+\begin{equation}
+	d_t \in \qty{\bmod \qty(d_i-30, \ldots, d_i + 30, 366)}
+\end{equation}
+where $\bmod$ is the modulo operator.
+The anomaly at a given time is then the observed field $X_i$ minus the climatological value for the corresponding julian day.
+Taking the climatology in this way is advantageous as it retains the seasonal cycle of observations while at the same time avoiding discontinuities in the climatology which is desirable for weather typing (see \cref{sec:methods}\ref{sec:weather-typing}).
 
-\subsection{Weather Typing}
+\subsection{Weather Typing} \label{sec:weather-typing}
 
-A cluster algorithm is used on daily data to help diagnose the mechanisms associated with the rainfall events of interest in this research.
+A cluster algorithm is used on daily data to diagnose mechanisms associated with the rainfall events of interest in this research.
 
-The clustering was performed on the daily NDJF $\SI{850}{\hecto\pascal}$ streamfunction field ($\psi$), calculated from the meridional and zonal wind fields using spherical harmonics as implemented in the \texttt{windspharm} python package \citep{Dawson:2016ws}, over the domain spanning \SIrange{15}{30}{\degree S} and \SIrange{65}{45}{\degree W} (\cref{fig:study-area}).
-To facilitate clustering (which tends to perform poorly in high-dimensional spaces), the anomaly field of $\psi$ (see \cref{sec:climatology-anomalies}) was projected  onto its four leading empirical orthogonal functions (EOFs), accounting for 95\% of the total observed variance (supplemental figure S1).
-The \texttt{eofs} python package was used to perform this dimension reduction \citep{Dawson:2016ge}.
-No additional time filtering was applied to the data before clustering, thus retaining the annual seasonal cycle, inter-annual, sub-seasonal and synoptic weather time scales for diagnosis.
-No latitudinal weighting was necessary as the selected domain is relatively small and it does not extend into high latitudes.
-Once the EOFs were calculated, the principal component time series were computed for each day and scaled to unit variance.
+The clustering was performed on the daily NDJF $\SI{850}{\hecto\pascal}$ streamfunction field ($\psi$), calculated by integrating the meridional and zonal wind fields using spherical harmonics \citep{Dawson:2016ws}, over the domain spanning \SIrange{15}{30}{\degree S} and \SIrange{65}{45}{\degree W} (\cref{fig:study-area}).
 
-Next, the $k$-means algorithm was used to assign a single cluster value to each day on record.
-The $k$-means technique is a partitioning method that classifies all days in study into a predefined number of clusters.
-It minimizes the intracluster sum of variance, $\mathcal{W}$, for a partition $\mathcal{P}$ defined as the Voronoi diagram generated by the classification process:
+To facilitate clustering (which tends to perform poorly in high-dimensional spaces), the NDJF anomaly field of $\psi_{850}$ (see \cref{sec:climatology-anomalies}) was projected  onto its four leading empirical orthogonal functions (EOFs), accounting for over 95\% of the total observed variance (supplemental figure S1), using the \texttt{eofs} python package \citep{Dawson:2016ge}.
+The loading patterns of these EOFs are shown in \cref{fig:eof-loading}.
+No meridional weighting was necessary as the selected domain is relatively small and does not extend into high latitudes.
+Once the EOFs were calculated, the principal component time series were computed for each day and scaled to unit variance as is typical for cluster analysis.
+
+Next, the $K$-means algorithm was used to assign a single cluster value to each day on record using the 4-dimension principal component time series computed in the previous step.
+The $K$-means technique is a partitioning method that classifies all days in the study into a predefined number of clusters.
+The algorithm proceeds as follows:
+\begin{enumerate}
+	\item Randomly choose $K$ cluster centers $\mu_1^{(0)}, \ldots, \mu_K^{(0)}$ (where $0$ refers to the $0$th iteration)
+	\item Iterate until convergence, indexing each iteration with $j$:
+	\begin{enumerate}
+		\item Assign each observation (day) $x_i$ to the nearest cluster center; we use the Euclidean distance:
 		\begin{equation}
-	\mathcal{W} \qty( \mathcal{P} ) = \sum_{i=1}^{k}\sum_{\vb{x} \in \vb{p}_{i}} \abs{\abs{\vb{x} - \vb{p}_i}}
+			m_i^{(j+1)} := \arg\min_{k\in 1,\ldots,K} \abs{\abs{x_i - \mu_k^{(j)}}}
 		\end{equation}
-where $k$ is the total number of classes in the solution, $\abs{\abs{\cdot}}$ is the Euclidan norm, $\vb{p}_i$ the different clusters in the partition and $\vb{x}$ the maps corresponding to cluster $i$.
+		\item Recompute the cluster centers as the mean of all points assigned to that cluster
+		\begin{equation}
+			\mu_k^{(j+1)} := \frac{1}{\abs{\qty{i \big| m_i^{(j+1)}=k}}} \sum_{i | m_i^{(j+1)}=k} x_i
+		\end{equation}
+		where $\abs{\cdot}$ denotes vector length.
+		\item Stop iteration if the change in centroids $\vec{\mu}^{(j+1)} - \vec{\mu}^{(j)}$ is less than a small but non-zero tolerance parameter $\tau$.
+	\end{enumerate}
+\end{enumerate}
+The cluster centroids (weather types) $\mu_k$ produced by the $K$-means algorithm can then be interpreted as a Voronoi decomposition of the phase space into $K$ regions.
+Intuitively, this means that an observation is assigned to the weather type whose centroid $\mu_k$ is closest to it in the phase space given by the leading 4 EOFs of $\psi_{850}.$
 
-The $k$-means algorithm is guaranteed to converge to at least a local minimum; to select the best partition, 250 simulations were created using the implementation in Python's \texttt{scikit-learn} package \citep{Pedregosa:2012tv}.
-The classifiability index of \citet{Michelangeli1995} was then used to select the optimal partition.
+The $K$-means algorithm is guaranteed to converge to a local minimum of intercluster variance; to select the best partition, 500 simulations were created using the implementation in Python's \texttt{scikit-learn} package \citep{Pedregosa:2012tv} and then the classifiability index of \citet{Michelangeli1995} was then used to select the optimal partition.
+This index can be conceptualized as a stability criterion as it selects the partitioning which is most like the other 249 partitions, as defined by Euclidean distance.
 Calculation of the classifiability index for several values of $k$ (supplemental figure S2) suggests that states with $k=4,5,6,7,8$ are all reasonable; we chose the solution $k=6$ because the weather types identified are qualitatively similar to those determined over southeastern South America \citep{Munoz2015,Munoz2016} and have an intuitive physical meaning, discussed further in the following sections.
@@
 \Cref{fig:streamflow} shows the streamflow time series at several gauges on the Paraguay River during NDJF 2015-16 in the context of their seasonality and decadal variability.
-Because no stage-discharge curves are available, we present only the river stage.
-While this is relevant from the perspective of flood damage, flow rates cannot be estimated without these curves (which are also highly variable in time and consequently must be updated frequently).
-In a very flat river basin such as the Lower Paraguay River Basin, such one might expect to expect that discharge scales approximately with $\text{stage}^3$.
-
 The most notable feature of \cref{fig:streamflow}(a) is that during November and December 2015, the river rose rapidly at Concepci\'on, Asunci\'on,  and Pilar but not at Bah\'ia Negra.
@@
 At $t=-2$ a mid-latitude baroclinic system approaches the South American continent, intensifying and moving to the East at $t=-1,0,1$.
-Along the cold front associated with this system, a low-level northerly jet advects of heat and moisture to the region, favored by a polar trough extending slightly to the east of Paraguay.
-This is in agreement with previous studies which find a close correspondence between the exit regions of the low-level jet and the observed rainfall \citep{Saulo:2007km,Salio:2007gd,Marengo2004,Velasco1987} as the moisture influx, energy influx, and strong vertical wind shear favor instability and intense convection \citep{Marengo2004,Silva2009}.
-Next, low-level convergence and upper-level divergence (not shown) associated with the eastward progression of the frontal system further favor rainfall over the Lower Paraguay River Basin.
-As the system progresses, the rainfall anomaly approximately follows the anomaly of the low-level wind.
+This system interacts with the sub-tropical low and the Andes Mountains (not shown) to produce an anticyclonic anomaly over Brazil.
+Along the cold front associated with this system, a low-level northerly jet advects of heat and moisture to the region.
+As the system progresses, the jet transitions from predominantly meridional flow (Chaco Jet; $t=-1$ to predominantly zonal flow below approximately \SI{20}{\degree N}.
+
+The rainfall largely follows the pattern of the low-level jet, in agreement with previous studies  \citep{Saulo:2007km,Salio:2007gd,Marengo2004,Velasco1987}.
 
@@
 
-We next use the weather typing algorithm to understand particular circulations and sequences of circulations associated with intense rainfall in the Lower Paraguay River Basin.
+We next use the weather typing algorithm outlined in \cref{sec:methods}\ref{sec:weather-typing} to understand particular circulations and sequences of circulations associated with intense rainfall in the Lower Paraguay River Basin.
 
 The first step of the weather typing algorithm is to identify leading EOFs of the \SI{850}{\hecto\pascal} streamfunction $\psi$.
-The EOF laodings are shown in \cref{fig:eof-loading}.
-Of these, EOF 1 explains by far the most variance ($\approx 67\%$) while EOFs 2, 3, and 4 collectively explain approximately $28\%$ of total variance (supplemental figure S2).
-To a first order of approximation, EOF 1 explains the large-scale anomalies induced by passing baroclinic waves (weaker in the West, likely due to the Andes Mountains).
-EOF 2 indicates the strength of the low-level jet circulation, while EOFs 3 and 4 control the exit region of this jet.
+The EOF loadings are shown in \cref{fig:eof-loading}.
+Of these, EOF 1 explains by far the most variance ($\approx 67\%$) while EOFs 2, 3, and 4 collectively explain approximately $28\%$ of total variance (supplemental figure S1).
+To a first order of approximation, EOF 1 explains the large-scale anomalies induced by passing baroclinic waves; the East-West gradient is likely a result of the influence of the Andes Mountains which are a source of friction for low-level flow.
+EOF 2 primarily controls the strength of the low-level jet circulation, while EOFs 3 and 4 control the exit region of this jet.
 Given these features in the EOF loadings, it is no surprise that composites of the composite \SI{850}{\hecto\pascal} circulations and rainfall associated with each weather type (\cref{fig:wt-composite}) reveal circulation patterns are associated with synoptic- and regional-scale circulation patterns.
 
-As predicted, there is a close correspondence between the exit regions of the low-level jet and the observed rainfall.
-Weather type 1 represnts a ``Chaco Jet'' event \citep{Salio:2002ev} in which the maximum wind speed penetrates past \SI{25}{\degree S}, leading to intense rainfall NE Argentina and Uruguay.
+Weather type 1 represents a ``Chaco Jet'' event \citep{Salio:2002ev} in which the maximum wind speed penetrates past \SI{25}{\degree S}, leading to intense rainfall NE Argentina and Uruguay.
 The other weather type associated with a SALLJ is weather type 4, the ``No-Chaco Jet'' event in which the wind turns to the East before \SI{25}{\degree S}, leading to heavy rainfall over Eastern Paraguay and SW Brazil.
@@
 \begin{figure*}
-	\noindent\includegraphics[width=6.5in]{./Streamflow.pdf}
+	\noindent\includegraphics[width=6.5in]{Streamflow.pdf}
 	\caption{
 		River stage (height; in \si{\meter}) for the Paraguay River at four gauges along the Paraguay River.
+		The station names are shortened versions of those shown in \cref{fig:study-area}.
 		(a) Seasonality (orange) and time series of 2015-16 observations (black) at each stream gauge.
@@
     	Monthly composite anomalies observed during NDJF 2015-16.
-      Top row shows anomalies of streamfunction $\psi$ at \SI{850}{\hecto\pascal}.
-      Bottom row shows monthly rainfall anomalies, in units of \si{\milli\meter\per\day}.
+      Top row shows absolute monthly mean values (color) and monthly-mean anomalies (contour) of streamfunction at \SI{850}{\hecto\pascal} $\psi_{850}$.
+			Contour interval for anomalies is \SI{1e6}{\meter\squared\per\second}.
+      Bottom row shows monthly-mean rainfall anomalies, in units of \si{\milli\meter\per\day}.
 	}
@@
     	Leading four EOF loadings of daily NDJF \SI{850}{\hecto\pascal} streamfunction $\psi$ over chosen domain.
+			EOFs are calculated after removing the climatology; see \cref{sec:methods}\ref{sec:climatology-anomalies} for details.
 	}
@@
 		Composite anomalies associated with each weather type (column).
-    Top row shows anomalies of streamfunction $\psi$ at \SI{850}{\hecto\pascal}.
+		Top row shows absolute monthly mean values (color) and monthly-mean anomalies (contour) of streamfunction at \SI{850}{\hecto\pascal} $\psi_{850}$.
+		Contour interval for anomalies is \SI{1e6}{\meter\squared\per\second}.
     Bottom row shows precipitation anomalies in units of \si{\milli\meter\per\day}.
@@
 	\caption{
-		Time series of area-averaged rainfall in the Lower Paraguay River Basin (\cref{fig:study-area}) for each day of NDJF  2015-16 (lines).
-		The weather types corresponding to each day are indicated in both color and adjacent label.
+		Time series of area-averaged rainfall in the Lower Paraguay River Basin (\cref{fig:study-area}) for each day of NDJF  2015-16 (lines and points).
+		The weather types corresponding to each day are indicated in both color and in an adjacent text label.
 		Dashed lines indicate the climatological 50th, 90th, and 99th percentiles of NDJF area-averaged rain over the Lower Paragauy River basin.
@@
 		Lagged composites are shown, by column, for $t = $ \SIlist{-2;-1;0;1}{\day} relative to the date of intense rainfall.
-		Top row shows anomalies of streamfunction $\psi$ at \SI{850}{\hecto\pascal}.
+		Top row shows absolute monthly mean values (color) and monthly-mean anomalies (contour) of streamfunction at \SI{850}{\hecto\pascal} $\psi_{850}$.
+		Contour interval for anomalies is \SI{1e6}{\meter\squared\per\second}.
     Bottom row shows precipitation anomalies in units of \si{\milli\meter\per\day}.

### 2017-09-22 config/PlotParameters.py

@@ -6,7 +6,4 @@ import numpy as np
 cmap = {'psi': 'PuOr', 'psi_a': 'PuOr', 'rain': 'Greens', 'rain_a': 'BrBG'}
 
 # Some extents for plotting
-extent = {'SAm': [275, 330, 5, -45], 'SH': [-140,10,-60,10]}
-
-# Some levels
-levels = {'psi': np.linspace(-2.4e7, 2.4e7, 13), 'psi_a': np.linspace(-3.5e6, 3.5e6, 8)}
+extent = {'SAm': [275, 330, 5, -40], 'SH': [-120, 0, 5, -50]}

### 2017-09-22 scripts/access/StreamfuncYear.py

@@ -5,8 +5,6 @@ Data from NCAR/NCEP Reanalysis I 6-hourly
 
 import numpy as np
 import xarray as xr
-import os
-from paraguayfloodspy.pars  import GetPars
 from windspharm.xarray import VectorWind
 import argparse
 

### 2017-09-25 writeup/Supplement.tex
Use Reanalysis II data

@@
 	siunitx, % units
-	cleveref,% cross-references
-	physics, % better physics notation
-	url, % for URLs
-	amsmath, % for math
-	array, % for fixed-width tables
   graphicx, % for images
+	natbib
 }
-
 \usepackage[letterpaper]{geometry}
 \renewcommand{\thefigure}{S\arabic{figure}}
+\usepackage[breaklinks=true]{hyperref}
 
@@
 \begin{figure}
+	\centering
+  \includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio=true]{gpw-v4-2015.png}
+	\caption{
+		Gridded estimate of population density (color; in units of persons per square kilometer) from \citet{GPWv4}.
+	}
+\end{figure}
+
+\begin{figure}
   \includegraphics[width=\textwidth]{../figs/PSI_Var_Explained.pdf}
@@
 		Variance explained for each EOF loading of the \SI{850}{\hecto\pascal} streamfunction $\psi$.
-    Left figure shows the individual variance explained for each EOF and right figure shows cumulative variance explained.
+    (L): the individual variance explained for each EOF.
+		(R): cumulative variance explained.
 	}
@@
 	\caption{
-		Classifiability index (see Methods) calculated for several chosen values of $k$, representing the number of weather types used for clustering.
-		Only $n=50$ simulations were used for each value of $k$.
+		Classifiability index \citep[see Methods and][]{Michelangeli1995} calculated for several chosen values of $K$, representing the number of weather types produced.
+		Results are produced with only 50 simulations for each $K$.
 	}
@@
     Both values represent area-averaged rainfall time series.
-    Expected rainfall is calculated by using a nearest-neighbors regression with 100 neighbors.
-    Because this methodology performs poorly under extrapolation (where there are few observations), the predicted domain is limited to include only regions with high density of observations.
+    Expected rainfall is calculated by using a nearest-neighbors regression with 100 neighbors in the two-dimensional phase space of EOF2 and EOF3.
+    Because nearest-neighbors approaches tend to perform poorly under extrapolation (where there are few observations), the domain shown is limited to regions with high density of observations.
   }
@@
 
+\clearpage
+\bibliographystyle{ametsoc2014}
+\bibliography{library}
+
 \end{document}

### 2017-09-25 writeup/paper.tex
Use Reanalysis II data

@@
 
-% THIS PARAGRAPH NEEDS SOME WORK
-% COME BACK TO IT LATER
-During the austral summer of 2015-16, repeated intense rainfall events led to severe flooding on the Paraguay River, displacing approximately \num{170000} people \citep{BrackenridgeDFO} and generating tremendous property damage.
-The Paraguay River (\cref{fig:study-area}) is the primary tributary of the Paran\'{a} River and an important contributor to the La Plata basin, and its flooding directly impacts the Paraguayan cities of Asunci\'{o}n, Concepci\'{o}n, and Pilar, and the Argentine cities of Formosa and Corrientes.
-Intense rainfall and flooding along the Paraguay River also has important implications for hydropower generation along the Paran\'{a} River, for agriculture, and for regional water resource management.
-Until now, the drivers of the November-February (NDJF) 2015-16 rainfall events and the extent to which they could have been skilfully predicted has not been analyzed in the scientific literature.
-The goal of this paper is to address these issues.
+During the austral summer of 2015-16, repeated intense rainfall events led to severe flooding in the Paraguay River basin, displacing approximately \num{170000} people \citep{BrackenridgeDFO} and causing tremendous damage to property and infrastructure \citep{MinisteriodeObrasPublicasyComunicacion:2016tq}.
+Because population in South America tends to concentrate along coasts and rivers (see supplemental figure S1), flooding in the Paraguay River basin directly affects not only much of the population of Paraguay, but also of populations in Argentina and Uruguay who lie along the Paran\'{a} and la Plata rivers into which the Paraguay River drains.
+Intense rainfall and flooding in the Paraguay River basin also has important implications for hydropower generation, for agriculture, and for regional water resource management.
+The aim of this paper is to diagnose the drivers of the November-February (NDJF) 2015-16 rainfall and flooding events and analyze the extent to which they were skillfully predicted.
 
-% Climatology and past floods
 The rainfall climatology of the Paraguay River basin varies strongly by season, with extratropical characteristics in the winter and monsoonal characteristics in the summer.
-In the summer season (NDJF), the focus of this study, circulation is dominated by the upper-tropospheric Bolivian High, the lower-level subtropical highs, the Chaco Low over northern Argentina, the South Atlantic Convergence Zone (SACZ), and the South American Low-Level Jet, or SALLJ \citep{Grimm2009,Marengo:2012cm}.
-Rainfall in the Lower Paraguay River Basin peaks around \SI{5}{\milli\meter\per\day} during the warm months (October-May) and reaches a minimum near \SI{2}{\milli\meter\per\day} in July and August.
+In the warm season (NDJF) which is the focus of this study, circulation is dominated by the upper-tropospheric Bolivian High, the lower-level subtropical highs, the Chaco Low over northern Argentina, the South Atlantic Convergence Zone (SACZ), and the South American Low-Level Jet, or SALLJ \citep{Grimm2009,Marengo:2012cm}.
+Rainfall peaks around \SI{5}{\milli\meter\per\day} during the warm months (October-May) and reaches a minimum near \SI{2}{\milli\meter\per\day} in July and August.
 However, the flat topography limits the ability to carry the summer runoff, causing seasonal innundation of the Pantanal and distributing the river discharge in time \citep{Bravo:2011et,Barros:2004bn}.
-Thus, the streamflow maxima typically occurs in phase with precipitation upstream of the Pantanal, while downstream of the Pantanal -- over the Lower Paraguay River -- the annual peak, including for extreme flood events, typically occurs between April and July \citep{Barros:2004bn}.
+Thus, the streamflow maxima typically occurs in phase with precipitation upstream of the Pantanal, while downstream of the Pantanal -- which we define as the Lower Paraguay River Basin in \cref{sec:methods} -- the annual peak typically occurs between April and July \citep{Barros:2004bn}.
 
-% SALLJ and MCS
 A large fraction of rainfall, and nearly all intense rainfall, in the Paraguay River Basin during the summer is associated with mesoscale convection \citep{Velasco1987}.
@@
 The strength and direction of this moisture transport varies substantially between events; SALLJ exit regions range from central Argentina \citep[``Chaco Jet Events'';][]{Salio:2002ev} to southeastern Brazil.
-In addition, the relationship between the low-level jet and convection involves two-way feedback \citep{Saulo:2007km}.
 
-% S2S Drivers
 At sub-seasonal timescales, convection in the Paraguay River Basin is modulated by a variety of drivers, notably including the SACZ and the Madden-Julien Oscillation (MJO).
 During SACZ (NoSACZ) conditions, strong (weak) convergence is observed over the Amazon basin with divergence (convergence) over southwestern Brazil, northern Argentina and Paraguay \citep{Herdies:2002jy,Carvalho2010}.
-SACZ occurrence is related to westerly wind regimes, as well as ``active'' and ``break'' periods of the South American Monsoon \citep{Marengo2004}.
+SACZ occurrence is related to westerly wind regimes, as well as ``active'' and ``break'' periods of the South American Monsoon System \citep{Marengo2004}.
 \citet{Carvalho2004} and \citet{Jones2002} demonstrate that mid-latitude dynamics influence low-level wind anomalies on many time scales, but also find that tropical-extratropical interactions complicate attempts to find one-way casual relationships.
 
-% Seasonal Drivers
 At seasonal timescales, El Ni\~no-Southern Oscillation (ENSO) is the dominant driver of convection variability in the Paraguay River Basin.
 During El Ni\~no years, a low-level anticyclonic anomaly over central Brazil enhances occurrence of the low-level jet, favoring the development of mesoscale convective systems \citep{Velasco1987}.
-The impact of ENSO events is felt partly by changes in the SACZ, which change due to ENSO-induced extratropical Rossby wave trains \citep{Grimm:2011vp}.
-There is also intraseasonal variability of rainfall during El Ni\~no years, typically including a reversal between November and January of the following year, which are likely dominated by land-surface interactions \citep[see][]{Grimm2009,Grimm2003}.
-These findings suggest that while interannual variability during the spring may be driven by remote processes,  during the summer local processes dominate.
-In particular, regional land-surface feedbacks can cause regions that exhibit wet anomalies in the spring to experience, on average, more summer precipitation \citep{Grimm:2009bq}.
+% So a key feature to note is that the intensity and the precise extent of this anomaly will be relevant for ENSO impacts
+\citet{Grimm2009,Grimm2003} also note substantial intraseasonal variability of rainfall during El Ni\~no years, including a reversal of rainfall anomalies between November[0] and January[+], likely dominated by land-surface interactions.
+Even beyond El Ni\~no years, regional land-surface feedbacks can cause regions that exhibit wet anomalies in the spring to experience more summer precipitation on average \citep{Grimm:2009bq}.
 
-The paper continues as follows.
-In \cref{sec:data} we describe our data sources, while in \cref{sec:methods} we provide an overview of technical methods carried out.
-In \cref{sec:diag} we diagnose the circulation patterns associated with intense rainfall during NDJF 2015-16, using composite and weather typing approaches.
-In \cref{sec:fcsts} we examine forecasts of rainfall in the Lower Paraguay River Basin and assess forecast skill, with special attention to the first week of December, when the most important flooding events started.
-In \cref{sec:discussion} we discuss potential implications of our findings and potential future work.
-Finally, in \cref{sec:summary} we summarize our results.
+We proceed as follows.
+We begin by describing our data sources in \cref{sec:data} and our methods in \cref{sec:methods}.
+In \cref{sec:flood} we highlight observed flooding and contextualize it within a long river stage time series.
+In \cref{sec:diagnostics} we use composites and a weather typing analysis to diagnose the circulation patterns associated with intense rainfall during NDJF 2015-16.
+Having built a this diagnostif framwork, we then turn in \cref{sec:fcsts} to the question of whether the observed rainfall was successfully predicted.
+To carry out this analysis we study both forecasts targeting the entire series for a limited area, and also forecasts targeting a large spatial area for only the first week of December, when the most important flooding events started.
+We also explore the impact on forecasts of several popular bias correction schemes.
+Finally in \cref{sec:discussion} we discuss limitations and potential implications of our findings and potential future work, which we summarize in \cref{sec:summary}.
 
@@
 
-The analysis involves observations and model forecasts.
+The core analysis presented involves both observations and model forecasts.
 
@@
 
-The period analyzed for diagnostics purposes is from 01 Nov 1979 through 29 Feb 2016.
+The period analyzed for diagnostics purposes is from 01 Nov 1979 through 28 Feb 2016.
+\Cref{fig:study-area} shows the study area and delineates several spatial domains which are discussed in this section and the next.
 
@@
 Spatial resolution is \SI{0.5}{\degree}, and temporal resolution is daily.
-Both the retro and real-time datasets are used.
+The retro and real-time datasets are blended and treated as a single dataset.
 
-Atmospheric circulations are diagnosed using daily data from the NCAR-NCEP Reanalysis I data set \citep{Kalnay1996}, available online at \url{https://www.esrl.noaa.gov/psd/thredds/dodsC/Datasets/ncep.reanalysis/}.
+Atmospheric circulations are diagnosed using daily data from the NCAR-NCEP Reanalysis II data set \citep{Kanamitsu:2002kk}, available online at \url{https://www.esrl.noaa.gov/psd/thredds/dodsC/Datasets/ncep.reanalysis2/}.
 Spatial resolution is \SI{2.5}{\degree}, which is adequate for synoptic-scale analyses.
-
-Because the end-of-day time for the rainfall data is 12:00GMT over most of South America, we use six-hour reanalysis data, and shift by twelve hours before re-sampling to the daily time step.
+Because the end-of-day time for the rainfall data is 12:00GMT over most of South America \citep{xie2010cpc}, we use six-hour reanalysis data, and shift by twelve hours before re-sampling to the daily time step.
 This ensures that the time steps in the reanalysis and rainfall data sets are the same, but means that a day is defined as beginning at 12:00 GMT.
@@
 Locations of streamflow gauges are shown in \cref{fig:study-area}.
-Because no stage-discharge curves are available, we present only the river stage values.
-While this is relevant from the perspective of flood damage, flow rates cannot be estimated without these curves (which vary in time and consequently must be updated frequently).
-In a very flat river basin such as the Lower Paraguay River Basin, such one might expect that discharge scales approximately as $\sim O\qty(\text{stage}^3)$.
-
-\subsection{Climate Indices}
+Because no stage-discharge curves are available, we present only the river stage values; while this is relevant from the perspective of flood damage, flow rates cannot be estimated without these curves (which are difficult to reconstruct as river geometry changes over time).
+In a very flat river basin such as this, such one might expect that discharge scales approximately as $\sim O\qty(\text{stage}^{3/2})$.
+% THIS MAY NOT BE RIGHT
+% MANU SAYS LOOK AT BROAD WEIR
 
@@
 
-This study analyzes probabilistic seasonal and sub-seasonal forecasts of extremely wet rainfall events, defined in the same way as in observations (90th percentile).
+This study analyzes probabilistic seasonal and sub-seasonal forecasts of intense rainfall events, defined as for observations by exceedance of the 90th percentile of modeled rainfall.
 
-The sub-seasonal forecasts used were isseud by the European Centre for Medium-Range Weather Forecasts (ECMWF) using a coupled model, available via the Subseasonal-to-Seasonal Prediction Project Database \citep{Vitart2016}.
-The horizontal resolution is \SI{1.5}{\degree} deg.
+The sub-seasonal forecasts used were issed by the European Centre for Medium-Range Weather Forecasts (ECMWF) using a coupled model, available via the Subseasonal-to-Seasonal Prediction Project Database \citep{Vitart2016} at \SI{1.5}{\degree} resolution.
 Forecasts consider the period starting in Dec 2015 until Mar 2016 and hindcasts to assess the real-time predictive skill consider the period Dec 1-7, 1995-2014.
-There is a total of 51 ensemble members for each forecast, and 11 ensemble members for each one of the 20 hindcasts (Dec 1-7, 1995-2014).
+There is a total of 51 ensemble members for each forecast, plus 11 ensemble members for each of the 20 hindcasts (Dec 1-7, 1995-2014).
 
@@
 Different types of analyses are used to diagnose the causes of the extreme rainfall events, and to bias-correct and verify the forecasts.
+All codes used for data and analyis are available at \url{https://github.com/jdossgollin/PYFloods}.
+Computation with was performed in the python environment using the \texttt{xarray} \citep{hoyer2017xarray} and \texttt{numpy} \citep{vanderWalt:2011dp} packages for gridded data and the \texttt{pandas} \citep{McKinney:2010un} package for tabular data; plotting was performed with the \texttt{matplotlib} \citep{Hunter:2007ux} and \texttt{cartopy} \citep{Cartopy} packages.
 
-All codes used for data and analyis are available at \url{https://github.com/jdossgollin/PYFloods}.
-Computation with multidimensional data was performed using the \texttt{xarray} \citep{hoyer2017xarray} and \texttt{numpy} \citep{vanderWalt:2011dp} packages of the python computing environment; computations on tabular data utilized the \texttt{pandas} \citep{McKinney:2010un} package; and plotting was performed with the \texttt{matplotlib} \citep{Hunter:2007ux} and \texttt{cartopy} \citep{Cartopy} packages.
+Given the behavior of the Paraguay River discussed above, we define the Lower Paraguay River Basin as the region bounded by \SIrange{-59.75}{-55.75}{\degree W} and \SIrange{26.75}{22.75}{\degree S}, as shown in \cref{fig:study-area}.
+This region corresponds to the region over which, given topography and previous studies \citep{Barros:2004bn,Bravo:2011et}, one might hypothesize rainfall inputs to most closely correspond to river levels at the stream gauges in \cref{fig:study-area}.
+Results are not sensitive to small variations in this domain.
 
@@
 We define the climatology of a circulation field $X$ (such as \SI{850}{\hecto\pascal} streamfunction) for each julian day using a moving-window approach.
-For each julian day $d = 1, \ldots, 366$, the climatological value of the field $\overline{X}(d)$ is defined as the mean value of all observations $X_t, t=1,\ldots,T$ whose julian day $d_t$ satisfies
-\begin{equation}
-	d_t \in \qty{\bmod \qty(d_i-30, \ldots, d_i + 30, 366)}
-\end{equation}
-where $\bmod$ is the modulo operator.
+For each julian day $d = 1, \ldots, 366$, the climatological value of the field $\overline{X}(d)$ is defined as the (unweighted) mean value of all days whose julian day is within $\pm 30$ of $d$, accounting for the discontinuity at 01 January.
 The anomaly at a given time is then the observed field $X_i$ minus the climatological value for the corresponding julian day.
@@
 
-The clustering was performed on the daily NDJF $\SI{850}{\hecto\pascal}$ streamfunction field ($\psi$), calculated by integrating the meridional and zonal wind fields using spherical harmonics \citep{Dawson:2016ws}, over the domain spanning \SIrange{15}{30}{\degree S} and \SIrange{65}{45}{\degree W} (\cref{fig:study-area}).
+The clustering was performed on the daily NDJF $\SI{850}{\hecto\pascal}$ streamfunction field ($\psi$), calculated by integrating the meridional and zonal wind fields using spherical harmonics \citep{Dawson:2016ws,Adams:1999fb} over the domain spanning \SIrange{15}{30}{\degree S} and \SIrange{65}{45}{\degree W} (\cref{fig:study-area}).
 
-To facilitate clustering (which tends to perform poorly in high-dimensional spaces), the NDJF anomaly field of $\psi_{850}$ (see \cref{sec:climatology-anomalies}) was projected  onto its four leading empirical orthogonal functions (EOFs), accounting for over 95\% of the total observed variance (supplemental figure S1), using the \texttt{eofs} python package \citep{Dawson:2016ge}.
+To facilitate clustering (which tends to perform poorly in high-dimensional spaces), the NDJF anomaly field of $\psi_{850}$ (see \cref{sec:climatology-anomalies}) was projected  onto its four leading empirical orthogonal functions (EOFs), accounting for over 95\% of the total observed variance (supplemental figure S2), using the \texttt{eofs} python package \citep{Dawson:2016ge}.
 The loading patterns of these EOFs are shown in \cref{fig:eof-loading}.
-No meridional weighting was necessary as the selected domain is relatively small and does not extend into high latitudes.
-Once the EOFs were calculated, the principal component time series were computed for each day and scaled to unit variance as is typical for cluster analysis.
+No meridional weighting was applied as the selected domain is relatively small and does not extend into high latitudes.
+Once the EOFs were calculated, the principal component time series were computed for each day and scaled to unit variance.
+This rescaling is not a necessary step; its result is to treat all principal components as equally important for the weather typing, which provides relatively greater weight to EOFs 2, 3, and 4 than carrying out weather typing without re-scaling.
 
-Next, the $K$-means algorithm was used to assign a single cluster value to each day on record using the 4-dimension principal component time series computed in the previous step.
+Next, the $K$-means algorithm was used to assign a single cluster value to each day on record using the 4-dimension principal component time series.
 The $K$-means technique is a partitioning method that classifies all days in the study into a predefined number of clusters.
@@
 \end{enumerate}
-The cluster centroids (weather types) $\mu_k$ produced by the $K$-means algorithm can then be interpreted as a Voronoi decomposition of the phase space into $K$ regions.
-Intuitively, this means that an observation is assigned to the weather type whose centroid $\mu_k$ is closest to it in the phase space given by the leading 4 EOFs of $\psi_{850}.$
+The cluster centroids (weather types) $\mu_k$ produced by the $K$-means algorithm can then be interpreted as a Voronoi decomposition of the phase space into $K$ regions, and specifically as the Voronoi diagram which minimizes within-cluster variance.
 
-The $K$-means algorithm is guaranteed to converge to a local minimum of intercluster variance; to select the best partition, 500 simulations were created using the implementation in Python's \texttt{scikit-learn} package \citep{Pedregosa:2012tv} and then the classifiability index of \citet{Michelangeli1995} was then used to select the optimal partition.
-This index can be conceptualized as a stability criterion as it selects the partitioning which is most like the other 249 partitions, as defined by Euclidean distance.
-Calculation of the classifiability index for several values of $k$ (supplemental figure S2) suggests that states with $k=4,5,6,7,8$ are all reasonable; we chose the solution $k=6$ because the weather types identified are qualitatively similar to those determined over southeastern South America \citep{Munoz2015,Munoz2016} and have an intuitive physical meaning, discussed further in the following sections.
+The $K$-means algorithm is guaranteed to converge to a local minimum of intercluster variance; to select the best partition, 500 simulations were created using the implementation in Python's \texttt{scikit-learn} package \citep{Pedregosa:2012tv} and then the classifiability index of \citet{Michelangeli1995} was computed between each partition and the 249 others.
+The partition whose classifiability index, averaged for all 249 pairwise comparisons, was the highest was selected.
+Calculation of the classifiability index for several values of $K$ (supplemental figure S3) suggests that states with $k=4,5,6,7$ are all reasonable; we chose the solution $k=6$ because the weather types identified are qualitatively similar to those determined over southeastern South America \citep{Munoz2015,Munoz2016} and have an intuitive physical meaning, discussed further in the following sections.
 
@@
 For the purposes of MOS corrections, the predictand (or variable to forecast) is the observed rainfall for the target period of interest, and the predictor (or variable to be corrected) is the raw model rainfall for the same period.
-We use the same spatial domain [\SI{39}{\degree S}-\SI{17}{\degree S}; \SI{66}{\degree W}-\SI{49}{\degree W}] for both the predictor and the predictand, except for the PCR and CCA cases, in which a larger domain [\SI{0}{\degree S}-\SI{60}{\degree S}; \SI{80}{\degree W}-\SI{30}{\degree W}] was used to better capture the spatial patterns in the raw model field.
+We use the same spatial domain [\SIrange{39}{17}{\degree S}; \SIrange{66}{49}{\degree W}] for both the predictor and the predictand, except for the PCR and CCA cases, in which a larger domain [\SIrange{0}{60}{\degree S}; \SIrange{80}{30}{\degree W}] was used to better capture the spatial patterns in the raw model field.
 A variety of approaches for modeling the output of many ensemble members was explored; a summary of the predictors found to be most skillful for each MOS model is presented in \cref{tab:mos-methods}.
 
-To avoid artificial skill when building the statistical models, we employ a cross-validation approach with a 5-year-out window.
+To avoid artificial skill when building the statistical models, we employ a cross-validation approach with a 5-year window.
 In this framwork, five continuous years are left out of the record, the regression coefficients are computed with the remaining of the time series, and the resulting model is validated comparing the prediction for the third year left out (middle of the window) against observations.
@@
 where $p$ and $p_c$ represent the forecast probability and the climatological probability, respectively.
-This is analagous to the commonly used practice of showing probability of below-normal, normal, and above-normal precipitationand simplifies the comparison in a way that tends to be better understood by decision-makers.
+This is analagous to the commonly used practice of showing probability of below-normal, normal, and above-normal precipitation and simplifies the comparison in a way that tends to be better understood by decision-makers.
 
-The IRI's seasonal predictions are already provided with bias correction \citep{Barnston:2010ge}, and hence we did not perform any further MOS on seasonal rainfall fields.
+The IRI's seasonal predictions are already provided with bias correction \citep{Barnston:2010ge}, and thus we did not perform any further MOS on seasonal rainfall fields.
 
@@
 
-\section{Observed Flooding} \label{sec:diag}
+\section{Observed Flooding} \label{sec:flood}
 
 \Cref{fig:streamflow} shows the streamflow time series at several gauges on the Paraguay River during NDJF 2015-16 in the context of their seasonality and decadal variability.
-The most notable feature of \cref{fig:streamflow}(a) is that during November and December 2015, the river rose rapidly at Concepci\'on, Asunci\'on,  and Pilar but not at Bah\'ia Negra.
+During November and December 2015, the river rose rapidly at Concepci\'on, Asunci\'on, and Pilar though not at Bah\'ia Negra.
 As discussed in \citet{Bravo:2011et,Barros:2004bn}, the location of the Bah\'ia Negra gauge (see \cref{fig:study-area}) in the Pantanal region means that it responds very slowly to rainfall input.
-However, the three downstream gauges respond to the rainfall forcing with a slow but steady rise.
+The three downstream gauges, because they are located in the Lower Paraguay River basin respond to the rainfall forcing with a slow but steady rise.
 Despite several very intense storms (see following sections), the streamflow record at Asunci\'on and Pilar (which are downstream of Concepci\'on) indicates relatively little response to individual storms.
+Because the region is so flat (see topographic data in \cref{fig:study-area}), river levels at a particular point may be affected not only by rain in the catchment corresponding to that point, but also by elevated river levels downstream which reduce the pressure gradient available to drive flow.
 
 Interestingly, examination of \cref{fig:streamflow}(b) suggests multidecadal oscillation in the streamflow record.
-This is in agreement with \citet{Collischonn:2001bi,Carvalho2011} who find a changepoint in the 1970s possibly associated with changes in the Pacific Decadal Oscillation.
-However, because only river stage data (and not discharge) data are available, it is not possible to discern whether the observed changes in river stage are driven by sediment loading and local measurement characteristics or by large-scale climate fluctuations.
+This is in agreement with previous studies \citep{Collischonn:2001bi,Carvalho2011} who find a changepoint in the 1970s, possibly associated with low-frequency Pacific variability.
+Because only river stage data (and not discharge) data are available, it is not possible to discern whether the observed changes in river stage are driven by sediment loading and local measurement characteristics or by large-scale climate fluctuations.
 Further treatment of this question is beyond the scope of this paper.
@@
 Along the cold front associated with this system, a low-level northerly jet advects of heat and moisture to the region.
-As the system progresses, the jet transitions from predominantly meridional flow (Chaco Jet; $t=-1$ to predominantly zonal flow below approximately \SI{20}{\degree N}.
-
-The rainfall largely follows the pattern of the low-level jet, in agreement with previous studies  \citep{Saulo:2007km,Salio:2007gd,Marengo2004,Velasco1987}.
-
+As the system progresses, the jet transitions from predominantly meridional flow (``Chaco Jet''; $t=-1$ to predominantly zonal flow (``No-Chaco Jet'') below approximately \SI{20}{\degree N}.
 Similar analysis for the 90th or 99th percentiles of daily rainfall (not shown) yield similar results, implying that the synoptic mechanism for the most extreme events is not fundamentally distinct from the mechanism for moderate-intensity events.
+This mean field, like all composites, masks intra-event variation but exploration of individual events (not shown) shows that the core features identified are generally present.
 
@@
 The EOF loadings are shown in \cref{fig:eof-loading}.
-Of these, EOF 1 explains by far the most variance ($\approx 67\%$) while EOFs 2, 3, and 4 collectively explain approximately $28\%$ of total variance (supplemental figure S1).
+Of these, EOF 1 explains by far the most variance ($\approx 67\%$) while EOFs 2, 3, and 4 collectively explain approximately $28\%$ of total variance.
+Values of indiviual and cumulative variance explained are presented in supplemental figure S2.
 To a first order of approximation, EOF 1 explains the large-scale anomalies induced by passing baroclinic waves; the East-West gradient is likely a result of the influence of the Andes Mountains which are a source of friction for low-level flow.
@@
 Weather type 1 represents a ``Chaco Jet'' event \citep{Salio:2002ev} in which the maximum wind speed penetrates past \SI{25}{\degree S}, leading to intense rainfall NE Argentina and Uruguay.
-The other weather type associated with a SALLJ is weather type 4, the ``No-Chaco Jet'' event in which the wind turns to the East before \SI{25}{\degree S}, leading to heavy rainfall over Eastern Paraguay and SW Brazil.
-Weather types 2 and 3 resemble the weak and intense phases, respectively, of the South Atlantic Convergence Zone \citep[SACZ;][]{Carvalho2004} while weather types 5 and 6 represent southerly and easterly wind, respectively, over the Lower Paraguay River Basin.
-\texttt{ANY THOUGHTS ON WT5,6?}
+Weather type 4 is also associated with a SALLJ event, but is of the ``No-Chaco'' jet type in which the wind turns to the East before \SI{25}{\degree S}, leading to heavy rainfall over Eastern Paraguay and SW Brazil.
+Weather types 3 and 5 are the inverses of weather types 1 and 4, respectively, and are associated with dry anomalies over the Lower Paraguay River Basin.
+\texttt{ANY THOUGHTS ON WT 2,6? DO WE AGREE ON 1,3,4,5?}
 
@@
 
-To understand what happened specifically during NDJF 2015-16, we show both sequences of daily weather types and monthly-mean circulation anomalies.
+We next use monthly-mean circulation anomalies (spatial patterns) and weather type sequences (time patterns) to understand the specefic events of NDJF 2015-16.
 
-While the weather typing approach requires simplifying the dynamics of daily circulation patterns, its advantage is that it greatly facilitates the analysis of sequences of precipitation.
-\Cref{fig:rain-wt} shows a time series of area-averaged rainfall over the Lower Paraguay River Basin for NDJF 2015-16; this plot shows that intense rainfall concentrated over a period spanning from mid-November 2015 through early January 2016, with smaller peaks in late January and mid-February.
+While weather typing requires simplifying the dynamics of daily circulation patterns, its advantage is that it greatly facilitates the analysis of sequences of precipitation.
+\Cref{fig:rain-wt} shows a time series of area-averaged rainfall over the Lower Paraguay River Basin for NDJF 2015-16 and the corresponding weather types.
+This plot shows that intense rainfall concentrated over a period spanning from mid-November 2015 through early January 2016, with shorter peaks in late January and mid-February.
+
 As predicted by \cref{fig:wt-composite}, the most intense rainfall occurred during weather types 1 and 4.
@@
 
-At monthly scale (\cref{fig:anomalies}), a weak anticyclonic circulation set up over central Brazil during November 2015 and strengthened into the following month.
+Additionally, the weather types associated with different ``regimes'' in the season were different.
+The rainfall in late November 2015 occurred during weather type 1 and 4 days, with intermitent (and dry) weather type 5 occurrences.
+During mid to late December 2015, weather types 1 and 4 produced rainfall but the intermittent dry days were weather type 2.
+During early January 2016 weather types 3 and 6 led to persistent low rainfall.
+Interestingly, the sequence of weather type 4 days did not produce significant rainfall; this is likely because the lack of preceding weather type 1 (Chaco Jet event) days limited the moisture and energy available for severe convection.
+Finally, during February 2016 the observed sequences 1-4-4, 3-6-4, and even 1-5-2 led to intense rainfall.
+
+Transitioning from exploring the time evolution of the reduced-dimension system represented by the weather types, monthly-scale circulation anomalies (\cref{fig:anomalies}) show a weak anticyclonic circulation thta set up over central Brazil during November 2015 and strengthened into the following month.
 In January 2016 it weakened before returning in February 2016.
-When analyzing these anomalies it is important to consider that these anomalies are all relative to the climatology (supplemental figure S4) which exhibits substantial seasonality.
-The observed rainfall and circulation anomalies are consistent with the aggregation of the observed weather types shown in \cref{fig:rain-wt}; November 2015 most closely resembles weather type 1, December 2015 weather types 1 and 4, January 2016 weather type
-3, and February 2016 weather types 2 and 4.
+The observed rainfall and circulation anomalies are consistent with the aggregation of the observed weather types shown in \cref{fig:rain-wt} and discussed above.
 
@@
 	\caption{
-		Map of the study area.
+		Topographical map of the study area.
+		Colors indicate $\log_{10}$ of elevation, in \si{\meter}, from the Global Land One-Km Base Elevation project available at \texttt{http://iridl.ldeo.columbia.edu/SOURCES/.NOAA/.NGDC/.GLOBE/.topo/}.
 		(L): all of South America.
@@
 		As for (L), the ``Lower Paraguay River Basin'' is indicated.
-		Streamflows shown in \cref{fig:streamflow} were taken from the fours stations indicated in green.
+		Streamflows shown in \cref{fig:streamflow} were taken from the fours stations indicated.
 		The Paraguay River and its tributaries, from the Natural Earth database (\texttt{www.naturalearthdata.com}), are also shown.

### 2017-09-25 scripts/access/StreamfuncYear.py

@@ -4,6 +4,7 @@ Data from NCAR/NCEP Reanalysis I 6-hourly
 '''
 
 import numpy as np
+import pandas as pd
 import xarray as xr
 from windspharm.xarray import VectorWind
 import argparse
@@ -13,6 +14,17 @@ parser.add_argument('--level', type=int, nargs=1, help = 'Pressure level (hPa) f
 parser.add_argument('--year', nargs=1, type=int, help='The year of data to download')
 parser.add_argument('--outfile', nargs=1, help='The output file to save!')
 
+def read_reanalysis2(url, level):
+    import xarray as xr
+    import numpy as np
+    # reading in this data is a bit strange
+    ds = xr.open_dataarray(url, decode_cf=False).sel(level=level)
+    del ds.attrs['_FillValue']
+    scale = ds.attrs['scale_factor']
+    offset = ds.attrs['add_offset']
+    ds.values = ds.values * scale + offset # this is not done automatically
+    return(ds)
+
 def main():
     # parse the arguments
     args = parser.parse_args()
@@ -21,10 +33,10 @@ def main():
     outfile = args.outfile[0]
 
     # Get the U and V wind
-    url_u = 'https://www.esrl.noaa.gov/psd/thredds/dodsC/Datasets/ncep.reanalysis/pressure/uwnd.{}.nc'.format(year)
-    url_v = 'https://www.esrl.noaa.gov/psd/thredds/dodsC/Datasets/ncep.reanalysis/pressure/vwnd.{}.nc'.format(year)
-    U = xr.open_dataarray(url_u).sel(level=level).load()
-    V = xr.open_dataarray(url_v).sel(level=level).load()
+    url_u = 'https://www.esrl.noaa.gov/psd/thredds/dodsC/Datasets/ncep.reanalysis2/pressure/uwnd.{}.nc'.format(year)
+    url_v = 'https://www.esrl.noaa.gov/psd/thredds/dodsC/Datasets/ncep.reanalysis2/pressure/vwnd.{}.nc'.format(year)
+    U = read_reanalysis2(url_u, level=level)
+    V = read_reanalysis2(url_v, level=level)
 
     # Get streamfunction
     W = VectorWind(U, V)

### 2017-10-04 writeup/Supplement.tex
No 200hPa
Don't use 200 hPa data any more

@@
   graphicx, % for images
-	natbib
+	natbib,
+	booktabs
 }
@@
 \renewcommand{\thefigure}{S\arabic{figure}}
+\renewcommand{\thetable}{S\arabic{table}}
 \usepackage[breaklinks=true]{hyperref}
@@
 Further supporting information is available in the form of source code and \texttt{jupyter} notebooks at \url{github.com/jdossgollin/PYFloods}.
+
+\listoftables
 \listoffigures
+
+\clearpage
+
+\begin{table}
+	\centering
+	\begin{tabular}{lrr}
+		\toprule
+		{} &  NDJF 2015-16 &  Climatology \\
+		WT &           &              \\
+		\midrule
+		1     &  0.266667 &     0.213306 \\
+		2     &  0.216667 &     0.177793 \\
+		3     &  0.133333 &     0.160935 \\
+		4     &  0.116667 &     0.152619 \\
+		5     &  0.191667 &     0.151944 \\
+		6     &  0.075000 &     0.143403 \\
+		\bottomrule
+	\end{tabular}
+	\caption{Weather type (WT) occurrence fraction during NDJF 2015-16 and Climatology}
+\end{table}
+
 \clearpage
@@
 		Results are produced with only 50 simulations for each $K$.
-	}
-\end{figure}
-
-\begin{figure}
-  \includegraphics[width=\textwidth]{../figs/WT_Occurrence_Proportion.pdf}
-	\caption{
-		Occurrence fraction of NDJF weather types for all years (blue) and 2015-16 (orange).
 	}

### 2017-10-04 writeup/paper.tex
No 200hPa
Don't use 200 hPa data any more

@@
 	array, % for fixed-width tables
+	booktabs, % for nice tables
 }
@@
 
-As predicted by \cref{fig:wt-composite}, the most intense rainfall occurred during weather types 1 and 4.
-In particular, many (though not all) such events occured during a transition of WT1-WT4-WT5, which is consistent with the westerlay passage of a mid-latitude system as shown in \cref{fig:lagged-rain}.
-During NDJF 2015-16 weather type 1 occurred nearly twice as often (33\% of days) as its climatology (19\% of days), weather type 4 occurred slightly more often (18\% of days) than its climatology (16\% of days), and other weather types occurred less frequently than their climatology (see Figure S3).
-During mid-January, during a sequence of persistent low rainfall, weather type 3 featured persistently, leading to intense rainfall over central Brazil (not shown) and negligible rainfall over the Lower Paraguay River Basin.
+As predicted by \cref{fig:wt-composite}, the most intense rainfall occurred during weather types 1 and 5.
+During NDJF 2015-16 weather types 1 and 5 (Chaco and No-Chaco jet extensions, respectively) occurred more frequency than their climatology (supplemental table S1); weather type 2 also occurred more frequently than its climatology, largely due to a long sequence in February 2016.
+During mid-January 2015, during a sequence of persistent low rainfall, weather type 3 featured persistently, leading to intense rainfall over central Brazil (not shown) and negligible rainfall over the Lower Paraguay River Basin.
 Thus while the intensity and persistence of intense rainfall was atypical, the causal mechanism of the intense rainfall observed during this season was consistent with climatology.
 
-Additionally, the weather types associated with different ``regimes'' in the season were different.
+Inspection of \cref{fig:rain-wt} also suggests that at time scales of days to weeks, particular sequences of weather types tend to repeat themselves, and are associated with repeated rainfall storms.
 The rainfall in late November 2015 occurred during weather type 1 and 4 days, with intermitent (and dry) weather type 5 occurrences.
-During mid to late December 2015, weather types 1 and 4 produced rainfall but the intermittent dry days were weather type 2.
+During mid to late December 2015, weather types 1 and 5 produced rainfall but the intermittent dry days were weather type 2.
 During early January 2016 weather types 3 and 6 led to persistent low rainfall.
-Interestingly, the sequence of weather type 4 days did not produce significant rainfall; this is likely because the lack of preceding weather type 1 (Chaco Jet event) days limited the moisture and energy available for severe convection.
-Finally, during February 2016 the observed sequences 1-4-4, 3-6-4, and even 1-5-2 led to intense rainfall.
 
@@
 \label{sec:discussion}
+
+From the results obtained, we can see that the rainfall of NDJF 2015-16 was associated with the typical mechanisms discussed in the literature \citep{Saulo:2007km,Salio:2007gd,Marengo2004,Velasco1987}.
+In particular, the intensity and exit location of the low-level jet let to several sequences of intense rainfall during NDJF 2015-16.
+The approximately six-week period from mid-November 2015 through early January 2016 that led to the most severe flooding was exceptional primarily for the number of intense storms that directly impacted the Lower Paraguay River Basin.
+During this period a low-level cyclonic anomaly over Central Brazil, associated with hemispheric-scale circulation anomalies (\cref{fig:anomalies}) strongly favored weather type 1, associated with the greatest influx of moisture and energy to the Lower Paraguay River Basin, and weather type 5, associated with low-level wind shear and rainfall over the Lower Paraguay River Basin and Southern Brazil.
+Weather types 1 and 5 look like like EOFs 2 and 3 of the \SI{850}{\hecto\pascal} streamfunction field \cref{fig:eof-loading}; further analysis (supplemental figure S4) shows that when EOF2 is negative and EOF3 is positive, strong rainfall is expected in the Lower Paraguay River Basin, and when EOF2 is negative but EOF3 is also negative low rainfall is expected in the Lower Paraguay River Basin but substantial rainfall is expected further to the South and West.
 
@@
 \begin{tabular}{L{0.75in} L{1.15in} L{3.5in}}
-\hline\hline
+\toprule
 Model & Region (Predictor) & Final predictor(s) selected \\
 %
-\hline
+\midrule
 %
@@
 %
-\hline\hline
+\bottomrule
 %
@@
 		The Paraguay River and its tributaries, from the Natural Earth database (\texttt{www.naturalearthdata.com}), are also shown.
+		Full station names are: Bah\'ia Negra (Bne); Concepci\'on (Conc); Asunci\'on (Asu); Pilar (Pil).
 	}
@@
     	Monthly composite anomalies observed during NDJF 2015-16.
-      Top row shows absolute monthly mean values (color) and monthly-mean anomalies (contour) of streamfunction at \SI{850}{\hecto\pascal} $\psi_{850}$.
-			Contour interval for anomalies is \SI{1e6}{\meter\squared\per\second}.
+    	Top and middle rows show streamfunction anomalies at \SIlist{200;850}{\hecto\pascal}, respectively.
       Bottom row shows monthly-mean rainfall anomalies, in units of \si{\milli\meter\per\day}.
@@
 	\caption{
-    	Leading four EOF loadings of daily NDJF \SI{850}{\hecto\pascal} streamfunction $\psi$ over chosen domain.
+    	Leading EOF loadings of daily NDJF \SI{850}{\hecto\pascal} streamfunction $\psi$ over chosen domain.
 			EOFs are calculated after removing the climatology; see \cref{sec:methods}\ref{sec:climatology-anomalies} for details.
@@
 		Composite anomalies associated with each weather type (column).
-		Top row shows absolute monthly mean values (color) and monthly-mean anomalies (contour) of streamfunction at \SI{850}{\hecto\pascal} $\psi_{850}$.
-		Contour interval for anomalies is \SI{1e6}{\meter\squared\per\second}.
-    Bottom row shows precipitation anomalies in units of \si{\milli\meter\per\day}.
+		Top and middle rows show streamfunction anomalies at \SIlist{200;850}{\hecto\pascal}, respectively.
+		Bottom row shows monthly-mean rainfall anomalies, in units of \si{\milli\meter\per\day}.
 	}
@@
 		Lagged composites are shown, by column, for $t = $ \SIlist{-2;-1;0;1}{\day} relative to the date of intense rainfall.
-		Top row shows absolute monthly mean values (color) and monthly-mean anomalies (contour) of streamfunction at \SI{850}{\hecto\pascal} $\psi_{850}$.
-		Contour interval for anomalies is \SI{1e6}{\meter\squared\per\second}.
-    Bottom row shows precipitation anomalies in units of \si{\milli\meter\per\day}.
+		Top and middle rows show streamfunction anomalies at \SIlist{200;850}{\hecto\pascal}, respectively.
+		Bottom row shows monthly-mean rainfall anomalies, in units of \si{\milli\meter\per\day}.
 	}

### 2017-11-27 writeup/Supplement.tex
final version as submitted

@@
 \renewcommand{\thetable}{S\arabic{table}}
-\usepackage[breaklinks=true]{hyperref}
+\usepackage[breaklinks=true,hidelinks]{hyperref}
+\graphicspath{{../figs/}}
 
@@
 
-\title{Supplemental Information for\\``Extreme rainfall in Paraguay during the 2015-2016 austral summer: causes and subseasonal-to-seasonal predictive skill''}
-\author{James Doss-Gollin\and \'{A}ngel G. Mu\~{n}oz \and Simon J. Mason \and Max Past\'{e}n}
+\title{Supplemental Information for\\``Heavy rainfall in Paraguay during the 2015-2016 austral summer: causes and subseasonal-to-seasonal predictive skill''}
+\author{James Doss-Gollin\and \'{A}ngel G. Mu\~{n}oz  \and Simon J. Mason \and Max Past\'{e}n }
 \date{\today}
@@
 	\midrule
-		1     &  0.266667 &     0.213306 \\
-		2     &  0.216667 &     0.177793 \\
-		3     &  0.133333 &     0.160935 \\
-		4     &  0.116667 &     0.152619 \\
-		5     &  0.191667 &     0.151944 \\
-		6     &  0.075000 &     0.143403 \\
+	1     &  0.258 &     0.212 \\
+	2     &  0.242 &     0.187 \\
+	3     &  0.133 &     0.158 \\
+	4     &  0.192 &     0.155 \\
+	5     &  0.092 &     0.146 \\
+	6     &  0.083 &     0.142 \\
 	\bottomrule
@@
 	\centering
-  \includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio=true]{gpw-v4-2015.png}
+  \includegraphics[width=\textwidth,height=0.6\textheight,keepaspectratio=true]{gpw-v4-2015.png}
 	\caption{
@@
 \begin{figure}
-  \includegraphics[width=\textwidth]{../figs/PSI_Var_Explained.pdf}
+  \includegraphics[width=\textwidth]{PSI_Var_Explained.pdf}
 	\caption{
@@
 \begin{figure}
-  \includegraphics[width=\textwidth]{../figs/WT_Classifiability.pdf}
+  \includegraphics[width=\textwidth]{WT_Classifiability.pdf}
 	\caption{
@@
 \begin{figure}
-  \includegraphics[width=\textwidth]{../figs/knn_rainfall_EOF.pdf}
+  \includegraphics[width=\textwidth]{MJO_Time_Series.pdf}
 	\caption{
-    Expected rainfall over two regions conditional on the (simultaneous) values of EOF2 and EOF3 of the \SI{850}{\hecto\pascal} streamfunction $\psi$ shown in the main text.
-    The ``Chaco Jet Extension Region'' is defined as extending from \SIrange{299.25}{302.75}{\degree E} and \SIrange{34.75}{30.75}{\degree S} and the ``Lower Paraguay River Basin'' is defined in the main text.
-    Both values represent area-averaged rainfall time series.
-    Expected rainfall is calculated by using a nearest-neighbors regression with 100 neighbors in the two-dimensional phase space of EOF2 and EOF3.
-    Because nearest-neighbors approaches tend to perform poorly under extrapolation (where there are few observations), the domain shown is limited to regions with high density of observations.
+		Phase plot of the Madden-Julian Oscillation during NDJF 2015-16.
+		The $x$-axis shows RMM1 and the $y$-axis shows RMM2 -- gray lines separate MJO phases.
 	}
-\end{figure}
-
-\begin{figure}
-  \includegraphics[width=\textwidth]{../figs/ENSO_MJO_EOF_PRCP_Time_Series.pdf}
-	\caption{
-		Time series of S2S predictors and their influence on evolution of circulation and rainfall.
-    Top row shows two predictors shown to influence the circulation: monthly NINO3.4 index and daily MJO RMM1 index.
-    Blue dots show daily observations of MJO RMM1, while the blue line shows the MJO RMM1 time series with a 7-day rolling mean applied.
-    Bottom row shows the evolution of EOFS 2 and 3 of the \SI{850}{\hecto\pascal} field.
-    As for the MJO RMM1, dots show the daily fields and lines show the fields with a 7-day running mean applied.
-    Bottom rain shows time series of daily rainfall over the ``Lower Paraguay River Basin'' and ``Chaco Jet Extension Region'' defined above.
-	}
+	\label{fig:mjo-ts}
 \end{figure}

### 2017-11-30 reports/poster-agu-2017/blocks/conditional-simulation.tex
re-work directory

\begin{block}{Conditional Simulation}
  Goal: condition the RIP events (poorly represented in CM3) on EWD and MHC indices (better represented in CM3).
  Assume:
  \begin{itemize}
    \item Linear model for EWD and MHC indices
    \item MHC is secondary: if $\text{EWD}<0$, do not consider MHC
    \item Logistic regression model: $p(\text{RIP}_t) \sim f \qty(\text{EWD}_t, \text{MHC}_t)$
    \item Fit model on reanalysis temperature and pressure fields, and observed RIP events
    \item Simulate from model using CM3 temperature and pressure fields
  \end{itemize}
  \textbf{Comparing \cref{fig:historical-simulations} with \cref{fig:RIP-historic}, we see a marked improvement in model skill by conditioning on large-scale circulation features}
  \begin{figure}[ht]
    \centerline{\includegraphics[width=0.95\textwidth]{figs/RIP_historical_sim_MAM.pdf}}
    \caption{
    (a) 10-year moving average of the number of MAM RIP days by year for the observed record (black solid line), the five conditional simulated RIP records.
    (b) Counts for the number of MAM RIP days by year for the observed record (black solid line), the five conditional simulated RIP records based on the GFDL CM3 ensemble member's geopotential height and temperature fields (light dashed lines), and the ensemble mean (thick dashed line).
    }
    \label{fig:historical-simulations}
  \end{figure}
\end{block}

### 2017-11-30 reports/poster-agu-2017/blocks/future-simulation.tex
re-work directory

\begin{block}{Future Simulations}
  \begin{itemize}
    \item A conditional simulation model using GCM atmospheric circulation and temperature projections for 21st century shows increasing trends for such events and
    \item Conditional simulation model attributes increasing RIP events approximately 1/3 to circulation changes and 2/3 to temperature changes
    \item Results differ signicantly from RIP events based only on MAM precipitation field
  \end{itemize}
  \begin{figure}[ht!]
    \centerline{\includegraphics[width=0.95\textwidth]{figs/RIP_future_sim_MAM.pdf}}
    \caption{
      (Left) 10-year moving average projections from 2006 to 2100 of MAM RIP counts by year from CM3 precipitation field (black solid line), conditional simulation based on CM3 temperature and pressure fields (light dashed lines show five simulations based on the one ensemble field and the dark dashed line shows the mean of these simulations).
      (Right top) Distribution of RIP counts by year for 2006-2050.
      (Right bottom) As (Right top) but for 2056-2100.
      }
      \label{fig:future-simulations}
  \end{figure}
\end{block}

### 2017-11-30 reports/poster-agu-2017/blocks/indices.tex
re-work directory

\begin{block}{Defining Indices for RIP Circulations}
  \begin{figure}[ht]
    \centerline{\includegraphics[width=0.75\textwidth]{figs/dipole_MHC_box_and_time_MAM.pdf}}
    \caption{The regions of low and high pressure defining the Eastern U.S.-Western Atlantic Dipole index (EWD), and the region defining the moisture holding capacity index.}
    \label{fig:EWD-MHC-location-time}
  \end{figure}
  \begin{itemize}
    \item Define Eastern U.S.-Western Atlantic Dipole (EWD) index: average of high pressure box (red) minus low pressure box (green) (\cref{fig:EWD-MHC-location-time})
    \item Define moisture-holding capacity (MHC) index using average temperature (brown; \cref{fig:EWD-MHC-location-time})
      $\text{MHC}_t = 6.1 \exp \qty[\frac{17.6T'_{t}}{T'_{t} + 243}]$
    \item CM3 reasonably simulates the distributional (\cref{fig:cdf-comp}) and persistence (not shown) features of two atmospheric indices that modulate the likelihood of RIP events
  \end{itemize}
  \begin{figure}[ht]
    \centerline{\includegraphics[width=0.95\textwidth]{figs/EWD_MHC_cdf_comp_MAM.pdf}}
    \caption{
      (Left) Cumulative distribution function for the reanalysis based dipole (solid line) and the GFDL GCM ensemble members based dipole (dashed lines) for MAM.
      (Middle) The positive tail of (Left).
      (Right) Same as (Left) for the moisture holding capacity.
    }
    \label{fig:cdf-comp}
  \end{figure}
\end{block}

### 2017-11-30 reports/poster-agu-2017/blocks/methods-data.tex
re-work directory

\begin{block}{Methods \& Data}
  \begin{figure}[ht]
    \centerline{\includegraphics[width=0.95\textwidth]{figs/GCM_CPC_grids.pdf}}
    \caption{
      Map of study area and Geophysical Fluid Dynamics Laboratory CM3 coupled model cells (blue grid).
      The Climate Prediction Center precipitation data (red grid) was upscaled to match the grid of the CM3 coupled model (blue grid), by taking the spatial average of the red cells inside each of the blue cells.
      The shaded area indicates the Ohio River Basin ($\sim$ 530 000 km$^2$) as defined by the United States Geological Survey.
    }
    \label{fig:area-map}
  \end{figure}
  \begin{itemize}
    \item MAM season in the Ohio River Basin (\Cref{fig:area-map}).
    ``Historical'' from 01 March 1950 through 30 May 2005, ``future'' study period from 01 March 2006 through 30 May 2100.
    \item ``Dynamical Model'' referes to GFDL global coupled model \citep{Donner2011} (CM3).
    \item CPC US unified gauge-based precipitation data, upscaled from \SI{0.25}{\degree} by \SI{0.25}{\degree} to match CM3 (\SI{2.5}{\degree} longitude by \SI{2.0}{\degree} latitude).
    \item Regional Intense Precipitation (RIP) days were defined as any day when over 20\% of the region receives experiences a 99th percentile exceedance of precipitation.
    \item Reanalysis data from NCEP/NCAR Reanalysis I \citep{Kalnay1996}.
  \end{itemize}
\end{block}

### 2017-11-30 reports/poster-agu-2017/blocks/motivation.tex
re-work directory

\begin{block}{Conceptual Framework}
    \begin{itemize}
      \item Projections of future extreme precipitation remain highly uncertain \citep{IPCC2012}
      \item Rainfall (esp. extreme) poorly represented in GCMs \citep[\ie][]{Kendon2012}
      \item Precipitation fields often used for hydrological and decision models after bias correction \citep[e.g. quantile-quantile mapping][]{Gudmundsson2012}) is applied; however model deficiencies not identified, persistence characteristics in precipitation fields poorly represented \citep[e.g.][]{Piani2010}.
    \end{itemize}
    \emph{Is there a way to use credibly simulated state variables from GCM simulations to derive or simulate credible sequences of regional intense precipitation events associated with societal impacts?}
\end{block}

### 2017-11-30 reports/poster-agu-2017/blocks/questions.tex
re-work directory

\begin{block}{Research Questions}
  We hypothesize that GCMs may simulate the frequency and intensity  of atmospheric circulation patterns associated with regional intense precipitation events better than the statistics of the precipitation events themselves.
  \begin{description}
    \item[Q1] For the Ohio River basin, are the intense springtime precipitation events relevant for extreme floods well simulated by the GCM? If not, why?
    \item[Q2] If no to \textbf{Q1}, can we find atmospheric indices that are associated with the onset of the regional intense precipitation events?
    \item[Q3] If they are not, are suitably derived atmospheric indices associated with such events in atmospheric re-analysis relatively well simulated by the GCM?
    \item[Q4] If the GCM simulations of atmospheric indices are better, can they provide more credible projections of regional intense precipitation occurrence than the GCM can directly?
  \end{description}
\end{block}

### 2017-11-30 reports/poster-agu-2017/blocks/rip-circulations.tex
re-work directory

\begin{block}{Circulation Patterns Associated with RIP Events}
  GCMs represent observed dipole pattern, with some latitudinal bias in storm track and geometry of high pressure system
  \begin{figure}[ht]
    \centerline{\includegraphics[width=0.95\textwidth]{figs/CPC_RIPs_MAM_ZG_700_PR_WAT.pdf}}
    \caption{
      Daily composites of $Z_{700}$ anomalies (shades) and integrated precipitable water content anomaly (contours at \SI{3}{\kilo\gram\per\square\meter}) from four days before each spring (MAM) RIP event to one day following the event.
      Solid contours represent positive anomalies and dashed contours represent negative anomalies.
      An ``X'' indicates that at least 80\% of composite members had anomalies of the same sign in that location.
    }
    \label{fig:MAM-gph-pr-wat}
  \end{figure}
  \begin{figure}[ht]
    \centerline{\includegraphics[width=0.95\textwidth]{figs/GCM_OBS_RIPs_MAM_Z_700.pdf}}
    \caption{
      MAM RIP day composites of absolute $Z_{700}$ (shading) and $Z_{700}$ anomalies (contours in \SI{15}{\meter} increments) for 5 CM3 members and reanalysis (Obs).
    }
    \label{fig:MAM-mod-obs-gph}
  \end{figure}
\end{block}

### 2017-11-30 reports/poster-agu-2017/blocks/rip-gcm-obs.tex
re-work directory

\begin{block}{RIP Events in a GCM \& Observations}
  \begin{itemize}
    \item CM3: too many MAM RIP days, too few back-to-back RIP days
    \item Seasonality bias: too many (few) RIP days in MAM (JJA)
    \item When CM3 produces intense precipitation in any part of the study region, it has a tendency to simultaneously produce intense precipitation in several grid cells
    \item Discrepancy between the GCM runs and the observed RIP records is even more stark when the observed precipitation data is used to calculate the 99th percentile thresholds for the model and RIP records
  \end{itemize}
  Conclusion: \textbf{relevant precipitation events not well-simulated by CM3} -- no to \textbf{Q1}.
  \begin{figure}[ht]
    \centering
    \includegraphics[width=0.95\textwidth]{RIP_CPC_mod_MAM_year.pdf}
    \caption{
      (a) 10-year moving average of the number of MAM RIP days by year for the observed record (black solid line), the five GFDL CM3 ensemble members (light dashed lines), and the ensemble mean (heavy dashed line).
      (b) The counts for the the number of MAM RIP days by year for the observed record (black solid line), the five GFDL CM3 ensemble members (light dashed lines), and the ensemble mean (heavy dashed line).
    }
    \label{fig:RIP-historic}
  \end{figure}
\end{block}

### 2017-11-30 reports/poster-agu-2017/blocks/summary.tex
re-work directory

\begin{block}{Summary}
  \begin{itemize}
    \item Recurrent atmospheric circulation patterns that correspond to regional intense precipitation events in  the Ohio River Basin are identified
    \item GCM biases in predicting regional intense precipitation frequency are largely overcome by conditioning these events on GCM atmospheric circulation patterns
    \item GCM frequency bias in RIP record appears a manifestation of inflated spatial correlation of high intensity precipitation
  \end{itemize}
  This \emph{framework for conditional simulation} is an alternative to grid-scale statistical downscaling
\end{block}

### 2017-11-30 reports/poster-agu-2017/blocks/thanks.tex
re-work directory

\begin{block}{Acknowledgements}
  This research was supported by funding from NSF Sustainability Research Networks (award 1444745): ``Integrated Urban Infrastructure Solutions for Environmentally Sustainable, Healthy, and Livable Cities'' and  SERDP grant \#2516: ``Climate Informed Estimation of Hydrologic Extremes for Robust Adaptation to Non-Stationary Climate''.
  The authors thank Yochanan Kushnir and James Booth for conversations regarding this work, and Casey Brown, Melissa Bukovsky, Rachel McCrary, Seth McGinnis, Linea Mearns, and Katherine Schlef for collaboration and advice.
\end{block}

### 2017-11-30 reports/writeup/paper.tex
re-work directory

%-------------------------------------------------------------------------------
% Preamble
%-------------------------------------------------------------------------------

\documentclass[twocol]{ametsoc}
\journal{jcli}

%Citations should be of the form ``author year''  not ``author, year''
\bibpunct{(}{)}{;}{a}{}{,}

%Packages
\usepackage{
	siunitx, % units
	cleveref,% cross-references
	physics, % better physics notation
	url, % for URLs
	amsmath, % for math
	array, % for fixed-width tables
	booktabs, % for nice tables
	nth, % for \nth{90}
}
\newcolumntype{L}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\graphicspath{{../figs/}}

%-------------------------------------------------------------------------------
% Title and Authors
%-------------------------------------------------------------------------------

\title{Heavy rainfall in Paraguay during the 2015-2016 austral summer: causes and subseasonal-to-seasonal predictive skill}
\authors{James Doss-Gollin\correspondingauthor{James Doss-Gollin. Columbia Water Center. Dept. of Earth and Environmental Engineering, Columbia University, 500 W. 120th St., New York, NY. USA.}}
\affiliation{Columbia Water Center. Dept. of Earth and Environmental Engineering, Columbia University, 500 W. 120th St., New York, NY. USA \\ Columbia Water Center, Columbia University, 500 W. 120th St., New York, NY. USA}
\email{james.doss-gollin@columbia.edu}
\extraauthor{\'Angel G. Mu\~noz}
\extraaffil{Atmospheric and Oceanic Sciences (AOS). Princeton University. Princeton, NJ 08540-6649. USA  \\
	International Research Institute for Climate and Society (IRI). The Earth Institute. Columbia University. Palisades, NY. 10964-1000. USA}
\extraauthor{Simon J. Mason}
\extraaffil{International Research Institute for Climate and Society (IRI). The Earth Institute. Columbia University. Palisades, NY. 10964-1000. USA}
\extraauthor{Max Past\'{e}n}
\extraaffil{Direcci\'{o}n de Meteorolog\'{i}a e Hidrolog\'{i}a. Paraguay \\ Facultad Polit\'ecnica. Universidad Nacional de Asunci\'{o}n. Paraguay}

%-------------------------------------------------------------------------------
% Abstract
% Up to 250 words in length
%-------------------------------------------------------------------------------

\abstract{
	During the austral summer 2015-16, severe flooding displaced over \num{170000} people on the Paraguay River system in Paraguay, Argentina, and Southern Brazil.
	These floods were driven by repeated heavy rainfall events in the Lower Paraguay River Basin.
	Alternating sequences of enhanced moisture inflow from the South American Low-Level Jet and local convergence associated with baroclinic systems were conducive to mesoscale convective activity and enhanced precipitation.
	These circulation patterns were favored by cross-timescale interactions of a very strong El Ni\~{n}o event, an unusually persistent Madden-Julian Oscillation in phases four and five, and the presence of a dipolar SST anomaly in the central southern Atlantic Ocean.
	The simultaneous use of seasonal and sub-seasonal heavy rainfall predictions could have provided decision makers useful information about the start of these flooding events from at least two-to-four weeks in advance.
	Probabilistic seasonal forecasts available at the beginning of November successfully indicated heightened probability of heavy rainfall (\nth{90} percentile) over southern Paraguay and Brazil for December-February.
	Raw sub-seasonal forecasts of heavy rainfall exhibited limited skill at lead times beyond the first two predicted weeks, but a Model Output Statistics approach involving principal component regression substantially improved the spatial distribution of skill for Week 3 relative to other methods tested including extended logistic regressions.
	A continuous monitoring of climate drivers impacting rainfall in the region, and the use of bias-corrected heavy precipitation seasonal and sub-seasonal forecasts, may help improve flood preparedness for the austral summer season in this part of the world.
}

\begin{document}

\maketitle

%-------------------------------------------------------------------------------
% Introduction
%-------------------------------------------------------------------------------

\section{Introduction}

During the austral summer of 2015-16, repeated heavy rainfall events led to severe flooding in the Paraguay River basin, displacing approximately \num{170000} people \citep{BrackenridgeDFO} and causing tremendous damage to property and infrastructure \citep{MinisteriodeObrasPublicasyComunicacion:2016tq}.
Because population in South America tends to concentrate along coasts and rivers (see supplemental figure S1), flooding in the Paraguay River basin directly affects not only much of the population of Paraguay, but also of populations in Argentina and Uruguay who lie along the Paran\'{a} and la Plata rivers into which the Paraguay River drains.
Heavy rainfall and flooding in the basin also has important implications for hydropower generation, for agriculture, and for regional water resource management.
The aim of this paper is to diagnose the drivers of the November-February (NDJF) 2015-16 rainfall and flooding events, and to assess the skill of the associated subseasonal-to-seasonal predictions.

The rainfall climatology of the Paraguay River basin varies strongly by season, with extratropical characteristics in the winter and monsoonal characteristics in the summer.
In the warm season (NDJF) which is the focus of this study, circulation is dominated by the upper-tropospheric Bolivian High, the lower-level subtropical highs, the Chaco Low over northern Argentina, the South Atlantic Convergence Zone (SACZ), and the South American Low-Level Jet, or SALLJ \citep{Grimm2009,Marengo:2012cm}.
Rainfall peaks around \SI{5}{\milli\meter\per\day} during the warm months (October-May) and reaches a minimum near \SI{2}{\milli\meter\per\day} in July and August.
However, the flat topography limits the river's ability to carry the summer runoff, causing seasonal inundation of the Pantanal and distributing the river discharge in time \citep{Bravo:2011et,Barros:2004bn}.
Thus, the streamflow maxima typically occurs in phase with precipitation upstream of the Pantanal, while downstream of the Pantanal -- which we define as the Lower Paraguay River Basin; see \cref{fig:study-area} -- the annual peak typically occurs between April and July.

During the warm season (NDJF) a large fraction of rainfall, and nearly all heavy rainfall, in the Paraguay River Basin is associated with mesoscale convection \citep{Velasco1987}.
Previous studies of organized convection and precipitation across subtropical continental South America have found close correspondence to the exit region of the low-level jets \citep{Saulo:2007km,Salio:2007gd,Marengo2004,Velasco1987}, which is influenced in both summer and winter by midlatitude baroclinic wave trains that interact with the Andes topography to generate orographically bound cyclones and northerly low-level flow \citep{Campetella:2002hx,Seluchi:2006bi,Boers:2014gt}.
The strength and direction of this moisture transport varies substantially between events; SALLJ exit regions range from central Argentina \citep[``Chaco Jet Events'';][]{Salio:2002ev} to Paraguay and southeastern Brazil \citep[``No-Chaco Jet Events'';][]{Vera:2006ib}.

At sub-seasonal timescales, convection in the Paraguay River Basin is modulated by a variety of drivers, notably including the SACZ and the Madden-Julien Oscillation (MJO).
During SACZ conditions, strong convergence is observed over the Amazon basin with divergence over southwestern Brazil, northern Argentina and Paraguay \citep{Herdies:2002jy,Carvalho2010}; the opposite is true for so-called NoSACZ conditions.
SACZ occurrence is related to westerly wind regimes, as well as ``active'' and ``break'' periods of the South American Monsoon System \citep{Marengo2004}.

At seasonal timescales, El Ni\~no-Southern Oscillation (ENSO) is the dominant driver of convection variability in the Paraguay River Basin.
During El Ni\~no years, a low-level anticyclonic anomaly over central Brazil enhances occurrence of the low-level jet, favoring the development of mesoscale convective systems \citep{Velasco1987}.
The intensity and precise extent of this anomaly is relevant for the precise impact of ENSO events.
The region also exhibits substantial variability between seasons of rainfall during El Ni\~no years, including a reversal of rainfall anomalies between November of that year and January of the following one, influenced by land-surface interactions \citep{Grimm2009,Grimm2003}.
Even beyond El Ni\~no years, regional land-surface feedbacks can cause regions that exhibit wet anomalies in the spring to experience more summer precipitation on average \citep{Grimm:2009bq}.
Similarly, mid-latitude dynamics influence low-level wind anomalies on many time scales, though this relationship is complicated due to coupled tropical-extratropical interactions \citep{Carvalho2004,Jones2002}.
To address these potential interactions, a cross-timescale approach based on synoptic circulation types is employed here to diagnose the causes of the rainfall events; this method has been used in previous work for southeastern South America \citep{Munoz2015,Munoz2016} and other regions \citep{Moron2015}.

We begin by describing our data sources in \cref{sec:data} and our methods in \cref{sec:methods}.
In \cref{sec:diagnostics} we start our diagnosis highlighting the observed flooding and contextualizing it within a long river stage time series; then we use composites and a weather typing analysis to diagnose the circulation patterns associated with the heavy rainfall during NDJF 2015-16. We turn in \cref{sec:fcsts} to the question of whether the observed rainfall was successfully predicted by available models.
To carry out this analysis we study both forecasts targeting the entire series for a limited area, and also forecasts targeting a large spatial area for only the first week of December, when the most important flooding events started.
We also explore the impact on forecasts of several bias-correction schemes.
In \cref{sec:discussion} we discuss limitations and potential implications of our findings and potential future work, and we present our concluding remarks in \cref{sec:summary}.

%-------------------------------------------------------------------------------
% DATA
%-------------------------------------------------------------------------------

\section{Data} \label{sec:data}

The analysis presented involves both observations and model forecasts.

\subsection{Observations}

The period analyzed for diagnostic purposes is from 1 Nov 1979 through 28 Feb 2016.
\Cref{fig:study-area} shows the study area and defines several spatial domains which are discussed throughout the paper.

Rainfall data are taken from the CPC Unified Gauge-based Analysis of Global Daily Precipitation dataset \citep{xie2010cpc}.
Spatial resolution is \SI{0.5}{\degree}, and temporal resolution is daily.
The retro and real-time datasets are combined and treated as a single dataset.
We define ``heavy'' rainfall events to be exceedances of the \nth{90} percentile; while the value is different for each grid cell, the \nth{90} percentile of area-averaged rainfall over the Lower Paraguay River Basin is approximately \SI{15}{\milli\meter\per\day}.

Atmospheric circulations are diagnosed using daily data from the NCAR-NCEP Reanalysis II dataset \citep{Kanamitsu:2002kk}.
Spatial resolution is \SI{2.5}{\degree}.
Because the end-of-day time for the rainfall data is 12:00GMT over most of South America \citep{xie2010cpc}, we use six-hour reanalysis data, and shift by twelve hours before re-sampling to the daily time step.
This ensures that the time steps in the reanalysis and rainfall data sets are the same, but means that a day is defined as beginning at 12:00 GMT.
Since most summer rainfall in this region occurs overnight, this end-of-day time (which translates to approximately 8:00 AM locally depending on the exact time zone) tends to separate distinct events.
Ocean sea surface temperature (SST) patterns are explored at the monthly time step using the \SI{1}{\degree} NOAA OI.v2 dataset \citep{Reynolds:2002iy}.

Streamflow data was collected by the Paraguayan Navy and National Administration of Navigation and Ports of Paraguay, and was processed and distributed by the Paraguayan Directorate of Meteorology and Hydrology (DINAC-DMH).
Locations of streamflow gauges are shown in \cref{fig:study-area}.
Because no stage-discharge curves are available, we present only the river stage values; while this is relevant from the perspective of flood damage, flow rates cannot be estimated without these curves (which are difficult to reconstruct as river geometry changes over time).

This study also makes use of some climate indices.
Data on the El Ni\~{n}o-Southern Oscillation (ENSO), or NI\~NO3.4, came from a statistical-dynamical interpolation \citep{Kaplan:1998df}, which is constrained by relatively high-quality observations during the study period.
Data on the Madden-Julien Oscillation (MJO) came from the Australian Bureau of Meteorology \citep{Wheeler:2004ea}.

\subsection{Model forecasts}

This study analyzes probabilistic seasonal and sub-seasonal forecasts of heavy rainfall events, defined as for observations by exceedance of the \nth{90} percentile of modeled rainfall.

The seasonal predictions used were ``flexible format'' forecasts provided by the International Research Institute for Climate and Society (IRI).
This product provides the probability of exceedance (or non-exceedance) of a user-specified percentile of the climatological distribution.
The DJF 2015-2016 forecasts analyzed were produced in November 2015.
Due to the short record of flexible format forecasts (2012-2016), no verification was performed for seasonal predictions.
These forecasts are provided at a horizontal resolution of \SI{2.5}{\degree}.

The sub-seasonal forecasts used were issued by the European Centre for Medium-Range Weather Forecasts (ECMWF) using a coupled model, available via the Subseasonal-to-Seasonal Prediction Project Database \citep{Vitart2016} at \SI{1.5}{\degree} resolution.
Forecasts consider the period starting in Dec 2015 until Mar 2016 and hindcasts to assess the real-time predictive skill consider the period Dec 1-7, 1995-2014.
There is a total of 51 ensemble members for each forecast, and 11 ensemble members for each of the 20 hindcasts (Dec 1-7, 1995-2014).

Hindcasts were used to define the significant event threshold, and for probabilistic forecast verification; forecasts were used to analyze modeled rainfall during the entire NDJF 2015-16 season and in particular the week of Dec 1-7, 2015.
For probabilistic analysis of the rainfall during the week 1-7 December 2015, rainfall forecasts and hindcasts considered were initialized on November 12th and 16th, 2015.

Anomalies were calculated relative to the seasonal mean from November 1979 to February 2016, thus retaining the seasonal cycle.

%-------------------------------------------------------------------------------
% METHODS
%-------------------------------------------------------------------------------

\section{Methods} \label{sec:methods}

Several types of analyses are used to diagnose the causes of the heavy rainfall events, and to bias-correct and verify the forecasts.
All codes used for data and fully reproducible analysis are available at \url{https://github.com/jdossgollin/PYFloods}.
Computation was performed in the python environment using the \texttt{xarray} \citep{hoyer2017xarray} and \texttt{numpy} \citep{vanderWalt:2011dp} packages for gridded data and the \texttt{pandas} \citep{McKinney:2010un} package for tabular data; plotting was performed with the \texttt{matplotlib} \citep{Hunter:2007ux} and \texttt{cartopy} \citep{Cartopy} packages.

Given the behavior of the Paraguay River discussed above, we define the Lower Paraguay River Basin as the region bounded by \SIrange{-59.75}{-55.75}{\degree W} and \SIrange{26.75}{22.75}{\degree S}, as shown in \cref{fig:study-area}.
This region corresponds to the region over which, given topography and previous studies \citep{Barros:2004bn,Bravo:2011et}, one might hypothesize rainfall inputs to most closely correspond to river levels at the stream gauges in \cref{fig:study-area}.
Results are not sensitive to small variations in this domain.

\subsection{Weather Typing} \label{sec:weather-typing}

A cluster algorithm is used on daily data to diagnose mechanisms associated with the rainfall events of interest in this research.

The clustering was performed on the daily NDJF \SI{850}{\hecto\pascal} streamfunction field ($\psi$), calculated by integrating the meridional and zonal wind fields using spherical harmonics, as implemented in the \texttt{windspharm} package \citep{Dawson:2016ws,Adams:1999fb}, over the domain spanning \SIrange{15}{30}{\degree S} and \SIrange{65}{45}{\degree W} (\cref{fig:study-area}).

To facilitate clustering (which tends to perform poorly in high-dimensional spaces), the NDJF anomaly field of $\psi_{850}$ was projected  onto its four leading empirical orthogonal functions (EOFs), accounting for $>95\%$ of the total observed variance (supplemental figure S2), using the \texttt{eofs} package \citep{Dawson:2016ge}.
No meridional weighting was applied as the selected domain is relatively small and does not extend into high latitudes.
Once the EOFs were calculated, the principal component time series were computed for each day and scaled to unit variance.
This rescaling is not a necessary step; its effect is to treat all retained principal components as equally important for the weather typing, which provides relatively greater weight to EOFs 2, 3, and 4 than carrying out weather typing without re-scaling.
Though our approach of first selecting the number of EOFs to use and then choosing to scale them equally involves more subjective decisions than an approach without rescaling, in this case the resulting physical mechanisms are more interpretable.

Next, the $K$-means algorithm was used to assign a single cluster value to each day on record using the 4-dimension principal component time series.
The $K$-means technique is a partitioning method that classifies all days in the study into a predefined number of clusters.
The algorithm proceeds as follows:
\begin{enumerate}
	\item Randomly choose $K$ cluster centers $\mu_1^{(0)}, \ldots, \mu_K^{(0)}$ (where $0$ refers to the $0$th iteration)
	\item Iterate until convergence, indexing each iteration with $j$:
	\begin{enumerate}
		\item Assign each observation (day) $x_i$ to the nearest cluster center; we define this using the Euclidean distance but other measures, such as the Mahalanobis distance, could also be used:
		\begin{equation}
			m_i^{(j+1)} := \arg\min_{k\in 1,\ldots,K} \abs{\abs{x_i - \mu_k^{(j)}}}
		\end{equation}
		\item Recompute the cluster centers as the mean of all points assigned to that cluster
		\begin{equation}
			\mu_k^{(j+1)} := \frac{1}{\abs{\qty{i \big| m_i^{(j+1)}=k}}} \sum_{i | m_i^{(j+1)}=k} x_i
		\end{equation}
		where $\abs{\cdot}$ denotes vector length.
		\item Stop iteration if the change in centroids $\mu^{(j+1)} - \mu^{(j)}$ is less than a small but non-zero tolerance parameter $\tau$.
	\end{enumerate}
\end{enumerate}
The cluster centroids (weather types) $\mu_k$ produced by the $K$-means algorithm can then be interpreted as a Voronoi decomposition of the phase space into $K$ regions, and specifically as the Voronoi diagram which minimizes within-cluster variance.

The $K$-means algorithm is guaranteed to converge to a local minimum of intercluster variance; to select the best partition, 500 simulations were created using the implementation in Python's \texttt{scikit-learn} package \citep{Pedregosa:2012tv} and then the classifiability index of \citet{Michelangeli1995} was computed between each partition and the 499 others.
The partition whose classifiability index, averaged for all 499 pairwise comparisons, was the highest was selected.
Calculation of the classifiability index for several values of $K$ (supplemental figure S3) suggests that states with $K=5,6,\ldots,9$ are all reasonable; we chose the solution $K=6$ because the weather types identified are qualitatively similar to those determined over southeastern South America \citep{Munoz2015,Munoz2016} and have an intuitive physical meaning, discussed further in the following sections.

\subsection{Forecasts and Model Output Statistics (MOS)}

A wide variety of methods, generically known as Model Output Statistics \citep[MOS;][]{Glahn:1972vt}, have been proposed to correct for different types of bias in model outputs.
In this work, we analyze how well the rainfall events could have been predicted, both using the raw sub-seasonal forecasts and  MOS-adjusted sub-seasonal forecasts.
We use  four types of MOS techniques: the homoscedastic and heteroscedastic versions of the extended logistic regression (XLR and HXLR, respectively); principal component regression (PCR); and canonical correlation analysis (CCA).

Logistic regression models the conditional probability of binary events, and it has been used in MOS approaches for binary predictands \citep{Hamill:2004hk}.
Nonetheless, when using the logistic regression to address multiple thresholds via independent fits, the predicted probabilities are, in general, not mutually consistent \citep{Messner:2014gp}.
The (homoscedastic) XLR was designed to address this shortcoming \citep{Wilks:2009bk}, via the consideration of a transformation of the thresholds of interest as an additional predictor variable.
The HXLR, a heteroscedastic generalization of the XLR, was proposed to appropriately use the ensemble spread as predictor for the dispersion of the predictive distribution \citep{Messner:2014gp}.

CCA is a common statistical method frequently used to forecast rainfall using a purely empirical approach \citep{Mason:2008da,Barnston:2012ce,Jolliffe2012:vm,Barnston:1992gd,Wilks:2006fx}.
CCA identifies modes of co-variability, called canonical variates or canonical modes, by maximizing the correlation between linear combinations of the predictor and predictand's empirical orthogonal functions (EOFs).
The method forecasts spatial patterns of variability spanning across the region of interest rather than making forecasts for individual locations.
In PCR, a special case of CCA, each grid cell in the predictand field is estimated by regression using a linear combination of the predictor's EOFs \citep{Mason:2008da,Wilks:2006fx} rather than by identifying canonical modes.
Unlike the XLR and HXLR models, which perform bias correction independently for each grid cell, the CCA and PCR models can address biases in both the magnitude and the spatial distribution of the modeled precipitation patterns.

For the purposes of MOS corrections, the predictand (variable to forecast) is the observed rainfall for the target period of interest, and the predictor (variable to be corrected) is the raw model rainfall for the same period.
Exceedance of the \nth{90} percentile during the 1995-2014 period is used to define the heavy event cases.
We use the same spatial domain [\SIrange{39}{17}{\degree S}; \SIrange{66}{49}{\degree W}] for both the predictor and the predictand, except for the PCR and CCA cases, in which a larger domain [\SIrange{0}{60}{\degree S}; \SIrange{80}{30}{\degree W}] was used to better capture the spatial patterns in the raw model field.
A variety of domains and ways to combine initialization times were explored; the best results were selected in terms of the corresponding Kendall's $\tau$ rank correlation coefficient.
A summary of the final candidate predictors found to be most skillful for each MOS model is presented in \cref{tab:mos-methods}.

To avoid artificial skill when building the statistical models, we employ a cross-validation approach with a 5-year window.
In this framework, five continuous years are left out of the record, the regression coefficients are computed with the remaining of the time series, and the resulting model is validated comparing the prediction for the third year left out (middle of the window) against observations.
The 5-year-long window is redefined a year at a time, moving from the beginning of the record to its end.

To visualize the probability of heavy rainfall at each grid cell, we present all predictions in terms of odds relative to the climatological odds:
\begin{equation} \label{eq:odds-ratio}
	\text{odds}_{r} \equiv \frac{p}{\qty(1 - p)} \frac{\qty(1 - p_c)}{p_c}
\end{equation}
where $p$ and $p_c$ represent the forecast probability for the exceedance of the \nth{90} percentile, and the related climatological probability, respectively.

The IRI's seasonal predictions are already provided with bias correction \citep{Barnston:2010ge}, and thus we did not perform any further MOS on seasonal rainfall fields.

\subsection{Probabilistic Forecast Verification}

In addition to visually comparing predictions and observations to verify how well the heavy rainfall events could have been predicted, we use the Ignorance Score,
\begin{equation}\label{eq:ignorance}
	\text{IGN} \equiv - \log_2 p(Y),
\end{equation}
where $Y$ is the observed outcome and $p(Y)$ is the density function of the forecast distribution \citep{Good:1952ux,Roulston:2002jz,Brocker:2007er}.
The Ignorance score was introduced as the fundamental information-based verification measure,  decomposable into easily interpretable components: reliability, resolution and uncertainty \citep{Weijs:2010cl}.
Due to its close relationship to Shannon's information entropy, it is used to measure forecast utility, or the amount of information gain expected from a forecast \citep{Roulston:2002jz}.

We also compute the Generalized Relative Operating Characteristics (GROC), or 2AFC score \citep{Mason:2009kr}, to evaluate skill of probabilistic rainfall forecasts.
This score measures the ``proportion of all available pairs of observations of differing category whose probability forecasts are discriminated in the correct direction'' \citep{Mason:2009kr}.
It has an intuitive interpretation as an indication of how often the forecasts are correct.

These two metrics, measuring reliability, resolution, uncertainty and discrimination, are deemed here to be sufficient to characterize the forecast skill for our events of interest.
To conduct the verification in a consistent manner, we use the Climate Predictability Tool (CPT), developed by the International Research Institute for Climate and Society \citep{Mason:2017gg}.

%-------------------------------------------------------------------------------
% DIAGNOSTICS
%-------------------------------------------------------------------------------

\section{Diagnostics} \label{sec:diagnostics}

%-------------------------------------------------------------------------------
% OBSERVED FLOODS
%-------------------------------------------------------------------------------

\subsection{Observed Flooding} \label{sec:flood}

\Cref{fig:streamflow} shows the streamflow time series at several gauges on the Paraguay River during NDJF 2015-16 in the context of their seasonality and decadal variability.
During November and December 2015, the river rose rapidly at Concepci\'on, Asunci\'on, and Pilar though not at Bah\'ia Negra.
As discussed in \citet{Bravo:2011et,Barros:2004bn}, the location of the Bah\'ia Negra gauge (see \cref{fig:study-area}) in the Pantanal region means that it responds very slowly to rainfall input.
The three downstream gauges, because they are located in the Lower Paraguay River basin respond to the rainfall forcing with a slow but steady rise.
Despite several very heavy storms, the streamflow record at Asunci\'on and Pilar (which are downstream of Concepci\'on) indicates relatively little response to individual storms.
Because the region is so flat (see topographic data in \cref{fig:study-area}), river levels at a particular point may be affected not only by rain in the catchment corresponding to that point, but also by elevated river levels downstream which reduce the pressure gradient available to drive flow.

Examination of \cref{fig:streamflow}b suggests multidecadal oscillation in the streamflow record.
This is in agreement with previous studies \citep{Collischonn:2001bi,Carvalho2011} which find a changepoint in the 1970s, possibly associated with low-frequency Pacific variability.
Because only river stage data (and not discharge) are available, it is not possible to discern whether the observed changes in river stage are driven by sediment loading and local measurement characteristics or by large-scale climate fluctuations.
Further treatment of this question is beyond the scope of this paper.

\subsection{Heavy Rainfall: Climatological Drivers} \label{sec:rainfall-circulation}

To understand how circulation anomalies observed during NDJF 2015-2016 led to the observed floods it is helpful to first explore the atmospheric circulations which are typically associated with heavy rainfall in the lower Paraguay River.

\Cref{fig:lagged-rain} shows time-lagged anomalies up to and after heavy rainfall dates (when area-averaged daily rainfall in the lower Paraguay River Basin exceeds its NDJF \nth{90} percentile) and is consistent with previous findings \citep{Marengo2004,Salio:2007gd,Marwan2015}.
At $t=-2$ a mid-latitude baroclinic system approaches the South American continent, intensifying and moving to the East at $t=-1,0,1$.
This system interacts with the sub-tropical low and the Andes Mountains to produce an anticyclonic anomaly over Brazil.
Along the cold front associated with this system, a low-level northerly jet advects  heat and moisture to the region.
As the system progresses, the jet below \SI{20}{\degree S} transitions from predominantly meridional flow (``Chaco Jet''; $t=-1$ to predominantly zonal flow (``No-Chaco Jet'').
Analysis for the \nth{95} or 99th percentiles of daily rainfall (not shown) yield similar results, implying that the synoptic mechanism for the most heavy events is not fundamentally distinct from the mechanism for moderate-intensity events.
This mean field, like all composites, masks between-event variation but exploration of individual events (not shown) shows that the core features identified are generally present.

\subsection{Weather Type Analysis: Daily Circulation Patterns} \label{sec:weather-types}

We next use the weather typing algorithm outlined in \cref{sec:methods}\ref{sec:weather-typing} to understand particular circulations and sequences of circulations associated with heavy rainfall in the Lower Paraguay River Basin.

The first step of the weather typing algorithm is to identify leading EOFs of the \SI{850}{\hecto\pascal} streamfunction $\psi$.
The EOF loadings are shown in \cref{fig:eof-loading}.
Of these, EOF 1 explains a substantial amount of variance ($\approx 67\%$) while EOFs 2, 3, and 4 collectively explain approximately $28\%$ of total variance.
Values of individual and cumulative variance explained are presented in supplemental figure S2.
To a first order of approximation, EOF 1 explains the large-scale anomalies induced by passing baroclinic waves; the East-West gradient is likely a result of the influence of the Andes Mountains which are a source of friction and torque for low-level flow.
EOF 2 primarily controls the strength of the low-level jet, while EOFs 3 and 4 control the exit region of this jet.
Given these features in the EOF loadings, it is no surprise that composites of the  \SI{850}{\hecto\pascal} circulations and rainfall associated with each weather type (\cref{fig:wt-composite}) reveal patterns associated with synoptic- and regional-scale circulation regimes.

Weather type 1 represents a ``Chaco Jet'' event \citep{Salio:2002ev} in which the maximum wind speed penetrates past \SI{25}{\degree S}, leading to heavy rainfall over NE Argentina and Uruguay.
Weather type 4 is also associated with a SALLJ event, but is of the ``No-Chaco'' jet type in which the wind turns to the East before \SI{25}{\degree S}, leading to heavy rainfall over Eastern Paraguay and SW Brazil.

Weather types 5 and 3 are nearly inverses of weather types 1 and 4, respectively, and are associated with dry anomalies over the Lower Paraguay River Basin.
The fact that they are not exact inverses suggest important nonlinearities in the system.
Finally, weather types 2 and 6 are related to a high-pressure configuration bringing below-average rainfall over most of Brazil, and a dipolar pattern conducive to above-average rainfall over central Brazil, respectively (\cref{fig:wt-composite}).

\subsection{NDJF 2015-16: Circulation Sequences}

We next use monthly-mean circulation anomalies (spatial patterns) and weather type sequences (time patterns) to understand the specific events of NDJF 2015-16.

While weather typing requires simplifying the dynamics of daily circulation patterns, its advantage is that it greatly facilitates the analysis of sequences of precipitation.
\Cref{fig:rain-wt} shows a time series of area-averaged rainfall over the Lower Paraguay River Basin for NDJF 2015-16 and the corresponding weather types.
This plot shows that heavy rainfall concentrated over a period spanning from mid-November 2015 through early January 2016, with shorter peaks in late January and mid-February.

As indicated in \cref{fig:rain-wt}, the most heavy rainfall occurred during weather types 1 and 4.
During NDJF 2015-16 weather types 1 and 4 (Chaco and No-Chaco jet extensions, respectively),  occurred more frequency than their climatology (supplemental table S1); weather type 2 also occurred more frequently than its climatology, largely due to a long sequence in February 2016.
In mid-January 2016, during a sequence of persistent low rainfall, weather type 3 featured persistently, leading to heavy rainfall over central Brazil (not shown) and negative rainfall anomalies over the Lower Paraguay River Basin.
Thus while the intensity and persistence of heavy rainfall was atypical, the causal mechanism of the heavy rainfall observed during this season was consistent with climatology.

Inspection of \cref{fig:rain-wt} also suggests that at time scales of days to weeks, particular sequences of weather types tend to recur, and are associated with repeated rainfall storms.
The rainfall in late November 2015 occurred during weather type 1 and 4 days, with intermittent (and dry) weather type 5 occurrences.
During mid to late December 2015, weather types 1 and 5 produced rainfall but the intermittent dry days were weather type 2.
During early January 2016 weather type 3 led to persistent low rainfall.

Transitioning from exploring the time evolution of the reduced-dimension system represented by the weather types, monthly-scale circulation anomalies (\cref{fig:anomalies}) show a weak anticyclonic circulation that set up over central Brazil during November 2015 and strengthened into the following month.
In January 2016 it weakened before returning in February 2016.
The observed rainfall and circulation anomalies are consistent with the aggregation of the observed weather types shown in \cref{fig:rain-wt} and discussed above.

%-------------------------------------------------------------------------------
% FORECASTS
%-------------------------------------------------------------------------------

\section{Forecasts} \label{sec:fcsts}

In this section we analyze the extent to which forecasts were able to predict the persistent rainfall during summer of 2015-16.

There are advantages in simultaneously considering useful climate information at multiple timescales, rather than just focusing on one of them.
This multistage prediction system is known as the IRI's ``Ready-Set-Go" approach \citep{Hellmuth:2011vt,Goddard:2014kf}.
For the sake of concision, in this section we discuss only probabilistic seasonal (DJF 2015-2016) and sub-seasonal (Dec 1-7, 2015) forecasts.

\subsection{Seasonal Forecast}

Heavy rainfall over the region was forecast for the DJF 2015-2016 season since at least November 2015 (see \cref{fig:seas-prob-fcst}).
Relative odds as high as 7:1 are visible over southern Paraguay and Brazil, and northern Uruguay and Argentina, generally in agreement with observations.
The model predicted only very weakly increased odds of heavy rainfall in the Pantanal region (directly North of the Lower Paraguay River Basin) and in northern Argentina at $\approx \SI{65}{\degree W}$, and missed the heavy precipitation along most of the northeastern border of Paraguay.
However, the regionally elevated forecast of heavy rainfall could have been used for disaster preparedness at least one month in advance.

\subsection{Sub-seasonal Forecasts}

Sub-seasonal predictions are still too new to be used as operational tools, and their skill is normally not high enough to be useful for most decision-making \citep{Vigaud:2017gg}.
Nonetheless, the international S2S Prediction Project \citep{Vitart2016} has been providing since 2015 free access to almost-real-time sub-seasonal forecasts from multiple models, an opportunity to explore how well the heavy rainfall events of the first week of December 2015 could have been predicted.

\Cref{fig:chiclet} uses a Chiclet diagram \citep{Carbin:2016fx} to visualize, as a function of lead time, the time evolution of the uncorrected, ensemble-mean rainfall anomaly forecast, spatially averaged over the Lower Paraguay River Basin.
At times greater than about two weeks, the ensemble-mean forecast is slightly wetter than climatology.
At weather timescales (less than one week), the ensemble-mean successfully predicts the timing and amplitude of the area-averaged rainfall.
At intermediate timescales, the model successfully forecast the strongest breaks and pauses in the rainfall, such as the heavy rainfall during December 2015 and the dry period during mid-January 2016.

To examine these forecasts more closely, we turn to the 14-19 day forecast of the December 1-7 2015 period.
As seen in \cref{fig:subs-prob-fcst}, the raw (uncorrected) sub-seasonal forecast of the ECMF model for Dec 1-7 2015 indicated very high relative odds for occurrence of heavy rainfall but with important biases in the actual location and spatial pattern; for Paraguay, it confidently suggests occurrence of heavy rainfall to the south-southeast of the country, which was mostly not observed.
Overall, the 20-year skill of probabilistic forecasts for the first week of December -- even when using  multiple initializations -- is inferior to climatology at the lead times used (14 and 19 days) for most of Paraguay (\cref{fig:subs-prob-fcst}b).
Despite these biases, the model is capturing a signal, which suggests the use of Model Output Statistics to explore the extent to which corrections in the magnitudes and spatial patterns may improve the forecast.

In general, the use of extended logistic regression models does not improve the forecast for the week.
For example, with respect to the raw prediction, XLR tends to amplify the relative odds, and to cluster and shift the forecast location of the heavy rainfall events (\cref{fig:subs-prob-fcst}, top row, first two columns); the forecast tends to be better for Uruguay, but suggests heavy rainfall in the Paraguayan Chaco, which was not present in the raw prediction.
On the other hand, the use of the ensemble spread in the HLXR model does not help; this forecast tends to be over-confident on the events occurring in almost all the region of interest (\cref{fig:subs-prob-fcst}, top row, third column).

Comparison of long-term skill between the raw model output and both extended logistic regression models shows similar results. Reliability, resolution and uncertainty, as measured by the ignorance score (\cref{fig:subs-prob-fcst} middle row, first three columns), suggests slight skill improvement in southern Brazil, deterioration in Argentina and Uruguay, and basically the same as the raw model output for Paraguay and southeastern Bolivia.
Changes in forecast discrimination exhibited by the extended logistic models, as measured via the 2AFC score (\cref{fig:subs-prob-fcst} bottom row, first three columns), are  null.
These models operate on a gridbox-by-gridbox basis to recalibrate the probabilities, and so this recalibration happens monotonically.
Since the 2AFC score is insensitive to monotonic transformations of forecasts, the forecast discrimination is unchanged.

Better forecasts are obtained when both magnitude and spatial corrections are performed, although with relative odds considerably less confident than the ones in the raw forecast.
The Principal Component Regression model correctly shows high relative odds in most of the places where heavy rainfall was observed (\cref{fig:subs-prob-fcst}, first row, fourth column), although it also indicates heightened risk in areas where heavy rainfall did not occur, like zones of western Paraguay and northeastern Argentina.
The main problem with the Canonical Correlation Analysis model is its lack of discrimination between occurrence or non-occurrence of heavy rainfall in the region: the spatial distribution of odds is too homogeneous (\Cref{fig:subs-prob-fcst}, first row, fifth column).

The 20-year based skill maps of probabilistic forecasts (see \Cref{fig:subs-prob-fcst}, middle and bottom row) computed with these two EOF-based models are very similar to each other, both in terms of the reliability, resolution and uncertainty (Ignorance Score; second row, fourth and fifth columns) and discrimination (2AFC score; middle row, fourth and fifth columns).
Compared to the raw model output and the extended logistic regression models, PCR and CCA models perform quite well (\cref{fig:subs-prob-fcst}, middle and bottom row).
The enhanced skill is achieved through the spatial corrections via the EOF-based regressions, which -- in contrast with the extended logistic models -- use information from multiple grid-boxes, and thus the original forecasts are not (necessarily) calibrated monotonically.

Despite the particular errors in the Dec 1-7 2015 forecasts, both PCR and CCA verify considerably better than the raw, XLR, and HXLR predictions.
Yet despite the generally high skill score for these forecasts, there are still zones along the eastern part of Paraguay with negative skill relative to climatology.

%-------------------------------------------------------------------------------
% DISCUSSION
%-------------------------------------------------------------------------------

\section{Discussion}
\label{sec:discussion}

Although many of the individual rainfall events of NDJF 2015-16 were intense, they were nontheless driven by the climatological mechanism for heavy rainfall shown in \cref{fig:lagged-rain} and discussed in more detail in previous studies \citep{Saulo:2007km,Salio:2007gd,Marengo2004,Velasco1987}.
Consequently, the most striking hydrometeorlogical feature of this season, and the leading cause of the extreme flooding, was the persistence of the heavy rainfall and the manner in which it switched ``on'' and ``off'' during the season (\cref{fig:rain-wt}).
Understanding the mechanisms and possible cross-timescale interactions between climate drivers that drove this behavior has the potential of improving model forecast skill, particularly at sub-seasonal timescales, presently a growing area of active research.

As a starting point, we consider the roles of El Ni\~{n}o, which has been extensively studied in the literature \citep{Carvalho2004,Velasco1987,Salio:2002ev,Bravo:2011et,Grimm2009}, and the MJO.
As indicated earlier, examination of the \SI{850}{\hecto\pascal} streamfunction field (\cref{fig:eof-loading}) suggests that EOF 2 represents the intensity of the SALLJ, while EOF 3 describes the East-West extent of the jet exit region -- in other words, whether it is a Chaco or No-Chaco jet event.

These circulations are strongly modulated by ENSO and the MJO: the Spearman (rank) correlation coefficient between the monthly-averaged loading EOF 2 and the NI\~NO3.4 index is \num{-.369} ($p \ll 0.01$), and the Spearman correlation between the daily loading of EOF 3 the MJO RMM1 index is \num{0.130} ($p \ll 0.01$).
A key question is how the combined effects of potentially predictable seasonal to sub-seasonal climate mechanisms modify the occurrence of each weather type.
The typical behavior is summarized by \cref{fig:wt-mjo-enso}, following the approach of \citet{Munoz2016} and \citet{Moron2015}.
During El Ni\~{n}o years such as 2015-16, weather type 1 -- related to Chaco jet events -- occurs more frequently for almost all MJO phases, and particularly during phases 1, 2, and 5, consistent with observations of the year under study.
The occurrence of WT3, favored during MJO phases 8 and 1, in late January 2016 and of WT2, enhanced during MJO phases 4-7, in mid-February 2016 is also consistent with the observed MJO time series (supplemental figure S4) and observed weather type sequences (\cref{fig:rain-wt}).

While the occurrence of WT1 during NDJF 2015-16 is well explained by ENSO and MJO variability, these features alone do not explain the frequent occurrence of WT4, the ``No-Chaco'' jet event.
Previous studies on the South American Low-Level Jet \citep[e.g.,][]{Vera:2006ib} and the modulation of rainfall in southeastern South America by extratropical transient wave trains during El Ni\~no years \citep{Barreiro:2017ct} emphasize the importance of Pacific-Atlantic interaction for forecasting climate events in this (and other) region(s).
A persistent dipolar SST anomaly in the central southern Atlantic Ocean favors the occurrence of WT4 by blocking transient extratropical wave activity coming the Pacific, facilitating transitions from Chaco jet events (WT 1) to No-Chaco jet events (WT 4) via enhanced low-level wind circulation from southern Brazil towards the Atlantic, and back to north-east Brazil and the Amazon (see \cref{fig:chaco-nochaco}) due to land-sea temperature contrasts.
We illustrate this mechanism in \cref{fig:chaco-nochaco} and note that it is consistent with the mechanism found to produce heavy rainfall in the Lower Paraguay River Basin (\cref{fig:lagged-rain}) and with previous studies \citep[e.g.,][]{Vera:2006ib,Liebmann:2004kx,Salio:2002ev}.
This dipole is present in a SST anomaly composite of the 15 NDJF months with the most frequent occurrence of WT 4 (shown in \cref{fig:wt4-months}), and it does not change substantially if more or fewer months are used for the composite.
The pattern is also consistent with the observed dipolar structure in SST anomalies that was present in the central southern Atlantic Ocean since mid November 2015 to roughly the end of December 2015, associated with an anomalous anti-cyclonic configuration localized around \SI{35}{\degree W} and \SI{30}{\degree S}.

A potential source of dynamical model biases in the representation of rainfall events in the region is related to simulation of the extratropical wave activity mentioned before.
For example, in order to adequately forecast rainfall in certain parts of southeastern South America during El Ni\~no years, models need to reproduce not one but two stationary wave trains, originating from central and eastern Pacific, respectively \citep{Barreiro:2017ct}.
The frontal activity is associated with the extension, shape and variability of SACZ, and could be blocked by the presence of anomalous anti-cyclones off the coast of Brazil, like the ones described in reference to No-Chaco jet events.

Other mechanisms that have been known to modulate rainfall signals in this region include the South Atlantic Convergence Zone \citep{Carvalho2004,Munoz2015,Munoz2016} and land-biosphere-atmosphere interactions \citep{Grimm:2000bt,Grimm:2009bq} which also tend to be poorly represented in models \citep{Green:2017cs,Koster:2004ge}.
The analysis of these sources of systematic errors could require a cross-timescale diagnostic approach \citep{Munoz:2017ct}, and it is beyond the scope of the present paper.
Analysis of La Ni\~na and ENSO-neutral years (\cref{fig:wt-mjo-enso}) also suggests modulation of weather type occurrence by MJO-ENSO interactions, consistent with the cross-timescale hypothesis detailed in \citet{Munoz2015}.
This cross-timescale interaction also helps to explain why sub-seasonal forecasts were not substantially better than seasonal ones: although the ENSO signal was homogeneous throughout the season, studies of MJO predictability have found it to be limited beyond two to three weeks \citep{Ding:2010em,Vitart:2014bt,Kim:2014bt}.

The nonlinear dynamics introduced by the coupling of the land-biosphere-ocean-atmosphere system can lead to favored trajectories through particular regions of phase space; these can be understood as ``weather regimes'' \citep[see][for a comprehensive discussion]{Hannachi2017}.
Looking in southeastern South America (slightly South and East of the Lower Paraguay River Basin defined in this study), \citet{Munoz2016} used the sequences of weather types identified for each year to analyze predictability of heavy rainfall events.
Although this approach -- as with the weather typing used in this study -- simplifies the dynamics of the system, it can be a valuable tool for diagnosing and assessing model processes \citep{Munoz:2017ct}, and for communicating model forecasts to policy-makers.
Such an assessment is beyond the scope of this case study but offers promising opportunities for local public and private sectors to use climate model forecasts.

Although we motivated this work by describing the impacts of severe flooding in the Lower Paraguay River Basin, the analysis presented has focused on climate drivers of rainfall.
Indeed, rainfall and streamflow are related, but the link is nonlinear in space and time.
As explained in \cref{sec:diagnostics}\ref{sec:flood}, in this region the flat topography (\cref{fig:study-area}) means that the Lower Paraguay River reacts slowly to rainfall \citep{Bravo:2011et,Barros:2004bn}, explaining the slow but steady rise in river levels shown in \cref{fig:streamflow}.
Groundwater dynamics are also important in explaining this behavior \citep{Santos:2016td}.
The observed flood peaks during 2015-16 also seem to occur in the context of an active phase of a multi-decadal oscillation, possibly associated with low-frequency Pacific activity \citep{Collischonn:2001bi,Huang:2005ir}.
Parsing the relative impacts of deforestation and land use changes in the river basin, installation of hydroelectric generation at the Itaipu and Yacyreta sites, river channel modification, and climate variability on flood levels will require gathering improved hydrological data and building a comprehensive system model but all of these factors may influence the timing and magnitude of flooding in this region.

From a policy perspective, reducing flood risk exposure in this region is key to reducing flood losses; flood events not only in 2015-16 but also in 2014 and 2017 caused substantial damage, and highlight the need for flood risk management.
Doing so will require compiling information on the properties, businesses, and infrastructure that are vulnerable to flooding.
This study also suggests that proposed dredging of the upper Paraguay River Basin to facilitate navigation could lead to increased summertime streamflow from the Upper Paraguay River Basin (Pantanal), effectively coupling the phases of streamflow from the Upper and Lower Paraguay River Basins which currently have a time-delay \citep{Bravo:2011et}.

%-------------------------------------------------------------------------------
% Summary
%-------------------------------------------------------------------------------

\section{Summary} \label{sec:summary}

In this study we examined the regional climate drivers of the persistent and heavy NDJF 2015-16 rainfall over the Lower Paraguay River Basin, leading to severe flood impacts.
The observed flooding was made more likely by antecedent soil moisture left by the preceding rainy season and followed two years of substantial, though less extreme, flooding.
These antecedent conditions combined with the persistent heavy rainfall occurring in November and December 2015 to produce heavy flooding on the Lower Paraguay river.

The heavy rainfall observed was driven by enhanced moisture inflow from the low-level jet and convergence associated with baroclinic systems.
Repeated SALLJ events, particularly No-Chaco jet events, led to favorable conditions for mesoscale convective activity in this region.
This activity was modulated by cross-timescale dynamics including a strong El Ni\~{n}o and an active MJO in phases 4-5, which favored SALLJ occurrence, and the presence of a dipolar SST anomaly in the central southern Atlantic Ocean, which enhanced the probability of occurrence of No-Chaco jet events.

Numerical forecasts skillfully predicted enhanced risk of heavy rainfall at the seasonal scale, likely guided by ENSO signal, but biases in the spatial patterns suggest that models are not well capturing the physical interactions between the Pacific and the Atlantic basins.
At sub-seasonal time scales, uncorrected model forecasts of rainfall had limited skill beyond 10-15 days, though use of Model Output Statistics -- particularly the PCR and CCA methods that correct both spatial patterns and magnitudes -- substantially improved forecast skill.

%-------------------------------------------------------------------------------
% Acknowledgments
%-------------------------------------------------------------------------------

\acknowledgments
The authors thank David Farnham, Upmanu Lall, and Andrew Robertson for insightful conversations and guidance.
JDG thanks the NSF GRFP program for support (grant DGE 16-44869).
AGM was supported by the Atmospheric and Oceanic Sciences (AOS) Program at Princeton University.
Streamflow data shown in \cref{fig:streamflow} was collected by the Paraguayan Navy and National Administration of Navigation and Ports of Paraguay, and was processed and distributed by the Paraguayan Directorate of Meteorology and Hydrology (DINAC-DMH).
The authors also are thankful to the WWRP/WCRP Subseasonal-to-Seasonal Prediction Project Database, available via the IRI Data Library mirror at \url{https://iridl.ldeo.columbia.edu/SOURCES/.ECMWF/.S2S/}.

%-------------------------------------------------------------------------------
% REFERENCES
%-------------------------------------------------------------------------------

\bibliographystyle{ametsoc2014}
\bibliography{library}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% TABLES
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table*}
%
\caption{
	Model Output Statistics (MOS) methods used to correct the ECMF sub-seasonal forecasts.
	Spatial domain for predictand is always the same (\SIrange{39}{17}{\degree S}; \SIrange{66}{49}{\degree W}).
	Two initializations are used: Nov 12th and 16th, 2015.} \label{tab:mos-methods}
\begin{center}
\begin{tabular}{L{0.75in} L{1.15in} L{3.5in}}
\toprule
Model & Region (Predictor) & Final predictor(s) selected \\
%
\midrule
%
\emph{Raw} & \SIrange{39}{17}{\degree S}; \SIrange{66}{49}{\degree W} & Ensemble mean, computed using members from the  two initializations. No correction performed. \\
%
\emph{XLR} & \SIrange{39}{17}{\degree S}; \SIrange{66}{49}{\degree W} & Ensemble mean, computed using members from the  two initializations.  \\
%
\emph{HLXR} & \SIrange{39}{17}{\degree S}; \SIrange{66}{49}{\degree W} & Ensemble mean and spread, computed using  members from the two initializations.\\
%
\emph{PCR} & \SIrange{60}{0}{\degree S}; \SIrange{80}{30}{\degree W} & Linear combination of model's EOFs, computed using both initializations as independent predictors (10 EOFs).\\
%
\emph{CCA} & \SIrange{60}{0}{\degree S}; \SIrange{80}{30}{\degree W} & Canonical modes computed using both initializations as independent predictors. (10 predictor EOFs, 4 predictand EOFs, 4 canonical modes) \\
%
\bottomrule
%
\end{tabular}
\end{center}
\end{table*}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% FIGURES
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{figure*}
	\noindent\includegraphics[width=6.5in]{StudyArea}
	\caption{
		Topographical map of the study area.
		Colors indicate $\log_{10}$ of elevation, in \si{\meter}, from the Global Land One-Km Base Elevation project available at \texttt{http://iridl.ldeo.columbia.edu/SOURCES/.NOAA/.NGDC/.GLOBE/.topo/}.
		(L): all of South America.
		The domains of the ``Lower Paraguay River Basin'' and the domain used for weather typing are indicated in black and red, respectively.
		(R): The Paraguay River.
		As for (L), the ``Lower Paraguay River Basin'' is indicated.
		Streamflows shown in \cref{fig:streamflow} were taken from the fours stations indicated.
		The Paraguay River and its tributaries, from the Natural Earth database (\texttt{www.naturalearthdata.com}), are also shown.
		Full station names are: Bah\'ia Negra (Bne); Concepci\'on (Conc); Asunci\'on (Asu); Pilar (Pil).
	}
  \label{fig:study-area}
\end{figure*}

\begin{figure*}
	\noindent\includegraphics[width=6.5in]{Streamflow.pdf}
	\caption{
		River stage (height; in \si{\meter}) for the Paraguay River at four gauges along the Paraguay River.
		The station names are shortened versions of those shown in \cref{fig:study-area}.
		(a) Seasonality (orange) and time series of 2015-16 observations (black) at each stream gauge.
			The seasonality was fit using local polynomial regression as implemented in the \texttt{locfit} package in
			the \textbf{R} statistical programming environment \citep{Loader1999}.
		(b) Time series of daily stage measurements from 1929 to 2016 at each station.
	}
  \label{fig:streamflow}
\end{figure*}

\begin{figure*}	\noindent\includegraphics[width=6.5in]{LaggedRain.pdf}
	\caption{
		Circulation anomalies associated with heavy rainfall (\nth{90} percentile exceedance of area-averaged rainfall in the Lower Paraguay River Basin).
		Lagged composites are shown, by column, for $t = $ \SIlist{-2;-1;0;1}{\day} relative to the date of heavy rainfall.
		Top and middle rows show streamfunction anomalies at \SIlist{850}{\hecto\pascal}, respectively.
		Bottom row shows monthly-mean rainfall anomalies, in units of \si{\milli\meter\per\day}.
	}
  \label{fig:lagged-rain}
\end{figure*}

\begin{figure*}
	\noindent\includegraphics[width=6.5in]{PSI_EOF_Loadings.pdf}
	\caption{
    	Leading EOF loadings of daily NDJF \SI{850}{\hecto\pascal} streamfunction $\psi$ over chosen domain.
	}
  \label{fig:eof-loading}
\end{figure*}

\begin{figure*}
	\noindent\includegraphics[width=6.5in,height=9in,keepaspectratio=true]{WTComposite.pdf}
	\caption{
		Composite anomalies associated with each weather type (WT; column).
		Top row shows streamfunction anomalies at \SIlist{850}{\hecto\pascal}.
		Bottom row shows monthly-mean rainfall anomalies, in units of \si{\milli\meter\per\day}. The relative frequency of occurrence of each weather type (in days) is presented on the top of each column.
	}
	\label{fig:wt-composite}
\end{figure*}

\begin{figure*}
	\noindent\includegraphics[width=6.5in]{RainfallWeatherType}
	\caption{
		Time series of area-averaged rainfall in the Lower Paraguay River Basin (\cref{fig:study-area}) for each day of NDJF  2015-16 (lines and points).
        The weather types corresponding to each day are indicated in both color and in an adjacent text label.
		Dashed lines indicate the climatological 50th, \nth{90}, and 99th percentiles of NDJF area-averaged rain over the Lower Paraguay River basin.
	}
  \label{fig:rain-wt}
\end{figure*}

\begin{figure*}
	\noindent\includegraphics[width=6.5in]{NDJF201516Anomaly.pdf}
	\caption{
    	Monthly composite anomalies observed during NDJF 2015-16.
    	Top and middle rows show streamfunction anomalies at \SIlist{200;850}{\hecto\pascal}, respectively.
      Bottom row shows monthly-mean rainfall anomalies, in units of \si{\milli\meter\per\day}.
	}
  \label{fig:anomalies}
\end{figure*}

\begin{figure*}
	\noindent\includegraphics[width=3.5in]{SeasonalForecast.pdf}
	\caption{
	Seasonal model forecast for probability of DJF \nth{90} percentile exceedance.
	Color indicates the forecast probability of exceeding the \nth{90} percentile of climatological rainfall during DJF 2015-16 -- this is presented as the odds ratio as explained in \cref{eq:odds-ratio}.
	A value greater than 1 indicates that the model forecast greater-than-average odds of rainfall exceeding the \nth{90} percentile.
	Grid cells which observed an exceedance of the \nth{90} percentile of DJF rainfall are outlined in black.
	}
  \label{fig:seas-prob-fcst}
\end{figure*}

\begin{figure*}
	\noindent\includegraphics[width=6.5in]{Chiclet.pdf}
	\caption{
    Chiclet diagram \citep{Carbin:2016fx} of ensemble-mean precipitation anomaly forecasts over the Lower Paraguay River Basin (\cref{fig:study-area}) from ECMWF S2S forecast data, as a function of the forecast target date (horizontal axis) and lead time (vertical axis). Time series of CPC daily mean precipitation over the same area is plotted with $y$-axis inverted.
	}
  \label{fig:chiclet}
\end{figure*}

\begin{figure*}
	\noindent\includegraphics[width=6.4in]{ForecastSkill.pdf}
	\caption{
		MOS-adjusted S2S model forecasts and skill scores for the methods indicated in \cref{tab:mos-methods}.
		Top row shows the heavy rainfall forecast for 1-7 December 2015 as the odds ratio of \cref{eq:odds-ratio} over the target domain.
		A value greater than 1 indicates that the model forecast greater-than-average odds of rainfall exceeding the \nth{90} percentile.
        Second row shows the Ignorance score \cref{eq:ignorance}, with zero indicating a perfect forecast.
		Bottom row shows the 2AFC skill score for each grid cell; a value greater than 50 indicates that the model outperforms climatology.
		For all three rows, the grid cells which experienced a \nth{90} percentile exceedance for 1-7 December 2015 are outlined in black.
	}
  \label{fig:subs-prob-fcst}
\end{figure*}

\begin{figure*}
	\noindent\includegraphics[width=6.4in]{WT_MJO_ENSO.pdf}
	\caption{
		Anomalous probability of occurrence of each weather type concurrent with observance of each MJO phase.
		MJO phases are considered here only when the amplitude $\qty(\sqrt{\text{RMM1}^2 + \text{RMM2}^2})$ is greater than 1.
		When amplitude $< 1$, the phase is defined as neutral (0).
		Plots are shown separately for El Ni\~no, La Ni\~na, and neutral ENSO months, defined by the NINO 3.4 index having a value above 1 or below -1.
		Only values which are significant at $p<0.10$, calculated using a bootstrap with 3000 samples, are shown.
	}
  \label{fig:wt-mjo-enso}
\end{figure*}

\begin{figure*}
	\noindent\includegraphics[width=6.4in]{ChacoNoChacojet.pdf}
	\caption{
		Simple schematics of low-level jet events (red arrows) during austral summer and El Ni\~no years.
		In general, most jet events are of the Chaco type, particularly when SST anomalies in the central southern Atlantic Ocean (a, see green box) are weak.
		When a dipolar SST anomaly occurs in the central southern Atlantic (with the warmer pole equatorward) the meridional temperature gradient and sea-land temperature contrasts establish an anti-cyclonic circulation (dot-dashed line) conducive to increased occurrence of No-Chaco jet events (b).
		Other SST anomaly configurations tend to be present outside the green box (not shown).
		Winds in panels are typical for each case (at 850 hPa).
		Reference wind vector in $m s^{-1}$.
	}
  \label{fig:chaco-nochaco}
\end{figure*}

\begin{figure*}
	\noindent\includegraphics[width=6.4in]{WT4_Monthly_Composite.pdf}
	\caption{
		Composite sea surface temperature (SST) anomalies of the 15 NDJF months with the most occurrences of weather type 4.
		SST data, in \si{\celsius}, come from \citet{Reynolds:2002iy}.
		To match the SST data, only months since NDJF 1981 are considered.
	}
  \label{fig:wt4-months}
\end{figure*}

\end{document}

### 2017-11-30 config/PlotParameters.py

@@ -1,9 +1,0 @@
-import numpy as np
-# Specify some parameters for plotting
-# this will keep the layout and look the same
-
-# Define color maps
-cmap = {'psi': 'PuOr', 'psi_a': 'PuOr', 'rain': 'Greens', 'rain_a': 'BrBG'}
-
-# Some extents for plotting
-extent = {'SAm': [275, 330, 5, -40], 'SH': [-120, 0, 5, -50]}

### 2017-11-30 scripts/access/CPCRainYear.py

@@ -1,56 +1,0 @@
-'''
-Download one year of CPC Rainfall Data from the IRI Data Library.
-We only get rainfall estimate -- neglects information on number of stations used in analysis.
-Please see caveats and documentation on IRI Server at http://iridl.ldeo.columbia.edu/SOURCES/.NOAA/.NCEP/.CPC/.UNIFIED_PRCP/.GAUGE_BASED/.GLOBAL/.v1p0/
-'''
-
-import pandas as pd
-import numpy as np
-import xarray as xr
-import datetime
-import argparse
-
-parser = argparse.ArgumentParser()
-parser.add_argument('--year', nargs=1, type=int, help='The year of data to download')
-parser.add_argument('--outfile', nargs=1, help='The output file to save')
-
-# Work with Times from the IRI DL
-def convert_t_to_time(Tvec):
-    times = np.array([datetime.date(1960, 1, 1) + datetime.timedelta(np.int(ti)) for ti in Tvec])
-    return(np.array(times))
-def convert_time_to_t(date):
-    date_diff = date - datetime.date(1960,1,1)
-    T = date_diff.days
-    return(T)
-
-def main():
-    # parse the arguments
-    args = parser.parse_args()
-    year = args.year[0]
-    outfile = args.outfile[0]
-
-    # Get the URL
-    if year >= 1979 and year <= 2005:
-        url = 'http://iridl.ldeo.columbia.edu/SOURCES/.NOAA/.NCEP/.CPC/.UNIFIED_PRCP/.GAUGE_BASED/.GLOBAL/.v1p0/.RETRO/.rain/dods'
-    elif year >= 2006 and year <= 2016:
-        url = 'http://iridl.ldeo.columbia.edu/SOURCES/.NOAA/.NCEP/.CPC/.UNIFIED_PRCP/.GAUGE_BASED/.GLOBAL/.v1p0/.REALTIME/.rain/dods'
-    else:
-        raise ValueError('You have entered an invalid year. {} is outside range [1979, 2016]'.format(year))
-
-    # get the data
-    Tstart = convert_time_to_t(datetime.date(year, 1, 1))
-    Tend = convert_time_to_t(datetime.date(year, 12, 31))
-    ds = xr.open_dataarray(url, decode_times=False)
-    ds = ds.sel(T = slice(Tstart, Tend))
-    ds.load() # force it to download
-
-    # convert to more standard format
[…]

### 2017-11-30 scripts/access/DailyIndices.py

@@ -1,39 +1,0 @@
-'''
-Read in daily indices (here just MJO) and save to file
-'''
-
-import pandas as pd
-import numpy as np
-import argparse
-
-parser = argparse.ArgumentParser()
-parser.add_argument('--years', type=int, nargs=2, help = 'LONMIN, LATMAX')
-parser.add_argument('--outfile', nargs=1, help='The output file to save')
-
-def main():
-
-    args = parser.parse_args()
-    syear, eyear = np.min(args.years), np.max(args.years)
-    outfile=args.outfile[0]
-
-    raw_url = 'http://www.bom.gov.au/climate/mjo/graphics/rmm.74toRealtime.txt'
-    col_names = ['year', 'month', 'day', 'RMM1', 'RMM2', 'phase', 'amplitude', 'source']
-
-
-    # Now read in the data
-    df = pd.read_table(raw_url, delim_whitespace=True,index_col=None, skiprows=2, names=col_names)
-
-    # Convert the year-month-day structure to a standard time object and select only relevant columns
-    df['time'] = pd.to_datetime(df[['year', 'month', 'day']])
-    df.index = df['time']
-    df = df[['RMM1', 'RMM2', 'phase', 'amplitude']]
-    df = df['{}'.format(syear):'{}'.format(eyear)]
-
-    # Check
-    df.head(10)
-
-    # Save to file
-    df.to_csv(outfile)
-
-if __name__ == '__main__':
-    main()

### 2017-11-30 scripts/access/S2S-Data.py

@@ -1,41 +1,0 @@
-'''
-Get Area-Averaged S2S Forecast Rainfall
-'''
-
-import xarray as xr
-import numpy as np
-import pandas as pd
-import datetime
-import argparse
-
-parser = argparse.ArgumentParser()
-parser.add_argument('--latlim', type=float, nargs=2, help = 'LATMIN, LATMAX')
-parser.add_argument('--lonlim', type=float, nargs=2, help = 'LONMIN, LATMAX')
-parser.add_argument('--outfile', nargs=1, help='The output file to save')
-
-def GetURL(x0, x1, y0, y1):
-    url = 'http://iridl.ldeo.columbia.edu/home/.mbell/.ECMWF/.S2Stest/.S2S/.ECMF_ph2/.forecast/.perturbed/.sfc_precip/.tp/'
-    url += 'X/{}/{}/RANGEEDGES/'.format(x0, x1)
-    url += 'Y/{}/{}/RANGEEDGES/'.format(y0, y1)
-    url += '[X+Y]average/'
-    url += 'L+differences/'
-    url += 'dods'
-    return(url)
-
-def main():
-
-    args = parser.parse_args()
-    latmin, latmax = np.min(args.latlim), np.max(args.latlim)
-    lonmin, lonmax = np.min(args.lonlim), np.max(args.lonlim)
-
-    url = GetURL(x0=lonmin, x1=lonmax, y0=latmin, y1=latmax)
-    outfile = args.outfile[0]
-
-    print('Downloading S2S data from {}'.format(url))
-    ds = xr.open_dataarray(url)
-    ds = ds.sel(S = slice('2015-09-01', '2016-02-29'))
-    ds['L'] = ds['L']
-    ds.to_netcdf(outfile, format="NETCDF4")
-
-if __name__ == '__main__':
-    main()

### 2017-11-30 scripts/derive/Anomalies.py

@@ -1,69 +1,0 @@
-'''
-Subsetting, Anomalies, and Climatologies
-'''
-
-import xarray as xr
-import glob
-import numpy as np
-from paraguayfloodspy.xrutil import *
-import argparse
-
-parser = argparse.ArgumentParser()
-parser.add_argument('--mode', nargs=1, help='Either [rain] or [reanalysis]')
-parser.add_argument('--pattern', nargs=1, help='The pattern of files to read in')
-parser.add_argument('--latlim', type=float, nargs=2, help = 'LATMIN, LATMAX')
-parser.add_argument('--lonlim', type=float, nargs=2, help = 'LONMIN, LATMAX')
-parser.add_argument('--years', type=int, nargs=2, help = 'LONMIN, LATMAX')
-parser.add_argument('--outfile', nargs=1, help='The output file to save')
-
-def main():
-    args = parser.parse_args()
-    latmin, latmax = np.min(args.latlim), np.max(args.latlim)
-    lonmin, lonmax = np.min(args.lonlim), np.max(args.lonlim)
-    syear, eyear = np.min(args.years), np.max(args.years)
-
-    def transfer_fn(ds, mode, lonmin, lonmax, latmin, latmax):
-        '''
-        This function sub-sets data as we read it
-        '''
-        assert lonmax > lonmin, "lonmax must be greater than lonmin"
-        assert latmax > latmin, "latnmax must be greater than latmin"
-        if mode == 'rain':
-            ds = ds.sel(lon = slice(lonmin, lonmax), lat = slice(latmin, latmax))
-        elif mode == 'reanalysis':
-            ds = ds.sel(lon = slice(lonmin, lonmax), lat = slice(latmax, latmin))
-        else:
-            raise ValueError('mode must be rain or reanalysis')
-        return(ds)
-
-    # Read in raw data
-    fnames = glob.glob(args.pattern[0])
-    if len(fnames) < 1:
-        raise ValueError("Pattern {} doesn't match any files".format(args.pattern[0]))
-
-    # subset data when we read it in
-    trans_lam = lambda ds: transfer_fn(ds, mode=args.mode[0], lonmin=lonmin, lonmax=lonmax, latmin=latmin, latmax=latmax)
-    raw = read_netcdfs(fnames, dim='time', transform_func=trans_lam)
-    if args.mode[0] =='rain':
-        raw = raw['rain']
-    elif args.mode[0] == 'reanalysis':
[…]

### 2017-11-30 scripts/derive/AreaAveragedRain.py

@@ -1,34 +1,0 @@
-"""
-GetRioPYRain
-------------------
-
-This script takes the area-averaged rainfall and saves to csv
-
-"""
-
-import xarray as xr
-import numpy as np
-import pandas as pd
-import argparse
-
-parser = argparse.ArgumentParser()
-parser.add_argument('--infile', nargs=1, help='the input file to read from')
-parser.add_argument('--lonlim', nargs=2, type=float, help='the input file to read to')
-parser.add_argument('--latlim', nargs=2, type=float, help='the input file to read to')
-parser.add_argument('--outfile', nargs=1, help='the output file to save to')
-args = parser.parse_args()
-
-def main():
-    infile = args.infile[0]
-    outfile = args.outfile[0]
-    latmin, latmax = np.min(args.latlim), np.max(args.latlim)
-    lonmin, lonmax = np.min(args.lonlim), np.max(args.lonlim)
-    rain = xr.open_dataset(infile).sel(lon = slice(lonmin, lonmax), lat = slice(latmin, latmax))
-    rain_mean = rain.mean(dim=['lon', 'lat'])
-    print(rain_mean)
-
-    # Save to file
-    rain_mean.to_netcdf(outfile)
-
-if __name__ == '__main__':
-    main()

### 2017-11-30 src/WeatherTypes.py

@@ -1,53 +1,0 @@
-'''
-Get the weather types using the implementation in paraguayfloodspy
-'''
-
-import xarray as xr
-import pandas as pd
-import numpy as np
-from paraguayfloodspy.weather_type import *
-import argparse
-
-parser = argparse.ArgumentParser()
-parser.add_argument('--psi850', nargs=1, help='The 850 hPa streamfunction to read in')
-parser.add_argument('--latlim', type=float, nargs=2, help = 'LATMIN, LATMAX')
-parser.add_argument('--lonlim', type=float, nargs=2, help = 'LONMIN, LATMAX')
-parser.add_argument('--ncluster', type=int, nargs=1, help = 'how many clusters to have')
-parser.add_argument('--pcscaling', type=int, nargs=1, help = '0 or 1')
-parser.add_argument('--wtprop', type=float, nargs=1, help = 'what fraction of variance to retain')
-parser.add_argument('--nsim', type=int, nargs=1, help = 'how many simulatinos to run')
-parser.add_argument('--outfile', nargs=1, help='The output file to save')
-
-def main():
-    args = parser.parse_args()
-    latmin, latmax = np.min(args.latlim), np.max(args.latlim)
-    lonmin, lonmax = np.min(args.lonlim), np.max(args.lonlim)
-    outfile = args.outfile[0]
-    n_clusters = args.ncluster[0]
-    wt_prop = args.wtprop[0]
-    nsim = args.nsim[0]
-    pcscaling = args.pcscaling[0]
-
-    psi = xr.open_dataset(args.psi850[0])
-    psi = psi.sel(lon = slice(lonmin, lonmax), lat = slice(latmax, latmin))
-
-    # TO REPRODUCE OUR RESULTS
-    np.random.seed(22590)
-
-    # All the computation is in the paraguayfloodspy.weather_type module
-    best_centroid, best_ts, classifiability = XrEofCluster(
-        psi,
-        n_clusters=n_clusters,
-        prop=wt_prop, # What proportion of variance should be retained?
-        nsim=nsim, # How many random initializations to compute?
-        pcscaling=pcscaling,
-        variable='anomaly',
-        verbose = True # get useful info from the algorithm
-    )
-    print("Classifiability Index: {}".format(classifiability))
-
-    df = pd.DataFrame({'wtype': pd.Series(np.int_(best_ts), index=psi.time)})
-    df.to_csv(outfile)
-
-if __name__ == '__main__':
-    main()

### 2017-11-30 src/pyfloods/anomalies.py

@@ -1,0 +1,79 @@
+import xarray as xr
+import numpy as np
+import pandas as pd
+import os
+from typing import Dict,Any,List,Tuple
+from datetime import datetime
+
+from pyfloods.dataset import SingleFileDataSet,CPC
+from pyfloods.region import Region
+
+class Anomaly(SingleFileDataSet):
+    def __init__(self, superset: Any, fname: str, region: Region, sdate: datetime, edate: datetime, months: List[int]=[11, 12, 1, 2], verbose: bool=False) -> None:
+        self.superset = superset
+        params: Dict = superset.assets[0].params.copy()
+        params.pop('year', None); params.pop('level', None)
+        params.pop('coord_system', None); params.pop('var', None)
+        params.update({
+            'sdate': sdate, 'edate': edate,
+            'lonmin': region.lonmin.astype('float'),
+            'lonmax': region.lonmax.astype('float'),
+            'latmin': region.latmin.astype('float'),
+            'latmax': region.latmax.astype('float')
+        })
+        self.months = months
+        super().__init__(fname=fname, params=params, verbose=verbose)
+
+    def download_data(self) -> xr.DataArray:
+        raw = self.superset.get_data().to_array()
+        raw = raw.sortby('lat').sortby('lon')
+        raw = raw.sel(
+            lon = slice(self.params['lonmin'], self.params['lonmax']),
+            lat = slice(self.params['latmin'], self.params['latmax'])
+        )
+        year = raw['time.year']
+        month = raw['time.month']
+        year_adj = year
+        year_adj[np.in1d(month, [11, 12])] += 1
+        raw.coords['year_adj'] = year_adj
+        raw = raw.sel(time = slice(self.params.get('sdate'), self.params.get('edate')))
+        raw = raw.sel(time = np.in1d(raw['time.month'], self.months))
+        anomaly = raw - raw.mean(dim='time')
+        ds = xr.Dataset({'raw': raw, 'anomaly': anomaly})
+        ds.to_netcdf(self.fname, format='NETCDF4')
+        return ds
+
+    def read_data(self) -> Tuple[xr.DataArray, Dict]:
+        ds = xr.open_dataset(self.fname)
+        pars = {
+            'edate': pd.to_datetime(ds['time'].max().values),
+            'lonmin': ds['lon'].min().values.tolist(),
+            'lonmax': ds['lon'].max().values.tolist(),
[…]

### 2017-11-30 src/pyfloods/dataset/__init__.py

@@ -1,0 +1,59 @@
+"""
+This submodule defines the abstraction of a dataset and can work generically
+for any dataset whether it has one or more files, regardless of the
+type of file that it describes.
+In general, the dataset has the following elements:
+* a set of parameters that determine what the dataset contains
+* a filename or set of filenames that describe where the data is stored
+* a way to access the data from some remote source and save it to file
+* a way to read the data, including its parameters, in from files
+* a way to check whether the data that was read from file has the desired parameters
+"""
+
+from typing import Tuple,Dict,Any
+import os
+
+class SingleFileDataSet():
+    def __init__(self, fname: str, verbose: bool = False,
+                params: Dict[str, Any] = {}) -> None:
+        self.params = params
+        self.fname = fname
+        self.verbose = verbose
+
+    def download_data(self):
+        raise NotImplementedError
+
+    def read_data(self) -> Tuple[Any, Dict]:
+        raise NotImplementedError
+
+    def get_data(self):
+        """Try to read in the data, and if it has changed or if it cannot be
+        read in, then download the data.
+        """
+        need_download: bool = False
+        try:
+            data,data_params = self.read_data()
+            if self.params != data_params:
+                os.remove(self.fname)
+                need_download = 1
+                if self.verbose:
+                    print('Data was read from file but had different parameters')
+            if self.verbose:
+                print('Successfully read data from file')
+        except:
+            if self.verbose:
+                print('Unable to read data from file')
+            need_download = True
+        if need_download:
+            if self.verbose:
+                print('Downloading data from source')
+            data = self.download_data()
+        return(data)
+
+# Make it so we don't have to remember the file names when we do imports later
+from .cpc_rain import CPC
+from .elevation import NOAANGDC
+from .mjo_bom import MJOBOMD
+from .nino_34 import KaplanNINO34
+from .reanalysisv2 import ReanalysisV2
+from .s2s_area_avg import S2SAreaAvg

### 2017-11-30 src/pyfloods/dataset/cpc_rain.py

@@ -1,0 +1,79 @@
+"""
+This implements a single variable and single level of a reanalysis
+V2 variable
+https://www.esrl.noaa.gov/psd/thredds/dodsC/Datasets/ncep.reanalysis2
+"""
+import xarray as xr
+import numpy as np
+import datetime
+from typing import Dict,Tuple,List
+
+from pyfloods.dataset import SingleFileDataSet
+
+class CPCYear(SingleFileDataSet):
+    """This implements just a single year!
+    """
+
+    def __init__(self, fname:str, verbose: bool = False, params: Dict ={}) -> None:
+        self.year = params.get('year')
+        super().__init__(fname=fname, params=params, verbose=verbose)
+
+    def download_data(self) -> xr.DataArray:
+        def convert_t_to_time(Tvec) -> np.ndarray:
+            times = np.array([datetime.date(1960, 1, 1) + datetime.timedelta(np.int(ti)) for ti in Tvec])
+            return(np.array(times))
+
+        def convert_time_to_t(date)  -> np.ndarray:
+            date_diff = date - datetime.date(1960,1,1)
+            T = date_diff.days
+            return(T)
+
+        if self.year >= 1979 and self.year <= 2005:
+            url = 'http://iridl.ldeo.columbia.edu/SOURCES/.NOAA/.NCEP/.CPC/.UNIFIED_PRCP/.GAUGE_BASED/.GLOBAL/.v1p0/.RETRO/.rain/dods'
+        elif self.year >= 2006 and self.year <= 2016:
+            url = 'http://iridl.ldeo.columbia.edu/SOURCES/.NOAA/.NCEP/.CPC/.UNIFIED_PRCP/.GAUGE_BASED/.GLOBAL/.v1p0/.REALTIME/.rain/dods'
+        else:
+            raise ValueError('You have entered an invalid year. {} is outside range [1979, 2016]'.format(self.year))
+
+        Tstart = convert_time_to_t(datetime.date(self.year, 1, 1))
+        Tend = convert_time_to_t(datetime.date(self.year, 12, 31))
+
+        ds = xr.open_dataarray(url, decode_times=False)
+        ds = ds.sel(T = slice(Tstart, Tend)).load()
+        ds = ds.rename({'X': 'lon', 'Y': 'lat', 'T': 'time'})
+        ds['time'] = convert_t_to_time(ds['time'])
+        ds['time'] = ds['time'].astype('datetime64')
[…]

### 2017-11-30 src/pyfloods/dataset/elevation.py

@@ -1,0 +1,30 @@
+"""
+A SingleFileDataSet object for the NOAA NGDC GLOBE topographic dataset
+stored on the iridl data library:
+http://iridl.ldeo.columbia.edu/SOURCES/.NOAA/.NGDC/.GLOBE/.topo/
+"""
+import xarray as xr
+from datetime import datetime
+from typing import Dict,Tuple
+
+from pyfloods.dataset import SingleFileDataSet
+
+class NOAANGDC(SingleFileDataSet):
+    def __init__(self, fname:str, verbose: bool = False, params: Dict = {}) -> None:
+        """The params must be a dict of datetime.datetime objects
+        with names 'sdate' and 'edate'.
+        """
+        super().__init__(fname=fname, params=params, verbose=verbose)
+
+    def download_data(self) -> xr.DataArray:
+        url = 'http://iridl.ldeo.columbia.edu/SOURCES/.NOAA/.NGDC/.GLOBE/.topo/'
+        url += 'X/-180/0.1/180/GRID/Y/-90/0.1/90/GRID/'
+        url += 'dods'
+        ds = xr.open_dataarray(url)
+        ds.to_netcdf(self.fname, format='NETCDF4', mode='w')
+        return ds
+
+    def read_data(self) -> Tuple[xr.DataArray, Dict]:
+        ds = xr.open_dataarray(self.fname)
+        pars: Dict[str, datetime] = {}
+        return ds, pars

### 2017-11-30 src/pyfloods/dataset/mjo_bom.py

@@ -1,0 +1,37 @@
+"""
+A SingleFileDataSet object for the MJO indices from the Australian BOM from
+http://www.bom.gov.au/climate/mjo/graphics/rmm.74toRealtime.txt
+"""
+import pandas as pd
+import numpy as np
+from datetime import datetime
+from typing import Dict,Tuple
+
+from pyfloods.dataset import SingleFileDataSet
+
+class MJOBOMD(SingleFileDataSet):
+    def __init__(self, fname:str, verbose: bool = False, params = Dict[str, datetime]) -> None:
+        """The params must be a dict of datetime.datetime objects
+        with names 'sdate' and 'edate'.
+        """
+        super().__init__(fname=fname, params=params, verbose=verbose)
+        self.sdate = self.params.get('sdate')
+        self.edate = self.params.get('edate')
+
+    def download_data(self) -> pd.DataFrame:
+        raw_url = 'http://www.bom.gov.au/climate/mjo/graphics/rmm.74toRealtime.txt'
+        col_names = ['year', 'month', 'day', 'RMM1', 'RMM2', 'phase', 'amplitude', 'source']
+        df = pd.read_table(raw_url, delim_whitespace=True,index_col=None, skiprows=2, names=col_names)
+        df['time'] = pd.to_datetime(df[['year', 'month', 'day']])
+        df = df.set_index('time')
+        df = df[['RMM1', 'RMM2', 'phase', 'amplitude']]
+        df = df.iloc[np.where(np.logical_and(df.index >= self.sdate, df.index <= self.edate))]
+        df.to_csv(self.fname)
+        return(df)
+
+    def read_data(self) -> Tuple[pd.DataFrame, Dict]:
+        df = pd.read_csv(self.fname, index_col='time', parse_dates=True)
+        sdate = pd.to_datetime(df.index.min())
+        edate = pd.to_datetime(df.index.max())
+        pars: Dict[str, datetime] = {'sdate': sdate, 'edate': edate}
+        return df, pars

### 2017-11-30 src/pyfloods/dataset/nino_34.py

@@ -1,0 +1,42 @@
+"""
+A SingleFileDataSet object for the Kaplan NINO3.4 indices stored on the iridl
+data library:
+http://iridl.ldeo.columbia.edu/SOURCES/.Indices/.nino/.EXTENDED/.NINO34/
+"""
+import pandas as pd
+import numpy as np
+from datetime import datetime
+from typing import Dict,Tuple
+
+from pyfloods.dataset import SingleFileDataSet
+
+class KaplanNINO34(SingleFileDataSet):
+    def __init__(self, fname:str, verbose: bool = False, params = Dict[str, datetime]) -> None:
+        """The params must be a dict of datetime.datetime objects
+        with names 'sdate' and 'edate'.
+        """
+        super().__init__(fname=fname, params=params, verbose=verbose)
+        self.sdate = self.params.get('sdate')
+        self.edate = self.params.get('edate')
+
+    def download_data(self) -> pd.DataFrame:
+        raw_url = 'http://iridl.ldeo.columbia.edu/SOURCES/.Indices/.nino/.EXTENDED/.NINO34/gridtable.tsv'
+        col_name = 'nino_34'
+        df = pd.read_table(raw_url, delim_whitespace=True,index_col=None, skiprows=2, names=['time', '{}'.format(col_name)])
+        df['time'] = np.int_(np.floor(df['time']))
+        df['year'] = 1960 + df['time'] // 12
+        df['month'] = 1 + df['time'] % 12
+        df['day'] = 1
+        df['time'] = pd.to_datetime(df[['year', 'month', 'day']])
+        df.index = df['time']
+        df = df[['{}'.format(col_name)]]
+        df = df.iloc[np.where(np.logical_and(df.index >= self.sdate, df.index <= self.edate))]
+        df.to_csv(self.fname)
+        return(df)
+
+    def read_data(self) -> Tuple[pd.DataFrame, Dict]:
+        df = pd.read_csv(self.fname, index_col='time', parse_dates=True)
+        sdate = pd.to_datetime(df.index.min())
+        edate = pd.to_datetime(df.index.max())
+        pars: Dict[str, datetime] = {'sdate': sdate, 'edate': edate}
+        return df, pars

### 2017-11-30 src/pyfloods/dataset/s2s_area_avg.py

@@ -1,0 +1,64 @@
+"""
+A SingleFileDataSet object for the ECMWF S2S data
+Specifically it returns the area-averaged rainfall over some box
+http://iridl.ldeo.columbia.edu/home/.mbell/.ECMWF/.S2Stest/.S2S/.ECMF_ph2/.forecast/.perturbed/.sfc_precip/.tp/
+"""
+import xarray as xr
+import numpy as np
+import pandas as pd
+from datetime import datetime,timedelta
+from typing import Dict,Tuple
+
+from pyfloods.dataset import SingleFileDataSet
+
+class S2SAreaAvg(SingleFileDataSet):
+    def __init__(self, fname:str, verbose: bool = False, params: Dict = {}) -> None:
+
+        super().__init__(fname=fname, params=params, verbose=verbose)
+
+        self.sdate: datetime = self.params.get('sdate', datetime(2015, 9, 1))
+        self.edate: datetime = self.params.get('edate', datetime(2016, 2, 29))
+        self.lonmin: float = self.params.get('lonmin', -65)
+        self.lonmax: float = self.params.get('lonmax', -45)
+        self.latmin: float = self.params.get('latmin', -30)
+        self.latmax: float = self.params.get('latmax', -15)
+        self.params: Dict = {
+            'sdate': self.sdate, 'edate': self.edate,
+            'lonmin': self.lonmin, 'lonmax': self.lonmax,
+            'latmin': self.latmin, 'latmax': self.latmax
+        }
+
+    def download_data(self) -> xr.DataArray:
+        url = 'http://iridl.ldeo.columbia.edu/home/.mbell/.ECMWF/.S2Stest/.S2S/.ECMF_ph2/.forecast/.perturbed/.sfc_precip/.tp/'
+        url += 'X/{}/{}/RANGEEDGES/'.format(self.lonmin, self.lonmax)
+        url += 'Y/{}/{}/RANGEEDGES/'.format(self.latmin, self.latmax)
+        url += '[X+Y]average/'
+        url += 'L+differences/'
+        url += 'dods'
+        ds = xr.open_dataarray(url)
+        ds = ds.sel(S = slice(self.sdate, self.edate))
+        ds['L'] = ds['L']
+        ds.attrs['lonmin'] = self.lonmin
+        ds.attrs['lonmax'] = self.lonmax
+        ds.attrs['latmin'] = self.latmin
+        ds.attrs['latmax'] = self.latmax
+        ds.to_netcdf(self.fname, format='NETCDF4', mode='w')
[…]

### 2017-11-30 src/pyfloods/raw_data.py

@@ -1,0 +1,52 @@
+"""
+Here we will define some datasets that use the dataset package
+but are specifically defined for this project.
+"""
+
+import xarray as xr
+import numpy as np
+from datetime import datetime
+import os
+
+from pyfloods.dataset import MJOBOMD, KaplanNINO34, NOAANGDC, S2SAreaAvg, ReanalysisV2, CPC
+from pyfloods import paths
+from pyfloods import region
+
+verbose = False # universal
+
+mjo = MJOBOMD(
+    fname = os.path.abspath(os.path.join(paths.data_external, 'mjo.csv')),
+    params = {'sdate': region.sdate, 'edate': region.edate},
+    verbose = verbose
+)
+enso = KaplanNINO34(
+    fname = os.path.abspath(os.path.join(paths.data_external, 'nino_34.csv')),
+    params = {'sdate': region.sdate, 'edate': region.edate},
+    verbose = verbose
+)
+topo = NOAANGDC(
+    fname = os.path.abspath(os.path.join(paths.data_external, 'elevation_data.nc')),
+    verbose = verbose
+)
+s2s = S2SAreaAvg(
+    fname = os.path.abspath(os.path.join(paths.data_external, 's2s_area_avg.nc')),
+    verbose = verbose
+)
+years = np.arange(region.syear, region.eyear+1)
+u850 = ReanalysisV2(
+    fnames = [os.path.join(paths.data_external, 'reanalysisv2_uwnd_850_{}.nc'.format(y)) for y in years],
+    years = years,
+    verbose = verbose,
+    params={'coord_system': 'pressure', 'level': 850, 'var': 'uwnd'}
+)
+v850 = ReanalysisV2(
+    fnames = [os.path.join(paths.data_external, 'reanalysisv2_vwnd_850_{}.nc'.format(y)) for y in years],
+    years = years,
+    verbose = verbose,
+    params={'coord_system': 'pressure', 'level': 850, 'var': 'vwnd'}
+)
+cpc = CPC(
+    [os.path.join(paths.data_external, 'cpc_rain_{}.nc'.format(y)) for y in years],
+    years = years,
+    verbose = verbose, params={}
+)

### 2017-11-30 src/pyfloods/streamfunction.py

@@ -1,0 +1,71 @@
+import xarray as xr
+import numpy as np
+from typing import Dict,Tuple,List
+from windspharm.xarray import VectorWind
+import os
+
+from pyfloods.dataset.reanalysisv2 import RV2Year
+from pyfloods.dataset import SingleFileDataSet,ReanalysisV2
+from pyfloods import paths
+from pyfloods import raw_data
+from pyfloods import region
+
+class PsiYear(SingleFileDataSet):
+    """Turn a single year of uwind and vwind into streamfunction
+    """
+
+    def __init__(self, fname: str, uwnd: RV2Year, vwnd: RV2Year, verbose:bool = False) -> None:
+        params: Dict = uwnd.params.copy()
+        params.update({'var': 'psi'})
+        self.params = params
+        self.coord_system = params.get('coord_system')
+        self.var = params.get('var')
+        self.level = params.get('level')
+        self.year = params.get('year')
+        self.fname = fname
+        self.uwnd = uwnd
+        self.vwnd = vwnd
+        super().__init__(fname=self.fname, params=params, verbose=verbose)
+
+    def download_data(self) -> xr.DataArray:
+        U = self.uwnd.get_data()
+        V = self.vwnd.get_data()
+        W = VectorWind(U, V)
+        streamfunc = W.streamfunction()
+        streamfunc.to_netcdf(self.fname, format='NETCDF4')
+        streamfunc.attrs['coord_system'] = self.params.get('coord_system')
+        streamfunc.attrs['level'] = self.params.get('level')
+        streamfunc.attrs['year'] = self.params.get('year')
+        streamfunc.attrs['var'] = self.params.get('var')
+        streamfunc.to_netcdf(self.fname, format='NETCDF4')
+        return streamfunc
+
+    def read_data(self) -> Tuple[xr.DataArray, Dict]:
+        ds = xr.open_dataarray(self.fname)
+        pars = {
+            'coord_system': ds.attrs.get('coord_system'),
+            'var': ds.attrs.get('var'),
+            'level': ds.attrs.get('level'),
+            'year': ds.attrs.get('year')
+        }
+        return ds, pars
+
+class PsiMulti():
+    def __init__(self, uwnd: ReanalysisV2, vwnd: ReanalysisV2, fnames: List[str], years: List[int], verbose: bool = False) -> None:
+        self.assets: List[PsiYear] = []
[…]

### 2017-11-30 src/pyfloods/weather_type.py

@@ -1,0 +1,79 @@
+import xarray as xr # for reading in netcdf files
+import numpy as np # for calculations
+import sklearn.cluster as cl # k-means implemetation
+from eofs.xarray import Eof # EOF implementation that recognizes xarray
+import pandas as pd # time series
+
+def ReSort(x):
+        '''
+        A helper function to re-sort the cluster labels.
+        Takes in x, a vector of cluster labels, and re-orders them so that
+        the lowest number is the most common, and the highest number
+        the least common
+        '''
+        import numpy as np
+        x = np.int_(x)
+        x_from = np.unique(x).argsort()
+        counts = np.array([np.sum(x == xi) for xi in x_from])
+        orders = (-counts).argsort().argsort()
+        return(orders[x_from[x]])
+
+def Classifiability(P, Q):
+    '''
+    The Auto-Correlation function
+    '''
+    from scipy.stats import pearsonr as correl
+    k = P.shape[0]
+    assert Q.shape[0] is k, "P and Q must be same k"
+    Aij = np.ones([k, k])
+    for i in range(k):
+        for j in range(k):
+            Aij[i,j], _ = correl(P[i,:], Q[j,:])
+    Aprime = Aij.max(axis=0)
+    classifiability = Aprime.min()
+    return(classifiability)
+
+def XrEofCluster(ds, n_clusters=5, prop=0.90, nsim=100, verbose = True, pcscaling=0):
+    if verbose:
+        print('xarray-based classifiability for {} clusters'.format(n_clusters))
+        print('Performing EOF decomposition of data for dimension reduction.')
+
+    # EOF Decomposition
+    solver = Eof(ds, center=True)
+    var_frac = solver.varianceFraction()
+    cumvar = np.cumsum(var_frac.values)
+    eofs_keep = np.where(cumvar >= prop)[0].min() + 1
+    pc = solver.pcs(npcs=eofs_keep, pcscaling=pcscaling) # time series of PC
+    if verbose:
+        print("Number of EOFs retained is {}".format(eofs_keep))
+        print('Carrying out {} k-means simulations...'.format(nsim))
+
+    centroids = np.ones([nsim, n_clusters, pc.shape[1]])
+    clusters = np.ones([nsim, pc.shape[0]])
+    for i in range(0, nsim):
+        centroids[i,:,:], clusters[i, :], _ = cl.k_means(pc, n_clusters=n_clusters, n_init=1)
[…]

### 2017-11-30 src/weather_type.py

@@ -1,76 +1,0 @@
-import xarray as xr # for reading in netcdf files
-import numpy as np # for calculations
-import sklearn.cluster as cl # k-means implemetation
-from eofs.xarray import Eof # EOF implementation that recognizes xarray
-import pandas as pd # time series
-
-def ReSort(x):
-        '''
-        A helper function to re-sort the cluster labels.
-        Takes in x, a vector of cluster labels, and re-orders them so that
-        the lowest number is the most common, and the highest number
-        the least common
-        '''
-        import numpy as np
-        x = np.int_(x)
-        x_from = np.unique(x).argsort()
-        counts = np.array([np.sum(x == xi) for xi in x_from])
-        orders = (-counts).argsort().argsort()
-        return(orders[x_from[x]])
-
-def Classifiability(P, Q):
-    '''
-    The Auto-Correlation function -- probably needs to be fixed
-    '''
-    from scipy.stats import pearsonr as correl
-    k = P.shape[0]
-    assert Q.shape[0] is k, "P and Q must be same k"
-    Aij = np.ones([k, k])
-    for i in range(k):
-        for j in range(k):
-            Aij[i,j], _ = correl(P[i,:], Q[j,:])
-    Aprime = Aij.max(axis=0)
-    classifiability = Aprime.min()
-    return(classifiability)
-
-def XrEofCluster(ds, n_clusters=5, prop=0.90, nsim=100, variable='hgt', verbose = True, pcscaling=0):
-    '''
-    A function to translate an xarray dataset into EOF space, then carry out k-means clustering
-    n_clusters: how many clusters should be generated?
-    prop: what proportion of the variance should be captured by the EOFs (how many EOFs to use)
-    nsim: how many random initializations to rum
-    variable: the name of the xarray variable in the dataset
-    '''
-    if verbose: print('xarray-based classifiability for {} clusters'.format(n_clusters))
-    if verbose: print('Performing EOF decomposition of data for dimension reduction...')
-    solver = Eof(ds[variable], center=True)
-    var_frac = solver.varianceFraction()
-    cumvar = np.cumsum(var_frac.values)
-    eofs_keep = np.where(cumvar >= prop)[0].min() + 1
-    print("Number of EOFs retained is {}".format(eofs_keep))
-    pc = solver.pcs(npcs=eofs_keep, pcscaling=pcscaling) # time series of PC
-
[…]

### 2017-12-01 reports/writeup/paper.tex
james edits

@@
 \newcolumntype{L}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
-\graphicspath{{../figs/}}
+\graphicspath{{../../figures/}{./}}
 
@@
 \begin{figure*}
-	\noindent\includegraphics[width=6.5in]{StudyArea}
+	\noindent\includegraphics[width=6.5in]{study-area.jpg}
 	\caption{
@@
 
-\begin{figure*}	\noindent\includegraphics[width=6.5in]{LaggedRain.pdf}
+\begin{figure*}	\noindent\includegraphics[width=6.5in]{lagged-rainfall.pdf}
 	\caption{
@@
 \begin{figure*}
-	\noindent\includegraphics[width=6.5in]{PSI_EOF_Loadings.pdf}
+	\noindent\includegraphics[width=6.5in]{eof-loadings.pdf}
 	\caption{
@@
 \begin{figure*}
-	\noindent\includegraphics[width=6.5in,height=9in,keepaspectratio=true]{WTComposite.pdf}
+	\noindent\includegraphics[width=6.5in,height=9in,keepaspectratio=true]{weather-type-composite.pdf}
 	\caption{
@@
 \begin{figure*}
-	\noindent\includegraphics[width=6.5in]{RainfallWeatherType}
+	\noindent\includegraphics[width=6.5in]{wt-rain-time-series.pdf}
 	\caption{
@@
 \begin{figure*}
-	\noindent\includegraphics[width=6.5in]{NDJF201516Anomaly.pdf}
+	\noindent\includegraphics[width=6.5in]{circulation-NDJF-1516-anomaly.pdf}
 	\caption{
@@
 \begin{figure*}
-	\noindent\includegraphics[width=3.5in]{SeasonalForecast.pdf}
+	\noindent\includegraphics[width=3.5in]{seasonal-forecast.pdf}
 	\caption{
@@
 \begin{figure*}
-	\noindent\includegraphics[width=6.5in]{Chiclet.pdf}
+	\noindent\includegraphics[width=6.5in]{chiclet-s2s-area-averaged.pdf}
 	\caption{
@@
 \begin{figure*}
-	\noindent\includegraphics[width=6.4in]{ForecastSkill.pdf}
+	\noindent\includegraphics[width=6.4in]{s2s-forecast-mos.pdf}
 	\caption{
@@
 \begin{figure*}
-	\noindent\includegraphics[width=6.4in]{WT_MJO_ENSO.pdf}
+	\noindent\includegraphics[width=6.4in]{wt-mjo-enso.pdf}
 	\caption{
@@
 \begin{figure*}
-	\noindent\includegraphics[width=6.4in]{WT4_Monthly_Composite.pdf}
+	\noindent\includegraphics[width=6.4in]{ssta-composite-wt4.pdf}
 	\caption{

### 2017-12-01 reports/poster-agu-2017/blocks/at-a-glance.tex
poster halfway done

\begin{block}{At a Glance}
  Beginning in Nov. 2015, repeated intense rainfall events associated with mesoscale convective activity caused severe flooding along the Paraguay-Paran\'{a} river system (\cref{fig:study-area}, red box), \emph{displacing over 170000 people}.
  We use a weather typing approch within a diagnostic approach to show that:
  \begin{itemize}
    \item Moisture and energy advection via the South American Low-Level Jet \cite{Marengo:2012cm}, particularly during ``No-Chaco'' events \cite{Vera:2006ib} favored mesoscale convective activity
    \item Strong El Ni\~{n}o favored strong SALLJ but MJO and Atlantic dipole pattern contributed to jet exit region (thus No-Chaco SALLJ events)
    \item Raw S2S model forecasts of rainfall had limited skill beyond 10-15 days, but statistical methods -- particularly the PCR and CCA methods that correct both spatial patterns and magnitudes -- substantially improve forecast skill
  \end{itemize}
  \begin{mdframed}
  \begin{figure}
  	\noindent\includegraphics[width=0.45\textwidth]{asuncion-inundaciones.jpg}~
    \noindent\includegraphics[width=0.45\textwidth]{rio-py-banados.jpg}
  	\caption{
  		Asuncion, Paraguay, 2015-16
  	}
    \label{fig:floods}
  \end{figure}
  \end{mdframed}
\end{block}

### 2017-12-01 reports/poster-agu-2017/blocks/obs-1516.tex
poster halfway done

\begin{block}{Observed Circulation Anomalies}
  \begin{mdframed}
  \begin{figure}
    \caption{
  		Daily rainfall averaged over Lower Paraguay River Basin (\cref{fig:study-area}) and observed weather type for each day in NDJF 2015-16.
      Blue lines indicate the climatological 50th, \nth{90}, and 99th percentiles of NDJF area-averaged rain.
      \label{fig:rain-wt}
  	}
  	\noindent\includegraphics[width=0.925\textwidth]{wt-rain-time-series.pdf}
  \end{figure}
  \end{mdframed}

  During NDJF 2015-16, the most heavy rainfall occurred during weather types 1 and 4 (\cref{fig:rain-wt}).
  Monthly-scale circulation anomalies (\cref{fig:anomalies}) show a weak anticyclonic circulation that set up over central Brazil during November 2015 and strengthened into the following month.
  In January 2016 it weakened before returning in February 2016.
  The observed rainfall and circulation anomalies are consistent with the aggregation of the observed weather types shown in \cref{fig:rain-wt} and discussed above.
  Weather types 1 and 4  occurred more frequency than their climatology.

  \begin{mdframed}
  \begin{figure}
  	\noindent\includegraphics[width=0.925\textwidth]{circulation-NDJF-1516-anomaly-alt.pdf}
  	\caption{
      	Monthly composite anomalies observed during NDJF 2015-16.
        Colors: anomalies of rainfall associated with each weather type [\si{\milli\meter\per\day}].
        Contours: anomalies of \SI{850}{\hecto\pascal} [contour interval \SI{1e4}{\meter\squared\per\second}]
        \label{fig:anomalies}
  	}
  \end{figure}
  \end{mdframed}
\end{block}

### 2017-12-01 reports/poster-agu-2017/blocks/obs-weather-types.tex
poster halfway done

\begin{block}{Observations and Weather Types}
  Observations come from:
  \begin{itemize}
    \item Rainfall: CPC Global Unified \cite{xie2010cpc}
    \item Atmosphere: NCAR-NCEP Reanalysis II \cite{Kanamitsu:2002kk}
  \end{itemize}
  We use weather typing \cite{Munoz2015} to identify daily circulation patterns:
  \begin{enumerate}
    \item Calculate streamfunction $\Psi$ from meridional and zonal wind \cite{Dawson:2016ge}
    \item Project \SI{850}{\hecto\pascal} streamfunction onto leading 4 EOFs
    \item $K$-means clustering using classifiability index \cite{Michelangeli1995} to generate single weather type for each day
  \end{enumerate}
  Weather typing simplifies dynamics of daily rainfall but \emph{facilitates analysis of sequences of daily weather patterns}.
  They are associated with patterns that have been well described in the literature; particularly relevant are:
  \begin{itemize}
    \item WT1 represents ``Chaco'' jet event \cite{Salio:2002ev}
    \item WT4 represnts ``No-Chaco'' jet events \cite{Vera:2006ib}
  \end{itemize}
  \begin{mdframed}
  \begin{figure}
  	\noindent\includegraphics[width=0.925\textwidth]{weather-type-composite-alt.pdf}
  	\caption{
  		Colors: anomalies of rainfall associated with each weather type [\si{\milli\meter\per\day}].
      Contours: anomalies of \SI{850}{\hecto\pascal} [contour interval \SI{1e4}{\meter\squared\per\second}]
  	}
    \label{fig:weather-types}
  \end{figure}
  \end{mdframed}
% \item S2S forecasts issued by ECMWF \cite{Vitart2016}
\end{block}

### 2017-12-01 reports/poster-agu-2017/blocks/study-area.tex
poster halfway done

\begin{block}{Lower Paraguay River Basin}
  Flat topography limits the river's ability to carry the summer runoff, causing seasonal inundation of the Pantanal and distributing the river discharge in time \cite{Bravo:2011et,Barros:2004bn}.
  The river is thus far more sensitive to persistent rainfall than individual storm events.
  During NDJF 2015-16, river stage [height] throughout the Lower Paraguay River Basin reached nearly three times climatological levels (not shown).
  \begin{mdframed}
  \begin{figure}
  	\noindent\includegraphics[width=0.925\textwidth]{study-area.jpg}
  	\caption{
  		Topographical map of the study area.
  	}
    \label{fig:study-area}
  \end{figure}
  \end{mdframed}
\end{block}

### 2017-12-01 reports/poster-agu-2017/blocks/thanks.tex
poster halfway done

@@
 \begin{block}{Acknowledgements}
-  This research was supported by funding from NSF Sustainability Research Networks (award 1444745): ``Integrated Urban Infrastructure Solutions for Environmentally Sustainable, Healthy, and Livable Cities'' and  SERDP grant \#2516: ``Climate Informed Estimation of Hydrologic Extremes for Robust Adaptation to Non-Stationary Climate''.
-  The authors thank Yochanan Kushnir and James Booth for conversations regarding this work, and Casey Brown, Melissa Bukovsky, Rachel McCrary, Seth McGinnis, Linea Mearns, and Katherine Schlef for collaboration and advice.
+  The authors thank David Farnham, Upmanu Lall, and Andrew Robertson for insightful conversations and guidance.
+  JDG thanks the NSF GRFP program for support (grant DGE 16-44869).
+  AGM was supported by the Atmospheric and Oceanic Sciences (AOS) Program at Princeton University.
 \end{block}

### 2017-12-01 src/21-circulation-NDJF-1516.py

@@ -61,3 +61,41 @@ cbar1.set_label('Precip. Anomaly [mm/d]', rotation=270)
 cbar1.ax.get_yaxis().labelpad = 20
 
 fig.savefig(os.path.join(paths.figures, 'circulation-NDJF-1516-anomaly.pdf'), bbox_inches='tight')
+
+# MAKE THE PLOT WITH ALTERNATIVE SETUP
+fig, axes = plt.subplots(nrows=2, ncols=2, subplot_kw={'projection': map_proj}, figsize=(11, 6))
+for i,(month,year) in enumerate(zip(months_plot, years_plot)):
+    def selector(ds):
+        ds = ds.sel(time = ds['time.month']==month)
+        ds = ds.sel(time = ds['time.year'] == year)
+        return(ds.mean(dim='time'))
+
+    ax = viz.GetRowCol(i, axes)
+    ax.set_title('{} {}'.format(calendar.month_name[month], year))
+    C1 = selector(prcp).plot.contourf(
+        transform = ccrs.PlateCarree(), ax=ax,
+        cmap = 'BrBG', extend="both",
+        levels=np.linspace(-6, 6, 13),
+        add_colorbar=False, add_labels=False
+    )
+    psi_sub = selector(psi850)
+    X,Y = np.meshgrid(psi_sub.lon, psi_sub.lat)
+    C0 = ax.contour(X,Y,psi_sub.values,
+        transform = ccrs.PlateCarree(),
+        levels=np.linspace(-10e4, 10e4, 21),
+        colors='k'
+    )
+
+my_extent = region.south_america.as_extent()
+my_extent[0] -= 20
+my_extent[1] += 20
+viz.FormatAxes(axes, extent = my_extent)
+
+plt.tight_layout()
+fig.subplots_adjust(right=0.925)
+cax = fig.add_axes([0.965, 0.1, 0.02, 0.75])
+cbar = fig.colorbar(C1, cax = cax)
+cbar.set_label('Rainfall Anomaly [mm/day]', rotation=270)
+cbar.ax.get_yaxis().labelpad = 20
+
+fig.savefig(os.path.join(paths.figures, 'circulation-NDJF-1516-anomaly-alt.pdf'), bbox_inches='tight')

### 2017-12-01 src/32-weather-type.py

@@ -128,6 +128,47 @@ viz.FormatAxes(axes[0,:], extent = region.southern_hemisphere.as_extent())
 viz.FormatAxes(axes[1,:], extent = region.south_america.as_extent())
 
 fig.savefig(os.path.join(paths.figures, 'weather-type-composite.pdf'), bbox_inches='tight')
+# PLOT ANOMALIES WITH A DIFFERENT SETUP
+wt_all = np.unique(wt['wtype'])
+psi = streamfunction.get_data()
+prcp = rainfall.get_data()
+
+fig,axes = plt.subplots(nrows=2, ncols = np.int(np.ceil(wt_all.size/2)),
+                         subplot_kw={'projection': ccrs.PlateCarree()},
+                         figsize=[12,8], sharex=True, sharey=True)
+for i,w in enumerate(wt_all):
+    def selector(ds):
+        times = wt.loc[wt['wtype'] == w].index
+        ds = ds.sel(time = np.in1d(ds.time, times))
+        ds = ds.mean(dim = 'time')
+        return(ds)
+
+    ax = viz.GetRowCol(i, axes)
+    ax.set_title('WT {}'.format(w))
+    C1 = selector(prcp['anomaly']).plot.contourf(
+        transform = ccrs.PlateCarree(), ax=ax,
+        cmap = 'BrBG', extend="both",
+        levels=np.linspace(-6, 6, 13),
+        add_colorbar=False, add_labels=False
+    )
+    psi_sub = selector(psi['anomaly'])
+    X,Y = np.meshgrid(psi_sub.lon, psi_sub.lat)
+    C0 = ax.contour(X,Y,psi_sub.values,
+        transform = ccrs.PlateCarree(),
+        levels=np.linspace(-9e4, 9e4, 19),
+        colors='k'
+    )
+    #ax.add_patch(region.wtype.as_patch(color='black'))
+
+plt.tight_layout()
+fig.subplots_adjust(right=0.9)
+cax = fig.add_axes([0.95, 0.1, 0.02, 0.75])
+cbar = fig.colorbar(C1, cax = cax)
+cbar.set_label('Rainfall Anomaly [mm/day]', rotation=270)
+cbar.ax.get_yaxis().labelpad = 20
+viz.FormatAxes(axes, extent = region.south_america.as_extent())
+
+fig.savefig(os.path.join(paths.figures, 'weather-type-composite-alt.pdf'), bbox_inches='tight')
 
 
 # Plot WT Time Series

### 2017-12-01 src/pyfloods/visualize.py

@@ -18,13 +18,13 @@ def _FormatAxis(ax, coast=True, grid=False, border=True, river = False, extent=N
         ax.set_yticks(ticks[1], crs=ccrs.PlateCarree())
         lat_formatter = LatitudeFormatter()
         ax.yaxis.set_major_formatter(lat_formatter)
-    if extent is not None:
-        ax.set_extent(extent, crs = ccrs.PlateCarree())
     if len(feature_list) > 0:
         for f in feature_list:
             ax.add_feature(f)
     if states is not None:
         ax.add_feature(states, edgecolor='gray')
+    if extent is not None:
+        ax.set_extent(extent, crs = ccrs.PlateCarree())
 
 def FormatAxes(axes, states=False, **kwargs):
     import matplotlib.pyplot as plt

### 2017-12-02 reports/poster-agu-2017/blocks/s2s-model-forecasts.tex
update poster

\begin{block}{S2S  Model Forecasts}
  \begin{mdframed}
  \begin{figure}
    \caption{
      Chiclet diagram of ensemble-mean precipitation anomaly forecasts over the Lower Paraguay River Basin from ECMWF S2S forecast data, as a function of the forecast target date (horizontal axis) and lead time (vertical axis).
      Time series of CPC daily mean precipitation over the same area is plotted with $y$-axis inverted.
  	}
    \noindent\includegraphics[width=0.925\textwidth]{chiclet-s2s-area-averaged.pdf}
  	\label{fig:chiclet}
  \end{figure}
  \end{mdframed}

  \Cref{fig:chiclet} uses a Chiclet diagram to visualize, as a function of lead time, the time evolution of the uncorrected, ensemble-mean rainfall anomaly forecast, spatially averaged over the Lower Paraguay River Basin.
  At times greater than about two weeks, the ensemble-mean forecast is slightly wetter than climatology.
  At weather timescales (less than one week), the ensemble-mean successfully predicts the timing and amplitude of the area-averaged rainfall.
  At intermediate timescales, the model successfully forecast the strongest breaks and pauses in the rainfall, such as the heavy rainfall during December 2015 and the dry period during mid-January 2016.

  \vspace{0.5cm}

  We explore whether using Model Output Statistics \cite[MOS;][]{Glahn:1972vt} can improve the modeled representation of rainfall (\cref{fig:subs-prob-fcst}).
  Specifically, we use: the raw model output(Raw); extended logistic regression \cite[XLR;][]{Wilks:2009bk}; heteroscedastic \cite[HXLR;][]{Messner:2014gp}; principal component regression \cite[PCR;][]{Mason:2008da,Wilks:2006fx}; and canonical correlation analysis \cite[CCA;][]{Mason:2008da,Barnston:1992gd} using 20 years of ECMWF forecasts.

  \vspace{0.5cm}

  In general, \cref{fig:subs-prob-fcst} indicates that better forecasts are obtained when both magnitude and spatial corrections are performed (PCR and CCA).
  The enhanced skill is achieved through the spatial corrections via the EOF-based regressions, which -- in contrast with the extended logistic models -- use information from multiple grid-boxes,.

  \begin{mdframed}
  \begin{figure}
  	\noindent\includegraphics[width=0.925\textwidth]{s2s-forecast-mos.pdf}
  	\caption{
    MOS-adjusted S2S model forecasts and skill scores.
    Top row shows the heavy rainfall ($>$\nth{90} percentile exceedance) forecast for 1-7 December 2015 as the odds ratio $\text{odds}_{r} \equiv \frac{p}{\qty(1 - p)} \frac{\qty(1 - p_c)}{p_c}$.
    Second row shows the Ignorance score $\text{IGN} \equiv - \log_2 p(Y)$.
    Bottom row shows the 2AFC skill score for each grid cell.
    For all three rows, the grid cells which experienced a \nth{90} percentile exceedance for 1-7 December 2015 are outlined in black.
        \label{fig:subs-prob-fcst}
  	}
  \end{figure}
  \end{mdframed}
\end{block}

### 2017-12-02 reports/poster-agu-2017/blocks/s2s-sallj.tex
update poster

\begin{block}{S2S Drivers of Weather Patterns}
  \begin{mdframed}
  \begin{figure}
    \caption{
  		Anomalous probability of occurrence of each weather type concurrent with observance of each MJO and ENSO phase.
  		Only values which are significant at $p<0.10$ are shown.
      \label{fig:wt-mjo-enso}
  	}
  	\noindent\includegraphics[width=0.925\textwidth]{wt-mjo-enso.pdf}
  \end{figure}
  \end{mdframed}

  During El Ni\~{n}o years such as 2015-16, weather type 1 -- related to Chaco jet events -- occurs more frequently for almost all MJO phases, and particularly during phases 1, 2, and 5, consistent with observations of the year under study.
  This is consistent with previous studies \cite[i.e.][]{Velasco1987} which find that the intensity and exact location of the anomalous low-level anticylonic anomaly over central Brazil is relevant for the precise impact of ENSO events.

  \vspace{0.5cm}

  While the occurrence of WT1 during NDJF 2015-16 is well explained by ENSO and MJO variability, these features alone do not explain the occurrence of WT4, the ``No-Chaco'' jet event.
  Previous studies emphasize the importance of Pacific-Atlantic interaction for forecasting climate effects in this region \cite{Vera:2006ib,Barreiro:2017ct,Lima:2017hw}.
  A persistent SST dipole in the central southern Atlantic Ocean favors the occurrence of WT4 by blocking transient extratropical wave activity coming the Pacific, facilitating transitions from Chaco jet events (WT 1) to No-Chaco jet events (WT 4) via enhanced low-level wind circulation from southern Brazil towards the Atlantic, and back to north-east Brazil and the Amazon (see \cref{fig:chaco-nochaco}) due to land-sea temperature contrasts.

  \vspace{0.5cm}

  Composite analysis (not shown) of months with many WT4 occurrences is consistent with the schematic shown here.

  \begin{mdframed}
  \begin{figure}
  	\noindent\includegraphics[width=0.925\textwidth]{../writeup/ChacoNoChacojet.pdf}
    \caption{
      Simple schematics of low-level jet events (red arrows) during austral summer and El Ni\~no years.
      \label{fig:chaco-nochaco}
  	}

  \end{figure}
  \end{mdframed}
\end{block}

### 2017-12-02 reports/poster-agu-2017/blocks/at-a-glance.tex
update poster
possible move to final poster version

@@
 2015, repeated intense rainfall events associated with mesoscale convective activity caused severe flooding along the Paraguay-Paran\'{a} river system (\cref{fig:study-area}, red box), \emph{displacing over 170000 people}.
-  We use a weather typing approch within a diagnostic approach to show that:
+  We use a weather typing approach within a diagnostic framework to show that:
   \begin{itemize}
-    \item Moisture and energy advection via the South American Low-Level Jet \cite{Marengo:2012cm}, particularly during ``No-Chaco'' events \cite{Vera:2006ib} favored mesoscale convective activity
-    \item Strong El Ni\~{n}o favored strong SALLJ but MJO and Atlantic dipole pattern contributed to jet exit region (thus No-Chaco SALLJ events)
-    \item Raw S2S model forecasts of rainfall had limited skill beyond 10-15 days, but statistical methods -- particularly the PCR and CCA methods that correct both spatial patterns and magnitudes -- substantially improve forecast skill
+    \item Moisture and energy advection via the South American Low-Level Jet \cite[SALLJ;][]{Marengo:2012cm}, particularly during ``No-Chaco'' events \cite{Vera:2006ib} favored mesoscale convective activity
+    \item Numerical forecasts predicted enhanced risk of heavy rainfall at the seasonal scale, but biases in spatial patterns suggest difficulties representing Pacific-Atlantic interaction
+    \item Uncorrected sub-seasonal model forecasts of rainfall had limited skill beyond 10-15 days; use of Model Output Statistics -- particularly methods that correct both spatial patterns and magnitudes -- substantially improved forecast skill
+    \item Strong El Ni\~{n}o and active MJO favored a strong SALLJ but an Atlantic dipole pattern influenced the jet's exit region
   \end{itemize}

### 2017-12-02 reports/poster-agu-2017/blocks/obs-1516.tex
update poster
possible move to final poster version

@@
 
-  During NDJF 2015-16, the most heavy rainfall occurred during weather types 1 and 4 (\cref{fig:rain-wt}).
+  During austral summer (NDJF) 2015-16, the most heavy rainfall occurred during weather types 1 and 4 (\cref{fig:rain-wt}).
   Monthly-scale circulation anomalies (\cref{fig:anomalies}) show a weak anticyclonic circulation that set up over central Brazil during November 2015 and strengthened into the following month.
@@
   The observed rainfall and circulation anomalies are consistent with the aggregation of the observed weather types shown in \cref{fig:rain-wt} and discussed above.
-  Weather types 1 and 4  occurred more frequency than their climatology.
 
@@
         Colors: anomalies of rainfall associated with each weather type [\si{\milli\meter\per\day}].
-        Contours: anomalies of \SI{850}{\hecto\pascal} [contour interval \SI{1e4}{\meter\squared\per\second}]
+        Contours: anomalies of \SI{850}{\hecto\pascal} streamfunction [contour interval \SI{1e4}{\meter\squared\per\second}]
         \label{fig:anomalies}

### 2017-12-02 reports/poster-agu-2017/blocks/obs-weather-types.tex
update poster
possible move to final poster version

@@
   \end{itemize}
-  We use weather typing \cite{Munoz2015} to identify daily circulation patterns:
+  We use weather typing \cite{Munoz2015} to represent daily circulation patterns:
   \begin{enumerate}
@@
   		Colors: anomalies of rainfall associated with each weather type [\si{\milli\meter\per\day}].
-      Contours: anomalies of \SI{850}{\hecto\pascal} [contour interval \SI{1e4}{\meter\squared\per\second}]
+      Contours: anomalies of \SI{850}{\hecto\pascal} streamfunction [contour interval \SI{1e4}{\meter\squared\per\second}]
   	}

### 2017-12-02 reports/poster-agu-2017/blocks/s2s-model-forecasts.tex
possible move to final poster version

@@
       Chiclet diagram of ensemble-mean precipitation anomaly forecasts over the Lower Paraguay River Basin from ECMWF S2S forecast data, as a function of the forecast target date (horizontal axis) and lead time (vertical axis).
-      Time series of CPC daily mean precipitation over the same area is plotted with $y$-axis inverted.
+      Time series of daily precipitation over the same area is plotted with $y$-axis inverted.
   	}
@@
 
-  \Cref{fig:chiclet} uses a Chiclet diagram to visualize, as a function of lead time, the time evolution of the uncorrected, ensemble-mean rainfall anomaly forecast, spatially averaged over the Lower Paraguay River Basin.
-  At times greater than about two weeks, the ensemble-mean forecast is slightly wetter than climatology.
+  \Cref{fig:chiclet} uses a Chiclet diagram \cite{Carbin:2016fx} to visualize, as a function of lead time, the time evolution of the uncorrected, ensemble-mean rainfall anomaly forecast, spatially averaged over the Lower Paraguay River Basin.
+  At lead times greater than about two weeks, the ensemble-mean forecast is slightly wetter than climatology.
   At weather timescales (less than one week), the ensemble-mean successfully predicts the timing and amplitude of the area-averaged rainfall.
@@
 
-  \vspace{0.5cm}
+  \end{block}
+  \begin{block}{Model Output Statistics}
 
@@
 
-  In general, \cref{fig:subs-prob-fcst} indicates that better forecasts are obtained when both magnitude and spatial corrections are performed (PCR and CCA).
+  \Cref{fig:subs-prob-fcst} indicates that better forecasts are obtained when both magnitude and spatial corrections are performed (PCR and CCA).
   The enhanced skill is achieved through the spatial corrections via the EOF-based regressions, which -- in contrast with the extended logistic models -- use information from multiple grid-boxes,.

### 2017-12-02 reports/poster-agu-2017/blocks/s2s-sallj.tex
possible move to final poster version

@@
   During El Ni\~{n}o years such as 2015-16, weather type 1 -- related to Chaco jet events -- occurs more frequently for almost all MJO phases, and particularly during phases 1, 2, and 5, consistent with observations of the year under study.
-  This is consistent with previous studies \cite[i.e.][]{Velasco1987} which find that the intensity and exact location of the anomalous low-level anticylonic anomaly over central Brazil is relevant for the precise impact of ENSO events.
+  This agreesß with previous studies \cite[i.e.][]{Velasco1987} which find that the intensity and exact location of the anomalous low-level anticylonic anomaly over central Brazil is relevant for the precise impact of ENSO events.
 
@@
   While the occurrence of WT1 during NDJF 2015-16 is well explained by ENSO and MJO variability, these features alone do not explain the occurrence of WT4, the ``No-Chaco'' jet event.
-  Previous studies emphasize the importance of Pacific-Atlantic interaction for forecasting climate effects in this region \cite{Vera:2006ib,Barreiro:2017ct,Lima:2017hw}.
-  A persistent SST dipole in the central southern Atlantic Ocean favors the occurrence of WT4 by blocking transient extratropical wave activity coming the Pacific, facilitating transitions from Chaco jet events (WT 1) to No-Chaco jet events (WT 4) via enhanced low-level wind circulation from southern Brazil towards the Atlantic, and back to north-east Brazil and the Amazon (see \cref{fig:chaco-nochaco}) due to land-sea temperature contrasts.
+  Previous studies emphasize the importance of Pacific-Atlantic interaction for forecasting climate effects in this region \cite{Vera:2006ib,Barreiro:2017ct,Lima:2017hw,Munoz2015}.
+  A persistent SST dipole in the central southern Atlantic Ocean favors the occurrence of WT4 by blocking transient extratropical wave activity coming from the Pacific, facilitating transitions from Chaco jet events (WT 1) to No-Chaco jet events (WT 4) via enhanced low-level wind circulation from southern Brazil towards the Atlantic, and back to north-east Brazil and the Amazon (see \cref{fig:chaco-nochaco}) due to land-sea temperature contrasts.
 

### 2017-12-02 src/21-circulation-NDJF-1516.py

@@ -20,7 +20,9 @@ data_proj = ccrs.PlateCarree()
 months_plot = [11, 12, 1, 2]
 years_plot = [2015, 2015, 2016, 2016]
 
-# MAKE THE PLOT
+################################################################################
+# Plot the anomalies
+################################################################################
 fig, axes = plt.subplots(nrows=2, ncols=4, subplot_kw={'projection': map_proj}, figsize=(16, 5))
 for i,(month,year) in enumerate(zip(months_plot, years_plot)):
     def selector(ds):
@@ -62,8 +64,10 @@ cbar1.ax.get_yaxis().labelpad = 20
 
 fig.savefig(os.path.join(paths.figures, 'circulation-NDJF-1516-anomaly.pdf'), bbox_inches='tight')
 
-# MAKE THE PLOT WITH ALTERNATIVE SETUP
-fig, axes = plt.subplots(nrows=2, ncols=2, subplot_kw={'projection': map_proj}, figsize=(11, 6))
+################################################################################
+# Plot the anomalies using a different setup
+################################################################################
+fig, axes = plt.subplots(nrows=2, ncols=2, subplot_kw={'projection': map_proj}, figsize=(8.25, 4.5))
 for i,(month,year) in enumerate(zip(months_plot, years_plot)):
     def selector(ds):
         ds = ds.sel(time = ds['time.month']==month)

### 2017-12-02 src/32-weather-type.py

@@ -22,6 +22,9 @@ pars = {
     'pcscaling': 1
 }
 
+################################################################################
+# Calculate the weather types
+################################################################################
 recalculate = False
 try:
     cluster_ts = xr.open_dataarray(paths.file_weather_types)
@@ -67,7 +70,9 @@ wt_prop = pd.DataFrame({'climatology': wt_counts / np.sum(wt_counts),
               'NDJF1516': wt_counts2 / np.sum(wt_counts2)})
 print(wt_prop)
 
-# Plot Proportion of Days
+################################################################################
+# Plot the proportion of days of each weather type
+################################################################################
 plt.figure(figsize=(8.5, 4.5))
 plt.plot(wt_prop['climatology'], label = "Climatology")
 plt.plot(wt_prop['NDJF1516'], label = "2015-16")
@@ -78,7 +83,9 @@ plt.grid()
 plt.savefig(os.path.join(paths.figures, 'wt_occurrence_fraction.pdf'), bbox_inches='tight')
 
 
-# Plot Anomalies
+################################################################################
+# Plot the anomalies associated with each weather type
+################################################################################
 wt_all = np.unique(wt['wtype'])
 psi = streamfunction.get_data()
 prcp = rainfall.get_data()
@@ -128,14 +135,17 @@ viz.FormatAxes(axes[0,:], extent = region.southern_hemisphere.as_extent())
 viz.FormatAxes(axes[1,:], extent = region.south_america.as_extent())
 
 fig.savefig(os.path.join(paths.figures, 'weather-type-composite.pdf'), bbox_inches='tight')
-# PLOT ANOMALIES WITH A DIFFERENT SETUP
+
+################################################################################
+# Plot the anomalies associated with each weather type using a different setup
+################################################################################
 wt_all = np.unique(wt['wtype'])
 psi = streamfunction.get_data()
 prcp = rainfall.get_data()
 
 fig,axes = plt.subplots(nrows=2, ncols = np.int(np.ceil(wt_all.size/2)),
                          subplot_kw={'projection': ccrs.PlateCarree()},
-                         figsize=[12,8], sharex=True, sharey=True)
[…]

### 2017-12-02 src/41-chiclet-s2s-area-avgeraged.py

@@ -34,7 +34,7 @@ target_date = pd.to_datetime(forecast.target_date.values)
 obs = obs['2015-11-01':'2016-02-29']
 
 # MAKE PLOT
-fig,axes = plt.subplots(ncols=1, nrows=2, figsize=((15,8)), gridspec_kw = {'height_ratios':[3, 1]})
+fig,axes = plt.subplots(ncols=1, nrows=2, figsize=((9, 5)), gridspec_kw = {'height_ratios':[3, 1]})
 
 ax=axes[0] # Chiclet
 C1 = forecast.plot.pcolormesh(x='target_date', y='L', cmap = 'BrBG', ax=ax, add_colorbar=False, add_labels=False)

### 2017-12-02 src/43-s2s-MOS-forecasts.py

@@ -40,7 +40,7 @@ ncols = len(model_names)
 fig, axes = plt.subplots(
     ncols=len(model_names), nrows=3,
     subplot_kw={'projection': ccrs.PlateCarree()},
-    figsize=(14,10), sharex=True, sharey=True
+    figsize=(11,8), sharex=True, sharey=True
 )
 for i,m in enumerate(model_names):
     ax = axes[0, i] # Forecast for December 1-7 2015

### 2017-12-04 reports/poster-agu-2017/blocks/obs-1516.tex
final version of poster for printing

@@
     \caption{
-  		Daily rainfall averaged over Lower Paraguay River Basin (\cref{fig:study-area}) and observed weather type for each day in NDJF 2015-16.
+  		Daily rainfall averaged over Lower Paraguay River Basin and observed weather type for each day in NDJF 2015-16.
       Blue lines indicate the climatological 50th, \nth{90}, and 99th percentiles of NDJF area-averaged rain.
@@
 
-  During austral summer (NDJF) 2015-16, the most heavy rainfall occurred during weather types 1 and 4 (\cref{fig:rain-wt}).
+  During austral summer (NDJF) 2015-16, most heavy rainfall occurred during weather types 1 and 4 (\cref{fig:rain-wt}).
   Monthly-scale circulation anomalies (\cref{fig:anomalies}) show a weak anticyclonic circulation that set up over central Brazil during November 2015 and strengthened into the following month.
   In January 2016 it weakened before returning in February 2016.
-  The observed rainfall and circulation anomalies are consistent with the aggregation of the observed weather types shown in \cref{fig:rain-wt} and discussed above.
+  The observed rainfall and circulation anomalies are consistent with the aggregation of the observed weather types shown in \cref{fig:rain-wt}.
 

### 2017-12-04 reports/poster-agu-2017/blocks/s2s-model-forecasts.tex
final version of poster for printing

@@
   We explore whether using Model Output Statistics \cite[MOS;][]{Glahn:1972vt} can improve the modeled representation of rainfall (\cref{fig:subs-prob-fcst}).
-  Specifically, we use: the raw model output(Raw); extended logistic regression \cite[XLR;][]{Wilks:2009bk}; heteroscedastic \cite[HXLR;][]{Messner:2014gp}; principal component regression \cite[PCR;][]{Mason:2008da,Wilks:2006fx}; and canonical correlation analysis \cite[CCA;][]{Mason:2008da,Barnston:1992gd} using 20 years of ECMWF forecasts.
+  Specifically, we use: the raw model output (Raw); extended logistic regression \cite[XLR;][]{Wilks:2009bk}; heteroscedastic XLR \cite[HXLR;][]{Messner:2014gp}; principal component regression \cite[PCR;][]{Mason:2008da,Wilks:2006fx}; and canonical correlation analysis \cite[CCA;][]{Mason:2008da,Barnston:1992gd} using 20 years of ECMWF forecasts.
 

### 2017-12-04 src/21-circulation-NDJF-1516.py

@@ -67,7 +67,8 @@ fig.savefig(os.path.join(paths.figures, 'circulation-NDJF-1516-anomaly.pdf'), bb
 ################################################################################
 # Plot the anomalies using a different setup
 ################################################################################
-fig, axes = plt.subplots(nrows=2, ncols=2, subplot_kw={'projection': map_proj}, figsize=(8.25, 4.5))
+map_proj = ccrs.PlateCarree()
+fig, axes = plt.subplots(nrows=2, ncols=2, subplot_kw={'projection': map_proj}, figsize=(8, 5.25), sharex=True, sharey=True)
 for i,(month,year) in enumerate(zip(months_plot, years_plot)):
     def selector(ds):
         ds = ds.sel(time = ds['time.month']==month)
@@ -90,10 +91,9 @@ for i,(month,year) in enumerate(zip(months_plot, years_plot)):
         colors='k'
     )
 
-my_extent = region.south_america.as_extent()
-my_extent[0] -= 20
-my_extent[1] += 20
-viz.FormatAxes(axes, extent = my_extent)
+my_extent = [-100, -20, -45, 10]
+viz.FormatAxes(axes, extent = my_extent,
+    ticks=[np.linspace(-180, 180, 19), np.linspace(-90, 90, 19)])
 
 plt.tight_layout()
 fig.subplots_adjust(right=0.925)

### 2017-12-04 src/pyfloods/weather_type.py

@@ -34,6 +34,7 @@ def Classifiability(P, Q):
     return(classifiability)
 
 def XrEofCluster(ds, n_clusters=5, prop=0.90, nsim=100, verbose = True, pcscaling=0):
+    # DS should be an xarray data array (not dataset!)
     if verbose:
         print('xarray-based classifiability for {} clusters'.format(n_clusters))
         print('Performing EOF decomposition of data for dimension reduction.')

### 2018-02-06 reports/writeup/ReviewerComments.tex
start putting things in makefile

\documentclass{article}

%Packages
\usepackage{
	siunitx, % units
  graphicx, % for images
	natbib,
	booktabs,
	enumitem
}
\usepackage[letterpaper]{geometry}
\renewcommand{\thefigure}{S\arabic{figure}}
\renewcommand{\thetable}{S\arabic{table}}
\usepackage[breaklinks=true,hidelinks]{hyperref}

\usepackage{mdframed}
\newenvironment{answer}{\begin{mdframed}}{\end{mdframed}}

%-------------------------------------------------------------------------------
% Title and Authors
%-------------------------------------------------------------------------------

\title{Response to Reviewers}
\author{James Doss-Gollin\and \'{A}ngel G. Mu\~{n}oz  \and Simon J. Mason \and Max Past\'{e}n }
\date{\today}

\begin{document}

\maketitle

We would like to thank the editors and reviewers for their specific and helpful comments.
In particular, reviewer suggestions as to the presentation of results and discussion of literature have helped us clarify our thinking and improve the clarity of our writing.

Per request by two reviewers we have added additional fields (\texttt{WHICH}) to some figures which necesiates minor changes to the discussion of those figures and the data sub-section.
This also provided us with the opportunity to revise all figure captions; all three reviewers noted that several figure captions contained errors.

Reviewer 3 gave three major comments which led to some additional analysis.
\texttt{Describe more later}.

Please find our responses to the specific comments below.

\section{Response to Reviewer 1 Comments}

\begin{enumerate}
	\item Lines 317-321: This should be shown by other atmospheric fields. For instance, wind field at 850hPa and vertically integrated moisture flux.
	\begin{answer}
		The reviewer is suggesting that we incorporate other variables into our plots of figure 3 in order to better show the relationship to the established literature.
		Do we want to do it?
	\end{answer}
	\item Lines 333-334: It would be convenient to include these percentages as subtitles in each panel of Fig. 4.
	\begin{answer}
		The reviewer is refering to the percentage of variance explained.
		This is easy to do and we will do it.
	\end{answer}
	\item Lines 334-337: This should be further proved. Fig. 4 does not provide enough evidence to support these statements.
	\begin{answer}
		The reviewer is refering to the large-scale anomalies associated with each EOF.
		I think that we can change the language here -- we don't want or need to make any strong statements about what these EOFs mean, and were just trying to provide some physical intuition about them.
	\end{answer}
	\item Line 361: Table S1 is not included in the supplemental material.
	\begin{answer}
		This is an erroneous reference, address it.
	\end{answer}
	\item Lines 555-559: This is not shown or discussed in the present manuscript.
	\begin{answer}
		The reviewer is refering to the statement that antecedent conditions plus persistent, heavy rainfall led to heavy flooding.
		It is correct that we did not show or discuss this; we can say that the persistent and intense rainfall led to flooding and may have been influenced by the antecedent conditions -- can add hydrological references if needed but not really necessary.
	\end{answer}
	\item	Lines 671-673: Check the authors of this reference (Kim et al. 2014).
	\begin{answer}
		We thank the reviewer for identifying the duplication of the authors
	\end{answer}
	\item Lines 853-856 (Fig. 3): One panel is missing. Streamfunction at 200hPa?
	\begin{answer}
		The streamfunction is shown at \SI{850}{\hecto\pascal} and the misleading caption has been corrected.
	\end{answer}
	\item	Lines 860-863 (Fig. 6): Yellow labels are not easy to read.
	\begin{answer}
		Let's look for a better color map -- none will be ideal here.
	\end{answer}
	\item	Lines 864-866 (Fig. 7): Streamfunction at 200hPa is not shown.
	\begin{answer}
		As for Fig. 3, the caption is incorrect and has been corrected.
	\end{answer}
\end{enumerate}

\section{Response to Reviewer 2 Comments}

\begin{enumerate}
	\item Weather type some times appears as "WT" and other as " weather type".  It is suggest to define it (I have not found) and to use WT along the text.
	\begin{answer}
		Suggestion taken, will implement.
	\end{answer}
	\item L61 - "Grimm and Tedeschi 2009"  is not a appropriated to explain the warm season systems in the region.
	\begin{answer}
		Let's change the reference
	\end{answer}
	\item L80 and L470 - Please, to include a reference on the influence of MJO over the convection in the region.
	\begin{answer}
		Let's add a good one
	\end{answer}
	\item L84 - The correct citation here is Marengo et al. 2012. Please, check.
	\begin{answer}
		The reviewer is correct, that is the paper we intended to cite.
	\end{answer}
	\item L163 - Why "retaining the seasonal cycle"?
	\begin{answer}
		We can be more clear here: let's say that we retain intra-seasonal variability or similar.
	\end{answer}
	\item L166-171 - The phrase "All codes used … packages" may be moved to the supplementary material.
	\begin{answer}
		This is certainly a long several sentences that doesn't collectively add much to the analysis.
		At the same time, it's important to me to support reproducible research  and open science by making codes available and citing the authors of important open source packages.
		Let's keep it but make it much shorter.
	\end{answer}
	\item L174-176 - "This region corresponds to the region over which given topography and previous studies  (Barros et al. 2004; Bravo et al. 2011), one might  " should be " In this region, given topography and previous studies  (Barros et al. 2004; Bravo et al. 2011), one might..."
	\begin{answer}
		The reviewer has suggested a more clear wording.
	\end{answer}
	\item L195 -What is the meaning of the affirmation "physical mechanisms are more interpretable"?
	\begin{answer}
		I'll ask \'{A}ngel to answer, he may have strong feelings here
	\end{answer}
	\item L261 - change "empirical orthogonal functions (EOFs) ..."  to "EOFs ..." since it was defined in page 11.
	\begin{answer}
		I'm not seeing it on line 261, assuming the reviewer meant line 244 in which case agreed
	\end{answer}
	\item L855, L864, - "Top and middle rows …" should be "Top and bottom rows …"
	\begin{answer}
		Agreed, Reviewer 1 also noticed some errors in the figure captions which have been addressed.
	\end{answer}
	\item L342 - "past …" should be "southward …"
	\begin{answer}
		Yes, this is a more clear wording
	\end{answer}
	\item L344 - "before …" should be "northward …"
	\begin{answer}
		Agreed
	\end{answer}
	\item L345 - Looking Figure 5, WT3 resembles that one anomaly pattern generally associated with establishment or not of SACZ and called "seesaw" by some authors.  Please, see and cite, for example, Nogués-Peagle and Mo  (1997) and Liebmann et al. (2004).
	\begin{answer}
		Let's re-read these papers but at a glance this looks good.
		The question is how much to incorporate into discussion.
	\end{answer}
	\item L372 - It is not clear the phrase "During early January 2016 weather type 3 led to persistent low rainfall" since I noted in Fig. 6 persistent WT3 from middle to end January. Please, clarify.
	\begin{answer}
		Yes, this occurred from mid to late January, not early January 2016.
	\end{answer}
	\item L375 - The legend of Figure 7 says "200 hPa and 850 hPa". But, it is showing stream-function anomalies in only one atmospheric level. What is the level? Please, to correct the legend.
	\begin{answer}
		The reviewer is correct and the erroneous figure captions have been fixed.
	\end{answer}
	\item L412 and Figure 10 legend - The denominations of the sub-seasonal forecasts have different names. It is suggest to use "S2S model forecasts" for both text and Figure.
	\begin{answer}
		I believe that the reviewer is referring to the ``Raw'' model only?
	\end{answer}
	\item L418 - There is no "10b" in Figure 10. It is suggested to include the letters "(a), (b), (c), … (m), …" in Figure 10, which is clearer than that used, for example, in L423 "10, top row, first two columns …"
	\begin{answer}
		OK, I'm not a huge fan of this but here it may be helpful.
	\end{answer}
	\item L439 - change "Principal Component Regression …" to "PCRs" already defined previously.
	\begin{answer}
		Agreed
	\end{answer}
	\item L442 - change "Canonical Correlation Analysis" to "CCA".
	\begin{answer}
		Agreed
	\end{answer}
	\item L448 - "resolution and uncertainty (Ignorance Score; second row, fourth and fifth columns) and discrimination (2AFC score; middle row, fourth and fifth columns)" should be "resolution and uncertainty according Ignorance Score and discrimination according 2AFC (fig. 10)"
	\begin{answer}
		Yes, this is a more readable wording
	\end{answer}
	\item L481 -Check the reference to Marengo et al. (2004) in this part of text since there is not rainfall analysis in this paper.
	\begin{answer}
		This is true, though the Marengo paper shows OLR and similar.
		I think it's reaonable to argue that we have identified simlar mechanisms to what were identified in the literature but we should be more careful with wording as it's not only heavy rainfall but really heavy rainfall, thunderstorms, and mesoscale convective activity.
		For reference, this citation occurs at line 461 not 481.
	\end{answer}
	\item L464-466 - Consider to re-write the phrase.
	\begin{answer}
		The reviewer is refering to the sentence ``Understanding the mechanisms and possible...research''.
		Will re-write.
	\end{answer}
	\item L488 - and others - change "South American Low-Level Jet…" to "SALLJ" already previously defined.
	\begin{answer}
		Agree
	\end{answer}
	\item L503 - and others - change "anti-cyclonic" to "anticyclonic"
	\begin{answer}
		The reviewer's suggestion is more consistent with AMS guidelines \url{http://glossary.ametsoc.org/wiki/Anticyclonic_circulation} so we have changed it.
	\end{answer}
	\item L512 - change "South Atlantic" to "SACZ"
	\begin{answer}
		Agree
	\end{answer}
\end{enumerate}

\section{Response to Reviewer 3 Comments}

\subsection{Major Comments}

\begin{enumerate}
	\item The following claim in the paper is not well supported by the evidence provided (or maybe the evidence is not clearly reported).
	Line 29, Line 501 to 503 and Lines 564 to 565 - The presence of a dipolar SST anomaly in southern Atlantic during NDJF 2015-16 which enhances the frequency of occurrence of No-Chaco events during this flood period is not shown in any figure. These No-Chaco jet events incidentally are the most important weather types (along with Chaco jets) in causing the flood (Figure 6) so a figure supporting the existence of the dipolar SST anomaly is required.
	Not much further analysis is required, only a figure showing the SST anomalies in December 2015-16 will suffice.
	Also no evidence has been shown of a strong El Nino in this period. Just a monthly mean time series of Nino 3.4 will suffice.
	\begin{answer}
		We have some analysis that we can show.
		Agree that a figure supporting the existence of the dipolar SST anomaly is required -- this was intended to be Fig. 13.
		Discuss with \'{A}ngel whether further analysis is needed here?
		Can put an ENSO time series quite easily in the supplemental figures, it's already downloaded.
	\end{answer}

	\item The derivation of the weather types using the 4 EOFs is not very clear from the text. It will help the typical reader if the method of deriving the 6 weather types was explained in more detail and using physical arguments/analogs.
	\begin{answer}
		OK, we can add some clarity here.
		I'll defer to \'{A}ngel to draw up this argument.
	\end{answer}

	\item It is not clear which of the two types of atmospheric conditions - the synoptic weather type or the large scale type - is more important for the floods? In other words, does a strong El Nino always enhance the occurrence of WT1 and hence always result in a flood? If not are the  WT1 and WT4 more important? I guess the jets are not more important as they occur in no-flood years as well. So then the occurrence of the El Nino in combination with the MJO phases is crucial for the occurrence of these floods? Please explain.
	\begin{answer}
		This is intended to be a central conclusion of the paper, so we should definitely make sure we're being as clear as possible with regard to this point.
		Because of small sample size, it's hard to build a really powerful statistical model relating seasonal features to floods.
		However, we \emph{can} say (i) that ENSO favors jet events; (ii) that the Atlantic dipole favors no-Chaco jet events; (iii) that the MJO modulates the jet; and (iv) that the no-Chaco jet events of WT4 are particularly important for heavy rainfall in this region.
		This is related to Major Comment 1 so let's discuss a bit more.
	\end{answer}
\end{enumerate}

\subsection{Minor Comments}

\begin{enumerate}
	\item I think maps showing the monthly rainfall anomalies in NDJF (or monthly) right in the beginning of the paper are essential. I suggest adding a figure showing these spatial anomaly maps in the main text. These maps will also help validate your choice of analysis regions for weather typing, EOFs etc.
	\begin{answer}
		OK I was a bit hesitant to put in figures before we have described data and methods, but the reviewer thinks it will be really helpful and I agree that it lays the story out more clearly.
		We can alwas cross-reference the methods.
	\end{answer}
	\item Line 83 - westerly wind regime over which region?
	\begin{answer}
		Need to review paper and fill this in
	\end{answer}
	\item Line 127 - Maybe give a reference supporting that the rainfall is mostly nighttime in this region?
	\begin{answer}
		Agreed, will think of an appropriate reference
	\end{answer}
	\item Lines 145 to 151 - some more information about how these seasonal forecasts are produced is needed. For example how much data goes into generating the pdfs? Also what does flexible format mean? is flexible format relevant here? Also what is meant by short record of 2012-2016? do the forecasts themselves span this time period or are the forecasts derived from statistics collected over this time period? I think it is the former, but an explanation in the manuscript will be appreciated and will clear confusion. Since the authors find that the seasonal forecasts performed better than subseasonal forecasts, a greater discussion of how these seasonal forecasts are derived will help.
	\begin{answer}
		I will refer to \'{A}ngel here.
	\end{answer}
	\item Line 181  - Is there a reason for using the 850mb stream function? why not any other level?
	\begin{answer}
		We chose this level because it shows the low-level jet activity and has been shown to be the level at which most moisture transport occurs.
		We could add other levels, but I think we should just state this more clearly in the text.
	\end{answer}
	\item Line 184 - EOFs are in general dependent on the size/location of the domain. Is the EOF analysis performed here sensitive to the domain chosen?
	\begin{answer}
		This is a good point.
		The Leading EOFs are relatively robust to changes of a few degrees in domain.
		These small changes also do not lead to substantial changes in the weather types identified.
		If the domain is made much larger, then other patterns become important.
	\end{answer}
	\item Line 225 - Why are the seasonal forecasts not bias corrected? I see the authors state that the seasonal forecasts are already bias corrected by IRI (Line 268). Maybe bring that line to Line 225? Also the method that IRI uses for this model forecast bias correction hasn't been reported. Do they use some type of MOS?
	\begin{answer}
		Defer to \'{A}ngel here
	\end{answer}
	\item Line 229 - Probability conditional on what?
	\begin{answer}
		I think we can simplify the language around logistic regression.
		Here we mean conditional on a set of predictors.
	\end{answer}
	\item Line 255 - It is not clear how the Kendall's tau rank test helps chose the best result.
	\begin{answer}
		Defer to \'{A}ngel
	\end{answer}
	\item Lines 258 to 263 - Did the authors display this validation in the paper?
	\begin{answer}
		This is how the skill scores are computed.
	\end{answer}
	\item Section on line 309 - Why is this section named climatological drivers? The drivers analyzed here are specific to the 2015-16 flood. Or does this figure show composites over several years irrespective of whether they were flood years or not?
	\begin{answer}
		We should clarify our language here.
		These are for all rainfall events, not just 2015-16.
	\end{answer}
	\item Figure 3 - There are only two rows in this figure (also Figure 7) unlike what the caption says. Maybe the empty space in the figure should be utilized to add grid lines and longitude and latitude values for better readability. Overlay horizontal wind vectors and isobars for improving understandability for the reader.
	\begin{answer}
		As reviewers 1 and 2 pointed out, our captions are incorrect.
		We can overlay wind vectors and isobars -- let's discuss the best way to do that.
	\end{answer}
	\item Lines 317 to 321 - The discussion presented here is hard to follow without wind vectors and isobars in Figure 3. This information can be derived from the stream function but why not present it explicitly? These fields will be especially helpful when the authors comment on the roles of Chaco and No Chaco jets later in the article.
	\begin{answer}
		OK, this is a helpful suggestion.
		Although it makes the figures a bit more cluttered, it's also true that having some more information in the figure will be helpful.
	\end{answer}
	\item Lines 334 to 337 - the relationship between the baroclinic waves and the east-west gradient is not clear. What is the physical meaning of these EOFs? How did the authors deduce the physical relationships between these EOFs and the jets? the explanation given seems insufficient. A better explanation should be provided especially because the subsequent weather types are dependent on these EOFs.
	\begin{answer}
		Reviewer 1 also noted this, see response to Reviewer 1 Comment \#3.
	\end{answer}
	\item Line 371 - weather types 1 and 5 produce rainfall - i don't see WT5 in this period. Also isn't WT5 associated with dry conditions?
	\begin{answer}
		The WT5 days during this period were typically dry days; even with large-scale conditions very favorable to rainfall, we would expect some southerly circulation over the study area.
		Let's clarify our language.
	\end{answer}
	\item Figure 8 - Maybe reiterate in this figure that the forecast was made in November 2015 to remind the reader?
	\begin{answer}
		This would be a helpful thing to add in the figure caption, agreed.
	\end{answer}
	\item Lines 389 to 396 - What constitutes this high odds? That is, what kind of weather conditions captured by the seasonal forecast model could have resulted in these high odds? a seasonal forecast of high LLJ season or a seasonal forecast of a high El Nino and MJO phases 1 to 5?
	\begin{answer}
		We ought to think about what was going on here.
		We may not have a straightforward answer but then at least we ought to say so.
	\end{answer}
	\item Figure 9 - indicate in the figure that the data is not bias corrected.
	\begin{answer}
		Fair point to include.
	\end{answer}
	\item Lines 445 to 453 - How are the PCR and CCA better than others as their high Odds ratio is displaced from the spatial location where the high precipitation was observed?
	\begin{answer}
		``Better'' or ``worse'' is a fairly heuristic description, so let's be clear about what we're saying.
	\end{answer}
	\item Lines 470 to 473 - see comment 14.
	\begin{answer}
		We concur; this point is addressed above, as referenced in comment 14.
	\end{answer}
	\item Lines 474 to 477 - what about the correlations with other EOFs? also why not show a table of all correlations? The correlations reported here are rather small.
	\begin{answer}
		These correlations have been computed so it would be straightforward to present them in the supplemental materials.
		We note that correlations are likely to be lower at daily time steps than monthly ones.
	\end{answer}
	\item Line 476 - what is RMM1
	\begin{answer}
		This is a MJO index, let's make sure we describe this.
	\end{answer}
	\item Paragraph on Line 504 - The purpose of this discussion is not clear. it has too little information about something complex and hence warrants more explanation.
	\begin{answer}
		I agree that we can be a bit more clear about how we describe this.
		Perhaps breaking this section into sub-sections may be helpful?
	\end{answer}
	\item Lines 556 to 557 - The observed …. flooding. The authors have not shown anywhere in the article that - The observed flooding was made more likely by antecedent soil moisture left by the preceding rainy season.
	\begin{answer}
		Reviewer 1 also pointed this out.
		We concur with the reviewer; see our comments to Reviewer 1 comment 4.
	\end{answer}
	\item Panel numbers in figures will be helpful.
	\begin{answer}
		Reviewer 2 also requested these and we will implement them.
	\end{answer}
\end{enumerate}

\subsection{Typos}
\begin{enumerate}
	\item  Line 71 - Change to - correspondence with the exit region of the low-level jets.
	\begin{answer}
		Agreed, correspondence with not correspondence to.
	\end{answer}
	\item Line 361 - change frequency to frequently.
	\begin{answer}
		Agreed
	\end{answer}
	\item Line 418 - Did you mean Figure 10 a?
	\begin{answer}
		Need to add panel numbers in figures and then will make sure we are referencing this correctly.
	\end{answer}
	\item Line 493 - coming from the Pacific
	\begin{answer}
		Yes, that is what we meant.
	\end{answer}
\end{enumerate}
\end{document}

### 2018-02-06 src/calculate_streamfunction.py

@@ -1,0 +1,54 @@
+#!/usr/bin/env python
+# -*- coding: utf-8 -*-
+"""Download a single year of 6-Hour Reanalysis V2 data from
+https://www.esrl.noaa.gov/psd/thredds/dodsC/Datasets/ncep.reanalysis2
+"""
+
+import argparse
+import os
+import xarray as xr
+import numpy as np
+import pandas as pd
+from windspharm.xarray import VectorWind
+
+parser = argparse.ArgumentParser() #pylint: disable=C0103
+parser.add_argument("--outfile", help="the filename of the data to save")
+parser.add_argument("--uwnd", help="path to zonal wind file")
+parser.add_argument("--vwnd", help="path to meridional wind file")
+
+def calculate_streamfunction(uwnd, vwnd):
+    """Calculate the Streamfunction
+    """
+    uwnd_ds = xr.open_dataarray(uwnd)
+    vwnd_ds = xr.open_dataarray(vwnd)
+    wind = VectorWind(uwnd_ds, vwnd_ds)
+    psi = wind.streamfunction()
+    return psi
+
+def to_daily(hourly):
+    """Convert data to daily time step
+    """
+    old_time = pd.to_datetime(hourly['time'].values)
+    new_time = old_time + pd.DateOffset(hours=12) # OFFSET 12 HOURS
+    hourly['time'] = new_time
+    daily = hourly.resample(time='1D').mean(dim='time')
+    return daily
+
+def main():
+    """Run everything
+    """
+    args = parser.parse_args()
+    uwnd = os.path.abspath(args.uwnd)
+    vwnd = os.path.abspath(args.vwnd)
+    outfile = os.path.abspath(args.outfile)
+    psi = calculate_streamfunction(uwnd, vwnd)
+    longitudes = psi['lon'].values
+    longitudes[np.where(longitudes > 180)] -= 360
+    psi['lon'].values = longitudes
+    psi = to_daily(psi)
+    if os.path.isfile(outfile):
+        os.remove(outfile)
+    psi.to_netcdf(outfile, format='NETCDF4', mode='w')
+
+if __name__ == "__main__":
+    main()

### 2018-02-06 src/download_cpc_year.py

@@ -1,0 +1,78 @@
+#!/usr/bin/env python
+# -*- coding: utf-8 -*-
+"""This script reads in, parses, downloads, and saves to file the data for
+a single year from the CPC Global Unified Precipitation Data Set from
+http://iridl.ldeo.columbia.edu/SOURCES/.NOAA/.NCEP/.CPC/.UNIFIED_PRCP/.GAUGE_BASED/.GLOBAL/
+"""
+
+import argparse
+import datetime
+import os
+import xarray as xr
+import numpy as np
+
+parser = argparse.ArgumentParser() #pylint: disable=C0103
+parser.add_argument("--outfile", help="the filename of the data to save")
+parser.add_argument("--year", help="the year of data to download")
+
+def convert_t_to_time(time_vec):
+    """Parse the times from the IRI Data Library
+    """
+    time = np.array([datetime.date(1960, 1, 1) + datetime.timedelta(np.int(ti)) for ti in time_vec])
+    return np.array(time)
+
+def convert_time_to_t(date):
+    """Turn a date into the appropriate IRI Data Library Format
+    """
+    date_diff = date - datetime.date(1960, 1, 1)
+    time_dl = date_diff.days
+    return time_dl
+
+def download_data(year, outfile):
+    """Download data for a single year
+    """
+    base_url = 'http://iridl.ldeo.columbia.edu/SOURCES/.NOAA/.NCEP/.CPC/'
+    base_url += '.UNIFIED_PRCP/.GAUGE_BASED/.GLOBAL/.v1p0'
+    if year >= 1979 and year <= 2005:
+        url = base_url + '/.RETRO/.rain/dods'
+    elif year >= 2006 and year <= 2016:
+        url = base_url + '/.REALTIME/.rain/dods'
+    else:
+        raise ValueError('You have entered an invalid year {}'.format(year))
+
+    # get the start and end times as IRIDL strings
+    dt_start = convert_time_to_t(datetime.date(year, 1, 1))
+    dt_end = convert_time_to_t(datetime.date(year, 12, 31))
+
+    # Read in the raw data, rename the variables
+    rain_year = xr.open_dataarray(url, decode_times=False)
+    rain_year = rain_year.sel(T=slice(dt_start, dt_end)).load()
+    rain_year = rain_year.rename({'X': 'lon', 'Y': 'lat', 'T': 'time'})
+
+    # convert the time data
+    rain_year['time'] = convert_t_to_time(rain_year['time'])
[…]

### 2018-02-06 src/download_s2s_area_avg.py

@@ -1,0 +1,63 @@
+#!/usr/bin/env python
+# -*- coding: utf-8 -*-
+"""Download a single year of 6-Hour Reanalysis V2 data from
+https://www.esrl.noaa.gov/psd/thredds/dodsC/Datasets/ncep.reanalysis2
+"""
+
+import argparse
+import os
+import xarray as xr
+import numpy as np
+import pandas as pd
+from datetime import datetime
+
+parser = argparse.ArgumentParser() #pylint: disable=C0103
+parser.add_argument("--outfile", help="the filename of the data to save")
+parser.add_argument("--year", help="the year of data to download")
+parser.add_argument("--X0", help="lon min")
+parser.add_argument("--X1", help="lon max")
+parser.add_argument("--Y0", help="lat min")
+parser.add_argument("--Y1", help="lat max")
+
+def download_data(outfile, year, lonmin, lonmax, latmin, latmax):
+    # Get the URL
+    url = 'http://iridl.ldeo.columbia.edu/home/.mbell/.ECMWF/.S2Stest/.S2S/.ECMF_ph2/'
+    url += '.forecast/.perturbed/.sfc_precip/.tp/'
+    url += 'X/{}/{}/RANGEEDGES/'.format(lonmin, lonmax)
+    url += 'Y/{}/{}/RANGEEDGES/'.format(latmin, latmax)
+    url += '[X+Y]average/'
+    url += 'L+differences/'
+    url += 'dods'
+
+    sdate = datetime(year, 9, 1)
+    edate = datetime(year+1, 2, 29)
+
+    # Load the data
+    data = xr.open_dataarray(url)
+    data = data.sel(S=slice(sdate, edate))
+    data['S'] = pd.to_datetime(data['S'].values)
+    data['L'] = data['L']
+    if os.path.isfile(outfile):
+        os.remove(outfile)
+    data.to_netcdf(outfile, format='NETCDF4', mode='w')
+
+def main():
+    """Parse the command line arguments and run download_data().
+    """
+    args = parser.parse_args()
+    outfile = os.path.abspath(args.outfile)
+    lonmin = args.X0
+    lonmax = args.X1
+    latmin = args.Y0
+    latmax = args.Y1
+    year = int(args.year)
+
+    download_data(
+        outfile=outfile,
+        lonmin=lonmin, lonmax=lonmax,
+        latmin=latmin, latmax=latmax,
+        year=year
+    )
+
+if __name__ == "__main__":
+    main()

### 2018-02-06 src/download_ssta.py

@@ -1,0 +1,47 @@
+#!/usr/bin/env python
+# -*- coding: utf-8 -*-
+"""Download monthly sea surface temperature anomalies from
+http://iridl.ldeo.columbia.edu/expert/SOURCES/.NOAA/.NCEP/.EMC/.CMB/.GLOBAL/.Reyn_SmithOIv2/.monthly/.ssta/dods
+"""
+
+import argparse
+import os
+import xarray as xr
+import numpy as np
+import pandas as pd
+
+parser = argparse.ArgumentParser() #pylint: disable=C0103
+parser.add_argument("--outfile", help="the filename of the data to save")
+
+def download_data(outfile):
+    """Download the monthly SST anomaly data
+    Args:
+        outfile: the name of the file to save to
+    """
+    # read in the data
+    url = 'http://iridl.ldeo.columbia.edu/expert/SOURCES/.NOAA/.NCEP/.EMC/.CMB/.GLOBAL/'
+    url += '.Reyn_SmithOIv2/.monthly/.ssta/dods' # split lines
+    data = xr.open_dataarray(url, decode_times=False) # doesn't follow time conventions
+
+    # parse the time manually
+    time = np.int_(np.floor(data['T']))
+    year = 1960 + time // 12
+    month = 1 + time % 12
+    time = pd.to_datetime(year * 10000 + month * 100 + 1, format="%Y%m%d")
+    data['T'] = time
+    data.rename({'T': 'time'})
+
+    # save to file
+    if os.path.isfile(outfile):
+        os.remove(outfile)
+    data.to_netcdf(outfile, format='NETCDF4', mode='w')
+
+def main():
+    """Parse the command line arguments and run download_data().
+    """
+    args = parser.parse_args()
+    outfile = os.path.abspath(args.outfile)
+    download_data(outfile=outfile)
+
+if __name__ == "__main__":
+    main()

### 2018-02-06 src/make_anomaly.py

@@ -1,0 +1,94 @@
+#!/usr/bin/env python
+# -*- coding: utf-8 -*-
+"""Take a list of data files, subset them, and calculate anomalies
+"""
+
+import argparse
+import os
+from glob import glob
+import calendar
+import xarray as xr
+import numpy as np
+import pandas as pd
+
+parser = argparse.ArgumentParser() #pylint: disable=C0103
+parser.add_argument("--outfile", help="the filename of the data to save")
+parser.add_argument("--path", help="the files to read in")
+parser.add_argument("--syear", type=int, help="start year")
+parser.add_argument("--eyear", type=int, help="end year")
+parser.add_argument("--X0", type=float, help="lon min")
+parser.add_argument("--X1", type=float, help="lon max")
+parser.add_argument("--Y0", type=float, help="lat min")
+parser.add_argument("--Y1", type=float, help="lat max")
+
+def read_netcdfs(files, dim, transform_func=None):
+    """From http://xarray.pydata.org/en/stable/io.html#netcdf
+    """
+    def process_one_path(path):
+        """Helper
+        """
+        with xr.open_dataset(path) as ds:
+            if transform_func is not None:
+                ds = transform_func(ds)
+            ds.load()
+            return ds
+
+    paths = sorted(glob(files))
+    datasets = [process_one_path(p) for p in paths]
+    combined = xr.concat(datasets, dim)
+    return combined
+
+def calc_anomaly(path, outfile, syear, eyear, lonmin, lonmax, latmin, latmax):
+    """Decompose into anomaly and subset geographically
+    """
+    # Read in the data
+    def subset_function(ds):
+        # HARD-CODED FOR NOW: subset NDJF, calculate "adjusted year"
+        sub = ds.sortby('lon').sortby('lat').sortby('time').sel(
+            lon=slice(lonmin, lonmax),
+            lat=slice(latmin, latmax)
+        )
+        sub = sub.sel(time=np.in1d(sub['time.month'], np.array([11, 12, 1, 2])))
+        return sub
+
+    raw = read_netcdfs(path, dim='time', transform_func=subset_function)
+    varname = list(raw.data_vars.keys())[0]
+    raw = raw[varname]
+    year = raw['time.year']
+    month = raw['time.month']
+    year_adj = year
+    year_adj[np.in1d(month, [11, 12])] += 1
[…]

### 2018-02-06 src/pyfloods/weather_type.py

@@ -1,80 +1,0 @@
-import xarray as xr # for reading in netcdf files
-import numpy as np # for calculations
-import sklearn.cluster as cl # k-means implemetation
-from eofs.xarray import Eof # EOF implementation that recognizes xarray
-import pandas as pd # time series
-
-def ReSort(x):
-        '''
-        A helper function to re-sort the cluster labels.
-        Takes in x, a vector of cluster labels, and re-orders them so that
-        the lowest number is the most common, and the highest number
-        the least common
-        '''
-        import numpy as np
-        x = np.int_(x)
-        x_from = np.unique(x).argsort()
-        counts = np.array([np.sum(x == xi) for xi in x_from])
-        orders = (-counts).argsort().argsort()
-        return(orders[x_from[x]])
-
-def Classifiability(P, Q):
-    '''
-    The Auto-Correlation function
-    '''
-    from scipy.stats import pearsonr as correl
-    k = P.shape[0]
-    assert Q.shape[0] is k, "P and Q must be same k"
-    Aij = np.ones([k, k])
-    for i in range(k):
-        for j in range(k):
-            Aij[i,j], _ = correl(P[i,:], Q[j,:])
-    Aprime = Aij.max(axis=0)
-    classifiability = Aprime.min()
-    return(classifiability)
-
-def XrEofCluster(ds, n_clusters=5, prop=0.90, nsim=100, verbose = True, pcscaling=0):
-    # DS should be an xarray data array (not dataset!)
-    if verbose:
-        print('xarray-based classifiability for {} clusters'.format(n_clusters))
-        print('Performing EOF decomposition of data for dimension reduction.')
-
-    # EOF Decomposition
-    solver = Eof(ds, center=True)
-    var_frac = solver.varianceFraction()
-    cumvar = np.cumsum(var_frac.values)
-    eofs_keep = np.where(cumvar >= prop)[0].min() + 1
-    pc = solver.pcs(npcs=eofs_keep, pcscaling=pcscaling) # time series of PC
-    if verbose:
-        print("Number of EOFs retained is {}".format(eofs_keep))
-        print('Carrying out {} k-means simulations...'.format(nsim))
-
-    centroids = np.ones([nsim, n_clusters, pc.shape[1]])
-    clusters = np.ones([nsim, pc.shape[0]])
-    for i in range(0, nsim):
[…]

### 2018-02-06 src_old/32-weather-type.py

@@ -25,7 +25,7 @@ pars = {
 ################################################################################
 # Calculate the weather types
 ################################################################################
-recalculate = False
+recalculate = True
 try:
     cluster_ts = xr.open_dataarray(paths.file_weather_types)
     data_pars = {
@@ -49,17 +49,19 @@ if recalculate:
     psi = psi.sel(time = np.in1d(psi['time.month'], [11, 12, 1, 2]))
     psi = psi.sel(time = slice('1979-11-01', '2016-02-29'))
     np.random.seed(22591)
-    best_centroid, cluster_ts, classifiability = weather_type.XrEofCluster(
-        ds=psi['anomaly'],
+    best_centroid, cluster_ts, classifiability = weather_type.cluster_xr_eof(
+        data_array=psi['anomaly'],
         n_clusters=pars.get('n_clusters'),
         prop=pars.get('wt_prop'),
         nsim=pars.get('nsim'),
         pcscaling=pars.get('pcscaling'),
-        verbose = True
+        verbose=True
     )
     print("Classifiability Index: {}".format(classifiability))
     cluster_ts.attrs.update(pars)
     cluster_ts.attrs.update({'classifiability': classifiability})
+    if os.path.isfile(paths.file_weather_types):
+        os.remove(paths.file_weather_types)
     cluster_ts.to_netcdf(paths.file_weather_types, format='NETCDF4')
 
 wt = cluster_ts.to_dataframe(name='wtype')

### 2018-02-07 src/analyze/print_wt_prop_table.py
get proportion of weather types each year

@@ -1,0 +1,48 @@
+#!/usr/bin/env python
+# -*- coding: utf-8 -*-
+"""Get proportion of weather types each year and save a latex table
+"""
+
+import argparse
+import xarray as xr
+import numpy as np
+import pandas as pd
+import matplotlib.pyplot as plt
+
+parser = argparse.ArgumentParser() #pylint: disable=C0103
+parser.add_argument("--outfile", help="the filename of the data to save")
+parser.add_argument("--wt", help="The Weather Type sequence")
+
+def get_year_adj(time):
+    """Given a datetime index, return the "year_adj" variable.
+
+    This is the year that the NDJF season ended, so the period
+    Nov 1979 to Feb 1980 has a "year_adj" of 1980.
+    """
+    month = time.month.values
+    year_adj = time.year.values
+    year_adj[np.in1d(month, [11, 12])] += 1
+    return year_adj
+
+def main():
+    """Get the proportion of weather types and save a latex table
+    """
+    args = parser.parse_args()
+    w_type = xr.open_dataarray(args.wt).to_dataframe()
+    w_type['year_adj'] = get_year_adj(w_type.index)
+
+    # Get prop of weather types by year
+    wt_prop = w_type.groupby(['year_adj', 'wtype']).size().to_xarray()
+    wt_prop = wt_prop / wt_prop.sum(dim='wtype')
+    wt_prop = wt_prop.to_pandas()
+
+    # Get NDJF 2015-16 and NDJF Climatology
+    wt_prop_1516 = wt_prop.loc[2016]
+    wt_prop_clim = wt_prop.mean()
+    wt_prop_tex = pd.DataFrame({'climatology': wt_prop_clim, 'NDJF 2015-16': wt_prop_1516})
+
+    # Save to latex table
+    wt_prop_tex.to_latex(args.outfile)
+
+if __name__ == "__main__":
+    main()

### 2018-02-27 src/analyze/plot_anom_1516.py
use new code base

@@ -90,8 +90,8 @@ def main():
     cbar1.set_label('Precip. Anomaly [mm/d]', rotation=270)
     cbar1.ax.get_yaxis().labelpad = 20
 
-    for i,ax in enumerate(axes.flat):
-        ax.set_title(string.ascii_lowercase[i])
+    #for i,ax in enumerate(axes.flat):
+    #    ax.set_title(string.ascii_lowercase[i])
 
     fig.savefig(args.outfile, bbox_inches='tight')
 

### 2018-02-27 src/analyze/plot_classifiability.py

@@ -1,0 +1,38 @@
+#!/usr/bin/env python
+# -*- coding: utf-8 -*-
+"""Plot classifiability as a function of # weather types
+"""
+
+import argparse
+import xarray as xr
+import numpy as np
+import pandas as pd
+import matplotlib.pyplot as plt
+from glob import glob
+
+parser = argparse.ArgumentParser() #pylint: disable=C0103
+parser.add_argument("--outfile", help="the filename of the data to save")
+parser.add_argument("--infiles", help="The classifiability indices")
+
+def main():
+    """Run everything
+    """
+    args = parser.parse_args()
+    
+    fnames = glob(args.infiles)
+    arrays = [xr.open_dataarray(fn) for fn in fnames]
+    classifiability = np.array([ds.attrs['class_idx'] for ds in arrays])
+    n_wtype = np.array([np.unique(ds.values).size for ds in arrays])
+
+    df = pd.DataFrame({'classifiability': classifiability, 'n_wtype': n_wtype}).sort_values('n_wtype').set_index('n_wtype')
+
+    # Plot
+    df['classifiability'].plot(figsize=(15, 4.5), grid=True)
+    plt.xlabel('Number of Weather Types')
+    plt.ylabel('Classifiability Index')
+
+    # Save the plot
+    plt.savefig(args.outfile, bbox_inches='tight')
+
+if __name__ == "__main__":
+    main()

### 2018-02-27 src/analyze/plot_sst_anomalies.py

@@ -36,7 +36,7 @@ def main():
     wt = wt.sort_values(4, ascending=False).head()
 
     # Plot options
-    n_months = 15
+    n_months = 10
     map_proj = ccrs.PlateCarree(central_longitude=-60)
     data_proj = ccrs.PlateCarree()
     figsize = (10, 8)
@@ -44,7 +44,8 @@ def main():
     cmap = cc.cm['coolwarm']
 
     # Times to plot
-    times = wt.sort_values(4, ascending=False).head(n_months).index.to_timestamp()
+    wt['SALLJ'] = wt[4] + wt[1]
+    times = wt.sort_values('SALLJ', ascending=False).head(n_months).index.to_timestamp()
     composite_wt4 = sea_temp.sel(T = np.in1d(sea_temp['T'], times)).mean(dim='T')
 
     # Set up 2 plots
@@ -58,20 +59,20 @@ def main():
     C1 = composite_wt4.plot.contourf(
         ax=ax, transform=data_proj,
         cmap=cmap,
-        levels=np.linspace(-0.6, 0.6, 13),
+        levels=np.linspace(-1.25, 1.25, 13),
         extend='both',
         add_colorbar=True,
         add_labels=False
     )
     ax.add_patch(dipole.as_patch(color='black'))
-
+    ax.set_title('{} Months with Most No-Chaco Jet Events'.format(n_months))
 
     # Second: Plot December 2015
     ax = axes[1]
     C2 = sea_temp.sel(T = '2015-12-01').plot.contourf(
         ax=ax, transform=data_proj,
         cmap=cmap,
-        levels=np.linspace(-2.5, 2.5, 11),
+        levels=np.linspace(-2, 2, 11),
         extend='both',
         add_colorbar=True,
         add_labels=False
@@ -79,6 +80,7 @@ def main():
     ax.add_patch(dipole.as_patch(color='black'))
     ax.set_xlim([-120, 120])
     ax.set_ylim([-70, 15])
+    ax.set_title('December 2015')
 
     # Format axes
     viz.format_axes(

### 2018-02-27 src/analyze/plot_wt_composite.py

@@ -20,6 +20,10 @@ parser.add_argument("--outfile", help="the filename of the data to save")
 parser.add_argument("--wt", help="The weather type data")
 parser.add_argument("--psi", help="The streamfunction data file")
 parser.add_argument("--rain", help="The rainfall data file")
+parser.add_argument("--WTX0", type=float)
+parser.add_argument("--WTX1", type=float)
+parser.add_argument("--WTY0", type=float)
+parser.add_argument("--WTY1", type=float)
 
 def main():
     """Run everything
@@ -28,6 +32,8 @@ def main():
     psi = xr.open_dataset(args.psi)
     prcp = xr.open_dataset(args.rain)
     w_type = xr.open_dataarray(args.wt).to_dataframe(name='wtype')
+
+    wt_region = Region(lon=[args.WTX0, args.WTX1], lat=[args.WTY0, args.WTY1])
 
     # Plot Options
     map_proj = ccrs.PlateCarree()#Orthographic(-60, -10)
@@ -64,7 +70,8 @@ def main():
             add_colorbar=False,
             add_labels=False
         )
-        #ax.add_patch(region.wtype.as_patch(color='black'))
+        ax.add_patch(wt_region.as_patch(label='Weather Typing Region', color='black'))
+
         # Bottom row: rainfall anomalies
         ax = axes[1, i]
         C1 = selector(prcp['anomaly']).plot.contourf(

### 2018-02-27 src/process/make_weather_type.py

@@ -5,6 +5,7 @@
 
 import argparse
 import os
+from collections import OrderedDict
 import xarray as xr
 import numpy as np
 import pandas as pd
@@ -88,13 +89,12 @@ def resort_labels(old_labels):
 def main():
     """Parse the command line arguments and run download_data().
     """
-    np.random.seed(2061834) # set seed from
+    np.random.seed(1085) # set seed from
     args = parser.parse_args()
     psi = xr.open_dataset(args.infile)['anomaly']
 
     # carry out PCA
     psi_stacked = psi.stack(grid=['lon', 'lat'])
-    #psi_stacked = StandardScaler().fit_transform(psi_stacked.values) # Standardize each grid cell
     pca = PCA().fit(psi_stacked)
     cum_var = pca.explained_variance_ratio_.cumsum()
     n_components_keep = np.where(cum_var > args.var_xpl)[0].min() + 1 #compensate for zero indexing
@@ -111,6 +111,7 @@ def main():
     best_wt = wtypes[best_part, :]
     best_wt = pd.Series(resort_labels(best_wt), index=psi['time']).to_xarray()
     best_wt.name = 'wtype'
+    best_wt.attrs = OrderedDict(class_idx = class_idx)
 
     if os.path.isfile(args.outfile):
         os.remove(args.outfile)

### 2018-02-27 src_old/02-derived-data.py

@@ -1,6 +1,0 @@
-from pyfloods.streamfunction import psi
-from pyfloods.anomalies import rainfall,streamfunction
-
-psi.get_data()
-streamfunction.get_data()
-rainfall.get_data()

### 2018-02-27 src_old/11-study-area.py

@@ -1,75 +1,0 @@
-import xarray as xr
-import numpy as np
-import matplotlib.pyplot as plt
-import cartopy.crs as ccrs
-import cartopy.feature
-import colorcet as cc
-import pandas as pd
-from cartopy.io import shapereader
-import matplotlib.patches as patches
-import os.path
-
-from pyfloods import region,paths,raw_data
-import pyfloods.visualize as viz
-
-# Figure Options
-map_proj = ccrs.PlateCarree()
-data_proj = ccrs.PlateCarree()
-elev_cmap = cc.cm['bgyw']
-
-rivers = shapereader.natural_earth(
-    category='physical',
-    name='rivers_lake_centerlines',
-    resolution='10m'
-)
-rivers = shapereader.Reader(rivers)
-
-# Station Descriptions
-desc = pd.read_csv(paths.file_station_description)
-
-# Elevation
-elev = raw_data.topo.get_data().sel(X=slice(-85, -32.5), Y=slice(-50,15))
-elev = np.log10(elev)
-elev_min = -0.3
-elev_max = 3.7
-contours = np.linspace(elev_min, elev_max, 21)
-levels = np.linspace(elev_min, elev_max, 41)
-
-# Create the plot!
-fig, axes = plt.subplots(nrows=1, ncols=2, subplot_kw={'projection': map_proj}, figsize=[10, 7.5])
-ax = axes[0]
-C = elev.plot.contourf(ax=ax, transform=data_proj, add_colorbar=False, add_labels=False, levels=levels, cmap=elev_cmap)
-ax.add_patch(region.wtype.as_patch(label='Weather Typing Region', color='blue'))
-ax.add_patch(region.lower_py_river.as_patch(label='Lower PY River Basin', color='red'))
-ax.legend(loc='lower right')
-
-ax = axes[1]
-C = elev.plot.contourf(ax=ax, transform=data_proj, add_colorbar=False, add_labels=False, levels=levels, cmap=elev_cmap)
-elev.plot.contour(ax=ax, transform=data_proj, add_colorbar=False, add_labels=False, colors='k', levels=contours, alpha=0.3, linewidths=0.4)
-ax.add_patch(region.lower_py_river.as_patch(label='Lower PY River Basin', color='red'))
-
-rivernum_plot = [36, 1032, 1125, 294]
-for rec in rivers.records():
[…]

### 2018-02-27 src_old/22-NDJF-1516-MJO.py

@@ -1,28 +1,0 @@
-import numpy as np
-import matplotlib.pyplot as plt
-import os
-import seaborn as sns
-
-from pyfloods import region,paths,raw_data
-
-
-mjo = raw_data.mjo.get_data()['2015-11-01':'2016-02-29']
-mjo['month'] = mjo.index.month
-
-# MAKE THE FIGURE
-plt.figure(figsize=(7,5))
-sns.lmplot(
-    x='RMM1', y='RMM2', data=mjo,
-    fit_reg=False, hue='month',
-    scatter_kws={"marker": "D", "s": 100}, size = 8
-)
-plt.plot(mjo['RMM1'], mjo['RMM2'], color='gray')
-plt.axvline(0, color='gray')
-plt.axhline(0, color='gray')
-plt.plot((100,-100), (100, -100), color='gray')
-plt.plot((-100,100), (100, -100), color='gray')
-plt.ylim(-3,3)
-plt.xlim(-3,3)
-plt.xlabel('RMM1')
-plt.ylabel('RMM2')
-plt.savefig(os.path.join(paths.figures, 'mjo-time-series.pdf'), bbox_inches='tight')

### 2018-02-27 src_old/33-EOF-plots.py

@@ -1,58 +1,0 @@
-import numpy as np
-import matplotlib.pyplot as plt
-import cartopy.crs as ccrs
-from eofs.xarray import Eof
-import os
-
-from pyfloods.anomalies import streamfunction
-from pyfloods.region import wtype
-import pyfloods.paths
-import pyfloods.visualize as viz
-
-psi = streamfunction.get_data().sel(
-    lon = slice(wtype.lonmin, wtype.lonmax),
-    lat = slice(wtype.latmin, wtype.latmax))['anomaly']
-
-solver = Eof(psi, center=True)
-var_frac = solver.varianceFraction()
-cumvar = np.cumsum(var_frac.values)
-
-n_eof = np.arange(cumvar.size) + 1
-
-# PLOT VARIANCE EXPLAINED
-fig,axes = plt.subplots(nrows=1, ncols=2, sharey=False, figsize=(12, 4))
-
-ax = axes[0]
-ax.plot(n_eof, var_frac)
-ax.scatter(n_eof, var_frac)
-ax.set_ylabel('Variance Explained')
-ax.grid()
-
-ax = axes[1]
-plt.plot(n_eof, cumvar)
-plt.scatter(n_eof, cumvar)
-ax.set_ylabel('Cumulative Variance')
-ax.grid()
-for ax in axes.flat:
-    ax.set_xlim([0.8, 9.2])
-
-fig.tight_layout()
-fig.savefig(os.path.join(pyfloods.paths.figures, 'eof-variance-explained.pdf'), bbox_inches='tight')
-
-
-# plot the EOF loadings
-loading = solver.eofs(neofs=4)
-loading['mode'] += 1
-
-aspect = (loading.lon.max() - loading.lon.min()) / (loading.lat.max() - loading.lat.min())
-p = loading.plot.contourf(
-    transform=ccrs.PlateCarree(), col='mode', #figsize=(16, 3.5),
-    aspect=aspect,
-    subplot_kws={'projection': ccrs.PlateCarree()},
-    cmap = "PuOr", levels=np.linspace(-0.3, 0.3, 13))
-viz.FormatAxes(p.axes, coast=True, grid=False, border=True,
-               ticks=[np.linspace(-180, 180, 91), np.linspace(-90, 90, 37)],
-               extent=pyfloods.region.wtype.as_extent())
-for i,ax in enumerate(p.axes.flat):
-    ax.set_title('EOF {}'.format(i+1))
[…]

### 2018-02-27 src_old/34-composite-sst-WT4.py

@@ -1,47 +1,0 @@
-import pandas as pd
-import xarray as xr
-import numpy as np
-import matplotlib.pyplot as plt
-from matplotlib import patches
-import cartopy.crs as ccrs
-import cartopy.feature
-import os
-
-import pyfloods.visualize as viz
-from pyfloods import paths,region
-from pyfloods.raw_data import sst
-
-savefigs = True
-map_proj = ccrs.PlateCarree(central_longitude=-100)
-data_proj = ccrs.PlateCarree()
-
-wt = xr.open_dataarray(paths.file_weather_types).to_dataframe(name='wtype')
-wt['T'] = wt.index.to_period("M")
-wt = wt.groupby(['T', 'wtype']).size().unstack()
-wt.sort_values(4, ascending=False).head()
-
-sea_temp = sst.get_data()
-
-n_months = 15
-times = wt.sort_values(4, ascending=False).head(n_months).index.to_timestamp()
-composite_wt4 = sea_temp.sel(T = np.in1d(sea_temp['T'], times)).mean(dim='T')
-
-dipole = region.Region(lon=[-30,-10], lat=[-20,-50])
-
-# MAKE PLOT
-fig, ax = plt.subplots(nrows=1, ncols=1, subplot_kw={'projection': map_proj},
-                       figsize=(16, 5))
-composite_wt4.plot.contourf(
-    ax=ax, transform=data_proj,
-    levels=np.linspace(-0.6, 0.6, 13),
-    cmap='PuOr_r',
-    cbar_kwargs={'label': 'SST Anomaly [Degrees C]'}, extend='both'
-)
-ax.add_patch(dipole.as_patch(color='black'))
-viz.FormatAxes(
-    ax, coast=True, grid=False, border=True, river = False,
-    ticks=[np.linspace(-180, 180, 19), np.linspace(-90, 90, 19)]
-)
-ax.set_xlim([-120, 120])
-ax.set_ylim([-70, 15])
-fig.savefig(os.path.join(paths.figures, 'ssta-composite-wt4.pdf'), bbox_inches='tight')

### 2018-02-27 src_old/41-chiclet-s2s-area-avgeraged.py

@@ -1,60 +1,0 @@
-import xarray as xr
-import cartopy.crs as ccrs
-import matplotlib.pyplot as plt
-import numpy as np
-import pandas as pd
-import datetime
-import os.path
-
-from pyfloods import visualize as viz
-from pyfloods import paths,region
-from pyfloods.raw_data import s2s
-from pyfloods.anomalies import rainfall
-from pyfloods.region import lower_py_river as lpr
-
-modeled = s2s.get_data()
-modeled['S'] = pd.to_datetime(modeled['S'].values).date
-modeled['S'] = pd.to_datetime(modeled['S'].values)
-modeled['L'] = np.int_(modeled['L'].astype('timedelta64[D]') / np.timedelta64(1, 'D'))
-modeled = modeled.to_dataframe().dropna()
-modeled = modeled.reset_index()
-modeled['target_date'] = modeled['S'] + modeled['L'].apply(np.ceil).apply(lambda x: pd.Timedelta(x, unit='D'))
-modeled = modeled.loc[np.logical_and(modeled['target_date'] >= '2015-11-01', modeled['target_date'] <= '2016-02-29')]
-modeled = modeled.set_index(['target_date', 'L', 'M'])
-
-obs = rainfall.get_data()['raw'].sel(
-    lon = slice(lpr.lonmin, lpr.lonmax),
-    lat = slice(lpr.latmin, lpr.latmax)).mean(dim=['lon', 'lat']).to_pandas()
-climatology = np.mean(obs)
-
-modeled['tpa'] = modeled['tp'] - climatology
-forecast = modeled.to_xarray()['tpa'].mean(dim='M')
-ft, fl = np.meshgrid(np.arange(forecast.target_date.size), forecast.L.values)
-target_date = pd.to_datetime(forecast.target_date.values)
-obs = obs['2015-11-01':'2016-02-29']
-
-# MAKE PLOT
-fig,axes = plt.subplots(ncols=1, nrows=2, figsize=((10, 5)), gridspec_kw = {'height_ratios':[3, 1]}, sharex=True)
-
-ax=axes[0] # Chiclet
-C1 = forecast.plot.pcolormesh(x='target_date', y='L', cmap = 'BrBG', ax=ax, add_colorbar=False, add_labels=False)
-ax.set_ylabel("Lead Time (Days)")
-ax.set_xticks([])
[…]

### 2018-02-27 src_old/42-seasonal-forecasts.py

@@ -1,45 +1,0 @@
-import xarray as xr
-import pandas as pd
-import cartopy.crs as ccrs
-import matplotlib.pyplot as plt
-import numpy as np
-import os
-
-import pyfloods.visualize as viz
-import pyfloods.paths
-from pyfloods.anomalies import rainfall
-
-df = pd.read_csv(
-    os.path.join(pyfloods.paths.data_raw, 'SeasonalForecast.tsv'),
-    skiprows=2, delim_whitespace=True,
-    names=['lon', 'lat', 'prob_exceed'],
-    index_col=['lat', 'lon'])['prob_exceed'].to_xarray()
-
-prcp = rainfall.get_data().sel(
-    lon=slice(df.lon.min(), df.lon.max()),
-    lat=slice(df.lat.min(), df.lat.max()))['raw']
-prcp = prcp.sel(time = prcp['time.season'] == 'DJF')
-prcp_year = prcp.groupby('year_adj').mean(dim='time')
-prcp_q90 = prcp_year.quantile(0.90, dim='year_adj')
-prcp_exceed = prcp_year.sel(year_adj=2016) > prcp_q90
-
-X,Y = np.meshgrid(df.lon, df.lat)
-Xh,Yh = np.meshgrid(prcp_exceed.lon, prcp_exceed.lat)
-hatch = np.ma.masked_invalid(prcp_exceed.values)
-hatch = np.ma.masked_less(hatch, True)
-
-# MAKE PLOT
-fig, ax = viz.SetupAxes(figsize=(5.5,8), ncol=1, nax=1,proj=ccrs.PlateCarree())
-fig.subplots_adjust(right=0.925)
-cax1 = fig.add_axes([0.95, 0.15, 0.035, 0.7])
-sub = np.ma.masked_invalid((df/(1-df)) / (.1/(1-.1)))
-C = ax.pcolormesh(X,Y, sub, transform=ccrs.PlateCarree(),
-                  cmap="terrain_r", vmin=0, vmax=7)
-ax.pcolor(Xh, Yh, hatch, facecolor='none', edgecolors='k')
-cb = plt.colorbar(C, cax=cax1)
-cb.set_label("Odds Ratio", rotation=270)
-cb.ax.get_yaxis().labelpad = 15
-viz.FormatAxes(ax, ticks=[np.linspace(-180, 180, 73), np.linspace(-90, 90, 37)],
[…]

### 2018-02-27 src_old/43-s2s-MOS-forecasts.py

@@ -1,96 +1,0 @@
-import xarray as xr
-import cartopy.crs as ccrs
-import matplotlib.pyplot as plt
-import numpy as np
-import pandas as pd
-import datetime
-import os
-
-from pyfloods import visualize as viz # custom plotting library
-from pyfloods import paths
-
-obs_url = 'http://iridl.ldeo.columbia.edu/home/.agmunoz/.Paraguay/.Obs/.Obsrain_90thpctl/dods'
-obs_exceed = xr.open_dataarray(obs_url)
-obs_exceed = obs_exceed.isel(T=(obs_exceed['T'].size - 1))
-hatch0 = np.ma.masked_invalid(obs_exceed.values)
-hatch = np.ma.masked_less(hatch0, 0.9)
-Xh, Yh = np.meshgrid(obs_exceed.X, obs_exceed.Y)
-
-model_names = ['Raw', 'XLR', 'HXLR', 'PCR', 'CCA']
-
-def get_forecast(model, type):
-    base_url = 'http://iridl.ldeo.columbia.edu/ds%3A/home/.agmunoz/.Paraguay/'
-    url = base_url + '{}/'.format(model)
-    if type == 'forecast':
-        url += '.Odds_1-7Dec2015/dods'
-    elif type == 'skill':
-        url += '.2AFCProbFcsts/dods'
-    elif type == 'ignorance':
-        url += '.IGNProbFcsts/dods'
-    else:
-        raise ValueError('type is not valid')
-    ds = xr.open_dataarray(url)
-    ds = ds.sortby(['X', 'Y']) # be more consistent
-    if ds.ndim > 2:
-        ds = ds.isel(time=0).drop('time')
-    ds.coords['model'] = model
-    return(ds)
-
-ncols = len(model_names)
-fig, axes = plt.subplots(
-    ncols=len(model_names), nrows=3,
-    subplot_kw={'projection': ccrs.PlateCarree()},
-    figsize=(11,8), sharex=True, sharey=True
-)
-for i,m in enumerate(model_names):
-    ax = axes[0, i] # Forecast for December 1-7 2015
-    ax.set_title(m)
-    ds = get_forecast(model=m, type='forecast')
-    X,Y = np.meshgrid(ds.X, ds.Y)
-    sub = ds.values
[…]

### 2018-02-27 src_old/51-wt-mjo-enso.py

@@ -1,58 +1,0 @@
-import xarray as xr
-import matplotlib.pyplot as plt
-import numpy as np
-import pandas as pd
-import os
-
-import pyfloods.visualize as viz
-from pyfloods import weather_type
-from pyfloods import paths,region
-from pyfloods import raw_data
-
-wt = xr.open_dataarray(paths.file_weather_types).to_dataframe(name='wtype')
-mjo = raw_data.mjo.get_data()
-enso = raw_data.enso.get_data().resample('1D').ffill()
-mrg = wt.join(mjo).join(enso)
-mrg.loc[mrg['amplitude'] < 1, 'phase'] =  0
-mrg['enso_state'] = 0
-mrg.loc[mrg['nino_34'] < -1, 'enso_state'] = -1
-mrg.loc[mrg['nino_34'] > 1, 'enso_state'] = 1
-mrg.head()
-
-
-conditional = mrg.groupby(['phase', 'enso_state'])['wtype'].apply(lambda g: g.value_counts() / g.size)
-conditional = conditional.to_xarray()
-anomaly = conditional - conditional.mean(dim=['level_2'])
-
-# Bootstrap simulations
-n_sim = 3000
-anomalies = []
-for i in np.arange(n_sim):
-    wt_seq = np.random.choice(mrg['wtype'], size=mrg.shape[0], replace=True)
-    simulated = pd.DataFrame({'wtype': wt_seq, 'phase': mrg['phase'], 'enso_state': mrg['enso_state']})
-    conditional = simulated.groupby(['phase', 'enso_state'])['wtype'].apply(lambda g: g.value_counts() / g.size)
-    conditional = conditional.to_xarray()
-    anomalies.append(conditional - conditional.mean(dim=['level_2']))
-anomalies = xr.concat(anomalies, dim='temp')
-
-# Select significant simulations
-alpha = 0.10
-upper = anomalies.quantile(1-alpha/2, dim='temp')
-lower = anomalies.quantile(alpha/2, dim='temp')
-significant = np.logical_or((anomaly >= upper), (anomaly <= lower))
-
-
-# Plot
-enso_states = {-1: 'La Niña', 0: 'Neutral', 1: 'El Niño'}
-p = anomaly.where(significant).plot(
-    x='level_2', y='phase', col='enso_state',
-    cmap='PuOr',
-    figsize=(12, 5)
-)
-for i,es in enumerate(enso_states):
[…]

### 2018-02-27 src_old/pyfloods/dataset/__init__.py

@@ -1,60 +1,0 @@
-"""
-This submodule defines the abstraction of a dataset and can work generically
-for any dataset whether it has one or more files, regardless of the
-type of file that it describes.
-In general, the dataset has the following elements:
-* a set of parameters that determine what the dataset contains
-* a filename or set of filenames that describe where the data is stored
-* a way to access the data from some remote source and save it to file
-* a way to read the data, including its parameters, in from files
-* a way to check whether the data that was read from file has the desired parameters
-"""
-
-from typing import Tuple,Dict,Any
-import os
-
-class SingleFileDataSet():
-    def __init__(self, fname: str, verbose: bool = False,
-                params: Dict[str, Any] = {}) -> None:
-        self.params = params
-        self.fname = fname
-        self.verbose = verbose
-
-    def download_data(self):
-        raise NotImplementedError
-
-    def read_data(self) -> Tuple[Any, Dict]:
-        raise NotImplementedError
-
-    def get_data(self):
-        """Try to read in the data, and if it has changed or if it cannot be
-        read in, then download the data.
-        """
-        need_download: bool = False
-        try:
-            data,data_params = self.read_data()
-            if self.params != data_params:
-                os.remove(self.fname)
-                need_download = 1
-                if self.verbose:
-                    print('Data was read from file but had different parameters')
-            if self.verbose:
-                print('Successfully read data from file')
-        except:
-            if self.verbose:
-                print('Unable to read data from file')
-            need_download = True
-        if need_download:
-            if self.verbose:
-                print('Downloading data from source')
-            data = self.download_data()
-        return(data)
-
-# Make it so we don't have to remember the file names when we do imports later
-from .cpc_rain import CPC
-from .elevation import NOAANGDC
-from .mjo_bom import MJOBOMD
-from .nino_34 import KaplanNINO34
-from .reanalysisv2 import ReanalysisV2
-from .s2s_area_avg import S2SAreaAvg
-from .sst import CMBGLOBAL

### 2018-02-27 src_old/pyfloods/dataset/cpc_rain.py

@@ -1,79 +1,0 @@
-"""
-This implements a single variable and single level of a reanalysis
-V2 variable
-https://www.esrl.noaa.gov/psd/thredds/dodsC/Datasets/ncep.reanalysis2
-"""
-import xarray as xr
-import numpy as np
-import datetime
-from typing import Dict,Tuple,List
-
-from pyfloods.dataset import SingleFileDataSet
-
-class CPCYear(SingleFileDataSet):
-    """This implements just a single year!
-    """
-
-    def __init__(self, fname:str, verbose: bool = False, params: Dict ={}) -> None:
-        self.year = params.get('year')
-        super().__init__(fname=fname, params=params, verbose=verbose)
-
-    def download_data(self) -> xr.DataArray:
-        def convert_t_to_time(Tvec) -> np.ndarray:
-            times = np.array([datetime.date(1960, 1, 1) + datetime.timedelta(np.int(ti)) for ti in Tvec])
-            return(np.array(times))
-
-        def convert_time_to_t(date)  -> np.ndarray:
-            date_diff = date - datetime.date(1960,1,1)
-            T = date_diff.days
-            return(T)
-
-        if self.year >= 1979 and self.year <= 2005:
-            url = 'http://iridl.ldeo.columbia.edu/SOURCES/.NOAA/.NCEP/.CPC/.UNIFIED_PRCP/.GAUGE_BASED/.GLOBAL/.v1p0/.RETRO/.rain/dods'
-        elif self.year >= 2006 and self.year <= 2016:
-            url = 'http://iridl.ldeo.columbia.edu/SOURCES/.NOAA/.NCEP/.CPC/.UNIFIED_PRCP/.GAUGE_BASED/.GLOBAL/.v1p0/.REALTIME/.rain/dods'
-        else:
-            raise ValueError('You have entered an invalid year. {} is outside range [1979, 2016]'.format(self.year))
-
-        Tstart = convert_time_to_t(datetime.date(self.year, 1, 1))
-        Tend = convert_time_to_t(datetime.date(self.year, 12, 31))
-
-        ds = xr.open_dataarray(url, decode_times=False)
-        ds = ds.sel(T = slice(Tstart, Tend)).load()
-        ds = ds.rename({'X': 'lon', 'Y': 'lat', 'T': 'time'})
-        ds['time'] = convert_t_to_time(ds['time'])
-        ds['time'] = ds['time'].astype('datetime64')
[…]

### 2018-02-27 src_old/pyfloods/dataset/elevation.py

@@ -1,30 +1,0 @@
-"""
-A SingleFileDataSet object for the NOAA NGDC GLOBE topographic dataset
-stored on the iridl data library:
-http://iridl.ldeo.columbia.edu/SOURCES/.NOAA/.NGDC/.GLOBE/.topo/
-"""
-import xarray as xr
-from datetime import datetime
-from typing import Dict,Tuple
-
-from pyfloods.dataset import SingleFileDataSet
-
-class NOAANGDC(SingleFileDataSet):
-    def __init__(self, fname:str, verbose: bool = False, params: Dict = {}) -> None:
-        """The params must be a dict of datetime.datetime objects
-        with names 'sdate' and 'edate'.
-        """
-        super().__init__(fname=fname, params=params, verbose=verbose)
-
-    def download_data(self) -> xr.DataArray:
-        url = 'http://iridl.ldeo.columbia.edu/SOURCES/.NOAA/.NGDC/.GLOBE/.topo/'
-        url += 'X/-180/0.025/180/GRID/Y/-90/0.025/90/GRID/'
-        url += 'dods'
-        ds = xr.open_dataarray(url)
-        ds.to_netcdf(self.fname, format='NETCDF4', mode='w')
-        return ds
-
-    def read_data(self) -> Tuple[xr.DataArray, Dict]:
-        ds = xr.open_dataarray(self.fname)
-        pars: Dict[str, datetime] = {}
-        return ds, pars

### 2018-02-27 src_old/pyfloods/dataset/sst.py

@@ -1,35 +1,0 @@
-"""
-A SingleFileDataSet object for the sst data
-http://iridl.ldeo.columbia.edu/expert/SOURCES/.NOAA/.NCEP/.EMC/.CMB/.GLOBAL/.Reyn_SmithOIv2/.monthly/.ssta/dods
-"""
-import xarray as xr
-import numpy as np
-import pandas as pd
-from datetime import datetime
-from typing import Dict,Tuple
-
-from pyfloods.dataset import SingleFileDataSet
-
-class CMBGLOBAL(SingleFileDataSet):
-    def __init__(self, fname:str, verbose: bool = False, params: Dict = {}) -> None:
-        """The params must be a dict of datetime.datetime objects
-        with names 'sdate' and 'edate'.
-        """
-        super().__init__(fname=fname, params=params, verbose=verbose)
-
-    def download_data(self) -> xr.DataArray:
-        url = 'http://iridl.ldeo.columbia.edu/expert/SOURCES/.NOAA/.NCEP/.EMC/.CMB/.GLOBAL/.Reyn_SmithOIv2/.monthly/.ssta/dods'
-        ds = xr.open_dataarray(url, decode_times=False)
-        time =np.int_(np.floor(ds['T']))
-        year = 1960 + time // 12
-        month = 1 + time % 12
-        day = np.ones_like(month)
-        time = pd.to_datetime(year * 10000 + month * 100 + 1, format="%Y%m%d")
-        ds['T'] = time
-        ds.to_netcdf(self.fname, format='NETCDF4', mode='w')
-        return ds
-
-    def read_data(self) -> Tuple[xr.DataArray, Dict]:
-        ds = xr.open_dataarray(self.fname)
-        pars: Dict[str, datetime] = {}
-        return ds, pars

### 2018-02-27 src_old/pyfloods/paths.py

@@ -1,15 +1,0 @@
-import os
-
-project_dir = os.path.abspath(os.path.join(os.path.dirname(__file__), os.path.pardir, os.path.pardir))
-
-data_external = os.path.join(project_dir, 'data', 'external')
-data_raw = os.path.join(project_dir, 'data', 'raw')
-data_processed = os.path.join(project_dir, 'data', 'processed')
-figures = os.path.join(project_dir, 'figures')
-
-for dir in [data_external, data_raw, data_processed, figures]:
-    if not os.path.isdir(dir):
-        os.makedirs(dir)
-
-file_station_description = os.path.join(data_raw, 'station_description.csv')
-file_weather_types = os.path.join(data_processed, 'weather_types.nc')

### 2018-02-27 src_old/pyfloods/raw_data.py

@@ -1,56 +1,0 @@
-"""
-Here we will define some datasets that use the dataset package
-but are specifically defined for this project.
-"""
-
-import xarray as xr
-import numpy as np
-from datetime import datetime
-import os
-
-from pyfloods.dataset import MJOBOMD, KaplanNINO34, NOAANGDC, S2SAreaAvg, ReanalysisV2, CPC, CMBGLOBAL
-from pyfloods import paths
-from pyfloods import region
-
-verbose = False # universal
-
-mjo = MJOBOMD(
-    fname = os.path.abspath(os.path.join(paths.data_external, 'mjo.csv')),
-    params = {'sdate': region.sdate, 'edate': region.edate},
-    verbose = verbose
-)
-enso = KaplanNINO34(
-    fname = os.path.abspath(os.path.join(paths.data_external, 'nino_34.csv')),
-    params = {'sdate': region.sdate, 'edate': region.edate},
-    verbose = verbose
-)
-topo = NOAANGDC(
-    fname = os.path.abspath(os.path.join(paths.data_external, 'elevation_data.nc')),
-    verbose = verbose
-)
-sst = CMBGLOBAL(
-    fname = os.path.abspath(os.path.join(paths.data_external, 'cmb_sst_data.nc')),
-    verbose = verbose
-)
-s2s = S2SAreaAvg(
-    fname = os.path.abspath(os.path.join(paths.data_external, 's2s_area_avg.nc')),
-    verbose = verbose
-)
-years = np.arange(region.syear, region.eyear+1)
-u850 = ReanalysisV2(
-    fnames = [os.path.join(paths.data_external, 'reanalysisv2_uwnd_850_{}.nc'.format(y)) for y in years],
-    years = years,
-    verbose = verbose,
-    params={'coord_system': 'pressure', 'level': 850, 'var': 'uwnd'}
-)
-v850 = ReanalysisV2(
-    fnames = [os.path.join(paths.data_external, 'reanalysisv2_vwnd_850_{}.nc'.format(y)) for y in years],
-    years = years,
-    verbose = verbose,
-    params={'coord_system': 'pressure', 'level': 850, 'var': 'vwnd'}
-)
-cpc = CPC(
-    [os.path.join(paths.data_external, 'cpc_rain_{}.nc'.format(y)) for y in years],
-    years = years,
-    verbose = verbose, params={}
-)

### 2018-02-27 src_old/pyfloods/region.py

@@ -1,42 +1,0 @@
-# Define spatial and temporal boundaries of the analysis
-import datetime
-import cartopy.crs as ccrs
-
-class Region:
-    def __init__(self, lon, lat):
-        from numpy import min
-        from numpy import max
-        self.lonmin = min(lon)
-        self.lonmax = max(lon)
-        self.latmin = min(lat)
-        self.latmax = max(lat)
-
-    def as_extent(self):
-        extent = [self.lonmin, self.lonmax, self.latmin, self.latmax]
-        return(extent)
-
-    def as_patch(self, label='', color='blue', fill=None, linewidth=2):
-        from matplotlib import patches
-        dx = self.lonmax - self.lonmin
-        dy = self.latmax - self.latmin
-        rec = patches.Rectangle(
-            (self.lonmin, self.latmin), dx, dy,
-            color = color,
-            label = label,
-            fill = fill,
-            linewidth=linewidth,
-            transform = ccrs.PlateCarree()
-        )
-        return(rec)
-
-wtype = Region(lon = [-65, -45], lat = [-30, -15])
-lower_py_river = Region(lon = [-59.75, -55.75], lat = [-27.75, -22.75])
-south_america = Region(lon = [-85, -30], lat = [-40, 5])
-southern_hemisphere = Region(lon = [-120, 0], lat = [-50, 5])
-reanalysis = Region(lon = [-177.5, 180], lat = [-60, 10])
-rainfall = Region(lon = [-89.75, -30.25], lat = [-59.75, 9.75])
-
-syear = 1979
-eyear = 2016
-sdate = datetime.datetime(syear, 1, 1)
-edate = datetime.datetime(eyear, 12, 31)

### 2018-02-27 src_old/pyfloods/streamfunction.py

@@ -1,80 +1,0 @@
-import xarray as xr
-import numpy as np
-from typing import Dict,Tuple,List
-from windspharm.xarray import VectorWind
-import pandas as pd
-import os
-
-from pyfloods.dataset.reanalysisv2 import RV2Year
-from pyfloods.dataset import SingleFileDataSet,ReanalysisV2
-from pyfloods import paths
-from pyfloods import raw_data
-from pyfloods import region
-
-class PsiYear(SingleFileDataSet):
-    """Turn a single year of uwind and vwind into streamfunction
-    """
-
-    def __init__(self, fname: str, uwnd: RV2Year, vwnd: RV2Year, verbose:bool = False) -> None:
-        params: Dict = uwnd.params.copy()
-        params.update({'var': 'psi'})
-        self.params = params
-        self.coord_system = params.get('coord_system')
-        self.var = params.get('var')
-        self.level = params.get('level')
-        self.year = params.get('year')
-        self.fname = fname
-        self.uwnd = uwnd
-        self.vwnd = vwnd
-        super().__init__(fname=self.fname, params=params, verbose=verbose)
-
-    def download_data(self) -> xr.DataArray:
-
-        U = self.uwnd.get_data()
-        V = self.vwnd.get_data()
-        W = VectorWind(U, V)
-        streamfunc = W.streamfunction()
-
-        old_time = pd.to_datetime(streamfunc['time'].values)
-        new_time = old_time + pd.DateOffset(hours=12)
-        streamfunc['time'] = new_time
-        streamfunc = streamfunc.resample(time='1D').mean(dim='time')
-        streamfunc.attrs.update({
-            'coord_system': self.params.get('coord_system'),
-            'level': self.params.get('level'),
-            'year': self.params.get('year'),
-            'var': self.params.get('var')
-        })
-
-        streamfunc.to_netcdf(self.fname, format='NETCDF4')
-        return streamfunc
-
-    def read_data(self) -> Tuple[xr.DataArray, Dict]:
-        ds = xr.open_dataarray(self.fname)
-        pars = {
-            'coord_system': ds.attrs.get('coord_system'),
-            'var': ds.attrs.get('var'),
-            'level': ds.attrs.get('level'),
-            'year': ds.attrs.get('year')
-        }
-        return ds, pars
-
-class PsiMulti():
[…]

### 2018-02-27 src_old/pyfloods/weather_type.py

@@ -1,106 +1,0 @@
-"""Python functions for performing weather typing
-"""
-
-import xarray as xr # for reading in netcdf files
-import numpy as np # for calculations
-import sklearn.cluster as cl # k-means clustering
-from eofs.xarray import Eof # EOF implementation that recognizes xarray
-from numba import jit # this massively speeds up the classifiability index
-from scipy.stats import pearsonr as correl
-
-def resort_labels(old_labels):
-    """Re-sort cluster labels
-
-    Takes in x, a vector of cluster labels, and re-orders them so that
-    the lowest number is the most common, and the highest number
-    the least common
-
-    Args:
-        old_labels: the previous labels of the clusters
-    Returns:
-        new_labels: the new cluster labels, ranked by frequency of occurrence
-    """
-    old_labels = np.int_(old_labels)
-    labels_from = np.unique(old_labels).argsort()
-    counts = np.array([np.sum(old_labels == xi) for xi in labels_from])
-    orders = counts.argsort()[::-1].argsort()
-    new_labels = orders[labels_from[old_labels]]
-    return new_labels
-
-@jit
-def calc_classifiability(P, Q): #pylint: disable=C0103
-    """Implement the Michaelangeli (1995) Classifiability Index
-
-    The variable naming here is not pythonic but follows the notation in the 1995 paper
-    which makes it easier to follow what is going on. You shouldn't need to call
-    this function directly but it is called in cluster_xr_eof.
-
-    Args:
-        P: a cluster centroid
-        Q: another cluster centroid
-    """
-    k = P.shape[0]
-    assert Q.shape[0] is k, "P and Q must be same k"
-    Aij = np.ones([k, k]) #pylint: disable=C0103
-    for i in range(k):
-        for j in range(k):
-            Aij[i, j], _ = correl(P[i, :], Q[j, :])
-    Aprime = Aij.max(axis=0) #pylint: disable=C0103
-    classifiability = Aprime.min()
-    return classifiability
-
-def cluster_xr_eof(data_array, n_clusters=5, prop=0.90, nsim=100, verbose=True, pcscaling=0):
-    """Carry out k-means clustering on EOFs of an xarray data array
-
-    Args:
-        data_array: the xarray data array (not dataset!) on which to cluster
-        n_clusters: the number of clusters to create
[…]

### 2018-02-28 reports/writeup/paper.tex
update poster
Fix library
Fix figures for main paper

--- 2017-11-30 reports/writeup/paper.tex
@@
 
-\documentclass[twocol]{ametsoc}
+\documentclass{ametsoc}
 \journal{jcli}
@@
 \newcolumntype{L}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
-\graphicspath{{../figs/}}
+\graphicspath{{../../figs/}{./}}
 
@@
 %-------------------------------------------------------------------------------
-
 \section{Introduction}
@@
 \bibliographystyle{ametsoc2014}
-\bibliography{library}
+\bibliography{../library}
 
@@
 \begin{figure*}
-	\noindent\includegraphics[width=6.5in]{StudyArea}
+	\noindent\includegraphics[width=6.5in]{study_area.jpg}
 	\caption{
@@
 		Colors indicate $\log_{10}$ of elevation, in \si{\meter}, from the Global Land One-Km Base Elevation project available at \texttt{http://iridl.ldeo.columbia.edu/SOURCES/.NOAA/.NGDC/.GLOBE/.topo/}.
-		(L): all of South America.
-		The domains of the ``Lower Paraguay River Basin'' and the domain used for weather typing are indicated in black and red, respectively.
-		(R): The Paraguay River.
-		As for (L), the ``Lower Paraguay River Basin'' is indicated.
+		(a): all of South America.
+		The domains of the ``Lower Paraguay River Basin'' and the domain used for weather typing are indicated in red and blue, respectively.
+		(b): The Paraguay River.
+		As for (a), the ``Lower Paraguay River Basin'' is denoted with a red box.
 		Streamflows shown in \cref{fig:streamflow} were taken from the fours stations indicated.
@@
 		Full station names are: Bah\'ia Negra (Bne); Concepci\'on (Conc); Asunci\'on (Asu); Pilar (Pil).
-	}
-  \label{fig:study-area}
+	}\label{fig:study-area}
 \end{figure*}
@@
 		(b) Time series of daily stage measurements from 1929 to 2016 at each station.
-	}
-  \label{fig:streamflow}
+	}\label{fig:streamflow}
 \end{figure*}
 
-\begin{figure*}	\noindent\includegraphics[width=6.5in]{LaggedRain.pdf}
+\begin{figure*}	\noindent\includegraphics[width=6.5in]{lagged_rain.pdf}
 	\caption{
-		Circulation anomalies associated with heavy rainfall (\nth{90} percentile exceedance of area-averaged rainfall in the Lower Paraguay River Basin).
+		Composite anomalies associated with heavy rainfall (\nth{90} percentile exceedance of area-averaged rainfall in the Lower Paraguay River Basin).
 		Lagged composites are shown, by column, for $t = $ \SIlist{-2;-1;0;1}{\day} relative to the date of heavy rainfall.
-		Top and middle rows show streamfunction anomalies at \SIlist{850}{\hecto\pascal}, respectively.
-		Bottom row shows monthly-mean rainfall anomalies, in units of \si{\milli\meter\per\day}.
-	}
-  \label{fig:lagged-rain}
+		Top row (a-d) shows composite streamfunction anomalies at \SIlist{850}{\hecto\pascal}.
+		Bottom row (e-h) shows composite rainfall anomalies, in units of \si{\milli\meter\per\day}.
+	}\label{fig:lagged-rain}
 \end{figure*}
@@
 \begin{figure*}
-	\noindent\includegraphics[width=6.5in]{PSI_EOF_Loadings.pdf}
+	\noindent\includegraphics[width=6.5in]{eof_loadings.pdf}
 	\caption{
-    	Leading EOF loadings of daily NDJF \SI{850}{\hecto\pascal} streamfunction $\psi$ over chosen domain.
-	}
-  \label{fig:eof-loading}
+		Loadings of the four leading EOFs of daily NDJF \SI{850}{\hecto\pascal} streamfunction over the weather typing region shown in \cref{fig:study-area}.
+	}\label{fig:eof-loading}
 \end{figure*}
@@
 \begin{figure*}
-	\noindent\includegraphics[width=6.5in,height=9in,keepaspectratio=true]{WTComposite.pdf}
+	\noindent\includegraphics[width=6.5in,height=9in,keepaspectratio=true]{wt_composite.pdf}
 	\caption{
-		Composite anomalies associated with each weather type (WT; column).
-		Top row shows streamfunction anomalies at \SIlist{850}{\hecto\pascal}.
-		Bottom row shows monthly-mean rainfall anomalies, in units of \si{\milli\meter\per\day}. The relative frequency of occurrence of each weather type (in days) is presented on the top of each column.
+		Composite anomalies associated with each weather typ.
+		Top row (a-f) shows streamfunction anomalies at \SIlist{850}{\hecto\pascal}.
+		Bottom row (g-l) shows rainfall anomalies, in units of \si{\milli\meter\per\day}. 
+		The relative frequency of occurrence of each weather type (in days) is presented on the top of each column.
 	}
@@
 \begin{figure*}
-	\noindent\includegraphics[width=6.5in]{RainfallWeatherType}
+	\noindent\includegraphics[width=6.5in]{rain_wt_201516.pdf}
 	\caption{
-		Time series of area-averaged rainfall in the Lower Paraguay River Basin (\cref{fig:study-area}) for each day of NDJF  2015-16 (lines and points).
-        The weather types corresponding to each day are indicated in both color and in an adjacent text label.
-		Dashed lines indicate the climatological 50th, \nth{90}, and 99th percentiles of NDJF area-averaged rain over the Lower Paraguay River basin.
-	}
-  \label{fig:rain-wt}
+		Time series of area-averaged rainfall in the Lower Paraguay River Basin (\cref{fig:study-area}) for each day of NDJF  2015-16.
+		Lines indicate the raimfall value, in units of \si{\milli\meter\per\day}.
+        The weather type corresponding to each day is indicated in an adjacent text label.
+		Dashed lines blue indicate (from bottom to top) the climatological 50th, \nth{90}, and 99th percentiles of NDJF area-averaged rain over the Lower Paraguay River basin.
+	}\label{fig:rain-wt}
 \end{figure*}
@@
 \begin{figure*}
-	\noindent\includegraphics[width=6.5in]{NDJF201516Anomaly.pdf}
+	\noindent\includegraphics[width=6.5in]{anomalies_ndjf1516.pdf}
 	\caption{
 		Monthly composite anomalies observed during NDJF 2015-16.
-    	Top and middle rows show streamfunction anomalies at \SIlist{200;850}{\hecto\pascal}, respectively.
-      Bottom row shows monthly-mean rainfall anomalies, in units of \si{\milli\meter\per\day}.
-	}
-  \label{fig:anomalies}
+		Top row (a-d) shows streamfunction anomalies at \SI{850}{\hecto\pascal}.
+		Bottom row (e-h) shows rainfall anomalies, in units of \si{\milli\meter\per\day}.
+	}\label{fig:anomalies}
 \end{figure*}
@@
 \begin{figure*}
-	\noindent\includegraphics[width=3.5in]{SeasonalForecast.pdf}
+	\noindent\includegraphics[width=3.5in]{seasonal_forecast.pdf}
 	\caption{
-	Seasonal model forecast for probability of DJF \nth{90} percentile exceedance.
-	Color indicates the forecast probability of exceeding the \nth{90} percentile of climatological rainfall during DJF 2015-16 -- this is presented as the odds ratio as explained in \cref{eq:odds-ratio}.
+		Seasonal model forecast for probability of exceedance of \nth{90} percentile of DJF rainfall.
+		Color indicates the forecast probability of exceeding the \nth{90} percentile of climatological rainfall during DJF 2015-16 -- this is presented as the odds ratio as defined in \cref{eq:odds-ratio}.
 		A value greater than 1 indicates that the model forecast greater-than-average odds of rainfall exceeding the \nth{90} percentile.
 		Grid cells which observed an exceedance of the \nth{90} percentile of DJF rainfall are outlined in black.
-	}
-  \label{fig:seas-prob-fcst}
+	}\label{fig:seas-prob-fcst}
 \end{figure*}
@@
 \begin{figure*}
-	\noindent\includegraphics[width=6.5in]{Chiclet.pdf}
+	\noindent\includegraphics[width=6.5in]{chiclet.pdf}
 	\caption{
-    Chiclet diagram \citep{Carbin:2016fx} of ensemble-mean precipitation anomaly forecasts over the Lower Paraguay River Basin (\cref{fig:study-area}) from ECMWF S2S forecast data, as a function of the forecast target date (horizontal axis) and lead time (vertical axis).
-Time series of CPC daily mean precipitation over the same area is plotted with $y$-axis inverted.
-	}
-  \label{fig:chiclet}
+		Chiclet diagram \citep{Carbin:2016fx} of ensemble-mean precipitation anomaly forecasts over the Lower Paraguay River Basin (see \cref{fig:study-area}) from ECMWF S2S forecast data, as a function of the forecast target date (horizontal axis) and lead time (vertical axis).
+		Time series of CPC daily mean precipitation over the same area is plotted with $y$-axis inverted; horizontal black line denotes NDJF climatology.
+	}\label{fig:chiclet}
 \end{figure*}
@@
 \begin{figure*}
-	\noindent\includegraphics[width=6.4in]{ForecastSkill.pdf}
+	\noindent\includegraphics[width=6.4in]{mos_forecasts.pdf}
 	\caption{
 		MOS-adjusted S2S model forecasts and skill scores for the methods indicated in \cref{tab:mos-methods}.
-		Top row shows the heavy rainfall forecast for 1-7 December 2015 as the odds ratio of \cref{eq:odds-ratio} over the target domain.
+		Top row (a-e) shows the heavy rainfall forecast for 1-7 December 2015 as the odds ratio defined in \cref{eq:odds-ratio} over the target domain.
 		A value greater than 1 indicates that the model forecast greater-than-average odds of rainfall exceeding the \nth{90} percentile.
-        Second row shows the Ignorance score \cref{eq:ignorance}, with zero indicating a perfect forecast.
-		Bottom row shows the 2AFC skill score for each grid cell; a value greater than 50 indicates that the model outperforms climatology.
-		For all three rows, the grid cells which experienced a \nth{90} percentile exceedance for 1-7 December 2015 are outlined in black.
-	}
-  \label{fig:subs-prob-fcst}
+        Second row (f-j) shows the Ignorance score defined in \cref{eq:ignorance}, with zero indicating a perfect forecast.
+		Bottom row (k-o) shows the 2AFC skill score for each grid cell; a value greater than 50 indicates that the model outperforms climatology.
+		Columns separate different MOS models except ``Raw'', which indicates the uncorrected S2S model output.
+		For all rows the grid cells which experienced a \nth{90} percentile exceedance for 1-7 December 2015 are outlined in black.
+	}\label{fig:subs-prob-fcst}
 \end{figure*}
@@
 \begin{figure*}
-	\noindent\includegraphics[width=6.4in]{WT_MJO_ENSO.pdf}
+	\noindent\includegraphics[width=6.4in]{wt_mjo_enso.pdf}
 	\caption{
@@
 		Only values which are significant at $p<0.10$, calculated using a bootstrap with 3000 samples, are shown.
-	}
-  \label{fig:wt-mjo-enso}
+	}\label{fig:wt-mjo-enso}
 \end{figure*}
@@
 	\caption{
-		Simple schematics of low-level jet events (red arrows) during austral summer and El Ni\~no years.
-		In general, most jet events are of the Chaco type, particularly when SST anomalies in the central southern Atlantic Ocean (a, see green box) are weak.
-		When a dipolar SST anomaly occurs in the central southern Atlantic (with the warmer pole equatorward) the meridional temperature gradient and sea-land temperature contrasts establish an anti-cyclonic circulation (dot-dashed line) conducive to increased occurrence of No-Chaco jet events (b).
+		Schematics of low-level jet events (red arrows) during austral summer and El Ni\~no years.
+		Most jet events are of the ``Chaco'' type, particularly when SST anomalies in the central southern Atlantic Ocean (a, see green box) are weak.
+		When a dipolar SST anomaly occurs in the central southern Atlantic with the warmer pole equatorward, the meridional temperature gradient and sea-land temperature contrasts establish an anti-cyclonic circulation (dot-dashed line) conducive to increased occurrence of No-Chaco jet events (b).
 		Other SST anomaly configurations tend to be present outside the green box (not shown).
-		Winds in panels are typical for each case (at 850 hPa).
-		Reference wind vector in $m s^{-1}$.
-	}
-  \label{fig:chaco-nochaco}
+		Winds in panels are typical for each case (at \SI{850}{\hecto\pascal}).
+		Reference wind vector in \si{\meter\per\second}.
+	}\label{fig:chaco-nochaco}
 \end{figure*}
@@
 \begin{figure*}
-	\noindent\includegraphics[width=6.4in]{WT4_Monthly_Composite.pdf}
+	\noindent\includegraphics[width=6.4in]{sst_anomalies.pdf}
 	\caption{
-		Composite sea surface temperature (SST) anomalies of the 15 NDJF months with the most occurrences of weather type 4.
-		SST data, in \si{\celsius}, come from \citet{Reynolds:2002iy}.
-		To match the SST data, only months since NDJF 1981 are considered.
-	}
-  \label{fig:wt4-months}
+		Composite sea surface temperature (SST) anomalies from \citet{Reynolds:2002iy}.
+		(a): The Pearson correlation of each SST grid cell with the time series of average zonal wind over the region indicated by the blue box.
+		(b): Sea surface temperature (SST) anomalies during December 2015.
+		Both panels show the Central Southern Atlantic with a black box; this is the same region as shown in \cref{fig:chaco-nochaco}.
+	}\label{fig:wt4-months}
 \end{figure*}

### 2018-02-28 src/analyze/plot_lagged_rain.py

@@ -11,6 +11,7 @@ import pandas as pd
 import matplotlib.pyplot as plt
 import cartopy.crs as ccrs
 import datetime
+import string
 
 from region import Region
 import visualize as viz
@@ -86,6 +87,13 @@ def main():
     cbar1 = fig.colorbar(C1, cax=cax1)
     cbar1.set_label('Precip. Anomaly [mm/d]', rotation=270)
     cbar1.ax.get_yaxis().labelpad = 20
+
+    # Add plot labels
+    letters = string.ascii_lowercase
+    for i, ax in enumerate(axes.flat):
+        label = '({})'.format(letters[i])
+        t = ax.text(0.025, 0.9, label, fontsize=12, transform=ax.transAxes)
+        t.set_bbox(dict(facecolor='white', alpha=0.75, edgecolor='gray'))
 
     fig.savefig(args.outfile, bbox_inches='tight')
 

### 2018-02-28 src/analyze/plot_leading_eofs.py

@@ -10,6 +10,7 @@ import numpy as np
 import matplotlib.pyplot as plt
 import cartopy.crs as ccrs
 from sklearn.decomposition import PCA
+import string
 
 from region import Region
 import visualize as viz
@@ -63,6 +64,13 @@ def main():
     )
     for i,ax in enumerate(p.axes.flat):
        ax.set_title('EOF {}'.format(i+1))
+
+    # Add plot labels
+    letters = string.ascii_lowercase
+    for i, ax in enumerate(p.axes.flat):
+        label = '({})'.format(letters[i])
+        t = ax.text(0.05, 0.9, label, fontsize=11, transform=ax.transAxes)
+        t.set_bbox(dict(facecolor='white', edgecolor='gray'))
 
     # Save figure
     plt.savefig(args.outfile, bbox_inches='tight')

### 2018-02-28 src/analyze/plot_nino_sst.py

@@ -1,0 +1,78 @@
+#!/usr/bin/env python
+# -*- coding: utf-8 -*-
+"""Plot SST Anomalies
+"""
+
+import argparse
+import os
+
+import pandas as pd
+import xarray as xr
+import numpy as np
+import matplotlib.pyplot as plt
+from matplotlib import patches
+import cartopy.crs as ccrs
+import cartopy.feature
+import colorcet as cc
+import string
+
+from region import Region
+import visualize as viz
+
+parser = argparse.ArgumentParser() #pylint: disable=C0103
+parser.add_argument("--outfile", help="the filename of the figure to save")
+parser.add_argument("--sst", help="the filename of the SST data")
+
+def main():
+    """Run everything
+    """
+    args = parser.parse_args()
+
+    sea_temp = xr.open_dataarray(args.sst).rename({'X': 'lon', 'Y': 'lat', 'T': 'time'})
+    sea_temp = sea_temp.sel(time = np.in1d(sea_temp['time.month'], [11, 12, 1, 2]))
+    
+    nino_times = pd.to_datetime(['1982-12-01', '1997-12-01', '2015-12-01'])
+    sst_plot = sea_temp.sel(time=np.in1d(sea_temp['time'], nino_times))
+
+    # Plot options
+    map_proj = ccrs.PlateCarree(central_longitude=-60)
+    data_proj = ccrs.PlateCarree()
+    figsize = (10, 9)
+    dipole = Region(lon=[-30,-10], lat=[-10,-40])
+
+    cmap = cc.cm['coolwarm']
+    p = sst_plot.plot.contourf(
+       x='lon', y='lat',
+       transform=data_proj,
+       col='time',
+       col_wrap=1,
+       subplot_kws={'projection': map_proj},
+       cmap=cmap,
+       levels=np.linspace(-3.5, 3.5, 15),
+       figsize=figsize
+    )
+
+    # Add stuff to the axes
+    viz.format_axes(
+        p.axes,
+        coast=True, grid=False, border=True,
+        xticks=np.linspace(-180, 180, 19), yticks=np.linspace(-90, 90, 10)
+    )
+
+    dipole = Region(lon=[-30,-10], lat=[-10,-40])
+    for ax in p.axes.flat:
+        ax.set_xlim([-80, 80]) 
+        ax.set_ylim([-60, 10])
+        ax.add_patch(dipole.as_patch(color='black'))
+
+    # Add plot labels
[…]

### 2018-02-28 src/analyze/plot_wt_mjo_enso.py

@@ -1,0 +1,73 @@
+#!/usr/bin/env python
+# -*- coding: utf-8 -*-
+"""Plot WT Occurrence on ENSO and MJO
+"""
+
+import argparse
+import xarray as xr
+import numpy as np
+import pandas as pd
+import matplotlib.pyplot as plt
+import datetime
+import string
+
+from region import Region
+import visualize as viz
+
+parser = argparse.ArgumentParser() #pylint: disable=C0103
+parser.add_argument("--outfile", help="the filename of the data to save")
+parser.add_argument("--wt", help="The weather type data")
+parser.add_argument("--nino", help="The streamfunction data file")
+parser.add_argument("--mjo", help="The streamfunction data file")
+
+def main():
+    args = parser.parse_args()
+    wt = xr.open_dataarray(args.wt).to_dataframe(name='wtype')
+    mjo = xr.open_dataset(args.mjo).to_dataframe()
+    enso = xr.open_dataarray(args.nino).to_dataframe(name='nino_34')
+    mrg = wt.join(mjo).join(enso)
+    mrg.loc[mrg['amplitude'] < 1, 'phase'] =  0
+    mrg['enso_state'] = 0
+    mrg.loc[mrg['nino_34'] < -1, 'enso_state'] = -1
+    mrg.loc[mrg['nino_34'] > 1, 'enso_state'] = 1
+    
+    conditional = mrg.groupby(['phase', 'enso_state'])['wtype'].apply(lambda g: g.value_counts() / g.size)
+    conditional = conditional.to_xarray()
+    anomaly = conditional - conditional.mean(dim=['level_2'])
+
+    # Bootstrap simulations
+    n_sim = 3000
+    anomalies = []
+    for i in np.arange(n_sim):
+        wt_seq = np.random.choice(mrg['wtype'], size=mrg.shape[0], replace=True)
+        simulated = pd.DataFrame({'wtype': wt_seq, 'phase': mrg['phase'], 'enso_state': mrg['enso_state']})
+        conditional = simulated.groupby(['phase', 'enso_state'])['wtype'].apply(lambda g: g.value_counts() / g.size)
+        conditional = conditional.to_xarray()
+        anomalies.append(conditional - conditional.mean(dim=['level_2']))
+    anomalies = xr.concat(anomalies, dim='temp')
+
+    # Select significant simulations
+    alpha = 0.10
[…]

### 2018-02-28 src/process/calculate_streamfunction.py

@@ -32,11 +32,10 @@ def main():
     vwnd = os.path.abspath(args.vwnd)
     outfile = os.path.abspath(args.outfile)
     psi = calculate_streamfunction(uwnd, vwnd)
-    longitudes = psi['lon'].values
-    longitudes[np.where(longitudes > 180)] -= 360
-    psi['lon'].values = longitudes
+    
     if os.path.isfile(outfile):
         os.remove(outfile)
+    
     psi.to_netcdf(outfile, format='NETCDF4', mode='w')
 
 if __name__ == "__main__":

### 2018-03-01 reports/writeup/Supplement.tex
james edits
rename figure eof-variance-explained so it appears correctly in supplemental info
Update some more plots and work on writeup

--- 2017-11-27 writeup/Supplement.tex
@@
 \usepackage[breaklinks=true,hidelinks]{hyperref}
-\graphicspath{{../figs/}}
+\graphicspath{{../../figs/}{./}}
 
@@
 		\toprule
-	{} &  NDJF 2015-16 &  Climatology \\
-	WT &           &              \\
+		{} &  NDJF 2015-16 &  climatology \\
+		wtype &               &              \\
 		\midrule
-	1     &  0.258 &     0.212 \\
-	2     &  0.242 &     0.187 \\
-	3     &  0.133 &     0.158 \\
-	4     &  0.192 &     0.155 \\
-	5     &  0.092 &     0.146 \\
-	6     &  0.083 &     0.142 \\
+		1     &      0.272727 &     0.212849 \\
+		2     &      0.231405 &     0.184517 \\
+		3     &      0.107438 &     0.156342 \\
+		4     &      0.206612 &     0.156189 \\
+		5     &      0.132231 &     0.149665 \\
+		6     &      0.049587 &     0.140438 \\
 		\bottomrule
@@
 \begin{figure}
-  \includegraphics[width=\textwidth]{PSI_Var_Explained.pdf}
+  \includegraphics[width=\textwidth]{wt_classifiability.pdf}
 	\caption{
-		Variance explained for each EOF loading of the \SI{850}{\hecto\pascal} streamfunction $\psi$.
-    (L): the individual variance explained for each EOF.
-		(R): cumulative variance explained.
+		Classifiability index \citep[see Methods and][]{Michelangeli1995} calculated for several chosen values of $K$, representing the number of weather types produced.
+		Results are produced with only 75 simulations for each $K$.
 	}
@@
 \begin{figure}
-  \includegraphics[width=\textwidth]{WT_Classifiability.pdf}
+	\includegraphics[width=\textwidth]{nino_sst_anomalies.pdf}
 	\caption{
-		Classifiability index \citep[see Methods and][]{Michelangeli1995} calculated for several chosen values of $K$, representing the number of weather types produced.
-		Results are produced with only 50 simulations for each $K$.
+		Monthly SST anomalies during major El Ni\~{n}o events.
+		Months shown are (a) December 1982, (b) December 1997, and (c) December 2015.
+		Units are in \si{\celsius}.
 	}
-\end{figure}
-
-\begin{figure}
-  \includegraphics[width=\textwidth]{MJO_Time_Series.pdf}
-	\caption{
-		Phase plot of the Madden-Julian Oscillation during NDJF 2015-16.
-		The $x$-axis shows RMM1 and the $y$-axis shows RMM2 -- gray lines separate MJO phases.
-	}
-	\label{fig:mjo-ts}
 \end{figure}

### 2018-03-01 reports/writeup/paper.tex
Update some more plots and work on writeup

@@
 Of these, EOF 1 explains a substantial amount of variance ($\approx 67\%$) while EOFs 2, 3, and 4 collectively explain approximately $28\%$ of total variance.
-Values of individual and cumulative variance explained are presented in supplemental figure S2.
-To a first order of approximation, EOF 1 explains the large-scale anomalies induced by passing baroclinic waves; the East-West gradient is likely a result of the influence of the Andes Mountains which are a source of friction and torque for low-level flow.
-EOF 2 primarily controls the strength of the low-level jet, while EOFs 3 and 4 control the exit region of this jet.
-Given these features in the EOF loadings, it is no surprise that composites of the  \SI{850}{\hecto\pascal} circulations and rainfall associated with each weather type (\cref{fig:wt-composite}) reveal patterns associated with synoptic- and regional-scale circulation regimes.
+The resulting weather types, shown in \cref{fig:wt-composite}, reveal patterns associated with synoptic- and regional-scale circulation regimes.
+This is consistent with the hypothesis that the EOFs over the study area are associated with large-scale patterns.
 
@@
 
-In this study we examined the regional climate drivers of the persistent and heavy NDJF 2015-16 rainfall over the Lower Paraguay River Basin, leading to severe flood impacts.
-The observed flooding was made more likely by antecedent soil moisture left by the preceding rainy season and followed two years of substantial, though less extreme, flooding.
-These antecedent conditions combined with the persistent heavy rainfall occurring in November and December 2015 to produce heavy flooding on the Lower Paraguay river.
+In this study we examined the regional climate drivers of the persistent and heavy NDJF 2015-16 rainfall over the Lower Paraguay River Basin which was associated with severe flood events.
 
@@
 		Loadings of the four leading EOFs of daily NDJF \SI{850}{\hecto\pascal} streamfunction over the weather typing region shown in \cref{fig:study-area}.
+		Parentheses indicate percentage of total variance explained by each EOF.
 	}\label{fig:eof-loading}

### 2018-03-01 src/analyze/plot_leading_eofs.py

@@ -62,8 +62,12 @@ def main():
         extent=extent,
         crs = data_proj,
     )
+    
+    # Label Titles with Variance Explained
+    var_xpl = pca.explained_variance_
+    var_xpl /= var_xpl.sum()
     for i,ax in enumerate(p.axes.flat):
-       ax.set_title('EOF {}'.format(i+1))
+       ax.set_title('EOF {} ({:.1f}%)'.format(i+1, var_xpl[i]*100))
 
     # Add plot labels
     letters = string.ascii_lowercase

### 2018-03-01 src/analyze/plot_wt_mjo_enso.py

@@ -22,10 +22,12 @@ parser.add_argument("--mjo", help="The streamfunction data file")
 
 def main():
     args = parser.parse_args()
+
     wt = xr.open_dataarray(args.wt).to_dataframe(name='wtype')
     mjo = xr.open_dataset(args.mjo).to_dataframe()
-    enso = xr.open_dataarray(args.nino).to_dataframe(name='nino_34')
+    enso = xr.open_dataarray(args.nino).to_dataframe(name='nino_34').resample('1D').ffill()
     mrg = wt.join(mjo).join(enso)
+    
     mrg.loc[mrg['amplitude'] < 1, 'phase'] =  0
     mrg['enso_state'] = 0
     mrg.loc[mrg['nino_34'] < -1, 'enso_state'] = -1
@@ -36,7 +38,7 @@ def main():
     anomaly = conditional - conditional.mean(dim=['level_2'])
 
     # Bootstrap simulations
-    n_sim = 3000
+    n_sim = 5000
     anomalies = []
     for i in np.arange(n_sim):
         wt_seq = np.random.choice(mrg['wtype'], size=mrg.shape[0], replace=True)

### 2018-03-03 reports/writeup/paper.tex
james writing edits

@@
 
+\usepackage[acronym]{glossaries}
+\makeglossaries
+\newacronym{wt}{WT}{weather type}
+\newacronym{sallj}{SALLJ}{South American low-level jet}
+\newacronym{enso}{ENSO}{El Ni\~{n}o-Southern Oscillation}
+\newacronym{mjo}{MJO}{Madden-Julian Oscillation}
+\newacronym{sacz}{SACZ}{South Atlantic Convergence Zone}
+\newacronym{iri}{IRI}{International Research Institute for Climate and Society}
+\newacronym{ecmwf}{ECMWF}{European Centre for Medium-Range Weather Forecasts}
+\newacronym{s2s}{S2S}{sub-seasonal to seasonal}
+\newacronym{eof}{EOF}{empirical orthogonal function}
+\newacronym{mos}{MOS}{model output statistics}
+\newacronym{xlr}{XLR}{homoscedastic extended logistic regression}
+\newacronym{hxlr}{HXLR}{heteroscedastic extended logistic regression}
+\newacronym{pcr}{PCR}{princiap components regression}
+\newacronym{cca}{CCA}{canonical correlation analysis}
+\newacronym{sst}{SST}{sea surface temperature}
+
 %-------------------------------------------------------------------------------
@@
 
-\title{Heavy rainfall in Paraguay during the 2015-2016 austral summer: causes and subseasonal-to-seasonal predictive skill}
+\title{Heavy rainfall in Paraguay during the 2015-2016 austral summer: causes and sub-seasonal-to-seasonal predictive skill}
 \authors{James Doss-Gollin\correspondingauthor{James Doss-Gollin. Columbia Water Center. Dept. of Earth and Environmental Engineering, Columbia University, 500 W. 120th St., New York, NY. USA.}}
@@
 
-During the austral summer of 2015-16, repeated heavy rainfall events led to severe flooding in the Paraguay River basin, displacing approximately \num{170000} people \citep{BrackenridgeDFO} and causing tremendous damage to property and infrastructure \citep{MinisteriodeObrasPublicasyComunicacion:2016tq}.
-Because population in South America tends to concentrate along coasts and rivers (see supplemental figure S1), flooding in the Paraguay River basin directly affects not only much of the population of Paraguay, but also of populations in Argentina and Uruguay who lie along the Paran\'{a} and la Plata rivers into which the Paraguay River drains.
+During the austral summer of 2015-16, repeated heavy rainfall events led to severe flooding in the Paraguay River basin, displacing approximately \num{170000} people \citep{Brakenridge2016} and causing tremendous damage to property and infrastructure \citep{MinisteriodeObrasPublicasyComunicacion2016}.
+Because population in South America tends to concentrate along coasts and rivers (see supplemental figure S1)\footnote{check}, flooding in the Paraguay River basin directly affects not only much of the population of Paraguay, but also of populations in Argentina and Uruguay who lie along the Paran\'{a} and la Plata rivers into which the Paraguay River drains.
 Heavy rainfall and flooding in the basin also has important implications for hydropower generation, for agriculture, and for regional water resource management.
@@
 The rainfall climatology of the Paraguay River basin varies strongly by season, with extratropical characteristics in the winter and monsoonal characteristics in the summer.
-In the warm season (NDJF) which is the focus of this study, circulation is dominated by the upper-tropospheric Bolivian High, the lower-level subtropical highs, the Chaco Low over northern Argentina, the South Atlantic Convergence Zone (SACZ), and the South American Low-Level Jet, or SALLJ \citep{Grimm2009,Marengo:2012cm}.
+In the warm season (NDJF) which is the focus of this study, circulation is dominated by the upper-tropospheric Bolivian High, the lower-level subtropical highs, the Chaco Low over northern Argentina, the \gls{sacz}, and the \gls{sallj} \citep{Grimm2009,Marengo2012}.
 Rainfall peaks around \SI{5}{\milli\meter\per\day} during the warm months (October-May) and reaches a minimum near \SI{2}{\milli\meter\per\day} in July and August.
-However, the flat topography limits the river's ability to carry the summer runoff, causing seasonal inundation of the Pantanal and distributing the river discharge in time \citep{Bravo:2011et,Barros:2004bn}.
-Thus, the streamflow maxima typically occurs in phase with precipitation upstream of the Pantanal, while downstream of the Pantanal -- which we define as the Lower Paraguay River Basin; see \cref{fig:study-area} -- the annual peak typically occurs between April and July.
+However, the flat topography limits the river's ability to carry the summer runoff, causing seasonal inundation of the Pantanal and distributing the river discharge in time \citep{Bravo2011,Barros2004}.
+Thus, the streamflow maxima typically occurs in phase with precipitation upstream of the Pantanal, while downstream of the Pantanal - which we define as the Lower Paraguay River Basin; see \cref{fig:study-area} - the annual peak typically occurs between April and July.
 
 During the warm season (NDJF) a large fraction of rainfall, and nearly all heavy rainfall, in the Paraguay River Basin is associated with mesoscale convection \citep{Velasco1987}.
-Previous studies of organized convection and precipitation across subtropical continental South America have found close correspondence to the exit region of the low-level jets \citep{Saulo:2007km,Salio:2007gd,Marengo2004,Velasco1987}, which is influenced in both summer and winter by midlatitude baroclinic wave trains that interact with the Andes topography to generate orographically bound cyclones and northerly low-level flow \citep{Campetella:2002hx,Seluchi:2006bi,Boers:2014gt}.
-The strength and direction of this moisture transport varies substantially between events; SALLJ exit regions range from central Argentina \citep[``Chaco Jet Events'';][]{Salio:2002ev} to Paraguay and southeastern Brazil \citep[``No-Chaco Jet Events'';][]{Vera:2006ib}.
+Previous studies of organized convection and precipitation across subtropical continental South America have found close correspondence to the exit region of the low-level jets \citep{Saulo2007,Salio2007,Marengo2004,Velasco1987}, which is influenced in both summer and winter by midlatitude baroclinic wave trains that interact with the Andes topography to generate orographically bound cyclones and northerly low-level flow \citep{Campetella2002,Seluchi2006,Boers2013,Boers2014}.
+The strength and direction of this moisture transport varies substantially between events; \gls{sallj} exit regions range from central Argentina \citep[``Chaco Jet Events'';][]{Salio2002} to Paraguay and southeastern Brazil \citep[``No-Chaco Jet Events'';][]{Vera2006}.
 
-At sub-seasonal timescales, convection in the Paraguay River Basin is modulated by a variety of drivers, notably including the SACZ and the Madden-Julien Oscillation (MJO).
-During SACZ conditions, strong convergence is observed over the Amazon basin with divergence over southwestern Brazil, northern Argentina and Paraguay \citep{Herdies:2002jy,Carvalho2010}; the opposite is true for so-called NoSACZ conditions.
-SACZ occurrence is related to westerly wind regimes, as well as ``active'' and ``break'' periods of the South American Monsoon System \citep{Marengo2004}.
+At sub-seasonal timescales, convection in the Paraguay River Basin is modulated by a variety of drivers, notably including the \gls{sacz} and the \gls{mjo}  .
+During \gls{sacz} conditions, strong convergence is observed over the Amazon basin with divergence over southwestern Brazil, northern Argentina and Paraguay \citep{Herdies2002,Carvalho2010}; the opposite is true for so-called NoSACZ conditions.
+\Gls{sacz} occurrence is related to westerly wind regimes, as well as ``active'' and ``break'' periods of the South American Monsoon System \citep{Marengo2004}.
 
-At seasonal timescales, El Ni\~no-Southern Oscillation (ENSO) is the dominant driver of convection variability in the Paraguay River Basin.
+At seasonal timescales, \gls{enso} is the dominant driver of convection variability in the Paraguay River Basin.
 During El Ni\~no years, a low-level anticyclonic anomaly over central Brazil enhances occurrence of the low-level jet, favoring the development of mesoscale convective systems \citep{Velasco1987}.
-The intensity and precise extent of this anomaly is relevant for the precise impact of ENSO events.
+The intensity and precise extent of this anomaly is relevant for the precise impact of \gls{enso} events.
 The region also exhibits substantial variability between seasons of rainfall during El Ni\~no years, including a reversal of rainfall anomalies between November of that year and January of the following one, influenced by land-surface interactions \citep{Grimm2009,Grimm2003}.
-Even beyond El Ni\~no years, regional land-surface feedbacks can cause regions that exhibit wet anomalies in the spring to experience more summer precipitation on average \citep{Grimm:2009bq}.
+Even beyond El Ni\~no years, regional land-surface feedbacks can cause regions that exhibit wet anomalies in the spring to experience more summer precipitation on average \citep{Grimm2007}.
 Similarly, mid-latitude dynamics influence low-level wind anomalies on many time scales, though this relationship is complicated due to coupled tropical-extratropical interactions \citep{Carvalho2004,Jones2002}.
@@
 
-Rainfall data are taken from the CPC Unified Gauge-based Analysis of Global Daily Precipitation dataset \citep{xie2010cpc}.
+Rainfall data are taken from the CPC Unified Gauge-based Analysis of Global Daily Precipitation dataset \citep{Xie2010}.
 Spatial resolution is \SI{0.5}{\degree}, and temporal resolution is daily.
@@
 
-Atmospheric circulations are diagnosed using daily data from the NCAR-NCEP Reanalysis II dataset \citep{Kanamitsu:2002kk}.
+Atmospheric circulations are diagnosed using daily data from the NCAR-NCEP Reanalysis II dataset \citep{Kanamitsu2002}.
 Spatial resolution is \SI{2.5}{\degree}.
-Because the end-of-day time for the rainfall data is 12:00GMT over most of South America \citep{xie2010cpc}, we use six-hour reanalysis data, and shift by twelve hours before re-sampling to the daily time step.
+Because the end-of-day time for the rainfall data is 12:00 GMT over most of South America \citep{Xie2010}, we use six-hour reanalysis data, and shift by twelve hours before re-sampling to the daily time step.
 This ensures that the time steps in the reanalysis and rainfall data sets are the same, but means that a day is defined as beginning at 12:00 GMT.
 Since most summer rainfall in this region occurs overnight, this end-of-day time (which translates to approximately 8:00 AM locally depending on the exact time zone) tends to separate distinct events.
-Ocean sea surface temperature (SST) patterns are explored at the monthly time step using the \SI{1}{\degree} NOAA OI.v2 dataset \citep{Reynolds:2002iy}.
+Oceanic \gls{sst} patterns are explored at the monthly time step using the \SI{1}{\degree} NOAA OI.v2 dataset \citep{Reynolds2002}.
 
-Streamflow data was collected by the Paraguayan Navy and National Administration of Navigation and Ports of Paraguay, and was processed and distributed by the Paraguayan Directorate of Meteorology and Hydrology (DINAC-DMH).
+Streamflow data was collected by the Paraguayan Navy and National Administration of Navigation and Ports of Paraguay, and was processed and distributed by the Paraguayan Directorate of Meteorology and Hydrology.
 Locations of streamflow gauges are shown in \cref{fig:study-area}.
@@
 This study also makes use of some climate indices.
-Data on the El Ni\~{n}o-Southern Oscillation (ENSO), or NI\~NO3.4, came from a statistical-dynamical interpolation \citep{Kaplan:1998df}, which is constrained by relatively high-quality observations during the study period.
-Data on the Madden-Julien Oscillation (MJO) came from the Australian Bureau of Meteorology \citep{Wheeler:2004ea}.
+Data on \gls{enso}, specifically the NI\~NO3.4 index, came from a statistical-dynamical interpolation \citep{Kaplan1998}, which is constrained by relatively high-quality observations during the study period.
+Data on the \gls{mjo}   came from the Australian Bureau of Meteorology \citep{Wheeler2004}.
 
@@
 
-The seasonal predictions used were ``flexible format'' forecasts provided by the International Research Institute for Climate and Society (IRI).
+The seasonal predictions used were ``flexible format'' forecasts provided by the \gls{iri}.
 This product provides the probability of exceedance (or non-exceedance) of a user-specified percentile of the climatological distribution.
@@
 
-The sub-seasonal forecasts used were issued by the European Centre for Medium-Range Weather Forecasts (ECMWF) using a coupled model, available via the Subseasonal-to-Seasonal Prediction Project Database \citep{Vitart2016} at \SI{1.5}{\degree} resolution.
+The sub-seasonal forecasts used were issued by the \gls{ecmwf} using a coupled model, available via the \gls{s2s} Prediction Project Database \citep{Vitart2016} at \SI{1.5}{\degree} resolution.
 Forecasts consider the period starting in Dec 2015 until Mar 2016 and hindcasts to assess the real-time predictive skill consider the period Dec 1-7, 1995-2014.
@@
 
-Anomalies were calculated relative to the seasonal mean from November 1979 to February 2016, thus retaining the seasonal cycle.
+Anomalies were calculated relative to the seasonal mean from November 1979 to February 2016, and the anomalies thus contain information on intra-seasonal variability.
 
@@
 Several types of analyses are used to diagnose the causes of the heavy rainfall events, and to bias-correct and verify the forecasts.
-All codes used for data and fully reproducible analysis are available at \url{https://github.com/jdossgollin/PYFloods}.
-Computation was performed in the python environment using the \texttt{xarray} \citep{hoyer2017xarray} and \texttt{numpy} \citep{vanderWalt:2011dp} packages for gridded data and the \texttt{pandas} \citep{McKinney:2010un} package for tabular data; plotting was performed with the \texttt{matplotlib} \citep{Hunter:2007ux} and \texttt{cartopy} \citep{Cartopy} packages.
+Computation was performed in the python environment using stable open source packages \citep{Hoyer2017,vanderWalt2011,McKinney2010,Hunter2007}.
+All codes to reproduce or modify this analysis are available at \url{https://github.com/jdossgollin/PYFloods}.
 
 Given the behavior of the Paraguay River discussed above, we define the Lower Paraguay River Basin as the region bounded by \SIrange{-59.75}{-55.75}{\degree W} and \SIrange{26.75}{22.75}{\degree S}, as shown in \cref{fig:study-area}.
-This region corresponds to the region over which, given topography and previous studies \citep{Barros:2004bn,Bravo:2011et}, one might hypothesize rainfall inputs to most closely correspond to river levels at the stream gauges in \cref{fig:study-area}.
+In this region, given topography and previous studies \citep{Barros2004,Bravo2011}, one might hypothesize rainfall inputs to most closely correspond to river levels at the stream gauges in \cref{fig:study-area}.
 Results are not sensitive to small variations in this domain.
@@
 
-The clustering was performed on the daily NDJF \SI{850}{\hecto\pascal} streamfunction field ($\psi$), calculated by integrating the meridional and zonal wind fields using spherical harmonics, as implemented in the \texttt{windspharm} package \citep{Dawson:2016ws,Adams:1999fb}, over the domain spanning \SIrange{15}{30}{\degree S} and \SIrange{65}{45}{\degree W} (\cref{fig:study-area}).
+The clustering was performed on the daily NDJF \SI{850}{\hecto\pascal} streamfunction field ($\psi$), calculated by integrating the meridional and zonal wind fields using spherical harmonics, as implemented in the \texttt{windspharm} package \citep{Dawson2016}, over the domain spanning \SIrange{15}{30}{\degree S} and \SIrange{65}{45}{\degree W} (\cref{fig:study-area}).
 
-To facilitate clustering (which tends to perform poorly in high-dimensional spaces), the NDJF anomaly field of $\psi_{850}$ was projected  onto its four leading empirical orthogonal functions (EOFs), accounting for $>95\%$ of the total observed variance (supplemental figure S2), using the \texttt{eofs} package \citep{Dawson:2016ge}.
+To facilitate clustering (which tends to perform poorly in high-dimensional spaces), the NDJF anomaly field of $\psi_{850}$ was projected  onto its four leading \glspl{eof}, accounting for $>95\%$ of the total observed variance.
 No meridional weighting was applied as the selected domain is relatively small and does not extend into high latitudes.
-Once the EOFs were calculated, the principal component time series were computed for each day and scaled to unit variance.
-This rescaling is not a necessary step; its effect is to treat all retained principal components as equally important for the weather typing, which provides relatively greater weight to EOFs 2, 3, and 4 than carrying out weather typing without re-scaling.
-Though our approach of first selecting the number of EOFs to use and then choosing to scale them equally involves more subjective decisions than an approach without rescaling, in this case the resulting physical mechanisms are more interpretable.
+Once the \glspl{eof} were calculated, the principal component time series were computed for each day and scaled to unit variance.
+This rescaling is not a necessary step; its effect is to treat all retained principal components as equally important, which provides relatively greater weight to \glspl{eof} 2, 3, and 4 than carrying out the clustering without re-scaling.
+Though our approach of first selecting the number of \glspl{eof} to use and then choosing to scale them equally involves more subjective decisions than an approach without rescaling, in this case the resulting physical patterns represented by the \glspl{eof} more closely represent patterns identified in the literature; this is further discussed in \cref{sec:diagnostics}.
 
@@
 \end{enumerate}
-The cluster centroids (weather types) $\mu_k$ produced by the $K$-means algorithm can then be interpreted as a Voronoi decomposition of the phase space into $K$ regions, and specifically as the Voronoi diagram which minimizes within-cluster variance.
+The cluster centroids $\mu_k$ produced by the $K$-means algorithm can then be interpreted as a Voronoi decomposition of the phase space into $K$ regions, and specifically as the Voronoi diagram which minimizes within-cluster variance.
 
-The $K$-means algorithm is guaranteed to converge to a local minimum of intercluster variance; to select the best partition, 500 simulations were created using the implementation in Python's \texttt{scikit-learn} package \citep{Pedregosa:2012tv} and then the classifiability index of \citet{Michelangeli1995} was computed between each partition and the 499 others.
+The $K$-means algorithm is guaranteed to converge to a local minimum of intercluster variance; to select the best partition, 500 simulations were created using the implementation in Python's \texttt{scikit-learn} package \citep{Pedregosa2012} and then the classifiability index of \citet{Michelangeli1995} was computed between each partition and the 499 others.
 The partition whose classifiability index, averaged for all 499 pairwise comparisons, was the highest was selected.
-Calculation of the classifiability index for several values of $K$ (supplemental figure S3) suggests that states with $K=5,6,\ldots,9$ are all reasonable; we chose the solution $K=6$ because the weather types identified are qualitatively similar to those determined over southeastern South America \citep{Munoz2015,Munoz2016} and have an intuitive physical meaning, discussed further in the following sections.
+Calculation of the classifiability index for several values of $K$ (supplemental figure S3)\footnote{CHECK} suggests that states with $K=5,6,\ldots,9$ are all reasonable; we chose the solution $K=6$ because the clusters identified are qualitatively similar to those determined over southeastern South America \citep{Munoz2015,Munoz2016} and have an intuitive physical meaning, discussed further in the following sections.
+We refer to the resulting clusters as \glspl{wt}.
 
-\subsection{Forecasts and Model Output Statistics (MOS)}
+\subsection{Forecasts and Model Output Statistics}
 
-A wide variety of methods, generically known as Model Output Statistics \citep[MOS;][]{Glahn:1972vt}, have been proposed to correct for different types of bias in model outputs.
+A wide variety of methods, generically known as \gls{mos} \citep{Glahn1972}, have been proposed to correct for different types of bias in model outputs.
 In this work, we analyze how well the rainfall events could have been predicted, both using the raw sub-seasonal forecasts and  MOS-adjusted sub-seasonal forecasts.
-We use  four types of MOS techniques: the homoscedastic and heteroscedastic versions of the extended logistic regression (XLR and HXLR, respectively); principal component regression (PCR); and canonical correlation analysis (CCA).
+We use  four types of \gls{mos} techniques: the homoscedastic extended logistic regression \gls{xlr}; the \gls{hxlr}; \gls{pcr}; and \gls{cca}.
 
-Logistic regression models the conditional probability of binary events, and it has been used in MOS approaches for binary predictands \citep{Hamill:2004hk}.
-Nonetheless, when using the logistic regression to address multiple thresholds via independent fits, the predicted probabilities are, in general, not mutually consistent \citep{Messner:2014gp}.
-The (homoscedastic) XLR was designed to address this shortcoming \citep{Wilks:2009bk}, via the consideration of a transformation of the thresholds of interest as an additional predictor variable.
-The HXLR, a heteroscedastic generalization of the XLR, was proposed to appropriately use the ensemble spread as predictor for the dispersion of the predictive distribution \citep{Messner:2014gp}.
+Logistic regression models the conditional probability of binary events, and it has been widely used in \gls{mos} approaches for binary predictands \citep{Hamill2004}.
+Nonetheless, when using the logistic regression to address multiple thresholds via independent fits, the predicted probabilities are, in general, not mutually consistent \citep{Messner2014}.
+The \gls{xlr} was designed to address this shortcoming \citep{Wilks2009}, via the consideration of a transformation of the thresholds of interest as an additional predictor variable.
+The \gls{hxlr}, a heteroscedastic generalization of the \gls{xlr}, was proposed to appropriately use the ensemble spread as predictor for the dispersion of the predictive distribution \citep{Messner2014}.
 
-CCA is a common statistical method frequently used to forecast rainfall using a purely empirical approach \citep{Mason:2008da,Barnston:2012ce,Jolliffe2012:vm,Barnston:1992gd,Wilks:2006fx}.
-CCA identifies modes of co-variability, called canonical variates or canonical modes, by maximizing the correlation between linear combinations of the predictor and predictand's empirical orthogonal functions (EOFs).
+\Gls{cca} is a common statistical method frequently used to forecast rainfall using a purely empirical approach \citep{Mason2008,Barnston2012,Jolliffe2012,Barnston1992,Wilks2006}.
+\Gls{cca} identifies modes of co-variability, called canonical variates or canonical modes, by maximizing the correlation between linear combinations of the predictor and predictand's \glspl{eof}.
 The method forecasts spatial patterns of variability spanning across the region of interest rather than making forecasts for individual locations.
-In PCR, a special case of CCA, each grid cell in the predictand field is estimated by regression using a linear combination of the predictor's EOFs \citep{Mason:2008da,Wilks:2006fx} rather than by identifying canonical modes.
-Unlike the XLR and HXLR models, which perform bias correction independently for each grid cell, the CCA and PCR models can address biases in both the magnitude and the spatial distribution of the modeled precipitation patterns.
+In \gls{pcr}, a special case of \gls{cca}, each grid cell in the predictand field is estimated by regression using a linear combination of the predictor's \glspl{eof} \citep{Mason2008,Wilks2006} rather than by identifying canonical modes.
+Unlike the \gls{xlr} and \gls{hxlr} models, which perform bias correction independently for each grid cell, the \gls{cca} and \gls{pcr} models can address biases in both the magnitude and the spatial distribution of the modeled precipitation patterns.
 
-For the purposes of MOS corrections, the predictand (variable to forecast) is the observed rainfall for the target period of interest, and the predictor (variable to be corrected) is the raw model rainfall for the same period.
+For the purposes of \gls{mos} corrections, the predictand (variable to forecast) is the observed rainfall for the target period of interest, and the predictor (variable to be corrected) is the raw model rainfall for the same period.
 Exceedance of the \nth{90} percentile during the 1995-2014 period is used to define the heavy event cases.
-We use the same spatial domain [\SIrange{39}{17}{\degree S}; \SIrange{66}{49}{\degree W}] for both the predictor and the predictand, except for the PCR and CCA cases, in which a larger domain [\SIrange{0}{60}{\degree S}; \SIrange{80}{30}{\degree W}] was used to better capture the spatial patterns in the raw model field.
+We use the same spatial domain [\SIrange{39}{17}{\degree S}; \SIrange{66}{49}{\degree W}] for both the predictor and the predictand, except for the \gls{pcr} and \gls{cca} cases, in which a larger domain [\SIrange{0}{60}{\degree S}; \SIrange{80}{30}{\degree W}] was used to better capture the spatial patterns in the raw model field.
 A variety of domains and ways to combine initialization times were explored; the best results were selected in terms of the corresponding Kendall's $\tau$ rank correlation coefficient.
-A summary of the final candidate predictors found to be most skillful for each MOS model is presented in \cref{tab:mos-methods}.
+A summary of the final candidate predictors found to be most skillful for each \gls{mos} model is presented in \cref{tab:mos-methods}.
 
@@
 
-The IRI's seasonal predictions are already provided with bias correction \citep{Barnston:2010ge}, and thus we did not perform any further MOS on seasonal rainfall fields.
+The \gls{iri}'s seasonal predictions are already provided with bias correction \citep{Barnston2010}, and thus we did not perform any further \gls{mos} on seasonal rainfall fields.
 
@@
 \end{equation}
-where $Y$ is the observed outcome and $p(Y)$ is the density function of the forecast distribution \citep{Good:1952ux,Roulston:2002jz,Brocker:2007er}.
-The Ignorance score was introduced as the fundamental information-based verification measure,  decomposable into easily interpretable components: reliability, resolution and uncertainty \citep{Weijs:2010cl}.
-Due to its close relationship to Shannon's information entropy, it is used to measure forecast utility, or the amount of information gain expected from a forecast \citep{Roulston:2002jz}.
+where $Y$ is the observed outcome and $p(Y)$ is the density function of the forecast distribution \citep{Good1952,Roulston2002,Brocker2007}.
+The Ignorance Score was introduced as an information theory-based verification measure,  decomposable into easily interpretable components: reliability, resolution and uncertainty \citep{Weijs2010}.
+Due to its close relationship to Shannon's information entropy, it is used to measure forecast utility, or the amount of information gain expected from a forecast \citep{Roulston2002}.
 
-We also compute the Generalized Relative Operating Characteristics (GROC), or 2AFC score \citep{Mason:2009kr}, to evaluate skill of probabilistic rainfall forecasts.
-This score measures the ``proportion of all available pairs of observations of differing category whose probability forecasts are discriminated in the correct direction'' \citep{Mason:2009kr}.
+We also compute the Generalized Relative Operating Characteristics score, commonly referred to as the 2AFC score \citep{Mason2009}, to evaluate skill of probabilistic rainfall forecasts.
+This score measures the ``proportion of all available pairs of observations of differing category whose probability forecasts are discriminated in the correct direction'' \citep{Mason2009}.
 It has an intuitive interpretation as an indication of how often the forecasts are correct.
@@
 These two metrics, measuring reliability, resolution, uncertainty and discrimination, are deemed here to be sufficient to characterize the forecast skill for our events of interest.
-To conduct the verification in a consistent manner, we use the Climate Predictability Tool (CPT), developed by the International Research Institute for Climate and Society \citep{Mason:2017gg}.
+To conduct the verification in a consistent manner, we use the Climate Predictability Tool, developed by the \gls{iri} \citep{Mason2017}.
 
@@
 During November and December 2015, the river rose rapidly at Concepci\'on, Asunci\'on, and Pilar though not at Bah\'ia Negra.
-As discussed in \citet{Bravo:2011et,Barros:2004bn}, the location of the Bah\'ia Negra gauge (see \cref{fig:study-area}) in the Pantanal region means that it responds very slowly to rainfall input.
+As discussed in \citet{Bravo2011,Barros2004}, the location of the Bah\'ia Negra gauge (see \cref{fig:study-area}) in the Pantanal region means that it responds very slowly to rainfall input.
 The three downstream gauges, because they are located in the Lower Paraguay River basin respond to the rainfall forcing with a slow but steady rise.
@@
 Examination of \cref{fig:streamflow}b suggests multidecadal oscillation in the streamflow record.
-This is in agreement with previous studies \citep{Collischonn:2001bi,Carvalho2011} which find a changepoint in the 1970s, possibly associated with low-frequency Pacific variability.
+This is in agreement with previous studies \citep{Collischonn2001,Carvalho2011} which find a changepoint in the 1970s, possibly associated with low-frequency Pacific variability.
 Because only river stage data (and not discharge) are available, it is not possible to discern whether the observed changes in river stage are driven by sediment loading and local measurement characteristics or by large-scale climate fluctuations.
@@
 
-\Cref{fig:lagged-rain} shows time-lagged anomalies up to and after heavy rainfall dates (when area-averaged daily rainfall in the lower Paraguay River Basin exceeds its NDJF \nth{90} percentile) and is consistent with previous findings \citep{Marengo2004,Salio:2007gd,Marwan2015}.
+\Cref{fig:lagged-rain} shows time-lagged anomalies up to and after heavy rainfall dates (when area-averaged daily rainfall in the lower Paraguay River Basin exceeds its NDJF \nth{90} percentile) and is consistent with previous findings \citep{Marengo2004,Salio2007,Marwan2015}.
 At $t=-2$ a mid-latitude baroclinic system approaches the South American continent, intensifying and moving to the East at $t=-1,0,1$.
@@
 
-The first step of the weather typing algorithm is to identify leading EOFs of the \SI{850}{\hecto\pascal} streamfunction $\psi$.
-The EOF loadings are shown in \cref{fig:eof-loading}.
-Of these, EOF 1 explains a substantial amount of variance ($\approx 67\%$) while EOFs 2, 3, and 4 collectively explain approximately $28\%$ of total variance.
-The resulting weather types, shown in \cref{fig:wt-composite}, reveal patterns associated with synoptic- and regional-scale circulation regimes.
-This is consistent with the hypothesis that the EOFs over the study area are associated with large-scale patterns.
+The first step of the weather typing algorithm is to identify leading \glspl{eof} of the \SI{850}{\hecto\pascal} streamfunction $\psi$.
+The \gls{eof} loadings are shown in \cref{fig:eof-loading}.
+Of these, \gls{eof} 1 explains a substantial amount of variance ($\approx 67\%$) while \glspl{eof} 2, 3, and 4 collectively explain approximately $28\%$ of total variance.
+The resulting \glspl{wt}, shown in \cref{fig:wt-composite}, reveal patterns associated with synoptic- and regional-scale circulation regimes.
+This is consistent with the hypothesis that the \glspl{eof} over the study area are associated with large-scale patterns.
 
-Weather type 1 represents a ``Chaco Jet'' event \citep{Salio:2002ev} in which the maximum wind speed penetrates past \SI{25}{\degree S}, leading to heavy rainfall over NE Argentina and Uruguay.
-Weather type 4 is also associated with a SALLJ event, but is of the ``No-Chaco'' jet type in which the wind turns to the East before \SI{25}{\degree S}, leading to heavy rainfall over Eastern Paraguay and SW Brazil.
+\Gls{wt} 1 represents a ``Chaco Jet'' event \citep{Salio2002} in which the maximum wind speed penetrates southward of \SI{25}{\degree S}, leading to heavy rainfall over NE Argentina and Uruguay.
+\Gls{wt} 4 is also associated with a \gls{sallj} event, but is of the ``No-Chaco'' jet type in which the wind turns to the East northward of \SI{25}{\degree S}, leading to heavy rainfall over Eastern Paraguay and SW Brazil.
 
-Weather types 5 and 3 are nearly inverses of weather types 1 and 4, respectively, and are associated with dry anomalies over the Lower Paraguay River Basin.
+\Glspl{wt} 5 and 3 are nearly inverses of \glspl{wt} 1 and 4, respectively, and are associated with dry anomalies over the Lower Paraguay River Basin.
 The fact that they are not exact inverses suggest important nonlinearities in the system.
-Finally, weather types 2 and 6 are related to a high-pressure configuration bringing below-average rainfall over most of Brazil, and a dipolar pattern conducive to above-average rainfall over central Brazil, respectively (\cref{fig:wt-composite}).
+Finally, \glspl{wt} 2 and 6 are related to a high-pressure configuration bringing below-average rainfall over most of Brazil, and a dipolar pattern conducive to above-average rainfall over central Brazil, respectively (\cref{fig:wt-composite}).
 
@@
 
-As indicated in \cref{fig:rain-wt}, the most heavy rainfall occurred during weather types 1 and 4.
-During NDJF 2015-16 weather types 1 and 4 (Chaco and No-Chaco jet extensions, respectively),  occurred more frequency than their climatology (supplemental table S1); weather type 2 also occurred more frequently than its climatology, largely due to a long sequence in February 2016.
-In mid-January 2016, during a sequence of persistent low rainfall, weather type 3 featured persistently, leading to heavy rainfall over central Brazil (not shown) and negative rainfall anomalies over the Lower Paraguay River Basin.
+As indicated in \cref{fig:rain-wt}, the most heavy rainfall occurred during \glspl{wt} 1 and 4.
+During NDJF 2015-16 \glspl{wt} (Chaco and No-Chaco jet extensions, respectively),  occurred more frequency than their climatology (supplemental table S1)\footnote{check}; \gls{wt} 2 also occurred more frequently than its climatology, largely due to a long sequence in February 2016.
+In mid-January 2016, during a sequence of persistent low rainfall, \gls{wt} 3 featured persistently, leading to heavy rainfall over central Brazil (not shown) and negative rainfall anomalies over the Lower Paraguay River Basin.
 Thus while the intensity and persistence of heavy rainfall was atypical, the causal mechanism of the heavy rainfall observed during this season was consistent with climatology.
@@
 Inspection of \cref{fig:rain-wt} also suggests that at time scales of days to weeks, particular sequences of weather types tend to recur, and are associated with repeated rainfall storms.
-The rainfall in late November 2015 occurred during weather type 1 and 4 days, with intermittent (and dry) weather type 5 occurrences.
-During mid to late December 2015, weather types 1 and 5 produced rainfall but the intermittent dry days were weather type 2.
-During early January 2016 weather type 3 led to persistent low rainfall.
+The rainfall in late November 2015 occurred during \gls{wt} 1 and 4 days, with intermittent (and dry) \gls{wt} 5 occurrences.
+During mid to late December 2015, \glspl{wt} 1 and 5 produced rainfall but the intermittent dry days were \gls{wt} 2.
+During early January 2016 \gls{wt} 3 led to persistent low rainfall.
 
@@
 There are advantages in simultaneously considering useful climate information at multiple timescales, rather than just focusing on one of them.
-This multistage prediction system is known as the IRI's ``Ready-Set-Go" approach \citep{Hellmuth:2011vt,Goddard:2014kf}.
+This multistage prediction system is known as the \gls{iri}'s ``Ready-Set-Go" approach \citep{Hellmuth2011,Goddard2014}.
 For the sake of concision, in this section we discuss only probabilistic seasonal (DJF 2015-2016) and sub-seasonal (Dec 1-7, 2015) forecasts.
@@
 
-Sub-seasonal predictions are still too new to be used as operational tools, and their skill is normally not high enough to be useful for most decision-making \citep{Vigaud:2017gg}.
-Nonetheless, the international S2S Prediction Project \citep{Vitart2016} has been providing since 2015 free access to almost-real-time sub-seasonal forecasts from multiple models, an opportunity to explore how well the heavy rainfall events of the first week of December 2015 could have been predicted.
+Sub-seasonal predictions are still too new to be used as operational tools, and their skill is normally not high enough to be useful for most decision-making \citep{Vigaud2017}.
+Nonetheless, the international \gls{s2s} Prediction Project \citep{Vitart2016} has been providing since 2015 free access to almost-real-time sub-seasonal forecasts from multiple models, an opportunity to explore how well the heavy rainfall events of the first week of December 2015 could have been predicted.
 
-\Cref{fig:chiclet} uses a Chiclet diagram \citep{Carbin:2016fx} to visualize, as a function of lead time, the time evolution of the uncorrected, ensemble-mean rainfall anomaly forecast, spatially averaged over the Lower Paraguay River Basin.
+\Cref{fig:chiclet} uses a Chiclet diagram \citep{Carbin2016} to visualize, as a function of lead time, the time evolution of the uncorrected, ensemble-mean rainfall anomaly forecast, spatially averaged over the Lower Paraguay River Basin.
 At times greater than about two weeks, the ensemble-mean forecast is slightly wetter than climatology.
@@
 To examine these forecasts more closely, we turn to the 14-19 day forecast of the December 1-7 2015 period.
-As seen in \cref{fig:subs-prob-fcst}, the raw (uncorrected) sub-seasonal forecast of the ECMF model for Dec 1-7 2015 indicated very high relative odds for occurrence of heavy rainfall but with important biases in the actual location and spatial pattern; for Paraguay, it confidently suggests occurrence of heavy rainfall to the south-southeast of the country, which was mostly not observed.
-Overall, the 20-year skill of probabilistic forecasts for the first week of December -- even when using  multiple initializations -- is inferior to climatology at the lead times used (14 and 19 days) for most of Paraguay (\cref{fig:subs-prob-fcst}b).
+As seen in \cref{fig:subs-prob-fcst}, the raw (uncorrected) sub-seasonal forecast of the \gls{ecmwf} model for Dec 1-7 2015 indicated very high relative odds for occurrence of heavy rainfall but with important biases in the actual location and spatial pattern; for Paraguay, it confidently suggests occurrence of heavy rainfall to the south-southeast of the country, which was mostly not observed.
+Overall, the 20-year skill of probabilistic forecasts for the first week of December - even when using  multiple initializations - is inferior to climatology at the lead times used (14 and 19 days) for most of Paraguay (\cref{fig:subs-prob-fcst}b).\footnote{Check if it's b}
 Despite these biases, the model is capturing a signal, which suggests the use of Model Output Statistics to explore the extent to which corrections in the magnitudes and spatial patterns may improve the forecast.
@@
 In general, the use of extended logistic regression models does not improve the forecast for the week.
-For example, with respect to the raw prediction, XLR tends to amplify the relative odds, and to cluster and shift the forecast location of the heavy rainfall events (\cref{fig:subs-prob-fcst}, top row, first two columns); the forecast tends to be better for Uruguay, but suggests heavy rainfall in the Paraguayan Chaco, which was not present in the raw prediction.
+For example, with respect to the raw prediction, \gls{xlr} tends to amplify the relative odds, and to cluster and shift the forecast location of the heavy rainfall events (\cref{fig:subs-prob-fcst}, top row, first two columns); the forecast tends to be better for Uruguay, but suggests heavy rainfall in the Paraguayan Chaco, which was not present in the raw prediction.
 On the other hand, the use of the ensemble spread in the HLXR model does not help; this forecast tends to be over-confident on the events occurring in almost all the region of interest (\cref{fig:subs-prob-fcst}, top row, third column).
@@
 Reliability, resolution and uncertainty, as measured by the ignorance score (\cref{fig:subs-prob-fcst} middle row, first three columns), suggests slight skill improvement in southern Brazil, deterioration in Argentina and Uruguay, and basically the same as the raw model output for Paraguay and southeastern Bolivia.
-Changes in forecast discrimination exhibited by the extended logistic models, as measured via the 2AFC score (\cref{fig:subs-prob-fcst} bottom row, first three columns), are  null.
+Changes in forecast discrimination exhibited by the extended logistic models, as measured via the 2AFC score (\cref{fig:subs-prob-fcst} bottom row, first three columns)\footnote{update to letters}, are  null.
 These models operate on a gridbox-by-gridbox basis to recalibrate the probabilities, and so this recalibration happens monotonically.
@@
 Better forecasts are obtained when both magnitude and spatial corrections are performed, although with relative odds considerably less confident than the ones in the raw forecast.
-The Principal Component Regression model correctly shows high relative odds in most of the places where heavy rainfall was observed (\cref{fig:subs-prob-fcst}, first row, fourth column), although it also indicates heightened risk in areas where heavy rainfall did not occur, like zones of western Paraguay and northeastern Argentina.
-The main problem with the Canonical Correlation Analysis model is its lack of discrimination between occurrence or non-occurrence of heavy rainfall in the region: the spatial distribution of odds is too homogeneous (\Cref{fig:subs-prob-fcst}, first row, fifth column).
+The \gls{pcr} model correctly shows high relative odds in most of the places where heavy rainfall was observed (\cref{fig:subs-prob-fcst}, first row, fourth column), although it also indicates heightened risk in areas where heavy rainfall did not occur, like zones of western Paraguay and northeastern Argentina.
+The main problem with the \gls{cca} model is its lack of discrimination between occurrence or non-occurrence of heavy rainfall in the region: the spatial distribution of odds is too homogeneous (\cref{fig:subs-prob-fcst}, first row, fifth column).
 
-The 20-year based skill maps of probabilistic forecasts (see \Cref{fig:subs-prob-fcst}, middle and bottom row) computed with these two EOF-based models are very similar to each other, both in terms of the reliability, resolution and uncertainty (Ignorance Score; second row, fourth and fifth columns) and discrimination (2AFC score; middle row, fourth and fifth columns).
-Compared to the raw model output and the extended logistic regression models, PCR and CCA models perform quite well (\cref{fig:subs-prob-fcst}, middle and bottom row).
-The enhanced skill is achieved through the spatial corrections via the EOF-based regressions, which -- in contrast with the extended logistic models -- use information from multiple grid-boxes, and thus the original forecasts are not (necessarily) calibrated monotonically.
+The 20-year based skill maps of probabilistic forecasts (see \Cref{fig:subs-prob-fcst}, middle and bottom row) computed with these two \gls{eof}-based models are very similar to each other, both in terms of the reliability, resolution and uncertainty (Ignorance Score; second row, fourth and fifth columns) and discrimination (2AFC score; middle row, fourth and fifth columns).
+Compared to the raw model output and the logistic regression models, \gls{pcr} and \gls{cca} models perform quite well (\cref{fig:subs-prob-fcst}, middle and bottom row).\footnote{Letters}
+The enhanced skill is achieved through the spatial corrections via the \gls{eof}-based regressions, which - in contrast with the extended logistic models - use information from multiple grid-boxes, and thus the original forecasts are not (necessarily) calibrated monotonically.
 
-Despite the particular errors in the Dec 1-7 2015 forecasts, both PCR and CCA verify considerably better than the raw, XLR, and HXLR predictions.
+Despite the particular errors in the Dec 1-7 2015 forecasts, both \gls{pcr} and \gls{cca} verify considerably better than the raw, \gls{xlr}, and \gls{hxlr} predictions.
 Yet despite the generally high skill score for these forecasts, there are still zones along the eastern part of Paraguay with negative skill relative to climatology.
@@
 
-Although many of the individual rainfall events of NDJF 2015-16 were intense, they were nontheless driven by the climatological mechanism for heavy rainfall shown in \cref{fig:lagged-rain} and discussed in more detail in previous studies \citep{Saulo:2007km,Salio:2007gd,Marengo2004,Velasco1987}.
+Although many of the individual rainfall events of NDJF 2015-16 were intense, they were nontheless driven by the climatological mechanism for heavy rainfall shown in \cref{fig:lagged-rain} and discussed in more detail in previous studies \citep{Saulo2007,Salio2007,Marengo2004,Velasco1987}.
 Consequently, the most striking hydrometeorlogical feature of this season, and the leading cause of the extreme flooding, was the persistence of the heavy rainfall and the manner in which it switched ``on'' and ``off'' during the season (\cref{fig:rain-wt}).
@@
 
-As a starting point, we consider the roles of El Ni\~{n}o, which has been extensively studied in the literature \citep{Carvalho2004,Velasco1987,Salio:2002ev,Bravo:2011et,Grimm2009}, and the MJO.
-As indicated earlier, examination of the \SI{850}{\hecto\pascal} streamfunction field (\cref{fig:eof-loading}) suggests that EOF 2 represents the intensity of the SALLJ, while EOF 3 describes the East-West extent of the jet exit region -- in other words, whether it is a Chaco or No-Chaco jet event.
+As a starting point, we consider the roles of \gls{enso}, which has been extensively studied in the literature \citep{Carvalho2004,Velasco1987,Salio2002,Bravo2011,Grimm2000,Grimm2003,Grimm2009a}, and the \gls{mjo}.
+Analysis of \cref{fig:eof-loading} suggests that \gls{eof} 2 approximately describes the meridional wind over the region, consistent with \gls{sallj} activity, while \gls{eof} 3 the meridional wind over the region.
 
-These circulations are strongly modulated by ENSO and the MJO: the Spearman (rank) correlation coefficient between the monthly-averaged loading EOF 2 and the NI\~NO3.4 index is \num{-.369} ($p \ll 0.01$), and the Spearman correlation between the daily loading of EOF 3 the MJO RMM1 index is \num{0.130} ($p \ll 0.01$).
-A key question is how the combined effects of potentially predictable seasonal to sub-seasonal climate mechanisms modify the occurrence of each weather type.
+These circulations are strongly modulated by \gls{enso} and the MJO: the Spearman (rank) correlation coefficient between the monthly-averaged loading \gls{eof} 2 and the NI\~NO3.4 index is \num{-.369} ($p \ll 0.01$), and the Spearman correlation between the daily loading of \gls{eof} 3 the \gls{mjo}   RMM1 index is \num{0.130} ($p \ll 0.01$).
+A key question is how the combined effects of potentially predictable seasonal to sub-seasonal climate mechanisms modify the occurrence of each weather type.\footnote{THESE CORRELATIONS ARE NOT RIGHT}
 The typical behavior is summarized by \cref{fig:wt-mjo-enso}, following the approach of \citet{Munoz2016} and \citet{Moron2015}.
-During El Ni\~{n}o years such as 2015-16, weather type 1 -- related to Chaco jet events -- occurs more frequently for almost all MJO phases, and particularly during phases 1, 2, and 5, consistent with observations of the year under study.
-The occurrence of WT3, favored during MJO phases 8 and 1, in late January 2016 and of WT2, enhanced during MJO phases 4-7, in mid-February 2016 is also consistent with the observed MJO time series (supplemental figure S4) and observed weather type sequences (\cref{fig:rain-wt}).
+During El Ni\~{n}o years such as 2015-16, \gls{wt} 1 -- related to Chaco jet events -- occurs more frequently for almost all \gls{mjo}  phases, and particularly during phases 1, 2, and 5, consistent with observations of the year under study.
+The occurrence of \gls{wt} 3, favored during \gls{mjo} phases 8 and 1, in late January 2016 and of\gls{wt} 2, enhanced during \gls{mjo}  phases 4-7, in mid-February 2016 is also consistent with the observed \gls{mjo}  time series (supplemental figure S4)\footnote{check} and observed weather type sequences (\cref{fig:rain-wt}).
 
-While the occurrence of WT1 during NDJF 2015-16 is well explained by ENSO and MJO variability, these features alone do not explain the frequent occurrence of WT4, the ``No-Chaco'' jet event.
-Previous studies on the South American Low-Level Jet \citep[e.g.,][]{Vera:2006ib} and the modulation of rainfall in southeastern South America by extratropical transient wave trains during El Ni\~no years \citep{Barreiro:2017ct} emphasize the importance of Pacific-Atlantic interaction for forecasting climate events in this (and other) region(s).
-A persistent dipolar SST anomaly in the central southern Atlantic Ocean favors the occurrence of WT4 by blocking transient extratropical wave activity coming the Pacific, facilitating transitions from Chaco jet events (WT 1) to No-Chaco jet events (WT 4) via enhanced low-level wind circulation from southern Brazil towards the Atlantic, and back to north-east Brazil and the Amazon (see \cref{fig:chaco-nochaco}) due to land-sea temperature contrasts.
-We illustrate this mechanism in \cref{fig:chaco-nochaco} and note that it is consistent with the mechanism found to produce heavy rainfall in the Lower Paraguay River Basin (\cref{fig:lagged-rain}) and with previous studies \citep[e.g.,][]{Vera:2006ib,Liebmann:2004kx,Salio:2002ev}.
-This dipole is present in a SST anomaly composite of the 15 NDJF months with the most frequent occurrence of WT 4 (shown in \cref{fig:wt4-months}), and it does not change substantially if more or fewer months are used for the composite.
-The pattern is also consistent with the observed dipolar structure in SST anomalies that was present in the central southern Atlantic Ocean since mid November 2015 to roughly the end of December 2015, associated with an anomalous anti-cyclonic configuration localized around \SI{35}{\degree W} and \SI{30}{\degree S}.
+While the occurrence of \gls{wt} during NDJF 2015-16 is well explained by \gls{enso} and \gls{mjo}  variability, these features alone do not explain the frequent occurrence of \gls{wt} 4, the ``No-Chaco'' jet event.
+Previous studies on the South American Low-Level Jet \citep[e.g.,][]{Vera2006} and the modulation of rainfall in southeastern South America by extratropical transient wave trains during El Ni\~no years \citep{Barreiro2017} emphasize the importance of Pacific-Atlantic interaction for forecasting climate events in this (and other) region(s).
+A persistent dipolar \gls{sst} anomaly in the central southern Atlantic Ocean favors the occurrence of \gls{wt} 4 by blocking transient extratropical wave activity coming the Pacific, facilitating transitions from Chaco jet events (\gls{wt} 1) to No-Chaco jet events (\gls{wt} 4) via enhanced low-level wind circulation from southern Brazil towards the Atlantic, and back to north-east Brazil and the Amazon (see \cref{fig:chaco-nochaco}) due to land-sea temperature contrasts.
+We illustrate this mechanism in \cref{fig:chaco-nochaco} and note that it is consistent with the mechanism found to produce heavy rainfall in the Lower Paraguay River Basin (\cref{fig:lagged-rain}) and with previous studies \citep[e.g.,][]{Vera2006,Salio2002,Liebmann2004}.
+This dipole is present in a \gls{sst} anomaly composite of the 15 NDJF months with the most frequent occurrence of \gls{wt} 4 (shown in \cref{fig:wt4-months}), and it does not change substantially if more or fewer months are used for the composite.\footnote{EDIT THIS PARAGRAPH}
+The pattern is also consistent with the observed dipolar structure in \gls{sst} anomalies that was present in the central southern Atlantic Ocean since mid November 2015 to roughly the end of December 2015, associated with an anomalous anti-cyclonic configuration localized around \SI{35}{\degree W} and \SI{30}{\degree S}.
 
 A potential source of dynamical model biases in the representation of rainfall events in the region is related to simulation of the extratropical wave activity mentioned before.
-For example, in order to adequately forecast rainfall in certain parts of southeastern South America during El Ni\~no years, models need to reproduce not one but two stationary wave trains, originating from central and eastern Pacific, respectively \citep{Barreiro:2017ct}.
-The frontal activity is associated with the extension, shape and variability of SACZ, and could be blocked by the presence of anomalous anti-cyclones off the coast of Brazil, like the ones described in reference to No-Chaco jet events.
+For example, in order to adequately forecast rainfall in certain parts of southeastern South America during El Ni\~no years, models need to reproduce not one but two stationary wave trains, originating from central and eastern Pacific, respectively \citep{Barreiro2017}.
+The frontal activity is associated with the extension, shape and variability of the \gls{sacz}, and could be blocked by the presence of anomalous anti-cyclones off the coast of Brazil, like the ones described in reference to No-Chaco jet events.
 
-Other mechanisms that have been known to modulate rainfall signals in this region include the South Atlantic Convergence Zone \citep{Carvalho2004,Munoz2015,Munoz2016} and land-biosphere-atmosphere interactions \citep{Grimm:2000bt,Grimm:2009bq} which also tend to be poorly represented in models \citep{Green:2017cs,Koster:2004ge}.
-The analysis of these sources of systematic errors could require a cross-timescale diagnostic approach \citep{Munoz:2017ct}, and it is beyond the scope of the present paper.
-Analysis of La Ni\~na and ENSO-neutral years (\cref{fig:wt-mjo-enso}) also suggests modulation of weather type occurrence by MJO-ENSO interactions, consistent with the cross-timescale hypothesis detailed in \citet{Munoz2015}.
-This cross-timescale interaction also helps to explain why sub-seasonal forecasts were not substantially better than seasonal ones: although the ENSO signal was homogeneous throughout the season, studies of MJO predictability have found it to be limited beyond two to three weeks \citep{Ding:2010em,Vitart:2014bt,Kim:2014bt}.
+Other mechanisms that have been known to modulate rainfall signals in this region include the \gls{sacz} \citep{Carvalho2004,Munoz2015,Munoz2016} and land-biosphere-atmosphere interactions \citep{Grimm2000,Grimm2007} which also tend to be poorly represented in models \citep{Green2017,Koster2004}.
+The analysis of these sources of systematic errors could require a cross-timescale diagnostic approach \citep{Munoz2017}, and it is beyond the scope of the present paper.
+Analysis of La Ni\~na and \gls{enso}-neutral years (\cref{fig:wt-mjo-enso}) also suggests modulation of weather type occurrence by \gls{mjo}-\gls{enso} interactions, consistent with the cross-timescale hypothesis detailed in \citet{Munoz2015}.
+This cross-timescale interaction also helps to explain why sub-seasonal forecasts were not substantially better than seasonal ones: although the \gls{enso} signal was homogeneous throughout the season, studies of \gls{mjo}  predictability have found it to be limited beyond two to three weeks \citep{Ding2009,Vitart2014,Kim2014a}.
 
@@
 Looking in southeastern South America (slightly South and East of the Lower Paraguay River Basin defined in this study), \citet{Munoz2016} used the sequences of weather types identified for each year to analyze predictability of heavy rainfall events.
-Although this approach -- as with the weather typing used in this study -- simplifies the dynamics of the system, it can be a valuable tool for diagnosing and assessing model processes \citep{Munoz:2017ct}, and for communicating model forecasts to policy-makers.
+Although this approach - as with the weather typing used in this study - simplifies the dynamics of the system, it can be a valuable tool for diagnosing and assessing model processes \citep{Munoz2017}, and for communicating model forecasts to policy-makers.
 Such an assessment is beyond the scope of this case study but offers promising opportunities for local public and private sectors to use climate model forecasts.
@@
 Indeed, rainfall and streamflow are related, but the link is nonlinear in space and time.
-As explained in \cref{sec:diagnostics}\ref{sec:flood}, in this region the flat topography (\cref{fig:study-area}) means that the Lower Paraguay River reacts slowly to rainfall \citep{Bravo:2011et,Barros:2004bn}, explaining the slow but steady rise in river levels shown in \cref{fig:streamflow}.
-Groundwater dynamics are also important in explaining this behavior \citep{Santos:2016td}.
-The observed flood peaks during 2015-16 also seem to occur in the context of an active phase of a multi-decadal oscillation, possibly associated with low-frequency Pacific activity \citep{Collischonn:2001bi,Huang:2005ir}.
-Parsing the relative impacts of deforestation and land use changes in the river basin, installation of hydroelectric generation at the Itaipu and Yacyreta sites, river channel modification, and climate variability on flood levels will require gathering improved hydrological data and building a comprehensive system model but all of these factors may influence the timing and magnitude of flooding in this region.
+As explained in \cref{sec:diagnostics}\ref{sec:flood}, in this region the flat topography (\cref{fig:study-area}) means that the Lower Paraguay River reacts slowly to rainfall \citep{Bravo2011,Barros2004}, explaining the slow but steady rise in river levels shown in \cref{fig:streamflow}.
+Groundwater dynamics are also important in explaining this behavior \citep{Santos2016}.
+The observed flood peaks during 2015-16 also seem to occur in the context of an active phase of a multi-decadal oscillation, possibly associated with low-frequency Pacific activity \citep{Collischonn2001,Huang2005}.
+Parsing the relative impacts of deforestation and land use changes in the river basin, installation of hydroelectric generation at the Itaipu and Yacyreta sites, river channel modification, and climate variability on flood levels will require gathering improved hydrological data and building a comprehensive system model but all of these factors may influence the timing and magnitude of flooding in this region.\footnote{Perhaps add something about antecedent conditions here and inter-decadal variability}
 
@@
 Doing so will require compiling information on the properties, businesses, and infrastructure that are vulnerable to flooding.
-This study also suggests that proposed dredging of the upper Paraguay River Basin to facilitate navigation could lead to increased summertime streamflow from the Upper Paraguay River Basin (Pantanal), effectively coupling the phases of streamflow from the Upper and Lower Paraguay River Basins which currently have a time-delay \citep{Bravo:2011et}.
+This study also suggests that proposed dredging of the upper Paraguay River Basin to facilitate navigation could lead to increased summertime streamflow from the Upper Paraguay River Basin (Pantanal), effectively coupling the phases of streamflow from the Upper and Lower Paraguay River Basins which currently have a time-delay \citep{Bravo2011}.
 
@@
 The heavy rainfall observed was driven by enhanced moisture inflow from the low-level jet and convergence associated with baroclinic systems.
-Repeated SALLJ events, particularly No-Chaco jet events, led to favorable conditions for mesoscale convective activity in this region.
-This activity was modulated by cross-timescale dynamics including a strong El Ni\~{n}o and an active MJO in phases 4-5, which favored SALLJ occurrence, and the presence of a dipolar SST anomaly in the central southern Atlantic Ocean, which enhanced the probability of occurrence of No-Chaco jet events.
+Repeated \gls{sallj} events, particularly No-Chaco jet events, led to favorable conditions for mesoscale convective activity in this region.
+This activity was modulated by cross-timescale dynamics including a strong El Ni\~{n}o and an active \gls{mjo}  in phases 4-5, which favored \gls{sallj} occurrence, and the presence of a dipolar SST anomaly in the central southern Atlantic Ocean, which enhanced the probability of occurrence of No-Chaco jet events.
 
-Numerical forecasts skillfully predicted enhanced risk of heavy rainfall at the seasonal scale, likely guided by ENSO signal, but biases in the spatial patterns suggest that models are not well capturing the physical interactions between the Pacific and the Atlantic basins.
-At sub-seasonal time scales, uncorrected model forecasts of rainfall had limited skill beyond 10-15 days, though use of Model Output Statistics -- particularly the PCR and CCA methods that correct both spatial patterns and magnitudes -- substantially improved forecast skill.
+Numerical forecasts skillfully predicted enhanced risk of heavy rainfall at the seasonal scale, likely guided by \gls{enso} signal, but biases in the spatial patterns suggest that models are not well capturing the physical interactions between the Pacific and the Atlantic basins.
+At sub-seasonal time scales, uncorrected model forecasts of rainfall had limited skill beyond 10-15 days, though use of Model Output Statistics - particularly the \gls{pcr} and \gls{cca} methods that correct both spatial patterns and magnitudes - substantially improved forecast skill.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{table*}
+\begin{table}
 %
 \caption{
-	Model Output Statistics (MOS) methods used to correct the ECMF sub-seasonal forecasts.
+	\Gls{mos} methods used to correct the \gls{ecmwf} sub-seasonal forecasts.
 	Spatial domain for predictand is always the same (\SIrange{39}{17}{\degree S}; \SIrange{66}{49}{\degree W}).
-	Two initializations are used: Nov 12th and 16th, 2015.} \label{tab:mos-methods}
+	Two initializations are used: Nov 12th and 16th, 2015.
+}\label{tab:mos-methods}
 \begin{center}
@@
 		%
-\emph{PCR} & \SIrange{60}{0}{\degree S}; \SIrange{80}{30}{\degree W} & Linear combination of model's EOFs, computed using both initializations as independent predictors (10 EOFs).\\
+		\emph{PCR} & \SIrange{60}{0}{\degree S}; \SIrange{80}{30}{\degree W} & Linear combination of model's \glspl{eof} computed using both initializations as independent predictors (10 \glspl{eof}).\\
 		%
 		\emph{CCA} & \SIrange{60}{0}{\degree S}; \SIrange{80}{30}{\degree W} & Canonical modes computed using both initializations as independent predictors.
-(10 predictor EOFs, 4 predictand EOFs, 4 canonical modes) \\
+(10 predictor \glspl{eof}, 4 predictand \glspl{eof}, 4 canonical modes) \\
 		%
@@
 \end{center}
-\end{table*}
+\end{table}
 
@@
 	\caption{
-		Loadings of the four leading EOFs of daily NDJF \SI{850}{\hecto\pascal} streamfunction over the weather typing region shown in \cref{fig:study-area}.
+		Loadings of the four leading \glspl{eof} of daily NDJF \SI{850}{\hecto\pascal} streamfunction over the weather typing region shown in \cref{fig:study-area}.
 		Parentheses indicate percentage of total variance explained by each EOF.
@@
 		Bottom row (k-o) shows the 2AFC skill score for each grid cell; a value greater than 50 indicates that the model outperforms climatology.
-		Columns separate different MOS models except ``Raw'', which indicates the uncorrected S2S model output.
+		Columns separate different \gls{mos} models except ``Raw'', which indicates the uncorrected S2S model output.
 		For all rows the grid cells which experienced a \nth{90} percentile exceedance for 1-7 December 2015 are outlined in black.
@@
 	\caption{
-		Anomalous probability of occurrence of each weather type concurrent with observance of each MJO phase.
+		Anomalous probability of occurrence of each weather type concurrent with observance of each \gls{mjo}  phase.
 		MJO phases are considered here only when the amplitude $\qty(\sqrt{\text{RMM1}^2 + \text{RMM2}^2})$ is greater than 1.
 		When amplitude $< 1$, the phase is defined as neutral (0).
-		Plots are shown separately for El Ni\~no, La Ni\~na, and neutral ENSO months, defined by the NINO 3.4 index having a value above 1 or below -1.
+		Plots are shown separately for El Ni\~no, La Ni\~na, and neutral \gls{enso} months, defined by the NINO 3.4 index having a value above 1 or below -1.
 		Only values which are significant at $p<0.10$, calculated using a bootstrap with 3000 samples, are shown.

### 2018-03-05 reports/writeup/paper.tex
update figures;

@@
 During the warm season (NDJF) a large fraction of rainfall, and nearly all heavy rainfall, in the Paraguay River Basin is associated with mesoscale convection \citep{Velasco1987}.
-Previous studies of organized convection and precipitation across subtropical continental South America have found close correspondence to the exit region of the low-level jets \citep{Saulo2007,Salio2007,Marengo2004,Velasco1987}, which is influenced in both summer and winter by midlatitude baroclinic wave trains that interact with the Andes topography to generate orographically bound cyclones and northerly low-level flow \citep{Campetella2002,Seluchi2006,Boers2013,Boers2014}.
+Previous studies of organized convection and precipitation across subtropical continental South America have found close correspondence withf the exit region of the low-level jets \citep{Saulo2007,Salio2007,Marengo2004,Velasco1987}, which is influenced in both summer and winter by midlatitude baroclinic wave trains that interact with the Andes topography to generate orographically bound cyclones and northerly low-level flow \citep{Campetella2002,Seluchi2006,Boers2013,Boers2014}.
 The strength and direction of this moisture transport varies substantially between events; \gls{sallj} exit regions range from central Argentina \citep[``Chaco Jet Events'';][]{Salio2002} to Paraguay and southeastern Brazil \citep[``No-Chaco Jet Events'';][]{Vera2006}.
@@
 During \gls{sacz} conditions, strong convergence is observed over the Amazon basin with divergence over southwestern Brazil, northern Argentina and Paraguay \citep{Herdies2002,Carvalho2010}; the opposite is true for so-called NoSACZ conditions.
-\Gls{sacz} occurrence is related to westerly wind regimes, as well as ``active'' and ``break'' periods of the South American Monsoon System \citep{Marengo2004}.
+\Gls{sacz} occurrence is related to westerly wind regimes over South East South America, as well as ``active'' and ``break'' periods of the South American Monsoon System \citep{Marengo2004}.
 
@@
 This ensures that the time steps in the reanalysis and rainfall data sets are the same, but means that a day is defined as beginning at 12:00 GMT.
-Since most summer rainfall in this region occurs overnight, this end-of-day time (which translates to approximately 8:00 AM locally depending on the exact time zone) tends to separate distinct events.
+Since most summer rainfall in this region occurs overnight \citep{Salio2007,Vera2006}, this end-of-day time (which translates to approximately 8:00 AM locally depending on the exact time zone) tends to separate distinct events.
 Oceanic \gls{sst} patterns are explored at the monthly time step using the \SI{1}{\degree} NOAA OI.v2 dataset \citep{Reynolds2002}.
@@
 
-The seasonal predictions used were ``flexible format'' forecasts provided by the \gls{iri}.
+The seasonal predictions used  are known as ``flexible format'' forecasts, provided by the \gls{iri} using a multi-model ensemble approach, with bias-corrected retrospective probabilistic forecasts produced using a total of 144 members forced by evolving sea-surface temperatures, and 68 members forced by persisted sea-surface temperatures; for details see \citet{Barnston2010}.
+Flexible format means that the user of these forecasts can arbitrarily choose  particular thresholds (percentiles) to compute the  probability of exceedance (or non-exceedance) from the complete  probability density function of the climatological distribution,  rather than using the more common tercile categories.
+The DJF 2015-2016 forecasts analyzed were produced in  November 2015.
+Due to the short sample of flexible format forecasts available (only for 2012-2016) at the time of writing  this paper, no verification was performed for these seasonal predictions.
+These forecasts are provided at a horizontal  resolution of \SI{2.5}{\degree}.
 This product provides the probability of exceedance (or non-exceedance) of a user-specified percentile of the climatological distribution.
@@
 Due to the short record of flexible format forecasts (2012-2016), no verification was performed for seasonal predictions.
-These forecasts are provided at a horizontal resolution of \SI{2.5}{\degree}.
 
@@
 We refer to the resulting clusters as \glspl{wt}.
+From a physical point of view, the K-means algorithm helps identify typical atmospheric circulation patterns in the EOF-filtered field via clustering of days with similar streamfunction configurations.
+These clusters can also be understood as proxies  of the available states of the system , or the most frequently visited trajectories in the phase space of the physical system \citep{Munoz2015,Munoz2016a,Munoz2017}.
 
@@
 
-For the purposes of \gls{mos} corrections, the predictand (variable to forecast) is the observed rainfall for the target period of interest, and the predictor (variable to be corrected) is the raw model rainfall for the same period.
+For the purposes of \gls{mos} corrections, the predictand (variable to forecast) is the observed rainfall for the target period of interest, and the predictor (variable to be corrected) is the uncorrected \gls{s2s} model forecast rainfall for the same period.
 Exceedance of the \nth{90} percentile during the 1995-2014 period is used to define the heavy event cases.
-We use the same spatial domain [\SIrange{39}{17}{\degree S}; \SIrange{66}{49}{\degree W}] for both the predictor and the predictand, except for the \gls{pcr} and \gls{cca} cases, in which a larger domain [\SIrange{0}{60}{\degree S}; \SIrange{80}{30}{\degree W}] was used to better capture the spatial patterns in the raw model field.
-A variety of domains and ways to combine initialization times were explored; the best results were selected in terms of the corresponding Kendall's $\tau$ rank correlation coefficient.
+We use the same spatial domain [\SIrange{39}{17}{\degree S}; \SIrange{66}{49}{\degree W}] for both the predictor and the predictand, except for the \gls{pcr} and \gls{cca} cases, in which a larger domain [\SIrange{0}{60}{\degree S}; \SIrange{80}{30}{\degree W}] was used to better capture the spatial patterns in the uncorrected \gls{s2s} model forecast field.
+A variety of domains and ways to combine initialization times were explored; the best results were selected in terms of the corresponding Kendall's $\tau$ rank correlation coefficient between observations and hindcasts.
 A summary of the final candidate predictors found to be most skillful for each \gls{mos} model is presented in \cref{tab:mos-methods}.
@@
 
-The \gls{iri}'s seasonal predictions are already provided with bias correction \citep{Barnston2010}, and thus we did not perform any further \gls{mos} on seasonal rainfall fields.
+The \gls{iri}'s seasonal forecasts are already provided with spatial \gls{mos} corrections of systematic errors of the individual models in the ensemble via \gls{cca} \citep{Barnston2010}, and thus we did not perform any further \gls{mos} on the seasonal rainfall fields.
 
@@
 As the system progresses, the jet below \SI{20}{\degree S} transitions from predominantly meridional flow (``Chaco Jet''; $t=-1$ to predominantly zonal flow (``No-Chaco Jet'').
-Analysis for the \nth{95} or 99th percentiles of daily rainfall (not shown) yield similar results, implying that the synoptic mechanism for the most heavy events is not fundamentally distinct from the mechanism for moderate-intensity events.
+The pattern resembles composites identified using one standard deviation exceedances of rainfall at \SI{60}{\degree W}, \SI{30}{\degree S} \citep{Liebmann2004} and analysis for the \nth{95} or 99th percentiles of daily rainfall (not shown) yield similar results, implying that the synoptic mechanism for the most heavy events is not fundamentally distinct from the mechanism for moderate-intensity events.
 This mean field, like all composites, masks between-event variation but exploration of individual events (not shown) shows that the core features identified are generally present.
@@
 
-\Glspl{wt} 5 and 3 are nearly inverses of \glspl{wt} 1 and 4, respectively, and are associated with dry anomalies over the Lower Paraguay River Basin.
-The fact that they are not exact inverses suggest important nonlinearities in the system.
+\Glspl{wt} 5 and 3 look loosely inverses of \glspl{wt} 1 and 4, respectively, and are associated with dry anomalies over the Lower Paraguay River Basin.
+The fact that they are not exact inverses suggest important non-linearities in the system.
+Weather types 1 and 5 resemble the two phases of the South American ``seesaw'' dipole, which is related to the \gls{sacz} \citep{Nogues-Paegle1997}.
 Finally, \glspl{wt} 2 and 6 are related to a high-pressure configuration bringing below-average rainfall over most of Brazil, and a dipolar pattern conducive to above-average rainfall over central Brazil, respectively (\cref{fig:wt-composite}).
@@
 As indicated in \cref{fig:rain-wt}, the most heavy rainfall occurred during \glspl{wt} 1 and 4.
-During NDJF 2015-16 \glspl{wt} (Chaco and No-Chaco jet extensions, respectively),  occurred more frequency than their climatology (supplemental table S1)\footnote{check}; \gls{wt} 2 also occurred more frequently than its climatology, largely due to a long sequence in February 2016.
+During NDJF 2015-16 \glspl{wt} (Chaco and No-Chaco jet extensions, respectively),  occurred more frequently than their climatology (supplemental table S1)\footnote{check}; \gls{wt} 2 also occurred more frequently than its climatology, largely due to a long sequence in February 2016.
 In mid-January 2016, during a sequence of persistent low rainfall, \gls{wt} 3 featured persistently, leading to heavy rainfall over central Brazil (not shown) and negative rainfall anomalies over the Lower Paraguay River Basin.
@@
 During mid to late December 2015, \glspl{wt} 1 and 5 produced rainfall but the intermittent dry days were \gls{wt} 2.
-During early January 2016 \gls{wt} 3 led to persistent low rainfall.
+During mid to late January 2016 \gls{wt} 3 led to persistent low rainfall.
 
@@
 
-Comparison of long-term skill between the raw model output and both extended logistic regression models shows similar results.
-Reliability, resolution and uncertainty, as measured by the ignorance score (\cref{fig:subs-prob-fcst} middle row, first three columns), suggests slight skill improvement in southern Brazil, deterioration in Argentina and Uruguay, and basically the same as the raw model output for Paraguay and southeastern Bolivia.
+Comparison of long-term skill between the uncorrected \gls{s2s} model forecast output and both extended logistic regression models shows similar results.
+Reliability, resolution and uncertainty, as measured by the ignorance score (\cref{fig:subs-prob-fcst} middle row, first three columns), suggests slight skill improvement in southern Brazil, deterioration in Argentina and Uruguay, and basically the same as the uncorrected \gls{s2s} model forecast for Paraguay and southeastern Bolivia.
 Changes in forecast discrimination exhibited by the extended logistic models, as measured via the 2AFC score (\cref{fig:subs-prob-fcst} bottom row, first three columns)\footnote{update to letters}, are  null.
@@
 
-The 20-year based skill maps of probabilistic forecasts (see \Cref{fig:subs-prob-fcst}, middle and bottom row) computed with these two \gls{eof}-based models are very similar to each other, both in terms of the reliability, resolution and uncertainty (Ignorance Score; second row, fourth and fifth columns) and discrimination (2AFC score; middle row, fourth and fifth columns).
-Compared to the raw model output and the logistic regression models, \gls{pcr} and \gls{cca} models perform quite well (\cref{fig:subs-prob-fcst}, middle and bottom row).\footnote{Letters}
+The 20-year based skill maps of probabilistic forecasts computed with these two \gls{eof}-based models are very similar to each other, both in terms of the reliability, resolution and uncertainty measured by the Ignorance Score, and discrimination measured by the 2AFC score (\cref{fig:subs-prob-fcst}.
+Compared to the uncorrected \gls{s2s} model forecast and the logistic regression models, \gls{pcr} and \gls{cca} models perform quite well (\cref{fig:subs-prob-fcst}, middle and bottom row).\footnote{Letters}
 The enhanced skill is achieved through the spatial corrections via the \gls{eof}-based regressions, which - in contrast with the extended logistic models - use information from multiple grid-boxes, and thus the original forecasts are not (necessarily) calibrated monotonically.
@@
 
-Although many of the individual rainfall events of NDJF 2015-16 were intense, they were nontheless driven by the climatological mechanism for heavy rainfall shown in \cref{fig:lagged-rain} and discussed in more detail in previous studies \citep{Saulo2007,Salio2007,Marengo2004,Velasco1987}.
+Although many of the individual rainfall events of NDJF 2015-16 were intense, they were nontheless driven by the climatological mechanism for heavy rainfall and intense convection shown in \cref{fig:lagged-rain} and discussed in more detail in previous studies \citep{Saulo2007,Salio2007,Marengo2004,Velasco1987}.
 Consequently, the most striking hydrometeorlogical feature of this season, and the leading cause of the extreme flooding, was the persistence of the heavy rainfall and the manner in which it switched ``on'' and ``off'' during the season (\cref{fig:rain-wt}).
-Understanding the mechanisms and possible cross-timescale interactions between climate drivers that drove this behavior has the potential of improving model forecast skill, particularly at sub-seasonal timescales, presently a growing area of active research.
+Improved understanding of the physical mechanisms, and their cross-timescale interactions, that led to this behavior may support both improved disaster preparedness \citep{Hellmuth2011,Goddard2014} and improved model forecast skill, particularly at sub-seasonal timescales which is an area of active research interest.
 
@@
 Previous studies on the South American Low-Level Jet \citep[e.g.,][]{Vera2006} and the modulation of rainfall in southeastern South America by extratropical transient wave trains during El Ni\~no years \citep{Barreiro2017} emphasize the importance of Pacific-Atlantic interaction for forecasting climate events in this (and other) region(s).
-A persistent dipolar \gls{sst} anomaly in the central southern Atlantic Ocean favors the occurrence of \gls{wt} 4 by blocking transient extratropical wave activity coming the Pacific, facilitating transitions from Chaco jet events (\gls{wt} 1) to No-Chaco jet events (\gls{wt} 4) via enhanced low-level wind circulation from southern Brazil towards the Atlantic, and back to north-east Brazil and the Amazon (see \cref{fig:chaco-nochaco}) due to land-sea temperature contrasts.
+A persistent dipolar \gls{sst} anomaly in the central southern Atlantic Ocean favors the occurrence of \gls{wt} 4 by blocking transient extratropical wave activity coming from the Pacific, facilitating transitions from Chaco jet events (\gls{wt} 1) to No-Chaco jet events (\gls{wt} 4) via enhanced low-level wind circulation from southern Brazil towards the Atlantic, and back to north-east Brazil and the Amazon (see \cref{fig:chaco-nochaco}) due to land-sea temperature contrasts.
 We illustrate this mechanism in \cref{fig:chaco-nochaco} and note that it is consistent with the mechanism found to produce heavy rainfall in the Lower Paraguay River Basin (\cref{fig:lagged-rain}) and with previous studies \citep[e.g.,][]{Vera2006,Salio2002,Liebmann2004}.
 This dipole is present in a \gls{sst} anomaly composite of the 15 NDJF months with the most frequent occurrence of \gls{wt} 4 (shown in \cref{fig:wt4-months}), and it does not change substantially if more or fewer months are used for the composite.\footnote{EDIT THIS PARAGRAPH}
-The pattern is also consistent with the observed dipolar structure in \gls{sst} anomalies that was present in the central southern Atlantic Ocean since mid November 2015 to roughly the end of December 2015, associated with an anomalous anti-cyclonic configuration localized around \SI{35}{\degree W} and \SI{30}{\degree S}.
+The pattern is also consistent with the observed dipolar structure in \gls{sst} anomalies that was present in the central southern Atlantic Ocean since mid November 2015 to roughly the end of December 2015, associated with an anomalous anticyclonic configuration localized around \SI{35}{\degree W} and \SI{30}{\degree S}.
 
@@
 \begin{figure*}
+	\noindent\includegraphics[width=6.5in]{anomalies_ndjf1516.pdf}
+	\caption{
+		Monthly composite anomalies observed during NDJF 2015-16.
+		Top row (a-d) shows streamfunction anomalies at \SI{850}{\hecto\pascal}.
+		Bottom row (e-h) shows rainfall anomalies, in units of \si{\milli\meter\per\day}.
+	}\label{fig:anomalies}
+\end{figure*}
+
+\begin{figure*}
 	\noindent\includegraphics[width=6.5in]{study_area.jpg}
@@
 \begin{figure*}
-	\noindent\includegraphics[width=6.5in]{anomalies_ndjf1516.pdf}
-	\caption{
-		Monthly composite anomalies observed during NDJF 2015-16.
-		Top row (a-d) shows streamfunction anomalies at \SI{850}{\hecto\pascal}.
-		Bottom row (e-h) shows rainfall anomalies, in units of \si{\milli\meter\per\day}.
-	}\label{fig:anomalies}
-\end{figure*}
-
-\begin{figure*}
 	\noindent\includegraphics[width=3.5in]{seasonal_forecast.pdf}
 	\caption{
-		Seasonal model forecast for probability of exceedance of \nth{90} percentile of DJF rainfall.
+		Seasonal model forecast for probability of exceedance of \nth{90} percentile of DJF rainfall, as issued in November 2015.
 		Color indicates the forecast probability of exceeding the \nth{90} percentile of climatological rainfall during DJF 2015-16 -- this is presented as the odds ratio as defined in \cref{eq:odds-ratio}.
@@
 	\caption{
-		Chiclet diagram \citep{Carbin:2016fx} of ensemble-mean precipitation anomaly forecasts over the Lower Paraguay River Basin (see \cref{fig:study-area}) from ECMWF S2S forecast data, as a function of the forecast target date (horizontal axis) and lead time (vertical axis).
+		Chiclet diagram \citep{Carbin:2016fx} of ensemble-mean precipitation anomaly forecasts over the Lower Paraguay River Basin (see \cref{fig:study-area}) from uncorrected \gls{ecmwf} \gls{s2s} model forecast data, as a function of the forecast target date (horizontal axis) and lead time (vertical axis).
 		Time series of CPC daily mean precipitation over the same area is plotted with $y$-axis inverted; horizontal black line denotes NDJF climatology.
@@
 		Most jet events are of the ``Chaco'' type, particularly when SST anomalies in the central southern Atlantic Ocean (a, see green box) are weak.
-		When a dipolar SST anomaly occurs in the central southern Atlantic with the warmer pole equatorward, the meridional temperature gradient and sea-land temperature contrasts establish an anti-cyclonic circulation (dot-dashed line) conducive to increased occurrence of No-Chaco jet events (b).
+		When a dipolar SST anomaly occurs in the central southern Atlantic with the warmer pole equatorward, the meridional temperature gradient and sea-land temperature contrasts establish an anticyclonic circulation (dot-dashed line) conducive to increased occurrence of No-Chaco jet events (b).
 		Other SST anomaly configurations tend to be present outside the green box (not shown).

### 2018-03-05 src/analyze/plot_mjo_ts.py

@@ -1,0 +1,58 @@
+#!/usr/bin/env python
+# -*- coding: utf-8 -*-
+"""Plot the MJO Time Series
+"""
+
+import argparse
+import xarray as xr
+import matplotlib.pyplot as plt
+import numpy as np
+import colorcet as cc
+
+parser = argparse.ArgumentParser() #pylint: disable=C0103
+parser.add_argument("--outfile", help="the filename of the data to save")
+parser.add_argument("--mjo", help="The MJO Data")
+
+def main():
+    # Read in raw data
+    args = parser.parse_args()
+    mjo = xr.open_dataset(args.mjo).sel(time=slice('2015-12-01', '2016-02-29'))
+    dt = np.arange(mjo['time'].size)
+
+    # Set up axes
+    fig, ax = plt.subplots(nrows=1, ncols=1, figsize=(8.5, 6.5))
+
+    # Add lines and circles for MJO phases
+    ax.axvline(0, c='gray')
+    ax.axhline(0, c='gray')
+    ax.plot([-100, 100], [-100, 100], c='gray')
+    ax.plot([-100, 100], [100, -100], c='gray')
+    circle = plt.Circle((0, 0), 1, color='gray', fill=False)
+    ax.add_artist(circle)
+
+    # Plot MJO
+    ax.plot(mjo['RMM1'], mjo['RMM2'], c='gray')
+    C = ax.scatter(mjo['RMM1'], mjo['RMM2'], c=dt, cmap=cc.cm['inferno'], s=45)
+    
+    # Set up axes
+    ax.set_aspect('equal')
+    ax.set_xlim([-3, 3])
+    ax.set_ylim([-3, 3])
+    plt.colorbar(C, label='Days Since 2015-11-01')
+    ax.set_xlabel('MJO RMM1')
+    ax.set_ylabel('MJO RMM2')
+
+    # Annotate phases
+    for phase in np.arange(1, 9):
+        theta = np.pi * 7/8 + np.pi * phase / 4 
+        radius = 3
+        x = radius * np.cos(theta)
+        y = radius * np.sin(theta)
+        ax.text(x, y, phase, fontsize=12)
+
+    # Save
+    fig.tight_layout()
+    fig.savefig(args.outfile, bbox_inches='tight')
+
+if __name__ == "__main__":
+    main()

### 2018-03-05 src/analyze/plot_nino34.py

@@ -1,0 +1,37 @@
+#!/usr/bin/env python
+# -*- coding: utf-8 -*-
+"""Plot the NINO 3.4 Time Series
+"""
+
+import argparse
+import xarray as xr
+import matplotlib.pyplot as plt
+
+parser = argparse.ArgumentParser() #pylint: disable=C0103
+parser.add_argument("--outfile", help="the filename of the data to save")
+parser.add_argument("--nino", help="The NINO 3.4 Data")
+
+def main():
+    # Read in raw data
+    args = parser.parse_args()
+    nino = xr.open_dataarray(args.nino)
+
+    # Set up figure
+    fig, ax = plt.subplots(nrows=1, ncols=1, figsize=(10, 3.5))
+
+    # Plot full time series on left hand side
+    nino.plot(ax=ax, c='blue')
+
+    # Plot Study Time with Dots
+    nino.sel(time=slice('2015-11-01', '2016-02-01')).plot(ax=ax, marker='o', c='blue')
+
+    # Label Axes
+    ax.set_xlabel('Time')
+    ax.set_ylabel('NINO 3.4 Index')
+
+    # Save
+    fig.tight_layout()
+    fig.savefig(args.outfile, bbox_inches='tight')
+
+if __name__ == "__main__":
+    main()

### 2018-03-05 reports/writeup/Supplement.tex
Finalize supplemental information

@@
 
-This document contains several supplemental figures described in the main text.
-Further supporting information is available in the form of source code and \texttt{jupyter} notebooks at \url{github.com/jdossgollin/PYFloods}.
+This document contains several supplemental figures referenced in the main text.
+Further supporting information is available in the form of source code at \url{github.com/jdossgollin/PYFloods}.
 
@@
 		{} &  NDJF 2015-16 &  climatology \\
-		wtype &               &              \\
+		WT &               &              \\
 		\midrule
@@
 \begin{figure}
-  \includegraphics[width=\textwidth]{wt_classifiability.pdf}
+	\includegraphics[width=\textwidth]{mjo_ts.pdf}
 	\caption{
-		Classifiability index \citep[see Methods and][]{Michelangeli1995} calculated for several chosen values of $K$, representing the number of weather types produced.
-		Results are produced with only 75 simulations for each $K$.
+		Time series of MJO evolution during NDJF 2015-16 season.
+		$x$-axis shows RMM1 and $y$-axis shows RMM2.
+		The associated phases are marked with gray lines and annotated in gray letters.
+		The inner circle denotes a ``neutral'' MJO, where the amplitude is less than 1/
+		The colored circles describe the time evolution of the system, from 2015-11-01 to 2016-02-29.
+		Data from Australian Bureau of Meteorology \citep{Wheeler2004}.
+	}
+\end{figure}
+
+\begin{figure}
+	\includegraphics[width=\textwidth]{nino_34_ts.pdf}
+	\caption{
+		Monthly NINO 3.4 time series during the study period.
+		Each month from November 2015 through February 2016 is specifically marked with a blue dot.
+		Data from \citet{Kaplan1998}.
 	}
@@
 	\caption{
-		Monthly SST anomalies during major El Ni\~{n}o events.
+		Monthly SST anomalies during December of several major El Ni\~{n}o events.
 		Months shown are (a) December 1982, (b) December 1997, and (c) December 2015.
 		Units are in \si{\celsius}.
+		Data from \citet{Reynolds2002}.
+	}
+\end{figure}
+
+\begin{figure}
+	\includegraphics[width=\textwidth]{predictors_wtypes.pdf}
+	\caption{
+		Pearson correlation coefficient between several S2S predictors and weather type occurrence.
+		For daily correlation (L), correlations are computed between the binary 0-1 occurrence of each weather type and the predictor.
+		Predictors are the 8 MJO phases, encoded as a multinomial variable, and the NINO 3.4 time series, interpolated from monthly to daily values by using the most recent monthly value (no interpolation).
+		For monthly correlation (R), correlations are computed between the proportion of occurrence of each weather type and the predictor.
+	}
+\end{figure}
+
+\begin{figure}
+	\includegraphics[width=\textwidth]{wt_classifiability.pdf}
+	  \caption{
+		  Classifiability index (see Methods) calculated for several chosen values of $K$, representing the number of weather types produced.
+		  Results are produced with only 75 simulations for each $K$.
 	  }

### 2018-03-05 src/analyze/plot_mjo_ts.py

@@ -16,7 +16,7 @@ parser.add_argument("--mjo", help="The MJO Data")
 def main():
     # Read in raw data
     args = parser.parse_args()
-    mjo = xr.open_dataset(args.mjo).sel(time=slice('2015-12-01', '2016-02-29'))
+    mjo = xr.open_dataset(args.mjo).sel(time=slice('2015-11-01', '2016-02-29'))
     dt = np.arange(mjo['time'].size)
 
     # Set up axes

### 2018-03-06 src/analyze/plot_predictors_wtypes.py
final figures pending minor cosmetic tweaks

@@ -10,12 +10,15 @@ import numpy as np
 import pandas as pd
 import seaborn as sns
 import matplotlib.pyplot as plt
+from sklearn.decomposition import PCA
 
 parser = argparse.ArgumentParser() #pylint: disable=C0103
 parser.add_argument("--outfile", help="the filename of the data to save")
 parser.add_argument("--mjo", help="The MJO data")
 parser.add_argument("--nino", help="The NINO data")
-parser.add_argument("--wt", help="The Weather Type sequence")
+parser.add_argument("--psi_wtype", help="The streamfunction over the weather typing region")
+parser.add_argument("--sst", help="The SST anomaly data")
+parser.add_argument("--n_eof", type=int)
 
 def main():
     # Read in raw data
@@ -24,39 +27,49 @@ def main():
     mjo = xr.open_dataset(args.mjo).to_dataframe()[['phase']]
     mjo = pd.get_dummies(mjo['phase'])
     mjo.columns = ['MJO' + str(c) for c in mjo.columns]
+    
     nino = xr.open_dataarray(args.nino).to_dataframe(name='NINO 3.4')
     nino_daily = nino.resample('1D').ffill()
-    wt = xr.open_dataarray(args.wt).to_dataframe(name='Weather Type')
-    wt = pd.get_dummies(wt['Weather Type'])
-    wt.columns = ['WT ' + str(w) for w in wt.columns]
+    
+    psi = xr.open_dataset(args.psi_wtype)['anomaly']
+    psi_stacked = psi.stack(grid=['lon', 'lat'])
+    pca = PCA(n_components=args.n_eof).fit(psi_stacked)
+    pc_ts = pca.transform(psi_stacked)
+    pc_ts = pd.DataFrame(pc_ts, index=psi_stacked['time'])
+    pc_ts.columns = ['EOF{}'.format(i) for i in np.arange(1, args.n_eof + 1)]
+
+    sst = xr.open_dataarray(args.sst).rename({'X': 'lon', 'Y': 'lat', 'T': 'time'})
+    sst = sst.sel(lon=slice(360-30, 360-10), lat=slice(-40, -15)).mean(dim=['lon', 'lat'])
+    sst = sst.to_dataframe(name='Dipole').resample('1D').ffill()
 
     # Join everything together
-    daily = mjo.join(nino_daily).join(wt).dropna()
+    daily = mjo.join(nino_daily).join(pc_ts).join(sst).dropna()
     monthly = daily.resample('1M').mean().dropna()
 
[…]

### 2018-03-06 src/analyze/plot_sst_anomalies.py

@@ -11,6 +11,7 @@ import xarray as xr
 import numpy as np
 import matplotlib.pyplot as plt
 from matplotlib import patches
+from sklearn.decomposition import PCA
 import cartopy.crs as ccrs
 import cartopy.feature
 import colorcet as cc
@@ -21,8 +22,13 @@ import visualize as viz
 
 parser = argparse.ArgumentParser() #pylint: disable=C0103
 parser.add_argument("--outfile", help="the filename of the figure to save")
-parser.add_argument("--sst", help="the filename of the SST data")
-parser.add_argument("--uwnd", help="the filename of the weather type data")
+parser.add_argument("--psi_wtype", help="The streamfunction over the weather typing region")
+parser.add_argument("--sst", help="The SST anomaly data")
+parser.add_argument("--n_eof", type=int)
+parser.add_argument("--WTX0", type=float)
+parser.add_argument("--WTX1", type=float)
+parser.add_argument("--WTY0", type=float)
+parser.add_argument("--WTY1", type=float)
 
 def covariance_gufunc(x, y):
     cov = (
@@ -55,45 +61,67 @@ def main():
     args = parser.parse_args()
 
     sea_temp = xr.open_dataarray(args.sst).rename({'X': 'lon', 'Y': 'lat', 'T': 'time'})
-    sea_temp = sea_temp.sel(time = np.in1d(sea_temp['time.month'], [11, 12, 1, 2]))
-    uwnd = xr.open_dataset(args.uwnd).mean(dim=['lon', 'lat'])
-    uwnd = daily_to_monthly(uwnd)['raw']
-    uwnd.name = 'uwnd'
-    mrg = xr.merge([sea_temp, uwnd], join='inner') 
-    uwnd_sst_cor = pearson_correlation(mrg['ssta'], mrg['uwnd'], dim='time')
+    sea_temp = sea_temp.sel(time = np.in1d(sea_temp['time.month'], [11, 12, 1, 2])).resample(time='1D').ffill()
+    psi = xr.open_dataset(args.psi_wtype)['anomaly']
+    psi_stacked = psi.stack(grid=['lon', 'lat'])
+    pca = PCA(n_components=args.n_eof).fit(psi_stacked)
+    pc_ts = pca.transform(psi_stacked)
+    pc_ts = pd.DataFrame(pc_ts, index=psi_stacked['time'])
[…]

### 2018-03-08 reports/writeup/Supplement.tex
final figures pending minor cosmetic tweaks
final version for coauthor review

--- 2018-03-01 reports/writeup/Supplement.tex
@@
 	natbib,
-	booktabs
+	booktabs,
+	rotating, % rotate figure
 }
@@
 \renewcommand{\thetable}{S\arabic{table}}
+\newcommand{\round}[1]{\num[round-mode=places,round-precision=3]{#1}}
 \usepackage[breaklinks=true,hidelinks]{hyperref}
@@
 		\toprule
-		{} &  NDJF 2015-16 &  climatology \\
-		wtype &               &              \\
+		{} &  NDJF 2015-16 &  Climatology \\
+		WT &               &              \\
 		\midrule
-		1     &      0.272727 &     0.212849 \\
-		2     &      0.231405 &     0.184517 \\
-		3     &      0.107438 &     0.156342 \\
-		4     &      0.206612 &     0.156189 \\
-		5     &      0.132231 &     0.149665 \\
-		6     &      0.049587 &     0.140438 \\
+		1     &         0.273 &        0.213 \\
+		2     &         0.231 &        0.185 \\
+		3     &         0.107 &        0.156 \\
+		4     &         0.207 &        0.156 \\
+		5     &         0.132 &        0.150 \\
+		6     &         0.050 &        0.140 \\
 		\bottomrule
@@
 	\caption{Weather type (WT) occurrence fraction during NDJF 2015-16 and Climatology}
+\end{table}
+
+\begin{table}
+	\centering
+	\begin{tabular}{lrrrr}
+		\toprule
+		{} &  EOF 1 &  EOF 2 &  EOF 3 &  EOF 4 \\
+		WT &        &        &        &        \\
+		\midrule
+		1  & -0.325 &  1.035 & -1.015 &  0.124 \\
+		2  & -0.887 &  0.519 &  0.647 & -0.812 \\
+		3  &  1.312 &  0.325 &  0.208 &  0.075 \\
+		4  &  0.190 & -0.807 & -0.120 & -0.844 \\
+		5  & -0.307 & -0.849 & -0.741 &  0.826 \\
+		6  & -0.334 &  0.057 &  1.093 &  1.033 \\
+		\bottomrule
+		\end{tabular}
+	\caption{
+		Centroids of each weather type in 4-dimensional phase space, where each dimension is an EOF of \SI{850}{\hecto\pascal} streamfunction over the weather typing region. 
+		These centroids are chosen by choosing the $K$-means partition which maximizes the Classifiability Index; see Methods.
+		The weather type for each day is computed by projecting the \SI{850}{\hecto\pascal} onto its leading four EOFs, re-scaling the EOFs as discussed in Methods, and identifying the nearest (Euclidean distance) centroid.
+	}
 \end{table}
@@
 \begin{figure}
-  \includegraphics[width=\textwidth]{wt_classifiability.pdf}
+	\includegraphics[width=\textwidth]{mjo_ts.pdf}
 	\caption{
-		Classifiability index \citep[see Methods and][]{Michelangeli1995} calculated for several chosen values of $K$, representing the number of weather types produced.
-		Results are produced with only 75 simulations for each $K$.
+		Evolution of the MJO during NDJF 2015-16.
+		Points are plotted on RMM1 ($x$-axis) and RMM2 ($y$-axis), derived from leading EOFs of OLR fields \citep{Wheeler2004}.
+		Gray lines divide the plot into the eight phases used in the text.
+		Colors show the time evolution of the system from 01 November 2015 (blue) to 29 February 2016 (yellow).
+		Gray circle indicates the area of neutral MJO activity with amplitude $<0$.
 	}
@@
 \begin{figure}
-	\includegraphics[width=\textwidth]{nino_sst_anomalies.pdf}
+	\includegraphics[width=\textwidth]{nino_34_ts.pdf}
 	\caption{
-		Monthly SST anomalies during major El Ni\~{n}o events.
-		Months shown are (a) December 1982, (b) December 1997, and (c) December 2015.
+		Monthly NINO 3.4 time series during the study period.
+		Each month from November 2015 through February 2016 is specifically marked with a blue dot.
+		Data from \citet{Kaplan1998}.
+	}
+\end{figure}
+
+\begin{sidewaysfigure}
+	\includegraphics[width=\textwidth, height=\textheight, keepaspectratio=true]{nino_sst_anomalies.pdf}
+	\caption{
+		Monthly SST anomalies during December of three major El Ni\~{n}o events.
+		Months shown are NDJF of (a,d,g,j): 1982-83, (b,e,h,k): 1997-98, and (c,f,i,l): 2015-16.
 		Units are in \si{\celsius}.
+		SST data from \citet{Reynolds2002}.
+		Also shown are rainfall anomalies over South America, from \citet{Xie2010}.
+		Rainfall contour intervals are \SI{1}{\milli\meter\per\day}.
+	}
+\end{sidewaysfigure}
+
+\begin{figure}
+	\noindent\includegraphics[width=6.4in]{sst_anomalies.pdf}
+	\caption{
+		Pearson correlation, at monthly time step, of  sea surface temperature (SST) anomalies \citep{Reynolds2002} with four leading EOFs of \SI{850}{\hecto\pascal} streamfunction over the weather typing region, shown with a black box.
+		Central Southern Atlantic Dipole is shown with a blue box.
+		(a): EOF 1; (b): EOF2; (c): EOF3; (d): EOF4.
+	}\label{fig:sst-anomalies}
+\end{figure}
+
+\begin{figure}
+	\includegraphics[width=\textwidth]{wt_classifiability.pdf}
+	\caption{
+		Classifiability index (see Methods) calculated for several chosen values of $K$, representing the number of weather types produced.
+		Results are produced with only 75 simulations for each $K$.
 	}

### 2018-03-08 reports/writeup/paper.tex
final version for coauthor review

@@
 
-\documentclass{ametsoc}
+\documentclass[twocol]{ametsoc}
 \journal{jcli}
@@
 \newacronym{sst}{SST}{sea surface temperature}
+\newacronym{scad}{SCAD}{South Central Atlantic Dipole}
+\newacronym{lprb}{LPRB}{Lower Paraguay River Basin}
 
@@
 	Alternating sequences of enhanced moisture inflow from the South American Low-Level Jet and local convergence associated with baroclinic systems were conducive to mesoscale convective activity and enhanced precipitation.
-	These circulation patterns were favored by cross-timescale interactions of a very strong El Ni\~{n}o event, an unusually persistent Madden-Julian Oscillation in phases four and five, and the presence of a dipolar SST anomaly in the central southern Atlantic Ocean.
+	These circulation patterns were favored by cross-timescale interactions of a very strong El Ni\~{n}o event, an unusually persistent Madden-Julian Oscillation in phases four and five, and the presence of a dipole SST anomaly in the central southern Atlantic Ocean.
 	The simultaneous use of seasonal and sub-seasonal heavy rainfall predictions could have provided decision makers useful information about the start of these flooding events from at least two-to-four weeks in advance.
@@
 
-During the austral summer of 2015-16, repeated heavy rainfall events led to severe flooding in the Paraguay River basin, displacing approximately \num{170000} people \citep{Brakenridge2016} and causing tremendous damage to property and infrastructure \citep{MinisteriodeObrasPublicasyComunicacion2016}.
-Because population in South America tends to concentrate along coasts and rivers (see supplemental figure S1)\footnote{check}, flooding in the Paraguay River basin directly affects not only much of the population of Paraguay, but also of populations in Argentina and Uruguay who lie along the Paran\'{a} and la Plata rivers into which the Paraguay River drains.
-Heavy rainfall and flooding in the basin also has important implications for hydropower generation, for agriculture, and for regional water resource management.
+During the austral summer of 2015-16, repeated heavy rainfall events led to severe flooding in the \gls{lprb}, displacing approximately \num{170000} people \citep{Brakenridge2016} and causing tremendous damage to property and infrastructure \citep{MinisteriodeObrasPublicasyComunicacion2016}.
+Because population in South America tends to concentrate along coasts and rivers (supplemental figure S1\footnote{check}), flooding in the \gls{lprb} directly affects not only much of the population of Paraguay, but also of populations in Argentina and Uruguay who lie along the Paran\'{a} and la Plata rivers into which the Paraguay River drains.
+Heavy rainfall and flooding in the \gls{lprb} also has important implications for hydropower generation, for agriculture, and for regional water resource management.
 The aim of this paper is to diagnose the drivers of the November-February (NDJF) 2015-16 rainfall and flooding events, and to assess the skill of the associated subseasonal-to-seasonal predictions.
 
-The rainfall climatology of the Paraguay River basin varies strongly by season, with extratropical characteristics in the winter and monsoonal characteristics in the summer.
+The rainfall climatology of the \gls{lprb} varies strongly by season, with extratropical characteristics in the winter and monsoonal characteristics in the summer.
 In the warm season (NDJF) which is the focus of this study, circulation is dominated by the upper-tropospheric Bolivian High, the lower-level subtropical highs, the Chaco Low over northern Argentina, the \gls{sacz}, and the \gls{sallj} \citep{Grimm2009,Marengo2012}.
@@
 However, the flat topography limits the river's ability to carry the summer runoff, causing seasonal inundation of the Pantanal and distributing the river discharge in time \citep{Bravo2011,Barros2004}.
-Thus, the streamflow maxima typically occurs in phase with precipitation upstream of the Pantanal, while downstream of the Pantanal - which we define as the Lower Paraguay River Basin; see \cref{fig:study-area} - the annual peak typically occurs between April and July.
+Thus, the streamflow maxima typically occurs in phase with precipitation upstream of the Pantanal, while downstream of the Pantanal - which we define as the \acrlong{lprb}; see \cref{fig:study-area} - the annual peak typically occurs between April and July.
 
-During the warm season (NDJF) a large fraction of rainfall, and nearly all heavy rainfall, in the Paraguay River Basin is associated with mesoscale convection \citep{Velasco1987}.
-Previous studies of organized convection and precipitation across subtropical continental South America have found close correspondence withf the exit region of the low-level jets \citep{Saulo2007,Salio2007,Marengo2004,Velasco1987}, which is influenced in both summer and winter by midlatitude baroclinic wave trains that interact with the Andes topography to generate orographically bound cyclones and northerly low-level flow \citep{Campetella2002,Seluchi2006,Boers2013,Boers2014}.
+During the warm season (NDJF) a large fraction of rainfall, and nearly all heavy rainfall, in the \gls{lprb} is associated with mesoscale convection \citep{Velasco1987}.
+Previous studies of organized convection and precipitation across subtropical continental South America have found close correspondence with the exit region of the low-level jets \citep{Saulo2007,Salio2007,Marengo2004,Velasco1987}, which is influenced in both summer and winter by mid-latitude baroclinic wave trains that interact with the Andes topography to generate orographically bound cyclones and northerly low-level flow \citep{Campetella2002,Seluchi2006,Boers2013,Boers2014}.
 The strength and direction of this moisture transport varies substantially between events; \gls{sallj} exit regions range from central Argentina \citep[``Chaco Jet Events'';][]{Salio2002} to Paraguay and southeastern Brazil \citep[``No-Chaco Jet Events'';][]{Vera2006}.
 
-At sub-seasonal timescales, convection in the Paraguay River Basin is modulated by a variety of drivers, notably including the \gls{sacz} and the \gls{mjo}  .
-During \gls{sacz} conditions, strong convergence is observed over the Amazon basin with divergence over southwestern Brazil, northern Argentina and Paraguay \citep{Herdies2002,Carvalho2010}; the opposite is true for so-called NoSACZ conditions.
+At sub-seasonal timescales, convection in the \gls{lprb} is modulated by a variety of drivers, notably including the \gls{sacz} and the \gls{mjo}  .
+During \gls{sacz} conditions, strong convergence is observed over the Amazon basin with divergence over southwestern Brazil, northern Argentina and Paraguay \citep{Herdies2002,Carvalho2010}; the opposite is true for so-called No-\gls{sacz} conditions.
 \Gls{sacz} occurrence is related to westerly wind regimes over South East South America, as well as ``active'' and ``break'' periods of the South American Monsoon System \citep{Marengo2004}.
 
-At seasonal timescales, \gls{enso} is the dominant driver of convection variability in the Paraguay River Basin.
+At seasonal timescales, \gls{enso} is the dominant driver of convection variability in the \gls{lprb}.
 During El Ni\~no years, a low-level anticyclonic anomaly over central Brazil enhances occurrence of the low-level jet, favoring the development of mesoscale convective systems \citep{Velasco1987}.
@@
 The retro and real-time datasets are combined and treated as a single dataset.
-We define ``heavy'' rainfall events to be exceedances of the \nth{90} percentile; while the value is different for each grid cell, the \nth{90} percentile of area-averaged rainfall over the Lower Paraguay River Basin is approximately \SI{15}{\milli\meter\per\day}.
+We define ``heavy'' rainfall events to be exceedances of the \nth{90} percentile; while the value is different for each grid cell, the \nth{90} percentile of area-averaged rainfall over the \gls{lprb} is approximately \SI{15}{\milli\meter\per\day}.
 
@@
 Since most summer rainfall in this region occurs overnight \citep{Salio2007,Vera2006}, this end-of-day time (which translates to approximately 8:00 AM locally depending on the exact time zone) tends to separate distinct events.
+The primary atmospheric variable  used was the \SI{850}{\hecto\pascal} streamfunction, calculated directly from the wind field as described in \cref{sec:methods}.
+The streamfunction is preferable to, for example, the geopotential height $\Phi$ because $\Phi$ has weak gradients near the equator, making it difficult to visualize circulations that span from the tropics to the extratropics.
+The \SI{850}{\hecto\pascal} height level was used because it is representative of \gls{sallj} activity and moisture transport in this region \citep{Marengo2004,Salio2007}.
+
 Oceanic \gls{sst} patterns are explored at the monthly time step using the \SI{1}{\degree} NOAA OI.v2 dataset \citep{Reynolds2002}.
@@
 
-Given the behavior of the Paraguay River discussed above, we define the Lower Paraguay River Basin as the region bounded by \SIrange{-59.75}{-55.75}{\degree W} and \SIrange{26.75}{22.75}{\degree S}, as shown in \cref{fig:study-area}.
+Given the behavior of the Paraguay River discussed above, we define the \acrlong{lprb} as the region bounded by \SIrange{-59.75}{-55.75}{\degree W} and \SIrange{26.75}{22.75}{\degree S}, as shown in \cref{fig:study-area}.
 In this region, given topography and previous studies \citep{Barros2004,Bravo2011}, one might hypothesize rainfall inputs to most closely correspond to river levels at the stream gauges in \cref{fig:study-area}.
@@
 
-The $K$-means algorithm is guaranteed to converge to a local minimum of intercluster variance; to select the best partition, 500 simulations were created using the implementation in Python's \texttt{scikit-learn} package \citep{Pedregosa2012} and then the classifiability index of \citet{Michelangeli1995} was computed between each partition and the 499 others.
+The $K$-means algorithm is guaranteed to converge to a local minimum of inter-cluster variance; to select the best partition, 500 simulations were created using the implementation in Python's \texttt{scikit-learn} package \citep{Pedregosa2012} and then the classifiability index of \citet{Michelangeli1995} was computed between each partition and the 499 others.
 The partition whose classifiability index, averaged for all 499 pairwise comparisons, was the highest was selected.
@@
 
-Logistic regression models the conditional probability of binary events, and it has been widely used in \gls{mos} approaches for binary predictands \citep{Hamill2004}.
+Logistic regression models the conditional probability of binary events, and it has been widely used in \gls{mos} approaches for binary outcomes.
 Nonetheless, when using the logistic regression to address multiple thresholds via independent fits, the predicted probabilities are, in general, not mutually consistent \citep{Messner2014}.
@@
 As discussed in \citet{Bravo2011,Barros2004}, the location of the Bah\'ia Negra gauge (see \cref{fig:study-area}) in the Pantanal region means that it responds very slowly to rainfall input.
-The three downstream gauges, because they are located in the Lower Paraguay River basin respond to the rainfall forcing with a slow but steady rise.
+The three downstream gauges, because they are located in the \gls{lprb} respond to the rainfall forcing with a slow but steady rise.
 Despite several very heavy storms, the streamflow record at Asunci\'on and Pilar (which are downstream of Concepci\'on) indicates relatively little response to individual storms.
@@
 
-\Cref{fig:lagged-rain} shows time-lagged anomalies up to and after heavy rainfall dates (when area-averaged daily rainfall in the lower Paraguay River Basin exceeds its NDJF \nth{90} percentile) and is consistent with previous findings \citep{Marengo2004,Salio2007,Marwan2015}.
+\Cref{fig:lagged-rain} shows time-lagged anomalies up to and after heavy rainfall dates (when area-averaged daily rainfall in the \gls{lprb} exceeds its NDJF \nth{90} percentile) and is consistent with previous analysis of heavy rainfall and intense convection in this region \citep{Liebmann2004,Marengo2004,Salio2007,Marwan2015}.
 At $t=-2$ a mid-latitude baroclinic system approaches the South American continent, intensifying and moving to the East at $t=-1,0,1$.
@@
 
-We next use the weather typing algorithm outlined in \cref{sec:methods}\ref{sec:weather-typing} to understand particular circulations and sequences of circulations associated with heavy rainfall in the Lower Paraguay River Basin.
+We next use the weather typing algorithm outlined in \cref{sec:methods}\ref{sec:weather-typing} to understand particular circulations and sequences of circulations associated with heavy rainfall in the \gls{lprb}.
 
@@
 
-\Gls{wt} 1 represents a ``Chaco Jet'' event \citep{Salio2002} in which the maximum wind speed penetrates southward of \SI{25}{\degree S}, leading to heavy rainfall over NE Argentina and Uruguay.
-\Gls{wt} 4 is also associated with a \gls{sallj} event, but is of the ``No-Chaco'' jet type in which the wind turns to the East northward of \SI{25}{\degree S}, leading to heavy rainfall over Eastern Paraguay and SW Brazil.
+\Gls{wt} 1 describes a \gls{sallj} event in which the strongest wind penetrates southward of \SI{25}{\degree S}, leading to heavy rainfall over NE Argentina and Uruguay; this has been called a ``Chaco Jet'' event \citep{Salio2002}.
+\Gls{wt} 4 is also associated with a \gls{sallj} event, but the wind turns to the East northward of \SI{25}{\degree S}, leading to heavy rainfall over Eastern Paraguay and SW Brazil; this has been called a ``No-Chaco'' Jet event \citep{Vera2006}.
+Supplemental Table S2\footnote{Check} shows the centroids of each cluster, in the 4-dimensional phase space of the leading \glspl{eof} of \SI{850}{\hecto\pascal} streamfunction.
+Consistent with the EOF loadings (\cref{fig:eof-loading}) and weather type composites (\cref{fig:wt-composite}), \glspl{wt} 1 and 4 have very similar (near-zero) projections onto \gls{eof} 1, but very different values of \glspl{eof} 2-4, particularly \gls{eof} 2.
 
-\Glspl{wt} 5 and 3 look loosely inverses of \glspl{wt} 1 and 4, respectively, and are associated with dry anomalies over the Lower Paraguay River Basin.
+\Glspl{wt} 5 and 3 look loosely inverses of \glspl{wt} 1 and 4, respectively, and are associated with dry anomalies over the \gls{lprb}.
 The fact that they are not exact inverses suggest important non-linearities in the system.
 Weather types 1 and 5 resemble the two phases of the South American ``seesaw'' dipole, which is related to the \gls{sacz} \citep{Nogues-Paegle1997}.
-Finally, \glspl{wt} 2 and 6 are related to a high-pressure configuration bringing below-average rainfall over most of Brazil, and a dipolar pattern conducive to above-average rainfall over central Brazil, respectively (\cref{fig:wt-composite}).
+Finally, \glspl{wt} 2 and 6 are related to a high-pressure configuration bringing below-average rainfall over most of Brazil, and a dipole pattern conducive to above-average rainfall over central Brazil, respectively (\cref{fig:wt-composite}).
 
@@
 While weather typing requires simplifying the dynamics of daily circulation patterns, its advantage is that it greatly facilitates the analysis of sequences of precipitation.
-\Cref{fig:rain-wt} shows a time series of area-averaged rainfall over the Lower Paraguay River Basin for NDJF 2015-16 and the corresponding weather types.
+\Cref{fig:rain-wt} shows a time series of area-averaged rainfall over the \gls{lprb} for NDJF 2015-16 and the corresponding weather types.
 This plot shows that heavy rainfall concentrated over a period spanning from mid-November 2015 through early January 2016, with shorter peaks in late January and mid-February.
@@
 During NDJF 2015-16 \glspl{wt} (Chaco and No-Chaco jet extensions, respectively),  occurred more frequently than their climatology (supplemental table S1)\footnote{check}; \gls{wt} 2 also occurred more frequently than its climatology, largely due to a long sequence in February 2016.
-In mid-January 2016, during a sequence of persistent low rainfall, \gls{wt} 3 featured persistently, leading to heavy rainfall over central Brazil (not shown) and negative rainfall anomalies over the Lower Paraguay River Basin.
+In mid-January 2016, during a sequence of persistent low rainfall, \gls{wt} 3 featured persistently, leading to heavy rainfall over central Brazil (not shown) and negative rainfall anomalies over the \gls{lprb}.
 Thus while the intensity and persistence of heavy rainfall was atypical, the causal mechanism of the heavy rainfall observed during this season was consistent with climatology.
@@
 Relative odds as high as 7:1 are visible over southern Paraguay and Brazil, and northern Uruguay and Argentina, generally in agreement with observations.
-The model predicted only very weakly increased odds of heavy rainfall in the Pantanal region (directly North of the Lower Paraguay River Basin) and in northern Argentina at $\approx \SI{65}{\degree W}$, and missed the heavy precipitation along most of the northeastern border of Paraguay.
+The model predicted only very weakly increased odds of heavy rainfall in the Pantanal region (directly north of the \gls{lprb}) and in northern Argentina at $\approx \SI{65}{\degree W}$, and missed the heavy precipitation along most of the northeastern border of Paraguay.
 However, the regionally elevated forecast of heavy rainfall could have been used for disaster preparedness at least one month in advance.
@@
 
-\Cref{fig:chiclet} uses a Chiclet diagram \citep{Carbin2016} to visualize, as a function of lead time, the time evolution of the uncorrected, ensemble-mean rainfall anomaly forecast, spatially averaged over the Lower Paraguay River Basin.
+\Cref{fig:chiclet} uses a Chiclet diagram \citep{Carbin2016} to visualize, as a function of lead time, the time evolution of the uncorrected, ensemble-mean rainfall anomaly forecast, spatially averaged over the \gls{lprb}.
 At times greater than about two weeks, the ensemble-mean forecast is slightly wetter than climatology.
@@
 For example, with respect to the raw prediction, \gls{xlr} tends to amplify the relative odds, and to cluster and shift the forecast location of the heavy rainfall events (\cref{fig:subs-prob-fcst}, top row, first two columns); the forecast tends to be better for Uruguay, but suggests heavy rainfall in the Paraguayan Chaco, which was not present in the raw prediction.
-On the other hand, the use of the ensemble spread in the HLXR model does not help; this forecast tends to be over-confident on the events occurring in almost all the region of interest (\cref{fig:subs-prob-fcst}, top row, third column).
+On the other hand, the use of the ensemble spread in the \gls{hxlr} model does not help; this forecast tends to be over-confident on the events occurring in almost all the region of interest (\cref{fig:subs-prob-fcst}, top row, third column).
 
@@
 Reliability, resolution and uncertainty, as measured by the ignorance score (\cref{fig:subs-prob-fcst} middle row, first three columns), suggests slight skill improvement in southern Brazil, deterioration in Argentina and Uruguay, and basically the same as the uncorrected \gls{s2s} model forecast for Paraguay and southeastern Bolivia.
-Changes in forecast discrimination exhibited by the extended logistic models, as measured via the 2AFC score (\cref{fig:subs-prob-fcst} bottom row, first three columns)\footnote{update to letters}, are  null.
+Changes in forecast discrimination exhibited by the extended logistic models, as measured via the 2AFC score (\cref{fig:subs-prob-fcst}(k-o)), are  null.
 These models operate on a gridbox-by-gridbox basis to recalibrate the probabilities, and so this recalibration happens monotonically.
@@
 The 20-year based skill maps of probabilistic forecasts computed with these two \gls{eof}-based models are very similar to each other, both in terms of the reliability, resolution and uncertainty measured by the Ignorance Score, and discrimination measured by the 2AFC score (\cref{fig:subs-prob-fcst}.
-Compared to the uncorrected \gls{s2s} model forecast and the logistic regression models, \gls{pcr} and \gls{cca} models perform quite well (\cref{fig:subs-prob-fcst}, middle and bottom row).\footnote{Letters}
+Compared to the uncorrected \gls{s2s} model forecast and the logistic regression models, \gls{pcr} and \gls{cca} models perform well (\cref{fig:subs-prob-fcst}(f-o)).
 The enhanced skill is achieved through the spatial corrections via the \gls{eof}-based regressions, which - in contrast with the extended logistic models - use information from multiple grid-boxes, and thus the original forecasts are not (necessarily) calibrated monotonically.
@@
 
-Although many of the individual rainfall events of NDJF 2015-16 were intense, they were nontheless driven by the climatological mechanism for heavy rainfall and intense convection shown in \cref{fig:lagged-rain} and discussed in more detail in previous studies \citep{Saulo2007,Salio2007,Marengo2004,Velasco1987}.
-Consequently, the most striking hydrometeorlogical feature of this season, and the leading cause of the extreme flooding, was the persistence of the heavy rainfall and the manner in which it switched ``on'' and ``off'' during the season (\cref{fig:rain-wt}).
-Improved understanding of the physical mechanisms, and their cross-timescale interactions, that led to this behavior may support both improved disaster preparedness \citep{Hellmuth2011,Goddard2014} and improved model forecast skill, particularly at sub-seasonal timescales which is an area of active research interest.
+Co-occurrence of \glspl{wt} 1 and 4, particularly in late November through late December 2015, favored advection of moisture and latent energy into the \glspl{lprb}, and low-level wind shear favored mesoscale convective activity, consistent with previous analyses in this region \citep{Velasco1987,Marengo2004,Saulo2007,Salio2007}
+Although many of the individual rainfall events of NDJF 2015-16 were intense, they were nonetheless driven by the climatological mechanism for heavy rainfall and intense convection shown in \cref{fig:lagged-rain}.
+Consequently, the most striking hydrometeorlogical feature of this season, likely a key driver of the observed flooding, was the persistence of the heavy rainfall and the manner in which it switched ``on'' and ``off'' over the study region (\cref{fig:rain-wt}).
+In fact, this apparent ``on'' and ``off'' switching was manifest principally as a spatial shift in the rainfall occurrence (\cref{fig:anomalies}) consistent with the increased occurrence of \gls{wt} 3 during mid-late January 2016 (\cref{fig:rain-wt,fig:wt-composite}); this pattern has been previously described as the South American ``seesaw'' pattern \citep{Nogues-Paegle1997}.
+Although many news reports blamed the flooding on El Ni\~{n}o\footnote{\url{http://www.bbc.com/news/world-latin-america-35184793}}, NDJF 2015-16 featured more intense rainfall than previous major El Ni\~no events, and this intense rainfall persisted for a longer time .
+While the link between El Ni\~{n}o and flooding in the \gls{lprb} is consistent with previous studies of ENSO and summertime rainfall in this region \citep{Velasco1987,Grimm2000,Salio2002,Grimm2003,Carvalho2004,Grimm2009a,Bravo2011}, both the ``on''-``off'' switching and the differences from previous major El Ni\~{n}o events suggest that other physical mechanisms, and their cross-timescale interactions, are relevant for understanding and predicting future events.
 
-As a starting point, we consider the roles of \gls{enso}, which has been extensively studied in the literature \citep{Carvalho2004,Velasco1987,Salio2002,Bravo2011,Grimm2000,Grimm2003,Grimm2009a}, and the \gls{mjo}.
-Analysis of \cref{fig:eof-loading} suggests that \gls{eof} 2 approximately describes the meridional wind over the region, consistent with \gls{sallj} activity, while \gls{eof} 3 the meridional wind over the region.
+By analyzing how the joint behavior physical mechanisms modulate the probability of occurrence of certain weather types, it may be possible to better understand the drivers of this and future extreme event(s).
+As a starting point, we consider the joint role of \gls{enso}, discussed above, and the \gls{mjo}, which  previous studies have linked to the South American ``seesaw'' pattern \citep{Nogues-Paegle1997,Paegle2000,Liebmann2004}.
+During NDJF 2015-16 The NINO 3.4 index was strongly positive, representing a strong El Ni\~{n}o state (Supplemental Figures S3, S4\footnote{check}).
+The \gls{mjo} began in November 2015 in a strong phase 3 and transitioned to phase 4 before losing amplitude around 21 November (Supplemental Figure S2)\footnote{check}.
+It stayed neutral until early December where it strengthened from a weak phase 4 to a strong phase 4 ten days later.
+Maintaining a high amplitude, it transitioned through phases 4-8 and reached phase 1 in mid January 2016.
+The \gls{mjo} then weakened slightly before emerging as a mid-strength phase 4 event in late January 2016 and moving through phases 5-7.
+\Cref{fig:wt-mjo-enso} shows that \gls{wt} 1 occurs more frequently during El Ni\~{n} o years for most \gls{mjo} phases, particularly during phase 2.
+During El Ni\~no years, \gls{wt} 3 - associated with dryness over the \gls{lprb} - occurs less frequently during \gls{mjo} phases 4, 7, and 7, and more often during \gls{mjo} phase 8; this is consistent with the lack of \gls{wt} 3 during December 2015 and the frequent \gls{wt} 3 occurrence in mid-January 2016 (\cref{fig:rain-wt}).
+Detailed consideration of the role of \gls{mjo}-\gls{enso} interaction with circulation patterns over the study region is beyond the scope of this paper, but these two patterns provided background conditions favorable for the weather type sequences observed during NDJF 2015-16.
 
-These circulations are strongly modulated by \gls{enso} and the MJO: the Spearman (rank) correlation coefficient between the monthly-averaged loading \gls{eof} 2 and the NI\~NO3.4 index is \num{-.369} ($p \ll 0.01$), and the Spearman correlation between the daily loading of \gls{eof} 3 the \gls{mjo}   RMM1 index is \num{0.130} ($p \ll 0.01$).
-A key question is how the combined effects of potentially predictable seasonal to sub-seasonal climate mechanisms modify the occurrence of each weather type.\footnote{THESE CORRELATIONS ARE NOT RIGHT}
-The typical behavior is summarized by \cref{fig:wt-mjo-enso}, following the approach of \citet{Munoz2016} and \citet{Moron2015}.
-During El Ni\~{n}o years such as 2015-16, \gls{wt} 1 -- related to Chaco jet events -- occurs more frequently for almost all \gls{mjo}  phases, and particularly during phases 1, 2, and 5, consistent with observations of the year under study.
-The occurrence of \gls{wt} 3, favored during \gls{mjo} phases 8 and 1, in late January 2016 and of\gls{wt} 2, enhanced during \gls{mjo}  phases 4-7, in mid-February 2016 is also consistent with the observed \gls{mjo}  time series (supplemental figure S4)\footnote{check} and observed weather type sequences (\cref{fig:rain-wt}).
+Of course, since a large fraction of the signal in \cref{fig:wt-mjo-enso} seems to come from the \gls{enso} signal, a logical question is why NDJF 2015-16 featured more persistent and intense rainfall in the \gls{lprb} than during other major El Ni\~no events (Supplemental Figure S4\footnote{check}).
+Previous studies of the \gls{sallj} \citep[e.g.,][]{Vera2006} and the modulation of rainfall in southeastern South America by extratropical transient wave trains during El Ni\~no years emphasize the importance of Pacific-Atlantic interaction for forecasting climate events in this (and other) region(s) \citep{Barreiro2017}.
+In particular, a persistent dipolar \gls{sst} anomaly in the central southern Atlantic Ocean may favor the occurrence of \gls{wt} 4 by blocking transient extratropical wave activity from the Pacific, facilitating transitions from ``Chaco'' jet events (\gls{wt} 1) to ``No-Chaco'' jet events (\gls{wt} 4) via enhanced low-level wind circulation from southern Brazil towards the Atlantic, and back to north-east Brazil and the Amazon (see \cref{fig:chaco-nochaco}) due to land-sea temperature contrasts.
+We illustrate a schematic of this mechanism in \cref{fig:chaco-nochaco} and note that it is consistent with the mechanism found to produce heavy rainfall in the \gls{lprb} (\cref{fig:lagged-rain}) and with previous studies \citep[e.g.,][]{Vera2006,Salio2002,Liebmann2004}.
+We refer to his mechanism as the \gls{scad} and measure it as the mean meridional \gls{sst} gradient over the box shown in \cref{fig:chaco-nochaco}.
+Supplemental figure S4\footnote{check} shows the \gls{sst} anomalies observed during NDJF 2015-16, and indeed the mechanism illustrated in \cref{fig:chaco-nochaco} was present -- particularly in December 2015 when the most intense rainfall occurred.
+Examination of \glspl{sst} correlations with the leading \glspl{eof} of \SI{850}{\hecto\pascal} streamfunction (Supplemental Figure S5)\footnote{check} show that this dipole pattern is negatively correlated with \glspl{eof} 2, 3, and 4.
+As supplemental table S2\footnote{check} shows, the centroid of \gls{wt} in phase space is a strongly negative value of \glspl{eof} 2 and 4, and a weakly negative value of \gls{eof} 3; by contrast \gls{wt} 1 is defined by a strongly positive value of \gls{eof} 2, strongly negative value of \gls{eof} 3, and weakly positive value of \gls{eof} 4.
+This suggests that not only did \gls{enso}-\gls{mjo} conditions favor \gls{sallj} activity, but Atlantic-Pacific conditions specifically favored  \gls{wt} 4 occurrence, helping to explain why the most intense rainfall anomalies occurred specifically in the \gls{lprb}.
 
-While the occurrence of \gls{wt} during NDJF 2015-16 is well explained by \gls{enso} and \gls{mjo}  variability, these features alone do not explain the frequent occurrence of \gls{wt} 4, the ``No-Chaco'' jet event.
-Previous studies on the South American Low-Level Jet \citep[e.g.,][]{Vera2006} and the modulation of rainfall in southeastern South America by extratropical transient wave trains during El Ni\~no years \citep{Barreiro2017} emphasize the importance of Pacific-Atlantic interaction for forecasting climate events in this (and other) region(s).
-A persistent dipolar \gls{sst} anomaly in the central southern Atlantic Ocean favors the occurrence of \gls{wt} 4 by blocking transient extratropical wave activity coming from the Pacific, facilitating transitions from Chaco jet events (\gls{wt} 1) to No-Chaco jet events (\gls{wt} 4) via enhanced low-level wind circulation from southern Brazil towards the Atlantic, and back to north-east Brazil and the Amazon (see \cref{fig:chaco-nochaco}) due to land-sea temperature contrasts.
-We illustrate this mechanism in \cref{fig:chaco-nochaco} and note that it is consistent with the mechanism found to produce heavy rainfall in the Lower Paraguay River Basin (\cref{fig:lagged-rain}) and with previous studies \citep[e.g.,][]{Vera2006,Salio2002,Liebmann2004}.
-This dipole is present in a \gls{sst} anomaly composite of the 15 NDJF months with the most frequent occurrence of \gls{wt} 4 (shown in \cref{fig:wt4-months}), and it does not change substantially if more or fewer months are used for the composite.\footnote{EDIT THIS PARAGRAPH}
-The pattern is also consistent with the observed dipolar structure in \gls{sst} anomalies that was present in the central southern Atlantic Ocean since mid November 2015 to roughly the end of December 2015, associated with an anomalous anticyclonic configuration localized around \SI{35}{\degree W} and \SI{30}{\degree S}.
+This Atlantic-Pacific interaction may also be a source of dynamical model biases in the representation of rainfall events in the region.
+In order to adequately forecast rainfall in certain parts of southeastern South America during El Ni\~no years, models need to reproduce not one but two stationary wave trains, originating from central and eastern Pacific, and their interactions \citep{Barreiro2017}.
+The frontal activity is associated with the extension, shape and variability of the \gls{sacz}, and could be blocked by the presence of anomalous anti-cyclones off the coast of Brazil, like the ones described in reference to ``No-Chaco'' jet events.
+Other mechanisms that have been known to modulate rainfall signals in this region include the \gls{sacz} \citep{Carvalho2004,Munoz2015,Munoz2016} and land-biosphere-atmosphere interactions \citep{Grimm2000,Grimm2007} which also tend to be poorly represented in models \citep{Green2017,Koster2004}.
 
-A potential source of dynamical model biases in the representation of rainfall events in the region is related to simulation of the extratropical wave activity mentioned before.
-For example, in order to adequately forecast rainfall in certain parts of southeastern South America during El Ni\~no years, models need to reproduce not one but two stationary wave trains, originating from central and eastern Pacific, respectively \citep{Barreiro2017}.
-The frontal activity is associated with the extension, shape and variability of the \gls{sacz}, and could be blocked by the presence of anomalous anti-cyclones off the coast of Brazil, like the ones described in reference to No-Chaco jet events.
+Finally it is of interest to consider the link between the observed rainfall events and the observed flooding.
+Although we motivated this work by describing the impacts of severe flooding in the \gls{lprb}, the analysis presented has focused on climate drivers of rainfall.
+As explained in \cref{sec:diagnostics}\ref{sec:flood}, in this region the flat topography (\cref{fig:study-area}) means that the Lower Paraguay River reacts slowly to rainfall \citep{Bravo2011,Barros2004}, explaining the slow but steady rise in river levels from mid November 2015 to early January 2016, as shown in \cref{fig:streamflow}.
+The observed flood peaks during 2015-16 also seem to occur in the context of an active phase of a multi-decadal oscillation, possibly associated with low-frequency Pacific activity \citep{Collischonn2001,Huang2005}; groundwater dynamics are also important in explaining this behavior \citep{Santos2016}.
+Parsing the relative impacts of deforestation and land use changes in the river basin, installation of hydroelectric generation at the Itaipu and Yacyreta sites, river channel modification, antecedent conditions, and climate variability on flood levels will require gathering improved hydrological data and building a comprehensive system model, which is beyond the scope of this paper.
 
-Other mechanisms that have been known to modulate rainfall signals in this region include the \gls{sacz} \citep{Carvalho2004,Munoz2015,Munoz2016} and land-biosphere-atmosphere interactions \citep{Grimm2000,Grimm2007} which also tend to be poorly represented in models \citep{Green2017,Koster2004}.
-The analysis of these sources of systematic errors could require a cross-timescale diagnostic approach \citep{Munoz2017}, and it is beyond the scope of the present paper.
-Analysis of La Ni\~na and \gls{enso}-neutral years (\cref{fig:wt-mjo-enso}) also suggests modulation of weather type occurrence by \gls{mjo}-\gls{enso} interactions, consistent with the cross-timescale hypothesis detailed in \citet{Munoz2015}.
-This cross-timescale interaction also helps to explain why sub-seasonal forecasts were not substantially better than seasonal ones: although the \gls{enso} signal was homogeneous throughout the season, studies of \gls{mjo}  predictability have found it to be limited beyond two to three weeks \citep{Ding2009,Vitart2014,Kim2014a}.
-
-The nonlinear dynamics introduced by the coupling of the land-biosphere-ocean-atmosphere system can lead to favored trajectories through particular regions of phase space; these can be understood as ``weather regimes'' \citep[see][for a comprehensive discussion]{Hannachi2017}.
-Looking in southeastern South America (slightly South and East of the Lower Paraguay River Basin defined in this study), \citet{Munoz2016} used the sequences of weather types identified for each year to analyze predictability of heavy rainfall events.
-Although this approach - as with the weather typing used in this study - simplifies the dynamics of the system, it can be a valuable tool for diagnosing and assessing model processes \citep{Munoz2017}, and for communicating model forecasts to policy-makers.
-Such an assessment is beyond the scope of this case study but offers promising opportunities for local public and private sectors to use climate model forecasts.
-
-Although we motivated this work by describing the impacts of severe flooding in the Lower Paraguay River Basin, the analysis presented has focused on climate drivers of rainfall.
-Indeed, rainfall and streamflow are related, but the link is nonlinear in space and time.
-As explained in \cref{sec:diagnostics}\ref{sec:flood}, in this region the flat topography (\cref{fig:study-area}) means that the Lower Paraguay River reacts slowly to rainfall \citep{Bravo2011,Barros2004}, explaining the slow but steady rise in river levels shown in \cref{fig:streamflow}.
-Groundwater dynamics are also important in explaining this behavior \citep{Santos2016}.
-The observed flood peaks during 2015-16 also seem to occur in the context of an active phase of a multi-decadal oscillation, possibly associated with low-frequency Pacific activity \citep{Collischonn2001,Huang2005}.
-Parsing the relative impacts of deforestation and land use changes in the river basin, installation of hydroelectric generation at the Itaipu and Yacyreta sites, river channel modification, and climate variability on flood levels will require gathering improved hydrological data and building a comprehensive system model but all of these factors may influence the timing and magnitude of flooding in this region.\footnote{Perhaps add something about antecedent conditions here and inter-decadal variability}
-
-From a policy perspective, reducing flood risk exposure in this region is key to reducing flood losses; flood events not only in 2015-16 but also in 2014 and 2017 caused substantial damage, and highlight the need for flood risk management.
+From a policy perspective, reducing flood risk exposure in this region is key to reducing flood losses; flood events not only in 2015-16 but also in 2014, 2017, and 2018 have caused substantial damage, and highlight the need for flood risk management strategies.
 Doing so will require compiling information on the properties, businesses, and infrastructure that are vulnerable to flooding.
@@
 		(a): all of South America.
-		The domains of the ``Lower Paraguay River Basin'' and the domain used for weather typing are indicated in red and blue, respectively.
-		(b): The Paraguay River.
-		As for (a), the ``Lower Paraguay River Basin'' is denoted with a red box.
-		Streamflows shown in \cref{fig:streamflow} were taken from the fours stations indicated.
+		The domains of the \acrlong{lprb} and the domain used for weather typing are indicated in red and blue, respectively.
+		(b): The \acrfull{lprb}.
+		As for (a), the \gls{lprb} is marked with a red box.
+		Streamflow time series shown in \cref{fig:streamflow} were taken from the fours stations indicated.
 		The Paraguay River and its tributaries, from the Natural Earth database (\texttt{www.naturalearthdata.com}), are also shown.
@@
 		The station names are shortened versions of those shown in \cref{fig:study-area}.
-		(a) Seasonality (orange) and time series of 2015-16 observations (black) at each stream gauge.
-			The seasonality was fit using local polynomial regression as implemented in the \texttt{locfit} package in
-			the \textbf{R} statistical programming environment \citep{Loader1999}.
-		(b) Time series of daily stage measurements from 1929 to 2016 at each station.
+		(a): Seasonality (orange) and time series of 2015-16 observations (black) at each stream gauge.
+		Seasonality was fit using local polynomial regression as implemented in the \texttt{locfit} package in the \textbf{R} statistical programming environment \citep{Loader1999}.
+		(b): Time series of daily stage measurements from 1929 to 2016 at each station.
 	}\label{fig:streamflow}
@@
 	\caption{
-		Composite anomalies associated with heavy rainfall (\nth{90} percentile exceedance of area-averaged rainfall in the Lower Paraguay River Basin).
+		Composite anomalies associated with heavy rainfall (\nth{90} percentile exceedance of area-averaged rainfall in the \acrlong{lprb}).
 		Lagged composites are shown, by column, for $t = $ \SIlist{-2;-1;0;1}{\day} relative to the date of heavy rainfall.
-		Top row (a-d) shows composite streamfunction anomalies at \SIlist{850}{\hecto\pascal}.
+		Top row (a-d) shows composite streamfunction and wind anomalies at \SIlist{850}{\hecto\pascal}.
+		Strongest 5\% of wind anomaly vectors between \SI{60}{\degree S} and \SI{10}{\degree N} (all longitudes) are also shown.
 		Bottom row (e-h) shows composite rainfall anomalies, in units of \si{\milli\meter\per\day}.
@@
 		Loadings of the four leading \glspl{eof} of daily NDJF \SI{850}{\hecto\pascal} streamfunction over the weather typing region shown in \cref{fig:study-area}.
-		Parentheses indicate percentage of total variance explained by each EOF.
+		Parentheses in sub-plot titles indicate the percentage of total variance explained by each \gls{eof}.
 	}\label{fig:eof-loading}
@@
 	\caption{
-		Composite anomalies associated with each weather typ.
-		Top row (a-f) shows streamfunction anomalies at \SIlist{850}{\hecto\pascal}.
+		Composite anomalies associated with each weather type.
+		Top row (a-f) shows streamfunction =anomalies at \SIlist{850}{\hecto\pascal}.
+		Strongest 20\% of wind anomaly vectors over the plot area are also shown.
 		Bottom row (g-l) shows rainfall anomalies, in units of \si{\milli\meter\per\day}. 
@@
 	\caption{
-		Time series of area-averaged rainfall in the Lower Paraguay River Basin (\cref{fig:study-area}) for each day of NDJF  2015-16.
-		Lines indicate the raimfall value, in units of \si{\milli\meter\per\day}.
+		Time series of area-averaged rainfall in the \acrlong{lprb} (\cref{fig:study-area}) for each day of NDJF  2015-16.
+		Lines indicate the rainfall value, in units of \si{\milli\meter\per\day}.
         The weather type corresponding to each day is indicated in an adjacent text label.
-		Dashed lines blue indicate (from bottom to top) the climatological 50th, \nth{90}, and 99th percentiles of NDJF area-averaged rain over the Lower Paraguay River basin.
+		Dashed lines blue indicate (from bottom to top) the climatological 50th, \nth{90}, and 99th percentiles of NDJF area-averaged rain over the \acrlong{lprb}.
 	}\label{fig:rain-wt}
@@
 	\caption{
-		Chiclet diagram \citep{Carbin:2016fx} of ensemble-mean precipitation anomaly forecasts over the Lower Paraguay River Basin (see \cref{fig:study-area}) from uncorrected \gls{ecmwf} \gls{s2s} model forecast data, as a function of the forecast target date (horizontal axis) and lead time (vertical axis).
+		Chiclet diagram \citep[see][]{Carbin2016} of ensemble-mean precipitation anomaly forecasts over the \acrlong{lprb} (see \cref{fig:study-area}) from uncorrected \gls{ecmwf} \gls{s2s} model forecast data, as a function of the forecast target date (horizontal axis) and lead time (vertical axis).
 		Time series of CPC daily mean precipitation over the same area is plotted with $y$-axis inverted; horizontal black line denotes NDJF climatology.
@@
 	\caption{
-		MOS-adjusted S2S model forecasts and skill scores for the methods indicated in \cref{tab:mos-methods}.
+		\Gls{mos}-adjusted \gls{s2s} model forecasts and skill scores for the methods indicated in \cref{tab:mos-methods}.
 		Top row (a-e) shows the heavy rainfall forecast for 1-7 December 2015 as the odds ratio defined in \cref{eq:odds-ratio} over the target domain.
@@
 		Bottom row (k-o) shows the 2AFC skill score for each grid cell; a value greater than 50 indicates that the model outperforms climatology.
-		Columns separate different \gls{mos} models except ``Raw'', which indicates the uncorrected S2S model output.
+		Columns separate different \gls{mos} models except for ``Raw'' (a,f,k), which indicates the uncorrected \gls{s2s} model output.
 		For all rows the grid cells which experienced a \nth{90} percentile exceedance for 1-7 December 2015 are outlined in black.
 	}\label{fig:subs-prob-fcst}
-\end{figure*}
-
-\begin{figure*}
-	\noindent\includegraphics[width=6.4in]{wt_mjo_enso.pdf}
-	\caption{
-		Anomalous probability of occurrence of each weather type concurrent with observance of each \gls{mjo}  phase.
-		MJO phases are considered here only when the amplitude $\qty(\sqrt{\text{RMM1}^2 + \text{RMM2}^2})$ is greater than 1.
-		When amplitude $< 1$, the phase is defined as neutral (0).
-		Plots are shown separately for El Ni\~no, La Ni\~na, and neutral \gls{enso} months, defined by the NINO 3.4 index having a value above 1 or below -1.
-		Only values which are significant at $p<0.10$, calculated using a bootstrap with 3000 samples, are shown.
-	}\label{fig:wt-mjo-enso}
 \end{figure*}
@@
 		Most jet events are of the ``Chaco'' type, particularly when SST anomalies in the central southern Atlantic Ocean (a, see green box) are weak.
-		When a dipolar SST anomaly occurs in the central southern Atlantic with the warmer pole equatorward, the meridional temperature gradient and sea-land temperature contrasts establish an anticyclonic circulation (dot-dashed line) conducive to increased occurrence of No-Chaco jet events (b).
+		When a dipole SST anomaly occurs in the central southern Atlantic with the warmer pole equatorward, the meridional temperature gradient and sea-land temperature contrasts establish an anticyclonic circulation (dot-dashed line) conducive to increased occurrence of No-Chaco jet events (b).
 		Other SST anomaly configurations tend to be present outside the green box (not shown).
@@
 \begin{figure*}
-	\noindent\includegraphics[width=6.4in]{sst_anomalies.pdf}
+	\includegraphics[width=\textwidth]{predictors_eofs.pdf}
 	\caption{
-		Composite sea surface temperature (SST) anomalies from \citet{Reynolds:2002iy}.
-		(a): The Pearson correlation of each SST grid cell with the time series of average zonal wind over the region indicated by the blue box.
-		(b): Sea surface temperature (SST) anomalies during December 2015.
-		Both panels show the Central Southern Atlantic with a black box; this is the same region as shown in \cref{fig:chaco-nochaco}.
-	}\label{fig:wt4-months}
+		Spearman (rank) correlation coefficient between several \gls{s2s} predictors and weather type occurrence at monthly time step.
+		These, correlations are computed between the proportion of occurrence of each weather type and the predictor monthly-mean time series.
+		For the \gls{mjo} phases, the monthly-mean time series is the proportion of occurrence of that phase for the given month.
+		Predictors are the 8 \gls{mjo} phases plus neutral \gls{mjo} (labeled MJO0), encoded as a multinomial variable, and the NINO 3.4 time series, up-sampled from monthly to daily values by using the most recent monthly value (no interpolation).
+		The \acrfull{scad} predictor, labeled Dipole, is the mean meridional gradient of \gls{sst} anomalies over the region \SIrange{30}{10}{\degree W} and \SIrange{40}{15}{\degree S}, up-sampled to daily values using the same method as the NINO 3.4 time series.
+	}\label{fig:predictor-eof}
+\end{figure*}
+
+\begin{figure*}
+	\includegraphics[width=\textwidth]{wt_mjo_enso.pdf}
+	\caption{
+		Anomalous probability of occurrence of each weather type concurrent with observance of each \gls{mjo} phase. 
+		When \gls{mjo} amplitude is less than zero, it is defined as neutral phase (0)/
+		Plots are shown separately for El Ni\~no (NINO 3.4 $>1$), La Ni\~na (NINO 3.4 $< -1$), and Neutral \gls{enso} phases.
+		Only values which are significant at $\alpha=0.10$, calculated with a bootstrap of 5000 samples, are shown.
+	}\label{fig:wt-mjo-enso}
 \end{figure*}

### 2018-03-08 src/analyze/plot_nino_sst.py

@@ -28,67 +28,102 @@ def main():
     """
     args = parser.parse_args()
 
-    sea_temp = xr.open_dataarray(args.sst).rename({'X': 'lon', 'Y': 'lat', 'T': 'time'})
+    sea_temp = xr.open_dataarray(args.sst)
     sea_temp = sea_temp.sel(time = np.in1d(sea_temp['time.month'], [11, 12, 1, 2]))
-    
-    nino_times = pd.to_datetime(['1982-12-01', '1997-12-01', '2015-12-01'])
-    sst_plot = sea_temp.sel(time=np.in1d(sea_temp['time'], nino_times))
-
     rain = xr.open_dataset('data/processed/rain.nc')['anomaly']
 
     # Plot options
     map_proj = ccrs.PlateCarree(central_longitude=-60)
     data_proj = ccrs.PlateCarree()
-    figsize = (10, 9)
+    figsize = (18, 9)
     dipole = Region(lon=[-30,-10], lat=[-10,-40])
 
-    cmap = cc.cm['coolwarm']
+    
+    nino_years = [1982, 1997, 2015]
 
-    p = sst_plot.plot.contourf(
-       x='lon', y='lat',
-       transform=data_proj,
-       col='time',
-       col_wrap=1,
-       subplot_kws={'projection': map_proj},
-       cmap=cmap,
-       levels=np.linspace(-3.5, 3.5, 15),
-       figsize=figsize
+    # Set up plots
+    fig, axes = plt.subplots(
+        nrows=4, ncols=len(nino_years), subplot_kw={'projection': map_proj},
+        figsize=figsize, sharex=True, sharey=True
     )
 
-    keywords = dict(add_colorbar=False, add_labels=False, vmin=-10, vmax=10, cmap='BrBG', transform=data_proj, levels=np.linspace(-10, 10, 21))
-    rain_sub = rain.sel(time=slice('1982-12-01', '1982-12-31')).mean(dim='time')
-    rain_sub.plot.contourf(ax=p.axes.flat[0], **keywords)
-    p.axes.flat[0].set_title('December 1982')
+    for i, yr in enumerate(nino_years):
+        sst_keywords = dict(
+            transform=data_proj,
+            cmap=cc.cm['coolwarm'],
+            levels=np.linspace(-3, 3, 13),
+            add_colorbar=True,
+            add_labels=False,
+            extend='both',
+        )
 
-    rain_sub = rain.sel(time=slice('1997-12-01', '1997-12-31')).mean(dim='time')
[…]

### 2018-03-08 src/analyze/plot_sst_anomalies.py

@@ -60,7 +60,7 @@ def main():
     """
     args = parser.parse_args()
 
-    sea_temp = xr.open_dataarray(args.sst).rename({'X': 'lon', 'Y': 'lat', 'T': 'time'})
+    sea_temp = xr.open_dataarray(args.sst)
     sea_temp = sea_temp.sel(time = np.in1d(sea_temp['time.month'], [11, 12, 1, 2])).resample(time='1D').ffill()
     psi = xr.open_dataset(args.psi_wtype)['anomaly']
     psi_stacked = psi.stack(grid=['lon', 'lat'])
@@ -68,75 +68,45 @@ def main():
     pc_ts = pca.transform(psi_stacked)
     pc_ts = pd.DataFrame(pc_ts, index=psi_stacked['time'])
     pc_ts.columns = ['EOF{}'.format(i) for i in np.arange(1, args.n_eof + 1)]
-    pc_ts = pc_ts.to_xarray()
     mrg = xr.merge([sea_temp, pc_ts], join='inner')
-    
+    mrg = daily_to_monthly(mrg)
+
+    corr1 = pearson_correlation(mrg['ssta'], mrg['EOF1'], dim='time')
+    corr1.name = 'Pearson Correlation with EOF 1'
     corr2 = pearson_correlation(mrg['ssta'], mrg['EOF2'], dim='time')
+    corr2.name = 'Pearson Correlation with EOF 2'
     corr3 = pearson_correlation(mrg['ssta'], mrg['EOF3'], dim='time')
+    corr3.name = 'Pearson Correlation with EOF 3'
+    corr4 = pearson_correlation(mrg['ssta'], mrg['EOF4'], dim='time')
+    corr4.name = 'Pearson Correlation with EOF 4'
 
     # Plot options
     map_proj = ccrs.PlateCarree(central_longitude=-60)
     data_proj = ccrs.PlateCarree()
-    figsize = (10, 6.5)
+    figsize = (10, 12)
     dipole = Region(lon=[-30,-10], lat=[-15,-40])
     wt_region = Region(lon=[args.WTX0, args.WTX1], lat=[args.WTY0, args.WTY1])
 
 
-    cmap = cc.cm['coolwarm']
-
-    # Set up 2 plots
+    # Set up plots
     fig, axes = plt.subplots(
-        nrows=3, ncols=1, subplot_kw={'projection': map_proj},
+        nrows=4, ncols=1, subplot_kw={'projection': map_proj},
[…]

### 2018-03-08 src/analyze/plot_study_area.py

@@ -19,6 +19,7 @@ import visualize as viz
 
 parser = argparse.ArgumentParser() #pylint: disable=C0103
 parser.add_argument("--outfile", help="the filename of the data to save")
+parser.add_argument("--elev", help="the path to elevation data")
 parser.add_argument("--LPRX0", type=float)
 parser.add_argument("--LPRX1", type=float)
 parser.add_argument("--LPRY0", type=float)
@@ -63,7 +64,7 @@ def main():
     wt_region = Region(lon=[args.WTX0, args.WTX1], lat=[args.WTY0, args.WTY1])
 
     # Read in raw data
-    elev = xr.open_dataarray('data/external/elevation.nc').sel(
+    elev = xr.open_dataarray(args.elev).sel(
         X=slice(-85, -32.5),
         Y=slice(-50, 15)
     )

### 2018-03-08 src/analyze/plot_wt_composite.py

@@ -21,6 +21,8 @@ parser.add_argument("--outfile", help="the filename of the data to save")
 parser.add_argument("--wt", help="The weather type data")
 parser.add_argument("--psi", help="The streamfunction data file")
 parser.add_argument("--rain", help="The rainfall data file")
+parser.add_argument("--uwnd", help="The zonal wind")
+parser.add_argument("--vwnd", help="The meridional wind")
 parser.add_argument("--WTX0", type=float)
 parser.add_argument("--WTX1", type=float)
 parser.add_argument("--WTY0", type=float)
@@ -30,8 +32,11 @@ def main():
     """Run everything
     """
     args = parser.parse_args()
-    psi = xr.open_dataset(args.psi)
+    psi = xr.open_dataset(args.psi).sel(lon=slice(-90, -30), lat=slice(-50, 10))
     prcp = xr.open_dataset(args.rain)
+    uwnd = xr.open_dataset(args.uwnd)['anomaly'].sel(lon=slice(-90, -30), lat=slice(-50, 10))
+    vwnd = xr.open_dataset(args.vwnd)['anomaly'].sel(lon=slice(-90, -30), lat=slice(-50, 10))
+    X, Y = np.meshgrid(uwnd.lon, uwnd.lat)
     w_type = xr.open_dataarray(args.wt).to_dataframe(name='wtype')
 
     wt_region = Region(lon=[args.WTX0, args.WTX1], lat=[args.WTY0, args.WTY1])
@@ -41,6 +46,7 @@ def main():
     data_proj = ccrs.PlateCarree()
     wt_unique = np.unique(w_type['wtype'])
     figsize = (14, 4.5)
+    plot_region = Region(lon = [-90, -30], lat = [-50, 10])
 
     # Get the proportion of weather types each day
     wt_counts = w_type.groupby('wtype').size().div(w_type['wtype'].size)
@@ -72,6 +78,13 @@ def main():
             add_labels=False
         )
         ax.add_patch(wt_region.as_patch(label='Weather Typing Region', color='black'))
+        
+        # add wind arrows
+        U = selector(uwnd).values
+        V = selector(vwnd).values
+        magnitude = np.sqrt(U**2 + V**2)
+        strongest = magnitude > np.percentile(magnitude, 80)
+        ax.quiver(X[strongest], Y[strongest], U[strongest], V[strongest], transform=data_proj, scale=80)
 
         # Bottom row: rainfall anomalies
[…]

### 2018-03-08 src/analyze/print_wt_prop_table.py

@@ -42,7 +42,7 @@ def main():
     wt_prop_tex = pd.DataFrame({'climatology': wt_prop_clim, 'NDJF 2015-16': wt_prop_1516})
 
     # Save to latex table
-    wt_prop_tex.to_latex(args.outfile)
+    wt_prop_tex.round(decimals=3).to_latex(args.outfile)
 
 if __name__ == "__main__":
     main()

### 2018-03-08 src/get/download_ssta.py

@@ -13,12 +13,11 @@ import pandas as pd
 parser = argparse.ArgumentParser() #pylint: disable=C0103
 parser.add_argument("--outfile", help="the filename of the data to save")
 
-def download_data(outfile):
-    """Download the monthly SST anomaly data
-    Args:
-        outfile: the name of the file to save to
+def main():
+    """Parse the command line arguments and run download_data().
     """
-    # read in the data
+    args = parser.parse_args()
+    
     url = 'http://iridl.ldeo.columbia.edu/expert/SOURCES/.NOAA/.NCEP/.EMC/.CMB/.GLOBAL/'
     url += '.Reyn_SmithOIv2/.monthly/.ssta/dods' # split lines
     data = xr.open_dataarray(url, decode_times=False) # doesn't follow time conventions
@@ -29,19 +28,19 @@ def download_data(outfile):
     month = 1 + time % 12
     time = pd.to_datetime(year * 10000 + month * 100 + 1, format="%Y%m%d")
     data['T'] = time
-    data.rename({'T': 'time'})
+    
+    data = data.rename({'X': 'lon', 'Y': 'lat', 'T': 'time'})
+    longitudes = data['lon'].values
+    longitudes[np.where(longitudes > 180)] -= 360
+    data['lon'].values = longitudes
+
+    # sort
+    data = data.sortby('lon').sortby('lat').sortby('time')
 
     # save to file
-    if os.path.isfile(outfile):
-        os.remove(outfile)
-    data.to_netcdf(outfile, format='NETCDF4', mode='w')
-
-def main():
-    """Parse the command line arguments and run download_data().
-    """
-    args = parser.parse_args()
-    outfile = os.path.abspath(args.outfile)
-    download_data(outfile=outfile)
+    if os.path.isfile(args.outfile):
+        os.remove(args.outfile)
+    data.to_netcdf(args.outfile, format='NETCDF4', mode='w')
 
 if __name__ == "__main__":
     main()

### 2018-03-08 src/process/make_weather_type.py

@@ -17,6 +17,7 @@ from numba import jit
 
 parser = argparse.ArgumentParser() #pylint: disable=C0103
 parser.add_argument("--outfile", help="the filename of the data to save")
+parser.add_argument("--table", help="the filename of the latex table to write", default=None)
 parser.add_argument("--infile", help="the input data")
 parser.add_argument("--var_xpl", type=float, help="Min amount of variance that must be retained")
 parser.add_argument("--n_cluster", type=int, help="Number of clusters to create")
@@ -107,7 +108,15 @@ def main():
 
     centroids, wtypes = loop_kmeans(X=psi_stacked, pc_ts=pc_ts, n_cluster=args.n_cluster, n_sim=args.n_sim, n_components_keep=n_components_keep)
     class_idx, best_part = matrix_classifiability(centroids)
-    best_centroid = centroids[best_part, :, :]
+
+    if args.table is not None:
+        best_centroid = centroids[best_part, :, :]
+        best_centroid = pd.DataFrame(best_centroid)
+        best_centroid.columns = ['EOF {}'.format(i) for i in np.arange(1, n_components_keep + 1)]
+        best_centroid['WT'] = np.arange(1, args.n_cluster + 1)
+        best_centroid.set_index('WT', inplace=True)
+        best_centroid.round(decimals=3).to_latex(args.table)
+
     best_wt = wtypes[best_part, :]
     best_wt = pd.Series(resort_labels(best_wt), index=psi['time']).to_xarray()
     best_wt.name = 'wtype'

### 2018-03-19 reports/writeup/paper.tex
incorporate angel comments

@@
 \newacronym{hxlr}{HXLR}{heteroscedastic extended logistic regression}
-\newacronym{pcr}{PCR}{princiap components regression}
+\newacronym{pcr}{PCR}{principal components regression}
 \newacronym{cca}{CCA}{canonical correlation analysis}
-\newacronym{sst}{SST}{sea surface temperature}
+\newacronym{sst}{SST}{sea-surface temperature}
 \newacronym{scad}{SCAD}{South Central Atlantic Dipole}
@@
 During the austral summer of 2015-16, repeated heavy rainfall events led to severe flooding in the \gls{lprb}, displacing approximately \num{170000} people \citep{Brakenridge2016} and causing tremendous damage to property and infrastructure \citep{MinisteriodeObrasPublicasyComunicacion2016}.
-Because population in South America tends to concentrate along coasts and rivers (supplemental figure S1\footnote{check}), flooding in the \gls{lprb} directly affects not only much of the population of Paraguay, but also of populations in Argentina and Uruguay who lie along the Paran\'{a} and la Plata rivers into which the Paraguay River drains.
+Because population in South America tends to concentrate along coasts and rivers (supplemental figure S1\footnote{check}), flooding in the \gls{lprb} directly affects not only much of the population of Paraguay, but also of populations in Argentina and Uruguay who lie along the Paran\'{a} and la Plata rivers, into which the Paraguay River drains.
 Heavy rainfall and flooding in the \gls{lprb} also has important implications for hydropower generation, for agriculture, and for regional water resource management.
@@
 The rainfall climatology of the \gls{lprb} varies strongly by season, with extratropical characteristics in the winter and monsoonal characteristics in the summer.
-In the warm season (NDJF) which is the focus of this study, circulation is dominated by the upper-tropospheric Bolivian High, the lower-level subtropical highs, the Chaco Low over northern Argentina, the \gls{sacz}, and the \gls{sallj} \citep{Grimm2009,Marengo2012}.
+In the warm season (NDJF), which is the focus of this study, circulation is dominated by the upper-tropospheric Bolivian High, the lower-level subtropical highs, the Chaco Low over northern Argentina, the \gls{sacz}, and the \gls{sallj} \citep{Grimm2009,Marengo2012}.
 Rainfall peaks around \SI{5}{\milli\meter\per\day} during the warm months (October-May) and reaches a minimum near \SI{2}{\milli\meter\per\day} in July and August.
@@
 
-During the warm season (NDJF) a large fraction of rainfall, and nearly all heavy rainfall, in the \gls{lprb} is associated with mesoscale convection \citep{Velasco1987}.
+During the warm season, a large fraction of rainfall, and nearly all heavy rainfall, in the \gls{lprb} is associated with mesoscale convection \citep{Velasco1987}.
 Previous studies of organized convection and precipitation across subtropical continental South America have found close correspondence with the exit region of the low-level jets \citep{Saulo2007,Salio2007,Marengo2004,Velasco1987}, which is influenced in both summer and winter by mid-latitude baroclinic wave trains that interact with the Andes topography to generate orographically bound cyclones and northerly low-level flow \citep{Campetella2002,Seluchi2006,Boers2013,Boers2014}.
@@
 
-At sub-seasonal timescales, convection in the \gls{lprb} is modulated by a variety of drivers, notably including the \gls{sacz} and the \gls{mjo}  .
-During \gls{sacz} conditions, strong convergence is observed over the Amazon basin with divergence over southwestern Brazil, northern Argentina and Paraguay \citep{Herdies2002,Carvalho2010}; the opposite is true for so-called No-\gls{sacz} conditions.
+At sub-seasonal timescales, convection in the \gls{lprb} is modulated by a variety of drivers, notably including the \gls{sacz} and the \gls{mjo}.
+During \gls{sacz} conditions, strong low-level convergence is observed over the Amazon basin with low-level divergence over southwestern Brazil, northern Argentina and Paraguay \citep{Herdies2002,Carvalho2010}; the opposite is true for so-called No-\gls{sacz} conditions.
 \Gls{sacz} occurrence is related to westerly wind regimes over South East South America, as well as ``active'' and ``break'' periods of the South American Monsoon System \citep{Marengo2004}.
+The \gls{mjo} has been associated with the South American ``seesaw'' pattern \citep{Nogues-Paegle1997,Paegle2000,Liebmann2004}, and has been identified as a source of rainfall predictability for the region \citep[e.g.][]{Munoz2015}.
 
@@
 Spatial resolution is \SI{0.5}{\degree}, and temporal resolution is daily.
-The retro and real-time datasets are combined and treated as a single dataset.
 We define ``heavy'' rainfall events to be exceedances of the \nth{90} percentile; while the value is different for each grid cell, the \nth{90} percentile of area-averaged rainfall over the \gls{lprb} is approximately \SI{15}{\milli\meter\per\day}.
@@
 
-This study analyzes probabilistic seasonal and sub-seasonal forecasts of heavy rainfall events, defined as for observations by exceedance of the \nth{90} percentile of modeled rainfall.
+This study analyzes probabilistic seasonal and sub-seasonal forecasts of heavy rainfall events, defined by exceedance of the \nth{90} percentile of modeled rainfall.
 
-The seasonal predictions used  are known as ``flexible format'' forecasts, provided by the \gls{iri} using a multi-model ensemble approach, with bias-corrected retrospective probabilistic forecasts produced using a total of 144 members forced by evolving sea-surface temperatures, and 68 members forced by persisted sea-surface temperatures; for details see \citet{Barnston2010}.
+The seasonal predictions used  are known as ``flexible format'' forecasts, provided by the \gls{iri}.
+These forecasts use a multi-model ensemble approach, with bias-corrected retrospective probabilistic forecasts produced using a total of (a) 144 members forced by evolving sea-surface temperatures, and (b) 68 members forced by persisted sea-surface temperatures; for details see \citet{Barnston2010}.
 Flexible format means that the user of these forecasts can arbitrarily choose  particular thresholds (percentiles) to compute the  probability of exceedance (or non-exceedance) from the complete  probability density function of the climatological distribution,  rather than using the more common tercile categories.
@@
 The DJF 2015-2016 forecasts analyzed were produced in November 2015.
-Due to the short record of flexible format forecasts (2012-2016), no verification was performed for seasonal predictions.
 
-The sub-seasonal forecasts used were issued by the \gls{ecmwf} using a coupled model, available via the \gls{s2s} Prediction Project Database \citep{Vitart2016} at \SI{1.5}{\degree} resolution.
-Forecasts consider the period starting in Dec 2015 until Mar 2016 and hindcasts to assess the real-time predictive skill consider the period Dec 1-7, 1995-2014.
+The sub-seasonal forecasts used were issued by the \gls{ecmwf} using the IFS cycle 41R1 coupled model.
+These forecasts are available via the \gls{s2s} Prediction Project Database \citep{Vitart2016} at \SI{1.5}{\degree} resolution.
+Forecasts consider the period starting in Dec 2015 until Mar 2016, and hindcasts to assess the real-time predictive skill consider the period Dec 1-7, 1995-2014.
 There is a total of 51 ensemble members for each forecast, and 11 ensemble members for each of the 20 hindcasts (Dec 1-7, 1995-2014).
@@
 In this region, given topography and previous studies \citep{Barros2004,Bravo2011}, one might hypothesize rainfall inputs to most closely correspond to river levels at the stream gauges in \cref{fig:study-area}.
-Results are not sensitive to small variations in this domain.
 
@@
 This rescaling is not a necessary step; its effect is to treat all retained principal components as equally important, which provides relatively greater weight to \glspl{eof} 2, 3, and 4 than carrying out the clustering without re-scaling.
-Though our approach of first selecting the number of \glspl{eof} to use and then choosing to scale them equally involves more subjective decisions than an approach without rescaling, in this case the resulting physical patterns represented by the \glspl{eof} more closely represent patterns identified in the literature; this is further discussed in \cref{sec:diagnostics}.
+Though our approach of first selecting the number of \glspl{eof} to use and then choosing to scale them equally involves more subjective decisions than an approach without rescaling, in this case the resulting physical patterns described by the \glspl{eof} more closely represent patterns identified in the literature; this is further discussed in \cref{sec:diagnostics}.
 
@@
 The partition whose classifiability index, averaged for all 499 pairwise comparisons, was the highest was selected.
-Calculation of the classifiability index for several values of $K$ (supplemental figure S3)\footnote{CHECK} suggests that states with $K=5,6,\ldots,9$ are all reasonable; we chose the solution $K=6$ because the clusters identified are qualitatively similar to those determined over southeastern South America \citep{Munoz2015,Munoz2016} and have an intuitive physical meaning, discussed further in the following sections.
+Calculation of the classifiability index for several values of $K$ (supplemental figure S3)\footnote{CHECK} suggests that states with $K=5,6,\ldots,8$ are all reasonable.
+We chose the solution $K=6$ because the clusters identified are qualitatively similar to those determined over southeastern South America \citep{Munoz2015,Munoz2016} and have an intuitive physical meaning, discussed further in the following sections.
 We refer to the resulting clusters as \glspl{wt}.
-From a physical point of view, the K-means algorithm helps identify typical atmospheric circulation patterns in the EOF-filtered field via clustering of days with similar streamfunction configurations.
+From a physical point of view, the $K$-means algorithm helps identify typical atmospheric circulation patterns in the \gls{eof}-filtered field via clustering of days with similar streamfunction configurations.
 These clusters can also be understood as proxies  of the available states of the system , or the most frequently visited trajectories in the phase space of the physical system \citep{Munoz2015,Munoz2016a,Munoz2017}.
@@
 The \gls{xlr} was designed to address this shortcoming \citep{Wilks2009}, via the consideration of a transformation of the thresholds of interest as an additional predictor variable.
-The \gls{hxlr}, a heteroscedastic generalization of the \gls{xlr}, was proposed to appropriately use the ensemble spread as predictor for the dispersion of the predictive distribution \citep{Messner2014}.
+The \gls{hxlr}, a generalization of the \gls{xlr}, was proposed to appropriately use the ensemble spread as predictor for the dispersion of the predictive distribution \citep{Messner2014}.
 
@@
 
-The \gls{iri}'s seasonal forecasts are already provided with spatial \gls{mos} corrections of systematic errors of the individual models in the ensemble via \gls{cca} \citep{Barnston2010}, and thus we did not perform any further \gls{mos} on the seasonal rainfall fields.
+As indicated earlier, the \gls{iri}'s seasonal forecasts are already provided with spatial \gls{mos} corrections of systematic errors of the individual models in the ensemble via \gls{cca} \citep{Barnston2010}, and thus we did not perform any further \gls{mos} on the seasonal rainfall fields.
 
@@
 
-We also compute the Generalized Relative Operating Characteristics score, commonly referred to as the 2AFC score \citep{Mason2009}, to evaluate skill of probabilistic rainfall forecasts.
+We also compute the Generalized Relative Operating Characteristics score, also known as the 2AFC score \citep{Mason2009}, to evaluate skill of probabilistic rainfall forecasts.
 This score measures the ``proportion of all available pairs of observations of differing category whose probability forecasts are discriminated in the correct direction'' \citep{Mason2009}.
@@
 These two metrics, measuring reliability, resolution, uncertainty and discrimination, are deemed here to be sufficient to characterize the forecast skill for our events of interest.
-To conduct the verification in a consistent manner, we use the Climate Predictability Tool, developed by the \gls{iri} \citep{Mason2017}.
+To conduct the verification in a consistent manner, we use the Climate Predictability Tool, developed and maintained by the \gls{iri} \citep{Mason2017}.
 
@@
 As discussed in \citet{Bravo2011,Barros2004}, the location of the Bah\'ia Negra gauge (see \cref{fig:study-area}) in the Pantanal region means that it responds very slowly to rainfall input.
-The three downstream gauges, because they are located in the \gls{lprb} respond to the rainfall forcing with a slow but steady rise.
+The three downstream gauges, because they are located in the \gls{lprb}, respond to the rainfall forcing with a slow but steady rise.
 Despite several very heavy storms, the streamflow record at Asunci\'on and Pilar (which are downstream of Concepci\'on) indicates relatively little response to individual storms.
@@
 \Cref{fig:lagged-rain} shows time-lagged anomalies up to and after heavy rainfall dates (when area-averaged daily rainfall in the \gls{lprb} exceeds its NDJF \nth{90} percentile) and is consistent with previous analysis of heavy rainfall and intense convection in this region \citep{Liebmann2004,Marengo2004,Salio2007,Marwan2015}.
-At $t=-2$ a mid-latitude baroclinic system approaches the South American continent, intensifying and moving to the East at $t=-1,0,1$.
+At $t=-2$ a mid-latitude baroclinic system approaches the South American continent, intensifying and moving to the East at $t=-1,0,1$ days.
 This system interacts with the sub-tropical low and the Andes Mountains to produce an anticyclonic anomaly over Brazil.
@@
 The pattern resembles composites identified using one standard deviation exceedances of rainfall at \SI{60}{\degree W}, \SI{30}{\degree S} \citep{Liebmann2004} and analysis for the \nth{95} or 99th percentiles of daily rainfall (not shown) yield similar results, implying that the synoptic mechanism for the most heavy events is not fundamentally distinct from the mechanism for moderate-intensity events.
-This mean field, like all composites, masks between-event variation but exploration of individual events (not shown) shows that the core features identified are generally present.
+This mean field, like all composites, masks between-event variation but exploration of individual events (not shown) indicates that the core features identified are generally present.
 
@@
 The \gls{eof} loadings are shown in \cref{fig:eof-loading}.
-Of these, \gls{eof} 1 explains a substantial amount of variance ($\approx 67\%$) while \glspl{eof} 2, 3, and 4 collectively explain approximately $28\%$ of total variance.
+Of these, \gls{eof} 1 explains a substantial amount of variance ($\approx 72\%$) while \glspl{eof} 2, 3, and 4 collectively explain approximately $27\%$ of total variance.
 The resulting \glspl{wt}, shown in \cref{fig:wt-composite}, reveal patterns associated with synoptic- and regional-scale circulation regimes.
@@
 Supplemental Table S2\footnote{Check} shows the centroids of each cluster, in the 4-dimensional phase space of the leading \glspl{eof} of \SI{850}{\hecto\pascal} streamfunction.
-Consistent with the EOF loadings (\cref{fig:eof-loading}) and weather type composites (\cref{fig:wt-composite}), \glspl{wt} 1 and 4 have very similar (near-zero) projections onto \gls{eof} 1, but very different values of \glspl{eof} 2-4, particularly \gls{eof} 2.
 
@@
 At weather timescales (less than one week), the ensemble-mean successfully predicts the timing and amplitude of the area-averaged rainfall.
-At intermediate timescales, the model successfully forecast the strongest breaks and pauses in the rainfall, such as the heavy rainfall during December 2015 and the dry period during mid-January 2016.
+At timescales of one to three weeks, the model successfully forecast the strongest breaks and pauses in the rainfall, such as the heavy rainfall during December 2015 and the dry period during mid-January 2016.
 
@@
 As seen in \cref{fig:subs-prob-fcst}, the raw (uncorrected) sub-seasonal forecast of the \gls{ecmwf} model for Dec 1-7 2015 indicated very high relative odds for occurrence of heavy rainfall but with important biases in the actual location and spatial pattern; for Paraguay, it confidently suggests occurrence of heavy rainfall to the south-southeast of the country, which was mostly not observed.
-Overall, the 20-year skill of probabilistic forecasts for the first week of December - even when using  multiple initializations - is inferior to climatology at the lead times used (14 and 19 days) for most of Paraguay (\cref{fig:subs-prob-fcst}b).\footnote{Check if it's b}
-Despite these biases, the model is capturing a signal, which suggests the use of Model Output Statistics to explore the extent to which corrections in the magnitudes and spatial patterns may improve the forecast.
+Overall, the 20-year skill of probabilistic forecasts for the first week of December is highest over southern Brazil, parts of Argentina and the western border of the domain under study (see \cref{fig:subs-prob-fcst} f,k), but not over Paraguay.
+These skill scores indicate that the model is capturing a signal, and suggest the use of \glspl{mos} to explore the extent to which corrections in the magnitudes and spatial patterns may improve the forecast.
 
@@
 Comparison of long-term skill between the uncorrected \gls{s2s} model forecast output and both extended logistic regression models shows similar results.
-Reliability, resolution and uncertainty, as measured by the ignorance score (\cref{fig:subs-prob-fcst} middle row, first three columns), suggests slight skill improvement in southern Brazil, deterioration in Argentina and Uruguay, and basically the same as the uncorrected \gls{s2s} model forecast for Paraguay and southeastern Bolivia.
-Changes in forecast discrimination exhibited by the extended logistic models, as measured via the 2AFC score (\cref{fig:subs-prob-fcst}(k-o)), are  null.
-These models operate on a gridbox-by-gridbox basis to recalibrate the probabilities, and so this recalibration happens monotonically.
+Reliability, resolution and uncertainty, as measured by the Ignorance Score (\cref{fig:subs-prob-fcst} middle row, first three columns), suggests slight skill improvement in southern Brazil, deterioration in Argentina and Uruguay, and basically the same as the uncorrected \gls{s2s} model forecast for Paraguay and southeastern Bolivia.
+Changes in forecast discrimination exhibited by the extended logistic models, as measured via the 2AFC score (\cref{fig:subs-prob-fcst}(k-m)), are  null.
+The extended logistic models operate on a gridbox-by-gridbox basis to recalibrate the probabilities, and so this recalibration happens monotonically.
 Since the 2AFC score is insensitive to monotonic transformations of forecasts, the forecast discrimination is unchanged.
@@
 Better forecasts are obtained when both magnitude and spatial corrections are performed, although with relative odds considerably less confident than the ones in the raw forecast.
-The \gls{pcr} model correctly shows high relative odds in most of the places where heavy rainfall was observed (\cref{fig:subs-prob-fcst}, first row, fourth column), although it also indicates heightened risk in areas where heavy rainfall did not occur, like zones of western Paraguay and northeastern Argentina.
-The main problem with the \gls{cca} model is its lack of discrimination between occurrence or non-occurrence of heavy rainfall in the region: the spatial distribution of odds is too homogeneous (\cref{fig:subs-prob-fcst}, first row, fifth column).
+The \gls{pcr} model correctly shows high relative odds in most of the places where heavy rainfall was observed (\cref{fig:subs-prob-fcst} d), although it also indicates heightened risk in areas where heavy rainfall did not occur, like zones of western Paraguay and northeastern Argentina.
+The main problem with the \gls{cca} model is its lack of discrimination between occurrence or non-occurrence of heavy rainfall in the region: the spatial distribution of odds is too homogeneous (\cref{fig:subs-prob-fcst} e).
 
-The 20-year based skill maps of probabilistic forecasts computed with these two \gls{eof}-based models are very similar to each other, both in terms of the reliability, resolution and uncertainty measured by the Ignorance Score, and discrimination measured by the 2AFC score (\cref{fig:subs-prob-fcst}.
-Compared to the uncorrected \gls{s2s} model forecast and the logistic regression models, \gls{pcr} and \gls{cca} models perform well (\cref{fig:subs-prob-fcst}(f-o)).
-The enhanced skill is achieved through the spatial corrections via the \gls{eof}-based regressions, which - in contrast with the extended logistic models - use information from multiple grid-boxes, and thus the original forecasts are not (necessarily) calibrated monotonically.
+The 20-year based skill maps of probabilistic forecasts computed with these two \gls{eof}-based models are very similar to each other, both in terms of the reliability, resolution and uncertainty measured by the Ignorance Score, and discrimination measured by the 2AFC score (\cref{fig:subs-prob-fcst} i,j,n,o).
+In terms of long-term skill for the regions of interest over Paraguay, outputs from the \gls{pcr}- and \gls{cca}-based \gls{mos} tend to outperform the raw forecasts and the extended logistic regression models, especially regarding discrimination (\cref{fig:subs-prob-fcst} 10k-o).
+The enhanced skill is achieved through the spatial corrections via the \gls{eof}-based regressions, which - in contrast with the extended logistic models - use information from multiple grid-boxes, and thus the original forecasts are not in general calibrated monotonically.
 
-Despite the particular errors in the Dec 1-7 2015 forecasts, both \gls{pcr} and \gls{cca} verify considerably better than the raw, \gls{xlr}, and \gls{hxlr} predictions.
-Yet despite the generally high skill score for these forecasts, there are still zones along the eastern part of Paraguay with negative skill relative to climatology.
+Despite the particular errors in the Dec 1-7 2015 forecasts, on the long term both \gls{pcr} and \gls{cca} verify considerably better than the raw, \gls{xlr}, and \gls{hxlr} predictions.
+Yet despite the generally high skill score for these forecasts, there are still zones along the eastern part of Paraguay with lower discrimination skill than that of climatology.
 
@@
 
-Co-occurrence of \glspl{wt} 1 and 4, particularly in late November through late December 2015, favored advection of moisture and latent energy into the \glspl{lprb}, and low-level wind shear favored mesoscale convective activity, consistent with previous analyses in this region \citep{Velasco1987,Marengo2004,Saulo2007,Salio2007}
-Although many of the individual rainfall events of NDJF 2015-16 were intense, they were nonetheless driven by the climatological mechanism for heavy rainfall and intense convection shown in \cref{fig:lagged-rain}.
+Co-occurrence of \glspl{wt} 1 and 4, particularly in late November through late December 2015, favored advection of moisture and moist static energy into the \glspl{lprb}, and low-level wind shear favored mesoscale convective activity, consistent with previous analyses in this region \citep{Velasco1987,Marengo2004,Saulo2007,Salio2007}.
+Although many of the individual rainfall events of NDJF 2015-16 were intense, they were nonetheless driven by the climatological mechanism for heavy rainfall and intense convection shown in \cref{fig:lagged-rain} rather than by some other extreme mechanism.
 Consequently, the most striking hydrometeorlogical feature of this season, likely a key driver of the observed flooding, was the persistence of the heavy rainfall and the manner in which it switched ``on'' and ``off'' over the study region (\cref{fig:rain-wt}).
@@
 As a starting point, we consider the joint role of \gls{enso}, discussed above, and the \gls{mjo}, which  previous studies have linked to the South American ``seesaw'' pattern \citep{Nogues-Paegle1997,Paegle2000,Liebmann2004}.
-During NDJF 2015-16 The NINO 3.4 index was strongly positive, representing a strong El Ni\~{n}o state (Supplemental Figures S3, S4\footnote{check}).
+During NDJF 2015-16 the NINO 3.4 index was strongly positive, representing a strong El Ni\~{n}o state (Supplemental Figures S3, S4\footnote{check}).
 The \gls{mjo} began in November 2015 in a strong phase 3 and transitioned to phase 4 before losing amplitude around 21 November (Supplemental Figure S2)\footnote{check}.
@@
 The \gls{mjo} then weakened slightly before emerging as a mid-strength phase 4 event in late January 2016 and moving through phases 5-7.
-\Cref{fig:wt-mjo-enso} shows that \gls{wt} 1 occurs more frequently during El Ni\~{n} o years for most \gls{mjo} phases, particularly during phase 2.
-During El Ni\~no years, \gls{wt} 3 - associated with dryness over the \gls{lprb} - occurs less frequently during \gls{mjo} phases 4, 7, and 7, and more often during \gls{mjo} phase 8; this is consistent with the lack of \gls{wt} 3 during December 2015 and the frequent \gls{wt} 3 occurrence in mid-January 2016 (\cref{fig:rain-wt}).
+\Cref{fig:wt-mjo-enso} shows that \gls{wt} 1 occurs more frequently during El Ni\~{n}o years for most \gls{mjo} phases, particularly during phase 2.
+During El Ni\~no years, \gls{wt} 3 - associated with dryness over the \gls{lprb} - occurs less frequently during \gls{mjo} phases 4, 6, and 7, and more often during \gls{mjo} phase 8; this is consistent with the lack of \gls{wt} 3 during December 2015 and the frequent \gls{wt} 3 occurrence in mid-January 2016 (\cref{fig:rain-wt}).
 Detailed consideration of the role of \gls{mjo}-\gls{enso} interaction with circulation patterns over the study region is beyond the scope of this paper, but these two patterns provided background conditions favorable for the weather type sequences observed during NDJF 2015-16.
@@
 Examination of \glspl{sst} correlations with the leading \glspl{eof} of \SI{850}{\hecto\pascal} streamfunction (Supplemental Figure S5)\footnote{check} show that this dipole pattern is negatively correlated with \glspl{eof} 2, 3, and 4.
-As supplemental table S2\footnote{check} shows, the centroid of \gls{wt} in phase space is a strongly negative value of \glspl{eof} 2 and 4, and a weakly negative value of \gls{eof} 3; by contrast \gls{wt} 1 is defined by a strongly positive value of \gls{eof} 2, strongly negative value of \gls{eof} 3, and weakly positive value of \gls{eof} 4.
+As supplemental table S2\footnote{check} shows, the centroid of \gls{wt} 4 in phase space is a strongly negative value of \glspl{eof} 2 and 4, and a weakly negative value of \gls{eof} 3; by contrast \gls{wt} 1 is defined by a strongly positive value of \gls{eof} 2, strongly negative value of \gls{eof} 3, and weakly positive value of \gls{eof} 4.
 This suggests that not only did \gls{enso}-\gls{mjo} conditions favor \gls{sallj} activity, but Atlantic-Pacific conditions specifically favored  \gls{wt} 4 occurrence, helping to explain why the most intense rainfall anomalies occurred specifically in the \gls{lprb}.
@@
 
-Numerical forecasts skillfully predicted enhanced risk of heavy rainfall at the seasonal scale, likely guided by \gls{enso} signal, but biases in the spatial patterns suggest that models are not well capturing the physical interactions between the Pacific and the Atlantic basins.
+Numerical forecasts skillfully predicted enhanced risk of heavy rainfall at the seasonal scale, likely guided by the \gls{enso} signal, but biases in the spatial patterns suggest that models are not well capturing the physical interactions between the Pacific and the Atlantic basins.
 At sub-seasonal time scales, uncorrected model forecasts of rainfall had limited skill beyond 10-15 days, though use of Model Output Statistics - particularly the \gls{pcr} and \gls{cca} methods that correct both spatial patterns and magnitudes - substantially improved forecast skill.
@@
 	\caption{
-		\Gls{mos}-adjusted \gls{s2s} model forecasts and skill scores for the methods indicated in \cref{tab:mos-methods}.
+		Raw and \gls{mos}-adjusted \gls{s2s} model forecasts and skill scores for the methods indicated in \cref{tab:mos-methods}.
 		Top row (a-e) shows the heavy rainfall forecast for 1-7 December 2015 as the odds ratio defined in \cref{eq:odds-ratio} over the target domain.
@@
 		Columns separate different \gls{mos} models except for ``Raw'' (a,f,k), which indicates the uncorrected \gls{s2s} model output.
-		For all rows the grid cells which experienced a \nth{90} percentile exceedance for 1-7 December 2015 are outlined in black.
+		For all rows the grid cells which observed a \nth{90} percentile exceedance for 1-7 December 2015 are outlined in black.
 	}\label{fig:subs-prob-fcst}
@@
 		Reference wind vector in \si{\meter\per\second}.
+		Green box shows location of \gls{scad}.
 	}\label{fig:chaco-nochaco}
@@
 		These, correlations are computed between the proportion of occurrence of each weather type and the predictor monthly-mean time series.
-		For the \gls{mjo} phases, the monthly-mean time series is the proportion of occurrence of that phase for the given month.
+		For the \gls{mjo} phases the monthly-mean time series is the proportion of occurrence of that phase for the given month.
 		Predictors are the 8 \gls{mjo} phases plus neutral \gls{mjo} (labeled MJO0), encoded as a multinomial variable, and the NINO 3.4 time series, up-sampled from monthly to daily values by using the most recent monthly value (no interpolation).

### 2018-03-19 reports/writeup/paper.tex
fix reference for CPC data

@@
 
-Rainfall data are taken from the CPC Unified Gauge-based Analysis of Global Daily Precipitation dataset \citep{Xie2010}.
+Rainfall data are taken from the CPC Unified Gauge-based Analysis of Global Daily Precipitation dataset \citep{Chen2008}.
 Spatial resolution is \SI{0.5}{\degree}, and temporal resolution is daily.
@@
 Spatial resolution is \SI{2.5}{\degree}.
-Because the end-of-day time for the rainfall data is 12:00 GMT over most of South America \citep{Xie2010}, we use six-hour reanalysis data, and shift by twelve hours before re-sampling to the daily time step.
-This ensures that the time steps in the reanalysis and rainfall data sets are the same, but means that a day is defined as beginning at 12:00 GMT.
+Because the end-of-day time for the rainfall data is 12:00 GMT over most of South America \citep{Chen2008}, we use six-hour reanalysis data, and shift by twelve hours before re-sampling to the daily time step.
+This ensures that the time steps in the reanalysis and ƒrainfall data sets are the same, but means that a day is defined as beginning at 12:00 GMT.
 Since most summer rainfall in this region occurs overnight \citep{Salio2007,Vera2006}, this end-of-day time (which translates to approximately 8:00 AM locally depending on the exact time zone) tends to separate distinct events.
@@
 		A value greater than 1 indicates that the model forecast greater-than-average odds of rainfall exceeding the \nth{90} percentile.
-        Second row (f-j) shows the Ignorance score defined in \cref{eq:ignorance}, with zero indicating a perfect forecast.
+        Second row (f-j) shows the Ignorance Score defined in \cref{eq:ignorance}, with zero indicating a perfect forecast.
 		Bottom row (k-o) shows the 2AFC skill score for each grid cell; a value greater than 50 indicates that the model outperforms climatology.
@@
 		Anomalous probability of occurrence of each weather type concurrent with observance of each \gls{mjo} phase. 
-		When \gls{mjo} amplitude is less than zero, it is defined as neutral phase (0)/
+		When \gls{mjo} amplitude is less than zero, it is defined as neutral phase (0).
 		Plots are shown separately for El Ni\~no (NINO 3.4 $>1$), La Ni\~na (NINO 3.4 $< -1$), and Neutral \gls{enso} phases.

### 2018-03-20 reports/writeup/paper.tex
James edits of Angel comments

@@
 	Raw sub-seasonal forecasts of heavy rainfall exhibited limited skill at lead times beyond the first two predicted weeks, but a Model Output Statistics approach involving principal component regression substantially improved the spatial distribution of skill for Week 3 relative to other methods tested including extended logistic regressions.
-	A continuous monitoring of climate drivers impacting rainfall in the region, and the use of bias-corrected heavy precipitation seasonal and sub-seasonal forecasts, may help improve flood preparedness for the austral summer season in this part of the world.
+	A continuous monitoring of climate drivers impacting rainfall in the region, and the use of bias-corrected heavy precipitation seasonal and sub-seasonal forecasts, may help improve flood preparedness in this and other regions.
 }
@@
 During the austral summer of 2015-16, repeated heavy rainfall events led to severe flooding in the \gls{lprb}, displacing approximately \num{170000} people \citep{Brakenridge2016} and causing tremendous damage to property and infrastructure \citep{MinisteriodeObrasPublicasyComunicacion2016}.
-Because population in South America tends to concentrate along coasts and rivers (supplemental figure S1\footnote{check}), flooding in the \gls{lprb} directly affects not only much of the population of Paraguay, but also of populations in Argentina and Uruguay who lie along the Paran\'{a} and la Plata rivers, into which the Paraguay River drains.
+Because population in South America tends to concentrate along coasts and rivers (supplemental figure S1), flooding in the \gls{lprb} directly affects not only much of the population of Paraguay, but also of populations in Argentina and Uruguay who lie along the Paran\'{a} and la Plata rivers, into which the Paraguay River drains.
 Heavy rainfall and flooding in the \gls{lprb} also has important implications for hydropower generation, for agriculture, and for regional water resource management.
@@
 However, the flat topography limits the river's ability to carry the summer runoff, causing seasonal inundation of the Pantanal and distributing the river discharge in time \citep{Bravo2011,Barros2004}.
-Thus, the streamflow maxima typically occurs in phase with precipitation upstream of the Pantanal, while downstream of the Pantanal - which we define as the \acrlong{lprb}; see \cref{fig:study-area} - the annual peak typically occurs between April and July.
+Thus, the streamflow maxima typically occurs in phase with precipitation upstream of the Pantanal, while downstream of the Pantanal - an area which we define in \cref{fig:study-area} as the \acrlong{lprb} - the annual peak typically occurs between April and July.
 
 During the warm season, a large fraction of rainfall, and nearly all heavy rainfall, in the \gls{lprb} is associated with mesoscale convection \citep{Velasco1987}.
-Previous studies of organized convection and precipitation across subtropical continental South America have found close correspondence with the exit region of the low-level jets \citep{Saulo2007,Salio2007,Marengo2004,Velasco1987}, which is influenced in both summer and winter by mid-latitude baroclinic wave trains that interact with the Andes topography to generate orographically bound cyclones and northerly low-level flow \citep{Campetella2002,Seluchi2006,Boers2013,Boers2014}.
-The strength and direction of this moisture transport varies substantially between events; \gls{sallj} exit regions range from central Argentina \citep[``Chaco Jet Events'';][]{Salio2002} to Paraguay and southeastern Brazil \citep[``No-Chaco Jet Events'';][]{Vera2006}.
+Previous studies of organized convection and precipitation across subtropical continental South America have found close correspondence with the exit region of the low-level jets \citep{Velasco1987,Marengo2004,Saulo2007,Salio2007}, which is influenced in both summer and winter by mid-latitude baroclinic wave trains that interact with the Andes topography to generate orographically bound cyclones and northerly low-level flow \citep{Campetella2002,Seluchi2006,Boers2013,Boers2014}.
+The strength and direction of this moisture transport varies substantially between events, and \gls{sallj} exit regions range from central Argentina \citep[``Chaco Jet Events'';][]{Salio2002} to Paraguay and southeastern Brazil \citep[``No-Chaco Jet Events'';][]{Vera2006}.
 
-At sub-seasonal timescales, convection in the \gls{lprb} is modulated by a variety of drivers, notably including the \gls{sacz} and the \gls{mjo}.
+At sub-seasonal timescales, heavy rainfall and convection in the \gls{lprb} is modulated by a variety of drivers, notably including the \gls{sacz} and the \gls{mjo}.
 During \gls{sacz} conditions, strong low-level convergence is observed over the Amazon basin with low-level divergence over southwestern Brazil, northern Argentina and Paraguay \citep{Herdies2002,Carvalho2010}; the opposite is true for so-called No-\gls{sacz} conditions.
@@
 The intensity and precise extent of this anomaly is relevant for the precise impact of \gls{enso} events.
-The region also exhibits substantial variability between seasons of rainfall during El Ni\~no years, including a reversal of rainfall anomalies between November of that year and January of the following one, influenced by land-surface interactions \citep{Grimm2009,Grimm2003}.
+The region also exhibits substantial variability between seasons of rainfall during El Ni\~no years, including a reversal of rainfall anomalies between November of that year and January of the following one, influenced by land-surface interactions \citep{Grimm2003,Grimm2009}.
 Even beyond El Ni\~no years, regional land-surface feedbacks can cause regions that exhibit wet anomalies in the spring to experience more summer precipitation on average \citep{Grimm2007}.
-Similarly, mid-latitude dynamics influence low-level wind anomalies on many time scales, though this relationship is complicated due to coupled tropical-extratropical interactions \citep{Carvalho2004,Jones2002}.
+Similarly, mid-latitude dynamics influence low-level wind anomalies on many time scales, though this relationship is complicated due to coupled tropical-extratropical interactions \citep{Jones2002,Carvalho2004}.
 To address these potential interactions, a cross-timescale approach based on synoptic circulation types is employed here to diagnose the causes of the rainfall events; this method has been used in previous work for southeastern South America \citep{Munoz2015,Munoz2016} and other regions \citep{Moron2015}.
@@
 We begin by describing our data sources in \cref{sec:data} and our methods in \cref{sec:methods}.
-In \cref{sec:diagnostics} we start our diagnosis highlighting the observed flooding and contextualizing it within a long river stage time series; then we use composites and a weather typing analysis to diagnose the circulation patterns associated with the heavy rainfall during NDJF 2015-16.
+In \cref{sec:diagnostics} we start our diagnosis highlighting the observed flooding and contextualizing it within a long river stage time series; we then use composites and a weather typing analysis to diagnose the circulation patterns associated with the heavy rainfall during NDJF 2015-16.
 We turn in \cref{sec:fcsts} to the question of whether the observed rainfall was successfully predicted by available models.
@@
 
-The analysis presented involves both observations and model forecasts.
+The analysis presented makes use of both observations and model forecasts.
 
@@
 Rainfall data are taken from the CPC Unified Gauge-based Analysis of Global Daily Precipitation dataset \citep{Chen2008}.
-Spatial resolution is \SI{0.5}{\degree}, and temporal resolution is daily.
+Spatial resolution is \SI{0.5}{\degree} and temporal resolution is daily.
 We define ``heavy'' rainfall events to be exceedances of the \nth{90} percentile; while the value is different for each grid cell, the \nth{90} percentile of area-averaged rainfall over the \gls{lprb} is approximately \SI{15}{\milli\meter\per\day}.
@@
 This ensures that the time steps in the reanalysis and ƒrainfall data sets are the same, but means that a day is defined as beginning at 12:00 GMT.
-Since most summer rainfall in this region occurs overnight \citep{Salio2007,Vera2006}, this end-of-day time (which translates to approximately 8:00 AM locally depending on the exact time zone) tends to separate distinct events.
+Since most summer rainfall in this region occurs overnight \citep{Vera2006,Salio2007}, this end-of-day time (which translates to approximately 8:00 AM locally depending on the exact time zone) tends to separate distinct events.
 The primary atmospheric variable  used was the \SI{850}{\hecto\pascal} streamfunction, calculated directly from the wind field as described in \cref{sec:methods}.
@@
 This study also makes use of some climate indices.
-Data on \gls{enso}, specifically the NI\~NO3.4 index, came from a statistical-dynamical interpolation \citep{Kaplan1998}, which is constrained by relatively high-quality observations during the study period.
+Data on \gls{enso}, specifically the NINO3.4 index, came from a statistical-dynamical interpolation \citep{Kaplan1998}, which is constrained by relatively high-quality observations during the study period.
 Data on the \gls{mjo}   came from the Australian Bureau of Meteorology \citep{Wheeler2004}.
@@
 The DJF 2015-2016 forecasts analyzed were produced in  November 2015.
-Due to the short sample of flexible format forecasts available (only for 2012-2016) at the time of writing  this paper, no verification was performed for these seasonal predictions.
+Due to the short sample of flexible format forecasts available (only for 2012-2016 at the time of writing this paper), no verification was performed for these seasonal predictions.
 These forecasts are provided at a horizontal  resolution of \SI{2.5}{\degree}.
-This product provides the probability of exceedance (or non-exceedance) of a user-specified percentile of the climatological distribution.
 The DJF 2015-2016 forecasts analyzed were produced in November 2015.
@@
 Several types of analyses are used to diagnose the causes of the heavy rainfall events, and to bias-correct and verify the forecasts.
-Computation was performed in the python environment using stable open source packages \citep{Hoyer2017,vanderWalt2011,McKinney2010,Hunter2007}.
+Computation was performed in the python environment using stable open source packages \citep{Hunter2007,McKinney2010,vanderWalt2011,Hoyer2017}.
 All codes to reproduce or modify this analysis are available at \url{https://github.com/jdossgollin/PYFloods}.
@@
 The partition whose classifiability index, averaged for all 499 pairwise comparisons, was the highest was selected.
-Calculation of the classifiability index for several values of $K$ (supplemental figure S3)\footnote{CHECK} suggests that states with $K=5,6,\ldots,8$ are all reasonable.
+Calculation of the classifiability index for several values of $K$ (supplemental figure S2) suggests that states with $K=5,6,\ldots,8$ are all reasonable.
 We chose the solution $K=6$ because the clusters identified are qualitatively similar to those determined over southeastern South America \citep{Munoz2015,Munoz2016} and have an intuitive physical meaning, discussed further in the following sections.
@@
 From a physical point of view, the $K$-means algorithm helps identify typical atmospheric circulation patterns in the \gls{eof}-filtered field via clustering of days with similar streamfunction configurations.
-These clusters can also be understood as proxies  of the available states of the system , or the most frequently visited trajectories in the phase space of the physical system \citep{Munoz2015,Munoz2016a,Munoz2017}.
+These clusters can also be understood as proxies  of the available states of the system, or the most frequently visited trajectories in the phase space of the physical system \citep{Munoz2015,Munoz2016a,Munoz2017}.
 
@@
 Logistic regression models the conditional probability of binary events, and it has been widely used in \gls{mos} approaches for binary outcomes.
-Nonetheless, when using the logistic regression to address multiple thresholds via independent fits, the predicted probabilities are, in general, not mutually consistent \citep{Messner2014}.
+Nonetheless, when using logistic regression to address multiple thresholds via independent fits, the predicted probabilities are, in general, not mutually consistent \citep{Messner2014}.
 The \gls{xlr} was designed to address this shortcoming \citep{Wilks2009}, via the consideration of a transformation of the thresholds of interest as an additional predictor variable.
@@
 
-To avoid artificial skill when building the statistical models, we employ a cross-validation approach with a 5-year window.
+To evaluate model skill, we use a cross-validation approach with a 5-year window.
 In this framework, five continuous years are left out of the record, the regression coefficients are computed with the remaining of the time series, and the resulting model is validated comparing the prediction for the third year left out (middle of the window) against observations.
@@
 \Cref{fig:streamflow} shows the streamflow time series at several gauges on the Paraguay River during NDJF 2015-16 in the context of their seasonality and decadal variability.
-During November and December 2015, the river rose rapidly at Concepci\'on, Asunci\'on, and Pilar though not at Bah\'ia Negra.
-As discussed in \citet{Bravo2011,Barros2004}, the location of the Bah\'ia Negra gauge (see \cref{fig:study-area}) in the Pantanal region means that it responds very slowly to rainfall input.
+During November and December 2015, the river rose rapidly at Concepci\'on, Asunci\'on, and Pilar, though not at Bah\'ia Negra.
+As discussed in \citet{Barros2004,Bravo2011}, the location of the Bah\'ia Negra gauge (see \cref{fig:study-area}) in the Pantanal region means that it responds very slowly to rainfall input.
 The three downstream gauges, because they are located in the \gls{lprb}, respond to the rainfall forcing with a slow but steady rise.
@@
 \Cref{fig:lagged-rain} shows time-lagged anomalies up to and after heavy rainfall dates (when area-averaged daily rainfall in the \gls{lprb} exceeds its NDJF \nth{90} percentile) and is consistent with previous analysis of heavy rainfall and intense convection in this region \citep{Liebmann2004,Marengo2004,Salio2007,Marwan2015}.
-At $t=-2$ a mid-latitude baroclinic system approaches the South American continent, intensifying and moving to the East at $t=-1,0,1$ days.
+At $t=\SI{-2}{\day}$ a mid-latitude baroclinic system approaches the South American continent, intensifying and moving to the East from \SIrange{-1}{1}{\day}.
 This system interacts with the sub-tropical low and the Andes Mountains to produce an anticyclonic anomaly over Brazil.
 Along the cold front associated with this system, a low-level northerly jet advects  heat and moisture to the region.
-As the system progresses, the jet below \SI{20}{\degree S} transitions from predominantly meridional flow (``Chaco Jet''; $t=-1$ to predominantly zonal flow (``No-Chaco Jet'').
+As the system progresses, the jet below \SI{20}{\degree S} transitions from predominantly meridional flow (``Chaco Jet''; $t=\SI{-1}{\day}$ to predominantly zonal flow (``No-Chaco Jet'').
 The pattern resembles composites identified using one standard deviation exceedances of rainfall at \SI{60}{\degree W}, \SI{30}{\degree S} \citep{Liebmann2004} and analysis for the \nth{95} or 99th percentiles of daily rainfall (not shown) yield similar results, implying that the synoptic mechanism for the most heavy events is not fundamentally distinct from the mechanism for moderate-intensity events.
@@
 \Gls{wt} 4 is also associated with a \gls{sallj} event, but the wind turns to the East northward of \SI{25}{\degree S}, leading to heavy rainfall over Eastern Paraguay and SW Brazil; this has been called a ``No-Chaco'' Jet event \citep{Vera2006}.
-Supplemental Table S2\footnote{Check} shows the centroids of each cluster, in the 4-dimensional phase space of the leading \glspl{eof} of \SI{850}{\hecto\pascal} streamfunction.
+Supplemental Table S1 shows the centroids of each cluster, in the 4-dimensional phase space of the leading \glspl{eof} of \SI{850}{\hecto\pascal} streamfunction.
 
@@
 
-We next use monthly-mean circulation anomalies (spatial patterns) and weather type sequences (time patterns) to understand the specific events of NDJF 2015-16.
+We next use monthly-mean circulation anomalies (spatial patterns) and weather type sequences (temporal patterns) to understand the specific events of NDJF 2015-16.
 
@@
 As indicated in \cref{fig:rain-wt}, the most heavy rainfall occurred during \glspl{wt} 1 and 4.
-During NDJF 2015-16 \glspl{wt} (Chaco and No-Chaco jet extensions, respectively),  occurred more frequently than their climatology (supplemental table S1)\footnote{check}; \gls{wt} 2 also occurred more frequently than its climatology, largely due to a long sequence in February 2016.
+During NDJF 2015-16, \glspl{wt} 1 and 4 (Chaco and No-Chaco jet extensions, respectively),  occurred more frequently than their climatology (supplemental table S2); \gls{wt} 2 also occurred more frequently than its climatology, largely due to a long sequence in February 2016.
 In mid-January 2016, during a sequence of persistent low rainfall, \gls{wt} 3 featured persistently, leading to heavy rainfall over central Brazil (not shown) and negative rainfall anomalies over the \gls{lprb}.
@@
 During mid to late December 2015, \glspl{wt} 1 and 5 produced rainfall but the intermittent dry days were \gls{wt} 2.
-During mid to late January 2016 \gls{wt} 3 led to persistent low rainfall.
+During mid to late January 2016, \gls{wt} 3 led to persistent low rainfall.
 
@@
 Heavy rainfall over the region was forecast for the DJF 2015-2016 season since at least November 2015 (see \cref{fig:seas-prob-fcst}).
-Relative odds as high as 7:1 are visible over southern Paraguay and Brazil, and northern Uruguay and Argentina, generally in agreement with observations.
+Relative odds as high as 9:1 are visible over southern Paraguay and Brazil, and northern Uruguay and Argentina, broadly in agreement with observations.
 The model predicted only very weakly increased odds of heavy rainfall in the Pantanal region (directly north of the \gls{lprb}) and in northern Argentina at $\approx \SI{65}{\degree W}$, and missed the heavy precipitation along most of the northeastern border of Paraguay.
@@
 At weather timescales (less than one week), the ensemble-mean successfully predicts the timing and amplitude of the area-averaged rainfall.
-At timescales of one to three weeks, the model successfully forecast the strongest breaks and pauses in the rainfall, such as the heavy rainfall during December 2015 and the dry period during mid-January 2016.
+At timescales of one to three weeks, the ensemble average successfully forecast the strongest breaks and pauses in the rainfall, such as the heavy rainfall during December 2015 and the dry period during mid-January 2016.
 
@@
 Overall, the 20-year skill of probabilistic forecasts for the first week of December is highest over southern Brazil, parts of Argentina and the western border of the domain under study (see \cref{fig:subs-prob-fcst} f,k), but not over Paraguay.
-These skill scores indicate that the model is capturing a signal, and suggest the use of \glspl{mos} to explore the extent to which corrections in the magnitudes and spatial patterns may improve the forecast.
+These skill scores indicate that the model is capturing a signal, and suggest the use of \gls{mos} methods to explore the extent to which corrections in the magnitudes and spatial patterns may improve the forecast.
 
@@
 By analyzing how the joint behavior physical mechanisms modulate the probability of occurrence of certain weather types, it may be possible to better understand the drivers of this and future extreme event(s).
-As a starting point, we consider the joint role of \gls{enso}, discussed above, and the \gls{mjo}, which  previous studies have linked to the South American ``seesaw'' pattern \citep{Nogues-Paegle1997,Paegle2000,Liebmann2004}.
-During NDJF 2015-16 the NINO 3.4 index was strongly positive, representing a strong El Ni\~{n}o state (Supplemental Figures S3, S4\footnote{check}).
-The \gls{mjo} began in November 2015 in a strong phase 3 and transitioned to phase 4 before losing amplitude around 21 November (Supplemental Figure S2)\footnote{check}.
+As a starting point, we consider the joint role of \gls{enso}, discussed above, and the \gls{mjo}.
+During NDJF 2015-16 the NINO 3.4 index was strongly positive, representing a strong El Ni\~{n}o state (Supplemental Figures S3, S4).
+The \gls{mjo} began in November 2015 in a strong phase 3 and transitioned to phase 4 before losing amplitude around 21 November (Supplemental Figure S5).
 It stayed neutral until early December where it strengthened from a weak phase 4 to a strong phase 4 ten days later.
@@
 The \gls{mjo} then weakened slightly before emerging as a mid-strength phase 4 event in late January 2016 and moving through phases 5-7.
+
 \Cref{fig:wt-mjo-enso} shows that \gls{wt} 1 occurs more frequently during El Ni\~{n}o years for most \gls{mjo} phases, particularly during phase 2.
@@
 
-Of course, since a large fraction of the signal in \cref{fig:wt-mjo-enso} seems to come from the \gls{enso} signal, a logical question is why NDJF 2015-16 featured more persistent and intense rainfall in the \gls{lprb} than during other major El Ni\~no events (Supplemental Figure S4\footnote{check}).
+Of course, since a large fraction of the signal in \cref{fig:wt-mjo-enso} seems to come from the \gls{enso} signal, a logical question is why NDJF 2015-16 featured more persistent and intense rainfall in the \gls{lprb} than during other major El Ni\~no events (Supplemental Figure S4).
 Previous studies of the \gls{sallj} \citep[e.g.,][]{Vera2006} and the modulation of rainfall in southeastern South America by extratropical transient wave trains during El Ni\~no years emphasize the importance of Pacific-Atlantic interaction for forecasting climate events in this (and other) region(s) \citep{Barreiro2017}.
 In particular, a persistent dipolar \gls{sst} anomaly in the central southern Atlantic Ocean may favor the occurrence of \gls{wt} 4 by blocking transient extratropical wave activity from the Pacific, facilitating transitions from ``Chaco'' jet events (\gls{wt} 1) to ``No-Chaco'' jet events (\gls{wt} 4) via enhanced low-level wind circulation from southern Brazil towards the Atlantic, and back to north-east Brazil and the Amazon (see \cref{fig:chaco-nochaco}) due to land-sea temperature contrasts.
-We illustrate a schematic of this mechanism in \cref{fig:chaco-nochaco} and note that it is consistent with the mechanism found to produce heavy rainfall in the \gls{lprb} (\cref{fig:lagged-rain}) and with previous studies \citep[e.g.,][]{Vera2006,Salio2002,Liebmann2004}.
+We illustrate a schematic of this mechanism in \cref{fig:chaco-nochaco} and note that it is consistent with the mechanism found to produce heavy rainfall in the \gls{lprb} (\cref{fig:lagged-rain}) and with previous studies \citep[e.g.,][]{Salio2002,Liebmann2004,Vera2006}.
 We refer to his mechanism as the \gls{scad} and measure it as the mean meridional \gls{sst} gradient over the box shown in \cref{fig:chaco-nochaco}.
-Supplemental figure S4\footnote{check} shows the \gls{sst} anomalies observed during NDJF 2015-16, and indeed the mechanism illustrated in \cref{fig:chaco-nochaco} was present -- particularly in December 2015 when the most intense rainfall occurred.
-Examination of \glspl{sst} correlations with the leading \glspl{eof} of \SI{850}{\hecto\pascal} streamfunction (Supplemental Figure S5)\footnote{check} show that this dipole pattern is negatively correlated with \glspl{eof} 2, 3, and 4.
-As supplemental table S2\footnote{check} shows, the centroid of \gls{wt} 4 in phase space is a strongly negative value of \glspl{eof} 2 and 4, and a weakly negative value of \gls{eof} 3; by contrast \gls{wt} 1 is defined by a strongly positive value of \gls{eof} 2, strongly negative value of \gls{eof} 3, and weakly positive value of \gls{eof} 4.
+Supplemental figure S4 shows the \gls{sst} anomalies observed during NDJF 2015-16, and indeed the mechanism illustrated in \cref{fig:chaco-nochaco} was present - particularly in December 2015 when the most intense rainfall occurred.
+Examination of \glspl{sst} correlations with the leading \glspl{eof} of \SI{850}{\hecto\pascal} streamfunction (Supplemental Figure S6) show that this dipole pattern is negatively correlated with \glspl{eof} 2, 3, and 4.
+As supplemental table S1 shows, the centroid of \gls{wt} 4 in phase space is a strongly negative value of \glspl{eof} 2 and 4, and a weakly negative value of \gls{eof} 3; by contrast \gls{wt} 1 is defined by a strongly positive value of \gls{eof} 2, strongly negative value of \gls{eof} 3, and weakly positive value of \gls{eof} 4.
 This suggests that not only did \gls{enso}-\gls{mjo} conditions favor \gls{sallj} activity, but Atlantic-Pacific conditions specifically favored  \gls{wt} 4 occurrence, helping to explain why the most intense rainfall anomalies occurred specifically in the \gls{lprb}.
@@
 This Atlantic-Pacific interaction may also be a source of dynamical model biases in the representation of rainfall events in the region.
-In order to adequately forecast rainfall in certain parts of southeastern South America during El Ni\~no years, models need to reproduce not one but two stationary wave trains, originating from central and eastern Pacific, and their interactions \citep{Barreiro2017}.
+In order to adequately forecast rainfall in certain parts of southeastern South America during El Ni\~no years, models need to reproduce not one but two stationary wave trains, originating the Pacific and the Atlantic, and their interactions \citep{Barreiro2017}.
 The frontal activity is associated with the extension, shape and variability of the \gls{sacz}, and could be blocked by the presence of anomalous anti-cyclones off the coast of Brazil, like the ones described in reference to ``No-Chaco'' jet events.
-Other mechanisms that have been known to modulate rainfall signals in this region include the \gls{sacz} \citep{Carvalho2004,Munoz2015,Munoz2016} and land-biosphere-atmosphere interactions \citep{Grimm2000,Grimm2007} which also tend to be poorly represented in models \citep{Green2017,Koster2004}.
+Other mechanisms that have been known to modulate rainfall signals in this region include the \gls{sacz} \citep{Carvalho2004,Munoz2015,Munoz2016} and land-biosphere-atmosphere interactions \citep{Grimm2000,Grimm2007} which also tend to be poorly represented in models \citep{Koster2004,Green2017}.
 
@@
 
-Numerical forecasts skillfully predicted enhanced risk of heavy rainfall at the seasonal scale, likely guided by the \gls{enso} signal, but biases in the spatial patterns suggest that models are not well capturing the physical interactions between the Pacific and the Atlantic basins.
+Numerical forecasts skillfully predicted enhanced risk of heavy rainfall at the seasonal scale, consistent with the observed \gls{enso} signal, but biases in the spatial patterns suggest that models are imperfectly capturing the physical interactions between the Pacific and the Atlantic basins.
 At sub-seasonal time scales, uncorrected model forecasts of rainfall had limited skill beyond 10-15 days, though use of Model Output Statistics - particularly the \gls{pcr} and \gls{cca} methods that correct both spatial patterns and magnitudes - substantially improved forecast skill.
@@
 \begin{figure*}
-	\noindent\includegraphics[width=6.5in]{anomalies_ndjf1516.pdf}
-	\caption{
-		Monthly composite anomalies observed during NDJF 2015-16.
-		Top row (a-d) shows streamfunction anomalies at \SI{850}{\hecto\pascal}.
-		Bottom row (e-h) shows rainfall anomalies, in units of \si{\milli\meter\per\day}.
-	}\label{fig:anomalies}
-\end{figure*}
-
-\begin{figure*}
 	\noindent\includegraphics[width=6.5in]{study_area.jpg}
@@
 	}\label{fig:study-area}
+\end{figure*}
+
+\begin{figure*}
+	\noindent\includegraphics[width=6.5in]{anomalies_ndjf1516.pdf}
+	\caption{
+		Monthly composite anomalies observed during NDJF 2015-16.
+		Top row (a-d) shows streamfunction anomalies at \SI{850}{\hecto\pascal}.
+		Bottom row (e-h) shows rainfall anomalies, in units of \si{\milli\meter\per\day}.
+	}\label{fig:anomalies}
 \end{figure*}
@@
 		Composite anomalies associated with each weather type.
-		Top row (a-f) shows streamfunction =anomalies at \SIlist{850}{\hecto\pascal}.
+		Top row (a-f) shows streamfunction anomalies at \SIlist{850}{\hecto\pascal}.
 		Strongest 20\% of wind anomaly vectors over the plot area are also shown.
@@
 	\caption{
-		Spearman (rank) correlation coefficient between several \gls{s2s} predictors and weather type occurrence at monthly time step.
-		These, correlations are computed between the proportion of occurrence of each weather type and the predictor monthly-mean time series.
+		Spearman (rank) correlation coefficient between several \gls{s2s} climate indices and weather type occurrence at monthly time step.
+		These, correlations are computed between the proportion of occurrence of each weather type and the monthly-mean time series of the climate index.
 		For the \gls{mjo} phases the monthly-mean time series is the proportion of occurrence of that phase for the given month.
-		Predictors are the 8 \gls{mjo} phases plus neutral \gls{mjo} (labeled MJO0), encoded as a multinomial variable, and the NINO 3.4 time series, up-sampled from monthly to daily values by using the most recent monthly value (no interpolation).
+		The NINO 3.4 time series was up-sampled from monthly to daily values by persisting the most recent monthly value (no interpolation).
 		The \acrfull{scad} predictor, labeled Dipole, is the mean meridional gradient of \gls{sst} anomalies over the region \SIrange{30}{10}{\degree W} and \SIrange{40}{15}{\degree S}, up-sampled to daily values using the same method as the NINO 3.4 time series.

### 2018-03-20 src/analyze/plot_predictors_wtypes.py

@@ -33,7 +33,7 @@ def main():
     wt = xr.open_dataarray(args.wt).to_dataframe(name='WT')
     wt = pd.get_dummies(wt['WT'],  prefix='WT')
 
-    dipole = xr.open_dataarray(args.dipole).to_dataframe(name='Dipole').resample('1D').ffill()
+    dipole = xr.open_dataarray(args.dipole).to_dataframe(name='SCAD').resample('1D').ffill()
 
     # Join everything together
     daily = mjo.join(nino_daily).join(dipole).join(wt).dropna()
@@ -41,7 +41,7 @@ def main():
 
     # Get the predictor, EOF, and WT columns
     predictors = mjo.columns.tolist()
-    predictors.append('Dipole')
+    predictors.append('SCAD')
     predictors.append('NINO 3.4')
     wts = wt.columns.tolist()
 

### 2018-03-20 reports/writeup/paper.tex
fix BBC ref

@@
 In fact, this apparent ``on'' and ``off'' switching was manifest principally as a spatial shift in the rainfall occurrence (\cref{fig:anomalies}) consistent with the increased occurrence of \gls{wt} 3 during mid-late January 2016 (\cref{fig:rain-wt,fig:wt-composite}); this pattern has been previously described as the South American ``seesaw'' pattern \citep{Nogues-Paegle1997}.
-Although many news reports blamed the flooding on El Ni\~{n}o\footnote{\url{http://www.bbc.com/news/world-latin-america-35184793}}, NDJF 2015-16 featured more intense rainfall than previous major El Ni\~no events, and this intense rainfall persisted for a longer time .
+Although many news reports blamed the flooding on El Ni\~{n}o \citep{BBC2015}, NDJF 2015-16 featured more intense rainfall than previous major El Ni\~no events, and this intense rainfall persisted for a longer time .
 While the link between El Ni\~{n}o and flooding in the \gls{lprb} is consistent with previous studies of ENSO and summertime rainfall in this region \citep{Velasco1987,Grimm2000,Salio2002,Grimm2003,Carvalho2004,Grimm2009a,Bravo2011}, both the ``on''-``off'' switching and the differences from previous major El Ni\~{n}o events suggest that other physical mechanisms, and their cross-timescale interactions, are relevant for understanding and predicting future events.

### 2018-03-21 reports/writeup/paper.tex
Edits from reviewer comments

@@
 Similarly, mid-latitude dynamics influence low-level wind anomalies on many time scales, though this relationship is complicated due to coupled tropical-extratropical interactions \citep{Jones2002,Carvalho2004}.
-To address these potential interactions, a cross-timescale approach based on synoptic circulation types is employed here to diagnose the causes of the rainfall events; this method has been used in previous work for southeastern South America \citep{Munoz2015,Munoz2016} and other regions \citep{Moron2015}.
+To address these potential interactions, a cross-timescale approach based on synoptic circulation types is employed here to diagnose the causes of the rainfall events.
+This method has been used in previous work for southeastern South America \citep{Munoz2015,Munoz2016} and other regions \citep{Moron2015}.
 
+The paper proceeds as follows.
 We begin by describing our data sources in \cref{sec:data} and our methods in \cref{sec:methods}.
@@
 
-This study analyzes probabilistic seasonal and sub-seasonal forecasts of heavy rainfall events, defined by exceedance of the \nth{90} percentile of modeled rainfall.
+This study analyzes probabilistic seasonal and sub-seasonal forecasts of heavy rainfall events, which we define as exceedance of the \nth{90} percentile of NDJF daily precipitation across all ensemble members and initializations.
 
 The seasonal predictions used  are known as ``flexible format'' forecasts, provided by the \gls{iri}.
-These forecasts use a multi-model ensemble approach, with bias-corrected retrospective probabilistic forecasts produced using a total of (a) 144 members forced by evolving sea-surface temperatures, and (b) 68 members forced by persisted sea-surface temperatures; for details see \citet{Barnston2010}.
+These forecasts use a multi-model ensemble approach, with bias-corrected retrospective probabilistic forecasts produced using a total of 144 members forced by evolving sea-surface temperatures and 68 members forced by persisted sea-surface temperatures; for details see \citet{Barnston2010}.
 Flexible format means that the user of these forecasts can arbitrarily choose  particular thresholds (percentiles) to compute the  probability of exceedance (or non-exceedance) from the complete  probability density function of the climatological distribution,  rather than using the more common tercile categories.
@@
 
-The $K$-means algorithm is guaranteed to converge to a local minimum of inter-cluster variance; to select the best partition, 500 simulations were created using the implementation in Python's \texttt{scikit-learn} package \citep{Pedregosa2012} and then the classifiability index of \citet{Michelangeli1995} was computed between each partition and the 499 others.
+The $K$-means algorithm is guaranteed to converge to a local minimum of inter-cluster variance; to select the best partition, 500 simulations were created using the implementation in Python's \texttt{scikit-learn} package \citep{Pedregosa2012}.
+Next, the classifiability index of \citet{Michelangeli1995} was computed between each partition and the 499 others.
 The partition whose classifiability index, averaged for all 499 pairwise comparisons, was the highest was selected.
 Calculation of the classifiability index for several values of $K$ (supplemental figure S2) suggests that states with $K=5,6,\ldots,8$ are all reasonable.
-We chose the solution $K=6$ because the clusters identified are qualitatively similar to those determined over southeastern South America \citep{Munoz2015,Munoz2016} and have an intuitive physical meaning, discussed further in the following sections.
+We chose the solution $K=6$ because the clusters identified are qualitatively similar to those determined over southeastern South America \citep{Munoz2015,Munoz2016} and have an intuitive physical meaning, which we discuss further in the following sections.
 We refer to the resulting clusters as \glspl{wt}.
@@
 
-Logistic regression models the conditional probability of binary events, and it has been widely used in \gls{mos} approaches for binary outcomes.
+Logistic regression models the conditional probability of binary events, given one or more predictors, and has been widely used in \gls{mos} approaches for binary outcomes.
 Nonetheless, when using logistic regression to address multiple thresholds via independent fits, the predicted probabilities are, in general, not mutually consistent \citep{Messner2014}.
@@
 
-To understand how circulation anomalies observed during NDJF 2015-2016 led to the observed floods it is helpful to first explore the atmospheric circulations which are typically associated with heavy rainfall in the lower Paraguay River.
+To understand how circulation anomalies observed during NDJF 2015-2016 led to the observed floods it is helpful to first explore the atmospheric circulations which are typically associated with heavy rainfall in the lower Paraguay River during the full observed record.
 
@@
 This multistage prediction system is known as the \gls{iri}'s ``Ready-Set-Go" approach \citep{Hellmuth2011,Goddard2014}.
-For the sake of concision, in this section we discuss only probabilistic seasonal (DJF 2015-2016) and sub-seasonal (Dec 1-7, 2015) forecasts.
+In this study we examine only probabilistic seasonal (DJF 2015-2016) and sub-seasonal (Dec 1-7, 2015) forecasts.
 
@@
 \Cref{fig:chiclet} uses a Chiclet diagram \citep{Carbin2016} to visualize, as a function of lead time, the time evolution of the uncorrected, ensemble-mean rainfall anomaly forecast, spatially averaged over the \gls{lprb}.
-At times greater than about two weeks, the ensemble-mean forecast is slightly wetter than climatology.
+At times greater than about two weeks, the ensemble-mean forecast is for slightly positive rainfall anomalies at nearly all initialization dates and lead times.
 At weather timescales (less than one week), the ensemble-mean successfully predicts the timing and amplitude of the area-averaged rainfall.
@@
 
-This Atlantic-Pacific interaction may also be a source of dynamical model biases in the representation of rainfall events in the region.
-In order to adequately forecast rainfall in certain parts of southeastern South America during El Ni\~no years, models need to reproduce not one but two stationary wave trains, originating the Pacific and the Atlantic, and their interactions \citep{Barreiro2017}.
-The frontal activity is associated with the extension, shape and variability of the \gls{sacz}, and could be blocked by the presence of anomalous anti-cyclones off the coast of Brazil, like the ones described in reference to ``No-Chaco'' jet events.
+This Atlantic-Pacific interaction may also explain the skill of dynamical models for representing heavy rainfall in the region.
+In order to adequately forecast rainfall in certain parts of southeastern South America during El Ni\~no years, models need to reproduce stationary wave trains originating in the Pacific and the Atlantic and their interactions \citep{Barreiro2017}.
 Other mechanisms that have been known to modulate rainfall signals in this region include the \gls{sacz} \citep{Carvalho2004,Munoz2015,Munoz2016} and land-biosphere-atmosphere interactions \citep{Grimm2000,Grimm2007} which also tend to be poorly represented in models \citep{Koster2004,Green2017}.
+The stationary wavetrain interactions, land-atmosphere interactions, and topography may explain why simulating heavy rainfall in this region is so difficult  \cref{fig:chiclet,fig:subs-prob-fcst}.
+Improving understanding of these phenomena is important opportunity for \gls{s2s} prediction, and is left for future work.
 

### 2018-03-21 reports/writeup/paper.tex
Minor paper revisions

@@
 Inspection of \cref{fig:rain-wt} also suggests that at time scales of days to weeks, particular sequences of weather types tend to recur, and are associated with repeated rainfall storms.
-The rainfall in late November 2015 occurred during \gls{wt} 1 and 4 days, with intermittent (and dry) \gls{wt} 5 occurrences.
-During mid to late December 2015, \glspl{wt} 1 and 5 produced rainfall but the intermittent dry days were \gls{wt} 2.
-During mid to late January 2016, \gls{wt} 3 led to persistent low rainfall.
+From mid November to late December 2015, nearly all days were weather types 1, 4, and 5, consistent with the anticyclonic anomaly observed over central Brazil during that time (\cref{fig:anomalies}).
+Nearly all of the heavy rainfall occurred during \glspl{wt} 1 and 4.
+During mid to late January 2016, repeated \gls{wt} 3 days led to persistent low rainfall, and in mid February 2016 frequent occurrence of \gls{wt} 2 led to frequent, though generally not intense, rainfall.
 

### 2018-05-06 reports/writeup/Supplement.tex
fix reference for CPC data
James edits of Angel comments
proposed james edits

@@
 	\centering
-	\begin{tabular}{lrr}
-		\toprule
-		{} &  NDJF 2015-16 &  Climatology \\
-		WT &               &              \\
-		\midrule
-		1     &         0.273 &        0.213 \\
-		2     &         0.231 &        0.185 \\
-		3     &         0.107 &        0.156 \\
-		4     &         0.207 &        0.156 \\
-		5     &         0.132 &        0.150 \\
-		6     &         0.050 &        0.140 \\
-		\bottomrule
-	\end{tabular}
-	\caption{Weather type (WT) occurrence fraction during NDJF 2015-16 and Climatology}
-\end{table}
-
-\begin{table}
-	\centering
 	\begin{tabular}{lrrrr}
@@
 
+\begin{table}
+	\centering
+	\begin{tabular}{lrr}
+		\toprule
+		{} &  NDJF 2015-16 &  Climatology \\
+		WT &               &              \\
+		\midrule
+		1     &         0.273 &        0.213 \\
+		2     &         0.231 &        0.185 \\
+		3     &         0.107 &        0.156 \\
+		4     &         0.207 &        0.156 \\
+		5     &         0.132 &        0.150 \\
+		6     &         0.050 &        0.140 \\
+		\bottomrule
+	\end{tabular}
+	\caption{Weather type (WT) occurrence fraction during NDJF 2015-16 and Climatology}
+\end{table}
+
 \clearpage
@@
 \begin{figure}
-	\includegraphics[width=\textwidth]{mjo_ts.pdf}
+	\includegraphics[width=\textwidth]{wt_classifiability.pdf}
 	\caption{
-		Evolution of the MJO during NDJF 2015-16.
-		Points are plotted on RMM1 ($x$-axis) and RMM2 ($y$-axis), derived from leading EOFs of OLR fields \citep{Wheeler2004}.
-		Gray lines divide the plot into the eight phases used in the text.
-		Colors show the time evolution of the system from 01 November 2015 (blue) to 29 February 2016 (yellow).
-		Gray circle indicates the area of neutral MJO activity with amplitude $<0$.
+		Classifiability index (see Methods) calculated for several chosen values of $K$, representing the number of weather types produced.
+		Results are produced with only 75 simulations for each $K$.
 	}
@@
 		SST data from \citet{Reynolds2002}.
-		Also shown are rainfall anomalies over South America, from \citet{Xie2010}.
+		Also shown are rainfall anomalies over South America, from \citet{Chen2008}.
 		Rainfall contour intervals are \SI{1}{\milli\meter\per\day}.
@@
 \end{sidewaysfigure}
+
+\begin{figure}
+	\includegraphics[width=\textwidth]{mjo_ts.pdf}
+	\caption{
+		Evolution of the MJO during NDJF 2015-16.
+		Points are plotted on RMM1 ($x$-axis) and RMM2 ($y$-axis), derived from leading EOFs of OLR fields \citep{Wheeler2004}.
+		Gray lines divide the plot into the eight phases used in the text.
+		Colors show the time evolution of the system from 01 November 2015 (blue) to 29 February 2016 (yellow).
+		Gray circle indicates the area of neutral MJO activity with amplitude $<0$.
+	}
+\end{figure}
 
@@
 \begin{figure}
-	\includegraphics[width=\textwidth]{wt_classifiability.pdf}
+	\includegraphics[width=\textwidth]{predictors_eofs.pdf}
 	\caption{
-		Classifiability index (see Methods) calculated for several chosen values of $K$, representing the number of weather types produced.
-		Results are produced with only 75 simulations for each $K$.
-	}
+		Spearman (rank) correlation coefficient between several \gls{s2s} climate indices and weather type occurrence at monthly time step.
+		These correlations are computed between the proportion of occurrence of each weather type and the monthly-mean time series of the climate index.
+		For the \gls{mjo} phases the monthly-mean time series is the proportion of occurrence of that phase for the given month.
+		The NINO 3.4 time series was up-sampled from monthly to daily values by persisting the most recent monthly value (no interpolation).
+		The \acrfull{scad} predictor, labeled Dipole, is the mean meridional gradient of \gls{sst} anomalies over the region \SIrange{30}{10}{\degree W} and \SIrange{40}{15}{\degree S}, up-sampled to daily values using the same method as the NINO 3.4 time series.
+	}\label{fig:predictor-eof}
 \end{figure}

### 2018-05-06 reports/writeup/paper.tex
proposed james edits

@@
 
-During the austral summer of 2015-16, repeated heavy rainfall events led to severe flooding in the \gls{lprb}, displacing approximately \num{170000} people \citep{Brakenridge2016} and causing tremendous damage to property and infrastructure \citep{MinisteriodeObrasPublicasyComunicacion2016}.
+During the austral summer of 2015-16, repeated heavy rainfall events led to severe flooding in the \gls{lprb} (\cref{fig:study-area,fig:anomalies}), displacing approximately \num{170000} people \citep{Brakenridge2016} and causing tremendous damage to property and infrastructure \citep{MinisteriodeObrasPublicasyComunicacion2016}.
 Because population in South America tends to concentrate along coasts and rivers (supplemental figure S1), flooding in the \gls{lprb} directly affects not only much of the population of Paraguay, but also of populations in Argentina and Uruguay who lie along the Paran\'{a} and la Plata rivers, into which the Paraguay River drains.
@@
 
-The rainfall climatology of the \gls{lprb} varies strongly by season, with extratropical characteristics in the winter and monsoonal characteristics in the summer.
-In the warm season (NDJF), which is the focus of this study, circulation is dominated by the upper-tropospheric Bolivian High, the lower-level subtropical highs, the Chaco Low over northern Argentina, the \gls{sacz}, and the \gls{sallj} \citep{Grimm2009,Marengo2012}.
+The climatology of the \gls{lprb} varies strongly by season, with extratropical characteristics in the winter and monsoonal characteristics in the summer.
+The most notable circulation features during the warm season (NDJF), which is the focus of this study, are the upper-tropospheric Bolivian High, the lower-level subtropical highs, the Chaco Low over northern Argentina, the \gls{sacz}, and the \gls{sallj} \citep{Grimm2009,Marengo2012}.
 Rainfall peaks around \SI{5}{\milli\meter\per\day} during the warm months (October-May) and reaches a minimum near \SI{2}{\milli\meter\per\day} in July and August.
@@
 The paper proceeds as follows.
-We begin by describing our data sources in \cref{sec:data} and our methods in \cref{sec:methods}.
-In \cref{sec:diagnostics} we start our diagnosis highlighting the observed flooding and contextualizing it within a long river stage time series; we then use composites and a weather typing analysis to diagnose the circulation patterns associated with the heavy rainfall during NDJF 2015-16.
+We first describe our data sources in \cref{sec:data} and our methods in \cref{sec:methods}.
+In \cref{sec:diagnostics} we start our diagnosis highlighting the observed flooding and contextualizing it within a long river stage time series; we then use composites and a weather typing analysis to diagnose the circulation patterns associated with the heavy rainfall during NDJF 201516.
 We turn in \cref{sec:fcsts} to the question of whether the observed rainfall was successfully predicted by available models.
-To carry out this analysis we study both forecasts targeting the entire series for a limited area, and also forecasts targeting a large spatial area for only the first week of December, when the most important flooding events started.
+To carry out this analysis we study both forecasts targeting the entire series for a limited area, and also forecasts targeting a large spatial area for only the first week of December, when the most important flooding events began.
 We also explore the impact on forecasts of several bias-correction schemes.
@@
 
-\section{Data} \label{sec:data}
+\section{Data}\label{sec:data}
 
@@
 Because the end-of-day time for the rainfall data is 12:00 GMT over most of South America \citep{Chen2008}, we use six-hour reanalysis data, and shift by twelve hours before re-sampling to the daily time step.
-This ensures that the time steps in the reanalysis and ƒrainfall data sets are the same, but means that a day is defined as beginning at 12:00 GMT.
+This ensures that the time steps in the reanalysis and rainfall data sets are the same, but means that a day is defined as beginning at 12:00 GMT.
 Since most summer rainfall in this region occurs overnight \citep{Vera2006,Salio2007}, this end-of-day time (which translates to approximately 8:00 AM locally depending on the exact time zone) tends to separate distinct events.
@@
 The seasonal predictions used  are known as ``flexible format'' forecasts, provided by the \gls{iri}.
-These forecasts use a multi-model ensemble approach, with bias-corrected retrospective probabilistic forecasts produced using a total of 144 members forced by evolving sea-surface temperatures and 68 members forced by persisted sea-surface temperatures; for details see \citet{Barnston2010}.
+These forecasts use a multi-model ensemble approach, with bias-corrected retrospective probabilistic forecasts produced using a total of 144 members forced by evolving \glspl{sst} and 68 members forced by persisted sea-surface temperatures; for details see \citet{Barnston2010}.
 Flexible format means that the user of these forecasts can arbitrarily choose  particular thresholds (percentiles) to compute the  probability of exceedance (or non-exceedance) from the complete  probability density function of the climatological distribution,  rather than using the more common tercile categories.
@@
 
-\section{Methods} \label{sec:methods}
+\section{Methods}\label{sec:methods}
 
@@
 
-\subsection{Weather Typing} \label{sec:weather-typing}
+\subsection{Weather Typing}\label{sec:weather-typing}
 
@@
 
-\section{Diagnostics} \label{sec:diagnostics}
+\section{Diagnostics}\label{sec:diagnostics}
 
@@
 
-\subsection{Observed Flooding} \label{sec:flood}
+\subsection{Observed Flooding}\label{sec:flood}
 
@@
 
-\subsection{Heavy Rainfall: Climatological Drivers} \label{sec:rainfall-circulation}
+\subsection{Heavy Rainfall: Climatological Drivers}\label{sec:rainfall-circulation}
 
@@
 
-\subsection{Weather Type Analysis: Daily Circulation Patterns} \label{sec:weather-types}
+\subsection{Weather Type Analysis: Daily Circulation Patterns}\label{sec:weather-types}
 
@@
 
-\section{Forecasts} \label{sec:fcsts}
+\section{Forecasts}\label{sec:fcsts}
 
 In this section we analyze the extent to which forecasts were able to predict the persistent rainfall during summer of 2015-16.
-
-There are advantages in simultaneously considering useful climate information at multiple timescales, rather than just focusing on one of them.
-This multistage prediction system is known as the \gls{iri}'s ``Ready-Set-Go" approach \citep{Hellmuth2011,Goddard2014}.
-In this study we examine only probabilistic seasonal (DJF 2015-2016) and sub-seasonal (Dec 1-7, 2015) forecasts.
+There are advantages in simultaneously considering useful climate information at multiple timescales, rather than just focusing on one of them \citep{Hellmuth2011,Goddard2014}.
+In this study we analyze probabilistic seasonal (DJF 2015-2016) and sub-seasonal (Dec 1-7, 2015) forecasts.
 
@@
 In general, the use of extended logistic regression models does not improve the forecast for the week.
-For example, with respect to the raw prediction, \gls{xlr} tends to amplify the relative odds, and to cluster and shift the forecast location of the heavy rainfall events (\cref{fig:subs-prob-fcst}, top row, first two columns); the forecast tends to be better for Uruguay, but suggests heavy rainfall in the Paraguayan Chaco, which was not present in the raw prediction.
-On the other hand, the use of the ensemble spread in the \gls{hxlr} model does not help; this forecast tends to be over-confident on the events occurring in almost all the region of interest (\cref{fig:subs-prob-fcst}, top row, third column).
+For example, with respect to the raw prediction, \gls{xlr} tends to amplify the relative odds, and to cluster and shift the forecast location of the heavy rainfall events (\cref{fig:subs-prob-fcst} a,b); the forecast tends to be better for Uruguay, but suggests heavy rainfall in the Paraguayan Chaco, which was not present in the raw prediction.
+On the other hand, the use of the ensemble spread in the \gls{hxlr} model does not help; this forecast tends to be over-confident on the events occurring in almost all the region of interest (\cref{fig:subs-prob-fcst} c).
 
 Comparison of long-term skill between the uncorrected \gls{s2s} model forecast output and both extended logistic regression models shows similar results.
-Reliability, resolution and uncertainty, as measured by the Ignorance Score (\cref{fig:subs-prob-fcst} middle row, first three columns), suggests slight skill improvement in southern Brazil, deterioration in Argentina and Uruguay, and basically the same as the uncorrected \gls{s2s} model forecast for Paraguay and southeastern Bolivia.
-Changes in forecast discrimination exhibited by the extended logistic models, as measured via the 2AFC score (\cref{fig:subs-prob-fcst}(k-m)), are  null.
+Reliability, resolution and uncertainty, as measured by the Ignorance Score (\cref{fig:subs-prob-fcst}f-h), suggests slight skill improvement in southern Brazil, deterioration in Argentina and Uruguay, and basically the same as the uncorrected \gls{s2s} model forecast for Paraguay and southeastern Bolivia.
+Changes in forecast discrimination exhibited by the extended logistic models, as measured via the 2AFC score (\cref{fig:subs-prob-fcst}k-m), are  null.
 The extended logistic models operate on a gridbox-by-gridbox basis to recalibrate the probabilities, and so this recalibration happens monotonically.
@@
 The 20-year based skill maps of probabilistic forecasts computed with these two \gls{eof}-based models are very similar to each other, both in terms of the reliability, resolution and uncertainty measured by the Ignorance Score, and discrimination measured by the 2AFC score (\cref{fig:subs-prob-fcst} i,j,n,o).
-In terms of long-term skill for the regions of interest over Paraguay, outputs from the \gls{pcr}- and \gls{cca}-based \gls{mos} tend to outperform the raw forecasts and the extended logistic regression models, especially regarding discrimination (\cref{fig:subs-prob-fcst} 10k-o).
+In terms of long-term skill for the regions of interest over Paraguay, outputs from the \gls{pcr}- and \gls{cca}-based \gls{mos} tend to outperform the raw forecasts and the extended logistic regression models, especially regarding discrimination (\cref{fig:subs-prob-fcst} k-o).
 The enhanced skill is achieved through the spatial corrections via the \gls{eof}-based regressions, which - in contrast with the extended logistic models - use information from multiple grid-boxes, and thus the original forecasts are not in general calibrated monotonically.
@@
 
-\section{Discussion}
-\label{sec:discussion}
+\section{Discussion}\label{sec:discussion}
 
@@
 In fact, this apparent ``on'' and ``off'' switching was manifest principally as a spatial shift in the rainfall occurrence (\cref{fig:anomalies}) consistent with the increased occurrence of \gls{wt} 3 during mid-late January 2016 (\cref{fig:rain-wt,fig:wt-composite}); this pattern has been previously described as the South American ``seesaw'' pattern \citep{Nogues-Paegle1997}.
-Although many news reports blamed the flooding on El Ni\~{n}o \citep{BBC2015}, NDJF 2015-16 featured more intense rainfall than previous major El Ni\~no events, and this intense rainfall persisted for a longer time .
+
+Although many news reports blamed the flooding on El Ni\~{n}o \citep{BBC2015}, NDJF 2015-16 featured more intense rainfall than previous major El Ni\~no events, and this intense rainfall persisted for a longer time.
 While the link between El Ni\~{n}o and flooding in the \gls{lprb} is consistent with previous studies of ENSO and summertime rainfall in this region \citep{Velasco1987,Grimm2000,Salio2002,Grimm2003,Carvalho2004,Grimm2009a,Bravo2011}, both the ``on''-``off'' switching and the differences from previous major El Ni\~{n}o events suggest that other physical mechanisms, and their cross-timescale interactions, are relevant for understanding and predicting future events.
+
+\Cref{fig:wt-mjo-enso} shows that \gls{wt} 1 occurs more frequently during El Ni\~{n}o years for most \gls{mjo} phases, particularly during phase 2.
+During El Ni\~no years, \gls{wt} 3 - associated with dryness over the \gls{lprb} - occurs less frequently during \gls{mjo} phases 4, 6, and 7, and more often during \gls{mjo} phase 8; this is consistent with the lack of \gls{wt} 3 during December 2015 and the frequent \gls{wt} 3 occurrence in mid-January 2016 (\cref{fig:rain-wt}).
+Detailed consideration of the role of \gls{mjo}-\gls{enso} interaction with circulation patterns over the study region is beyond the scope of this paper, but these two patterns provided background conditions favorable for the weather type sequences observed during NDJF 2015-16.
 
@@
 
-\Cref{fig:wt-mjo-enso} shows that \gls{wt} 1 occurs more frequently during El Ni\~{n}o years for most \gls{mjo} phases, particularly during phase 2.
-During El Ni\~no years, \gls{wt} 3 - associated with dryness over the \gls{lprb} - occurs less frequently during \gls{mjo} phases 4, 6, and 7, and more often during \gls{mjo} phase 8; this is consistent with the lack of \gls{wt} 3 during December 2015 and the frequent \gls{wt} 3 occurrence in mid-January 2016 (\cref{fig:rain-wt}).
-Detailed consideration of the role of \gls{mjo}-\gls{enso} interaction with circulation patterns over the study region is beyond the scope of this paper, but these two patterns provided background conditions favorable for the weather type sequences observed during NDJF 2015-16.
-
 Of course, since a large fraction of the signal in \cref{fig:wt-mjo-enso} seems to come from the \gls{enso} signal, a logical question is why NDJF 2015-16 featured more persistent and intense rainfall in the \gls{lprb} than during other major El Ni\~no events (Supplemental Figure S4).
 Previous studies of the \gls{sallj} \citep[e.g.,][]{Vera2006} and the modulation of rainfall in southeastern South America by extratropical transient wave trains during El Ni\~no years emphasize the importance of Pacific-Atlantic interaction for forecasting climate events in this (and other) region(s) \citep{Barreiro2017}.
+
 In particular, a persistent dipolar \gls{sst} anomaly in the central southern Atlantic Ocean may favor the occurrence of \gls{wt} 4 by blocking transient extratropical wave activity from the Pacific, facilitating transitions from ``Chaco'' jet events (\gls{wt} 1) to ``No-Chaco'' jet events (\gls{wt} 4) via enhanced low-level wind circulation from southern Brazil towards the Atlantic, and back to north-east Brazil and the Amazon (see \cref{fig:chaco-nochaco}) due to land-sea temperature contrasts.
@@
 We refer to his mechanism as the \gls{scad} and measure it as the mean meridional \gls{sst} gradient over the box shown in \cref{fig:chaco-nochaco}.
-Supplemental figure S4 shows the \gls{sst} anomalies observed during NDJF 2015-16, and indeed the mechanism illustrated in \cref{fig:chaco-nochaco} was present - particularly in December 2015 when the most intense rainfall occurred.
-Examination of \glspl{sst} correlations with the leading \glspl{eof} of \SI{850}{\hecto\pascal} streamfunction (Supplemental Figure S6) show that this dipole pattern is negatively correlated with \glspl{eof} 2, 3, and 4.
-As supplemental table S1 shows, the centroid of \gls{wt} 4 in phase space is a strongly negative value of \glspl{eof} 2 and 4, and a weakly negative value of \gls{eof} 3; by contrast \gls{wt} 1 is defined by a strongly positive value of \gls{eof} 2, strongly negative value of \gls{eof} 3, and weakly positive value of \gls{eof} 4.
-This suggests that not only did \gls{enso}-\gls{mjo} conditions favor \gls{sallj} activity, but Atlantic-Pacific conditions specifically favored  \gls{wt} 4 occurrence, helping to explain why the most intense rainfall anomalies occurred specifically in the \gls{lprb}.
+Examination of the \gls{sst} anomalies observed during NDJF 2015-16 (supplemental figure S4) indicates that the mechanism illustrated in \cref{fig:chaco-nochaco} was active - particularly in December 2015 when the most intense rainfall occurred.
+This suggests that not only did \gls{enso}-\gls{mjo} conditions favor \gls{sallj} activity, but Atlantic-Pacific interactions specifically favored  \gls{wt} 4 occurrence, helping to explain why the most intense rainfall anomalies occurred specifically in the \gls{lprb}.
 
-This Atlantic-Pacific interaction may also explain the skill of dynamical models for representing heavy rainfall in the region.
+This Atlantic-Pacific interaction may also help to explain spatial uncertainty in model-based estimates of heavy rainfall in the region.
 In order to adequately forecast rainfall in certain parts of southeastern South America during El Ni\~no years, models need to reproduce stationary wave trains originating in the Pacific and the Atlantic and their interactions \citep{Barreiro2017}.
 Other mechanisms that have been known to modulate rainfall signals in this region include the \gls{sacz} \citep{Carvalho2004,Munoz2015,Munoz2016} and land-biosphere-atmosphere interactions \citep{Grimm2000,Grimm2007} which also tend to be poorly represented in models \citep{Koster2004,Green2017}.
-The stationary wavetrain interactions, land-atmosphere interactions, and topography may explain why simulating heavy rainfall in this region is so difficult  \cref{fig:chiclet,fig:subs-prob-fcst}.
+The stationary wavetrain interactions, land-atmosphere interactions, and topography may explain why simulating heavy rainfall in this region is so difficult (\cref{fig:chiclet,fig:subs-prob-fcst}).
 Improving understanding of these phenomena is important opportunity for \gls{s2s} prediction, and is left for future work.
 
-Finally it is of interest to consider the link between the observed rainfall events and the observed flooding.
+Finally, it is of interest to consider the link between the observed rainfall events and the observed flooding.
 Although we motivated this work by describing the impacts of severe flooding in the \gls{lprb}, the analysis presented has focused on climate drivers of rainfall.
@@
 
-\section{Summary} \label{sec:summary}
+\section{Summary}\label{sec:summary}
 
@@
 
-The heavy rainfall observed was driven by enhanced moisture inflow from the low-level jet and convergence associated with baroclinic systems.
+Both enhanced moisture inflow from the low-level jet and convergence associated with baroclinic systems drove the observed heavy rainfall.
 Repeated \gls{sallj} events, particularly No-Chaco jet events, led to favorable conditions for mesoscale convective activity in this region.
-This activity was modulated by cross-timescale dynamics including a strong El Ni\~{n}o and an active \gls{mjo}  in phases 4-5, which favored \gls{sallj} occurrence, and the presence of a dipolar SST anomaly in the central southern Atlantic Ocean, which enhanced the probability of occurrence of No-Chaco jet events.
+Large-scale climate patterns at both seaonal and sub-seaonal scales favored the synoptic weather patterns observed.
+Notably, a strong El Ni\~{n}o and an active \gls{mjo} in phases 4-5 favored \gls{sallj} occurrence.
+The presence of a dipolar \gls{sst} anomaly in the central southern Atlantic Ocean also favored the occurrence of No-Chaco jet events.
 
-Numerical forecasts skillfully predicted enhanced risk of heavy rainfall at the seasonal scale, consistent with the observed \gls{enso} signal, but biases in the spatial patterns suggest that models are imperfectly capturing the physical interactions between the Pacific and the Atlantic basins.
-At sub-seasonal time scales, uncorrected model forecasts of rainfall had limited skill beyond 10-15 days, though use of Model Output Statistics - particularly the \gls{pcr} and \gls{cca} methods that correct both spatial patterns and magnitudes - substantially improved forecast skill.
+Numerical forecasts skillfully predicted enhanced risk of heavy rainfall at the seasonal scale, consistent with the observed \gls{enso} signal, but biases in the spatial patterns of forecast rainfall suggest that models imperfectly capture the physical interactions between the Pacific and the Atlantic basins.
+At sub-seasonal time scales, uncorrected model forecasts of rainfall had limited skill beyond 15 days, though use of \acrlong{mos} -- particularly the \gls{pcr} and \gls{cca} methods that correct both spatial patterns and magnitudes -- substantially improved forecast skill.
 
@@
 \acknowledgments
-The authors thank David Farnham, Upmanu Lall, and Andrew Robertson for insightful conversations and guidance.
+The authors thank David Farnham, Upmanu Lall, Laureline Josset, and Andrew Robertson for insightful conversations and guidance.
 JDG thanks the NSF GRFP program for support (grant DGE 16-44869).
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%-------------------------------------------------------------------------------
 % TABLES
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%-------------------------------------------------------------------------------
 \begin{table}
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%-------------------------------------------------------------------------------
 % FIGURES
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%-------------------------------------------------------------------------------
 
@@
 		The relative frequency of occurrence of each weather type (in days) is presented on the top of each column.
-	}
-	\label{fig:wt-composite}
+	}\label{fig:wt-composite}
 \end{figure*}
@@
 \begin{figure*}
+	\includegraphics[width=\textwidth]{wt_mjo_enso.pdf}
+	\caption{
+		Anomalous probability of occurrence of each weather type concurrent with observance of each \gls{mjo} phase.
+		When \gls{mjo} amplitude is less than 1, it is defined as neutral phase (0).
+		Plots are shown separately for El Ni\~no (NINO 3.4 $>1$), La Ni\~na (NINO 3.4 $< -1$), and Neutral \gls{enso} phases.
+		Only values which are significant at $\alpha=0.10$, calculated with a bootstrap of 5000 samples, are shown.
+	}\label{fig:wt-mjo-enso}
+\end{figure*}
+
+\begin{figure*}
 	\noindent\includegraphics[width=6.4in]{ChacoNoChacojet.pdf}
@@
 
-\begin{figure*}
-	\includegraphics[width=\textwidth]{predictors_eofs.pdf}
-	\caption{
-		Spearman (rank) correlation coefficient between several \gls{s2s} climate indices and weather type occurrence at monthly time step.
-		These, correlations are computed between the proportion of occurrence of each weather type and the monthly-mean time series of the climate index.
-		For the \gls{mjo} phases the monthly-mean time series is the proportion of occurrence of that phase for the given month.
-		The NINO 3.4 time series was up-sampled from monthly to daily values by persisting the most recent monthly value (no interpolation).
-		The \acrfull{scad} predictor, labeled Dipole, is the mean meridional gradient of \gls{sst} anomalies over the region \SIrange{30}{10}{\degree W} and \SIrange{40}{15}{\degree S}, up-sampled to daily values using the same method as the NINO 3.4 time series.
-	}\label{fig:predictor-eof}
-\end{figure*}
-
-\begin{figure*}
-	\includegraphics[width=\textwidth]{wt_mjo_enso.pdf}
-	\caption{
-		Anomalous probability of occurrence of each weather type concurrent with observance of each \gls{mjo} phase. 
-		When \gls{mjo} amplitude is less than zero, it is defined as neutral phase (0).
-		Plots are shown separately for El Ni\~no (NINO 3.4 $>1$), La Ni\~na (NINO 3.4 $< -1$), and Neutral \gls{enso} phases.
-		Only values which are significant at $\alpha=0.10$, calculated with a bootstrap of 5000 samples, are shown.
-	}\label{fig:wt-mjo-enso}
-\end{figure*}
-
 \end{document}

### 2018-05-06 reports/writeup/Supplement.tex
make supplemental figs fit on page

--- 2018-03-08 reports/writeup/Supplement.tex
@@
 \begin{figure}
-	\noindent\includegraphics[width=6.4in]{sst_anomalies.pdf}
+	\includegraphics[width=\textwidth]{mjo_ts.pdf}
+	\caption{
+		Evolution of the MJO during NDJF 2015-16.
+		Points are plotted on RMM1 ($x$-axis) and RMM2 ($y$-axis), derived from leading EOFs of OLR fields \citep{Wheeler2004}.
+		Gray lines divide the plot into the eight phases used in the text.
+		Colors show the time evolution of the system from 01 November 2015 (blue) to 29 February 2016 (yellow).
+		Gray circle indicates the area of neutral MJO activity with amplitude $<0$.
+	}
+\end{figure}
+
+\begin{figure}
+	\noindent\includegraphics[width=6in]{sst_anomalies.pdf}
 	\caption{
@@
 \begin{figure}
-	\includegraphics[width=\textwidth]{wt_classifiability.pdf}
+	\includegraphics[width=\textwidth]{predictors_eofs.pdf}
 	\caption{
-		Classifiability index (see Methods) calculated for several chosen values of $K$, representing the number of weather types produced.
-		Results are produced with only 75 simulations for each $K$.
-	}
+		Spearman (rank) correlation coefficient between several S2S climate indices and weather type occurrence at monthly time step.
+		These correlations are computed between the proportion of occurrence of each weather type and the monthly-mean time series of the climate index.
+		For the MJO phases the monthly-mean time series is the proportion of occurrence of that phase for the given month.
+		The NINO 3.4 time series was up-sampled from monthly to daily values by persisting the most recent monthly value (no interpolation).
+		The SCAD predictor, labeled Dipole, is the mean meridional gradient of SST anomalies over the region \SIrange{30}{10}{\degree W} and \SIrange{40}{15}{\degree S}, up-sampled to daily values using the same method as the NINO 3.4 time series.
+	}\label{fig:predictor-eof}
 \end{figure}

### 2018-05-07 reports/writeup/paper.tex
Proofreading

@@
 %-------------------------------------------------------------------------------
-
 \documentclass{ametsoc}
@@
 	These circulation patterns were favored by cross-timescale interactions of a very strong El Ni\~{n}o event, an unusually persistent Madden-Julian Oscillation in phases four and five, and the presence of a dipole SST anomaly in the central southern Atlantic Ocean.
-	The simultaneous use of seasonal and sub-seasonal heavy rainfall predictions could have provided decision makers useful information about the start of these flooding events from at least two-to-four weeks in advance.
+	The simultaneous use of seasonal and sub-seasonal heavy rainfall predictions could have provided decision makers useful information about the start of these flooding events from two to four weeks in advance.
 	Probabilistic seasonal forecasts available at the beginning of November successfully indicated heightened probability of heavy rainfall (\nth{90} percentile) over southern Paraguay and Brazil for December-February.
-	Raw sub-seasonal forecasts of heavy rainfall exhibited limited skill at lead times beyond the first two predicted weeks, but a Model Output Statistics approach involving principal component regression substantially improved the spatial distribution of skill for Week 3 relative to other methods tested including extended logistic regressions.
-	A continuous monitoring of climate drivers impacting rainfall in the region, and the use of bias-corrected heavy precipitation seasonal and sub-seasonal forecasts, may help improve flood preparedness in this and other regions.
+	Raw sub-seasonal forecasts of heavy rainfall exhibited limited skill at lead times beyond the first two predicted weeks, but a Model Output Statistics approach involving principal component regression substantially improved the spatial distribution of skill for week 3 relative to other methods tested including extended logistic regressions.
+	A continuous monitoring of climate drivers impacting rainfall in the region, and the use of statistically corrected heavy precipitation seasonal and sub-seasonal forecasts, may help improve flood preparedness in this and other regions.
 }
@@
 However, the flat topography limits the river's ability to carry the summer runoff, causing seasonal inundation of the Pantanal and distributing the river discharge in time \citep{Bravo2011,Barros2004}.
-Thus, the streamflow maxima typically occurs in phase with precipitation upstream of the Pantanal, while downstream of the Pantanal - an area which we define in \cref{fig:study-area} as the \acrlong{lprb} - the annual peak typically occurs between April and July.
+Thus, upstream of the Pantanal the streamflow maxima typically occurs in phase with precipitation, while downstream of the Pantanal - an area which we define in \cref{fig:study-area} as the \acrlong{lprb} - the annual peak typically occurs between April and July.
 
@@
 
-Logistic regression models the conditional probability of binary events, given one or more predictors, and has been widely used in \gls{mos} approaches for binary outcomes.
+Logistic regression models the probability of binary events, conditional on one or more predictors, and has been widely used in \gls{mos}.
 Nonetheless, when using logistic regression to address multiple thresholds via independent fits, the predicted probabilities are, in general, not mutually consistent \citep{Messner2014}.
-The \gls{xlr} was designed to address this shortcoming \citep{Wilks2009}, via the consideration of a transformation of the thresholds of interest as an additional predictor variable.
+The \gls{xlr} was designed to address this shortcoming via the consideration of a transformation of the thresholds of interest as an additional predictor variable \citep{Wilks2009}.
 The \gls{hxlr}, a generalization of the \gls{xlr}, was proposed to appropriately use the ensemble spread as predictor for the dispersion of the predictive distribution \citep{Messner2014}.
@@
 Along the cold front associated with this system, a low-level northerly jet advects  heat and moisture to the region.
-As the system progresses, the jet below \SI{20}{\degree S} transitions from predominantly meridional flow (``Chaco Jet''; $t=\SI{-1}{\day}$ to predominantly zonal flow (``No-Chaco Jet'').
+As the system progresses, the jet below \SI{20}{\degree S} transitions from predominantly meridional flow (``Chaco Jet''; $t=\SI{-1}{\day}$) to predominantly zonal flow (``No-Chaco Jet''; $t=\SI{0}{\day}$).
 The pattern resembles composites identified using one standard deviation exceedances of rainfall at \SI{60}{\degree W}, \SI{30}{\degree S} \citep{Liebmann2004} and analysis for the \nth{95} or 99th percentiles of daily rainfall (not shown) yield similar results, implying that the synoptic mechanism for the most heavy events is not fundamentally distinct from the mechanism for moderate-intensity events.
@@
 \Gls{wt} 1 describes a \gls{sallj} event in which the strongest wind penetrates southward of \SI{25}{\degree S}, leading to heavy rainfall over NE Argentina and Uruguay; this has been called a ``Chaco Jet'' event \citep{Salio2002}.
-\Gls{wt} 4 is also associated with a \gls{sallj} event, but the wind turns to the East northward of \SI{25}{\degree S}, leading to heavy rainfall over Eastern Paraguay and SW Brazil; this has been called a ``No-Chaco'' Jet event \citep{Vera2006}.
+\Gls{wt} 4 also shows \gls{sallj} activity, but the wind turns to the East northward of \SI{25}{\degree S}, leading to heavy rainfall over Eastern Paraguay and SW Brazil; this has been called a ``No-Chaco'' Jet event \citep{Vera2006}.
 Supplemental Table S1 shows the centroids of each cluster, in the 4-dimensional phase space of the leading \glspl{eof} of \SI{850}{\hecto\pascal} streamfunction.
@@
 Sub-seasonal predictions are still too new to be used as operational tools, and their skill is normally not high enough to be useful for most decision-making \citep{Vigaud2017}.
-Nonetheless, the international \gls{s2s} Prediction Project \citep{Vitart2016} has been providing since 2015 free access to almost-real-time sub-seasonal forecasts from multiple models, an opportunity to explore how well the heavy rainfall events of the first week of December 2015 could have been predicted.
+Nonetheless, the international \gls{s2s} Prediction Project \citep{Vitart2016} provides free access to almost-real-time sub-seasonal forecasts from multiple models, an opportunity to explore how well the heavy rainfall events of the first week of December 2015 could have been predicted.
 
@@
 As explained in \cref{sec:diagnostics}\ref{sec:flood}, in this region the flat topography (\cref{fig:study-area}) means that the Lower Paraguay River reacts slowly to rainfall \citep{Bravo2011,Barros2004}, explaining the slow but steady rise in river levels from mid November 2015 to early January 2016, as shown in \cref{fig:streamflow}.
-The observed flood peaks during 2015-16 also seem to occur in the context of an active phase of a multi-decadal oscillation, possibly associated with low-frequency Pacific activity \citep{Collischonn2001,Huang2005}; groundwater dynamics are also important in explaining this behavior \citep{Santos2016}.
+The observed flood peaks during 2015-16 also seem to occur in the context of an active phase of a multi-decadal oscillation, possibly associated with low-frequency Pacific activity \citep{Collischonn2001,Huang2005}.
+Groundwater dynamics are also important in explaining this behavior \citep{Santos2016}.
 Parsing the relative impacts of deforestation and land use changes in the river basin, installation of hydroelectric generation at the Itaipu and Yacyreta sites, river channel modification, antecedent conditions, and climate variability on flood levels will require gathering improved hydrological data and building a comprehensive system model, which is beyond the scope of this paper.
 
-From a policy perspective, reducing flood risk exposure in this region is key to reducing flood losses; flood events not only in 2015-16 but also in 2014, 2017, and 2018 have caused substantial damage, and highlight the need for flood risk management strategies.
+From a policy perspective, reducing flood risk exposure in this region is key to reducing flood losses.
+Flood events not only in 2015-16 but also in 2014, 2017, and 2018 have caused substantial damage, and highlight the need for flood risk management strategies.
 Doing so will require compiling information on the properties, businesses, and infrastructure that are vulnerable to flooding.

### 2018-09-12 reports/poster/blocks/at-a-glance.tex
Add poster for AGM to present

\begin{block}{Conceptual Framework}
    Beginning in Nov.  2015, repeated intense rainfall events associated with mesoscale convective activity caused severe flooding along the
    Paraguay-Paran ́a river system (fig. 5, red box), displacing over \num{170000} people.
    We use a weather typing approach within a diagnostic framework to show that:
    \begin{itemize}
        \item Moisture and energy advection via the South American Low-Level Jet (SALLJ)~\cite{Marengo:2004kr}, particularly during ``No-Chaco'' events~\cite{Vera:2006ib} favored mesoscale convective activity
        \item Strong El Ni\~{n}o and active MJO favored a strong SALLJ but an Atlantic dipole pattern influenced the jet’s exit region
        \item Numerical forecasts predicted enhanced risk of heavy rainfall at the seasonal scale, but biases in spatial patterns suggest difficulties representing Pacific-Atlantic interaction
        \item Uncorrected sub-seasonal model forecasts of rainfall had limited skill beyond 10-15 days; use of Model Output Statistics –         particularly methods that correct both spatial patterns and magnitudes – substantially improved forecast skill 
    \end{itemize}
    \begin{framed}
        \begin{figure}
            \centering
            \includegraphics[width=0.49\textwidth]{asuncion-inundaciones.jpg}\hfill\includegraphics[width=0.49\textwidth]{fig-lidia-perez.jpg}
            \caption{Asunci\'{o}n, Paraguay, 2015-16}
            \label{fig:asuncion-flooded}
        \end{figure}
    \end{framed}
\end{block}

### 2018-09-12 reports/poster/blocks/atlantic-pacific.tex
Add poster for AGM to present

\begin{block}{Atlantic-Pacific Interaction and WT4}
    While the occurrence of \gls{wt}1 during NDJF 2015-16 is well explained by ENSO and MJO variability, these features alone do not explain the occurrence of \gls{wt}4, the ``No-Chaco'' jet event.
    Previous studies emphasize the importance of Pacific-Atlantic interaction for forecasting climate effects in this region~\cite{Barreiro:2017ct}.
    A persistent SST dipole in the central southern Atlantic Ocean favors the occurrence of \gls{wt}4 by blocking transient extratropical wave activity coming from the Pacific, facilitating transitions from Chaco jet events (WT 1) to No-Chaco jet events (WT 4) via enhanced low-level wind circulation from southern Brazil towards the Atlantic, and back to north-east Brazil and the Amazon (see \cref{fig:atlantic-pacific}) due to land-sea temperature contrasts.
    Composite analysis (not shown) of months with many \gls{wt}4 occurrences is consistent with the schematic shown here.
    \begin{framed}
        \begin{figure}
            \centering
            \includegraphics[width=\textwidth]{ChacoNoChacojet.pdf}
            \caption{
                Schematics of low-level jet events (red arrows) during austral summer of El Ni\~{n}o years.
            }\label{fig:atlantic-pacific}
        \end{figure}
    \end{framed}
\end{block}

### 2018-09-12 reports/poster/blocks/mjo-enso.tex
Add poster for AGM to present

\begin{block}{ENSO, MJO, and Weather Types}
    \begin{framed}
        \begin{figure}
            \centering
            \caption{
                Anomalous probability of occurrence of each weather type concurrent with observance of each MJO and ENSO phase.
                Only values which are significant at $p < 0.10$ are shown.
            }\label{fig:wt-mjo-ensoflooded}
            \includegraphics[width=\textwidth]{wt_mjo_enso.pdf}
        \end{figure}
    \end{framed}
    \gls{wt} 1 occurs more frequently during El Ni\~{n}o years for most \gls{mjo} phases, particularly during phase 2.
    During El Ni\~no years, \gls{wt} 3 - associated with dryness over the \gls{lprb} - occurs less frequently during \gls{mjo} phases 4, 6, and 7, and more often during \gls{mjo} phase 8; this is consistent with the lack of \gls{wt} 3 during December 2015 and the frequent \gls{wt} 3 occurrence in mid-January 2016 (\cref{fig:fig:rain-wtype}).
\end{block}

### 2018-09-12 reports/poster/blocks/model-output-stats.tex
Add poster for AGM to present

\begin{block}{Model Output Statistics}
    We explore whether using \gls{mos}~\cite{Glahn:1972iw} can improve the modeled representation of rainfall (\cref{fig:subs-prob-fcst}).
    Specifically, we use: the raw model output (Raw); \gls{xlr}; \gls{hxlr}; \gls{pcr}; and \gls{cca} using 20 years of ECMWF forecasts.
    \Cref{fig:subs-prob-fcst} indicates that better forecasts are obtained when both magnitude and spatial corrections are performed (\gls{pcr} and \gls{cca}).
    The enhanced skill is achieved through the spatial corrections via the EOF-based regressions, which – in contrast with the extended logistic models – use information from multiple grid-boxes,.
    \begin{framed}
        \begin{figure}
            \noindent\includegraphics[width=0.9\textwidth]{mos_forecasts.pdf}
            \caption{
                \Gls{mos}-adjusted S2S model forecasts and skill scores.
                Top row shows the heavy rainfall ($>$\nth{90} percentile exceedance) forecast for 1-7 December 2015 as the odds ratio relative to climatology $\text{odds} = \frac{p}{1-p}\frac{1-p_c}{p_c}$.
                Second row shows the Ignorance score $\text{IGN}=-\log_2 p(Y)$.
                Bottom row shows the 2AFC skill score for each grid cell.
                For all three rows, the grid cells which experienced a \nth{90} percentile exceedance for 1-7 December 2015 are outlined in black.
            }\label{fig:subs-prob-fcst}
        \end{figure}
    \end{framed}
\end{block}

### 2018-09-12 reports/poster/blocks/observations-wts.tex
Add poster for AGM to present

\begin{block}{Observations and Weather Types}
    Observations come from:
    \begin{itemize}
        \item Rainfall:  CPC Global Unified~\cite{Xie:1996ga}
        \item Atmosphere:  NCAR-NCEP Reanalysis II~\cite{Kanamitsu:2002ig}
    \end{itemize}
    We use weather typing~\cite{Munoz:2015dc} to represent daily circulation patterns:
    \begin{enumerate}
        \item Calculate streamfunction $\Psi$ from meridional and zonal wind~\cite{Dawson:2016cu}
        \item Project \SI{850}{\hecto\pascal} streamfunction onto leading 4 EOFs
        \item $K$-means clustering using classifiability index~\cite{Michelangeli:1995es} to generate a single weather type for each day.\\
    \end{enumerate}
    Weather typing simplifies dynamics of daily rainfall but \emph{facilitates analysis of sequences of daily weather patterns}.
    They are associated with patterns that have been well described in the literature; particularly relevant are:
    \begin{enumerate}
        \item \gls{wt}1 represents ``Chaco'' jet event~\cite{Salio:2002ev}
        \item \gls{wt}4 represents ``No-Chaco'' jet events~\cite{Vera:2006ib}
    \end{enumerate}
    \begin{framed}
        \begin{figure}
            \centering
            \noindent\includegraphics[width=\textwidth]{wt_composite.pdf}
            \caption{
                Composite anomalies associated with each weather type.
                Top row (a-f) shows streamfunction anomalies at \SIlist{850}{\hecto\pascal}.
                Strongest 20\% of wind anomaly vectors over the plot area are also shown.
                Bottom row (g-l) shows rainfall anomalies, in units of \si{\milli\meter\per\day}.
                The relative frequency of occurrence of each weather type (in days) is presented on the top of each column.
            }
            \label{fig:weather-type-composite}
        \end{figure}
    \end{framed}
\end{block}

### 2018-09-12 reports/poster/blocks/observed-anomalies.tex
Add poster for AGM to present

\begin{block}{Observed Circulation Anomalies}
    \begin{framed}
        \begin{figure}
            \centering
            \caption{
                Time series of area-averaged rainfall in the \acrlong{lprb} (\cref{fig:study-area}) for each day of NDJF  2015-16.
                Lines indicate the rainfall value, in units of \si{\milli\meter\per\day}.
                The weather type corresponding to each day is indicated in an adjacent text label.
                Dashed lines blue indicate (from bottom to top) the climatological 50th, \nth{90}, and 99th percentiles of NDJF area-averaged rain over the \acrlong{lprb}.
            }
            \noindent\includegraphics[width=0.9\textwidth]{rain_wt_201516.pdf}
            \label{fig:rain-wtype}
        \end{figure}
    \end{framed}
    During austral summer (NDJF) 2015-16, most heavy rainfall occurred during \glspl{wt} 1 and 4 (\cref{fig:rain-wtype}). 
    Monthly-scale circulation anomalies (\cref{fig:anomaly-ndjf}) show a weak anticyclonic circulation that set up over central Brazil during November 2015 and strengthened into the following month.
    In January 2016 it weakened before returning in February 2016.
    The observed rainfall and circulation anomalies are consistent with the aggregation of the observed weather types shown in \cref{fig:rain-wtype}.
    \begin{framed}
        \begin{figure}
            \centering
            \noindent\includegraphics[width=0.9\textwidth]{anomalies_ndjf1516.pdf}
            \caption{
                Monthly composite anomalies observed during NDJF 2015-16.
                Top row shows streamfunction anomalies at \SI{850}{\hecto\pascal}.
                Bottom row shows rainfall anomalies, in units of \si{\milli\meter\per\day}.
            }
            \label{fig:anomaly-ndjf}
        \end{figure}
    \end{framed}
\end{block}

### 2018-09-12 reports/poster/blocks/s2s-forecasts.tex
Add poster for AGM to present

--- 2017-12-02 reports/poster-agu-2017/blocks/s2s-model-forecasts.tex
@@
 \begin{block}{S2S Model Forecasts}
-  \begin{mdframed}
+    \begin{framed}
         \begin{figure}
+            \centering
             \caption{
-      Chiclet diagram of ensemble-mean precipitation anomaly forecasts over the Lower Paraguay River Basin from ECMWF S2S forecast data, as a function of the forecast target date (horizontal axis) and lead time (vertical axis).
+                Chiclet diagram of ensemble-mean precipitation anomaly forecasts over the \gls{lprb} from ECMWF S2S forecast data, as a function of the forecast target date (horizontal axis) and lead time (vertical axis).
                 Time series of daily precipitation over the same area is plotted with $y$-axis inverted.
-  	}
-    \noindent\includegraphics[width=0.925\textwidth]{chiclet-s2s-area-averaged.pdf}
-  	\label{fig:chiclet}
+            }\label{fig:chiclet}
+            \noindent\includegraphics[width=\textwidth]{chiclet.pdf}
         \end{figure}
-  \end{mdframed}
-
-  \Cref{fig:chiclet} uses a Chiclet diagram \cite{Carbin:2016fx} to visualize, as a function of lead time, the time evolution of the uncorrected, ensemble-mean rainfall anomaly forecast, spatially averaged over the Lower Paraguay River Basin.
+    \end{framed}
+    \Cref{fig:chiclet} uses a Chiclet diagram~\cite{Carbin:2016fx} to visualize, as a function of lead time, the time evolution of the uncorrected, ensemble-mean rainfall anomaly forecast, spatially averaged over the \gls{lprb}.
     At lead times greater than about two weeks, the ensemble-mean forecast is slightly wetter than climatology.
@@
     At intermediate timescales, the model successfully forecast the strongest breaks and pauses in the rainfall, such as the heavy rainfall during December 2015 and the dry period during mid-January 2016.
-
 \end{block}
-  \begin{block}{Model Output Statistics}
-
-  We explore whether using Model Output Statistics \cite[MOS;][]{Glahn:1972vt} can improve the modeled representation of rainfall (\cref{fig:subs-prob-fcst}).
-  Specifically, we use: the raw model output(Raw); extended logistic regression \cite[XLR;][]{Wilks:2009bk}; heteroscedastic \cite[HXLR;][]{Messner:2014gp}; principal component regression \cite[PCR;][]{Mason:2008da,Wilks:2006fx}; and canonical correlation analysis \cite[CCA;][]{Mason:2008da,Barnston:1992gd} using 20 years of ECMWF forecasts.
-
-  \vspace{0.5cm}
-
-  \Cref{fig:subs-prob-fcst} indicates that better forecasts are obtained when both magnitude and spatial corrections are performed (PCR and CCA).
-  The enhanced skill is achieved through the spatial corrections via the EOF-based regressions, which -- in contrast with the extended logistic models -- use information from multiple grid-boxes,.
-
-  \begin{mdframed}
-  \begin{figure}
-  	\noindent\includegraphics[width=0.925\textwidth]{s2s-forecast-mos.pdf}
-  	\caption{
-    MOS-adjusted S2S model forecasts and skill scores.
-    Top row shows the heavy rainfall ($>$\nth{90} percentile exceedance) forecast for 1-7 December 2015 as the odds ratio $\text{odds}_{r} \equiv \frac{p}{\qty(1 - p)} \frac{\qty(1 - p_c)}{p_c}$.
-    Second row shows the Ignorance score $\text{IGN} \equiv - \log_2 p(Y)$.
-    Bottom row shows the 2AFC skill score for each grid cell.
-    For all three rows, the grid cells which experienced a \nth{90} percentile exceedance for 1-7 December 2015 are outlined in black.
-        \label{fig:subs-prob-fcst}
-  	}
-  \end{figure}
-  \end{mdframed}
-\end{block}

### 2018-09-12 reports/poster/blocks/study-area.tex
Add poster for AGM to present

\begin{block}{\acrlong{lprb}}
    Flat topography limits the river’s ability to carry the summer runoff, causing seasonal inundation of the Pantanal and distributing the river discharge in time~\cite{Barros:2004wg, Bravo:2012iq}.
    During NDJF 2015-16, river stage [height] throughout the Lower Paraguay River Basin reached nearly three times climatological levels (not shown).
    \begin{framed}
        \begin{figure}
            \noindent\includegraphics[width=0.7\textwidth]{study_area.jpg}
            \caption{
                Topographical map of the study area.
                Colors indicate $\log_{10}$ of elevation, in \si{\meter}.
                (a): all of South America.
                The domains of the \acrlong{lprb} and the domain used for weather typing are indicated in red and blue, respectively.
                (b): The \acrfull{lprb}.
                As for (a), the \gls{lprb} is marked with a red box.
            }\label{fig:study-area}
        \end{figure}
    \end{framed}
\end{block}

### 2018-09-12 reports/poster/poster.tex
Add poster for AGM to present

\documentclass[serif,mathserif,final]{beamer}
\mode<presentation>{\usetheme{Lankton}}

\usepackage[
  orientation=landscape,
  width=40in,height=30in,%size=a0, % paper size
  scale=1.05 % font scale factor
  ]{beamerposter}
 \setbeamertemplate{caption}[numbered] % number figs

% Graphics Path
\usepackage{graphicx}
\graphicspath{{../../figs/}{figs/}}

% OTHER PACKAGE CALLS
\usepackage[utf8]{inputenc}
\usepackage{
  amsmath,amsfonts,amssymb,
  xspace,
  framed,
  siunitx,
  nth,
  physics
}
\sisetup{round-mode = figures, round-precision = 3}
\usepackage{cleveref}

%------ USE BIBLATEX WITH EXTREMELY MINAMALIST STYLE -----
% Use BibLaTeX
\usepackage[
	backend=biber,
	url=false,
	isbn=false,
	doi=false,
	maxnames = 2,
	style=numeric-comp,
	uniquename=false,
	uniquelist=false,
	firstinits=true,
	sorting=none,
	natbib=false,
	isbn=false,
]{biblatex}
\addbibresource{../library.bib}
\renewbibmacro{in:}{}

% One-paragraph bibliography environment
\defbibenvironment{bibliography}
{\list%
  {\printtext[labelnumberwidth]{%
    \printfield{prefixnumber}%
    \printfield{labelnumber}}%
  \ifentrytype{article}{}{}}%
    {\setlength{\leftmargin}{0pt}%
    \setlength{\topsep}{0pt}}%
    \renewcommand*{\makelabel}[1]{##1}}%
  {\endlist}%
{\mkbibitem}

% \mkbibitem just prints item label and non-breakable space
\makeatletter
\newcommand{\mkbibitem}{\@itemlabel\addnbspace}
\makeatother

% Add breakable space between bibliography items
\renewcommand*{\finentrypunct}{\addperiod\space}

% Set fields to clear
\AtEveryBibitem{%
	\clearfield{month}%
	\clearfield{day}%
  \clearlist {language}%why this has to be clear list.....
  \clearfield{pages}%
	\clearfield{pagetotal}%
	\clearfield{eprinttype}%
	\clearfield{eprint}%
	\clearfield{number}%
	\clearfield{volume}%
	\clearfield{issue}%
	\clearfield{name:given}
	\clearfield{name:first}
	\ifentrytype{article}{%
		\clearfield{title}%
	}%
}

%------- GLOSSARIES PACKAGE -------
\usepackage[acronym]{glossaries}
\makeglossaries
\newacronym{wt}{WT}{weather type}
\newacronym{sallj}{SALLJ}{South American low-level jet}
\newacronym{enso}{ENSO}{El Ni\~{n}o-Southern Oscillation}
\newacronym{mjo}{MJO}{Madden-Julian Oscillation}
\newacronym{sacz}{SACZ}{South Atlantic Convergence Zone}
\newacronym{iri}{IRI}{International Research Institute for Climate and Society}
\newacronym{ecmwf}{ECMWF}{European Centre for Medium-Range Weather Forecasts}
\newacronym{s2s}{S2S}{sub-seasonal to seasonal}
\newacronym{eof}{EOF}{empirical orthogonal function}
\newacronym{mos}{MOS}{model output statistics}
\newacronym{xlr}{XLR}{homoscedastic extended logistic regression}
\newacronym{hxlr}{HXLR}{heteroscedastic extended logistic regression}
\newacronym{pcr}{PCR}{principal components regression}
\newacronym{cca}{CCA}{canonical correlation analysis}
\newacronym{sst}{SST}{sea-surface temperature}
\newacronym{scad}{SCAD}{South Central Atlantic Dipole}
\newacronym{lprb}{LPRB}{Lower Paraguay River Basin}

%-- Header and footer information ----------------------------------
\newcommand{\footleft}{Doss-Gollin, J., Á. G. Muñoz, S. J. Mason, and M. Pastén (2018), Heavy rainfall in Paraguay during the 2015-2016 austral summer: causes and sub-seasonal-to-seasonal predictive skill, Journal of Climate, doi:\href{dx.doi.org/10.1175/JCLI-D-17-0805.1}{10.1175/JCLI-D-17-0805.1}.}
\newcommand{\footright}{}
\title{Heavy Rainfall in Paraguay During the 2015-2016 Austral Summer: Causes and Sub-Seasonal-to-Seasonal Predictive Skill}
\author{James Doss-Gollin\inst{1} \quad \'{A}ngel G. Mu\~{n}oz\inst{2,3} \quad Simon J. Mason\inst{3} \quad Max Past\'{e}n\inst{4}}
\institute
{\inst{1} Columbia Water Center, Columbia University \inst{2} AOS, Princeton University \inst{3} International Research Institute for Climate and Society, Columbia University \inst{4} Direcci\'{o}n de Meteorolog\'{i}a e Hidrolog\'{i}a, Paraguay}

%-------------------------------------------------------------------

%-- Main Document --------------------------------------------------
\begin{document}
\begin{frame}{}
  \begin{columns}[t]

    %-- Column 1 ---------------------------------------------------
    \begin{column}{0.22\linewidth}

        \input{blocks/at-a-glance.tex}
        \input{blocks/observations-wts.tex}

    \end{column}%1

    %-- Column 2 ---------------------------------------------------
    \begin{column}{0.22\linewidth}

      \input{blocks/observed-anomalies.tex}
      \input{blocks/study-area.tex}

    \end{column}%2

    %-- Column 3 ---------------------------------------------------
    \begin{column}{0.22\linewidth}

      \input{blocks/s2s-forecasts.tex}
      \input{blocks/model-output-stats.tex}

    \end{column}%3

    %-- Column 4 ---------------------------------------------------
    \begin{column}{0.22\linewidth}

      \input{blocks/mjo-enso.tex}
      \input{blocks/atlantic-pacific.tex}

      %-- Block 4-2
      \begin{block}{References}
          \renewcommand*{\bibfont}{\footnotesize}
          \printbibliography[heading=none]
      \end{block}

    \end{column}%1

  \end{columns}
\end{frame}
\end{document}

### 2018-09-12 reports/poster/blocks/mjo-enso.tex
excise princeton and fix erroneous cref

@@
     \gls{wt} 1 occurs more frequently during El Ni\~{n}o years for most \gls{mjo} phases, particularly during phase 2.
-    During El Ni\~no years, \gls{wt} 3 - associated with dryness over the \gls{lprb} - occurs less frequently during \gls{mjo} phases 4, 6, and 7, and more often during \gls{mjo} phase 8; this is consistent with the lack of \gls{wt} 3 during December 2015 and the frequent \gls{wt} 3 occurrence in mid-January 2016 (\cref{fig:fig:rain-wtype}).
+    During El Ni\~no years, \gls{wt} 3 - associated with dryness over the \gls{lprb} - occurs less frequently during \gls{mjo} phases 4, 6, and 7, and more often during \gls{mjo} phase 8; this is consistent with the lack of \gls{wt} 3 during December 2015 and the frequent \gls{wt} 3 occurrence in mid-January 2016 (\cref{fig:rain-wtype}).
 \end{block}

### 2018-09-12 reports/poster/poster.tex
excise princeton and fix erroneous cref

@@
 \title{Heavy Rainfall in Paraguay During the 2015-2016 Austral Summer: Causes and Sub-Seasonal-to-Seasonal Predictive Skill}
-\author{James Doss-Gollin\inst{1} \quad \'{A}ngel G. Mu\~{n}oz\inst{2,3} \quad Simon J. Mason\inst{3} \quad Max Past\'{e}n\inst{4}}
+\author{James Doss-Gollin\inst{1} \quad \'{A}ngel G. Mu\~{n}oz\inst{2} \quad Simon J. Mason\inst{2} \quad Max Past\'{e}n\inst{3}}
 \institute
-{\inst{1} Columbia Water Center, Columbia University \inst{2} AOS, Princeton University \inst{3} International Research Institute for Climate and Society, Columbia University \inst{4} Direcci\'{o}n de Meteorolog\'{i}a e Hidrolog\'{i}a, Paraguay}
+{\inst{1} Columbia Water Center, Columbia University \inst{2} International Research Institute for Climate and Society, Columbia University \inst{3} Direcci\'{o}n de Meteorolog\'{i}a e Hidrolog\'{i}a, Paraguay}
 

### 2019-01-27 reports/preprint/preprint.tex
add codes for preprint on eartharxiv

--- 2018-05-07 reports/writeup/paper.tex
@@
-%-------------------------------------------------------------------------------
-% Preamble
-%-------------------------------------------------------------------------------
-\documentclass{ametsoc}
-\journal{jcli}
+\documentclass[12pt,letterpaper]{article}
 
-%Citations should be of the form ``author year''  not ``author, year''
-\bibpunct{(}{)}{;}{a}{}{,}
+% Page setup
+\usepackage[
+    margin=1in
+]{geometry}
 
@@
 \usepackage{
-	siunitx, % units
-	cleveref,% cross-references
-	physics, % better physics notation
-	url, % for URLs
 	amsmath, % for math
 	array, % for fixed-width tables
+	authblk, % for authors
 	booktabs, % for nice tables
+	graphicx, % figures
+	natbib, % references
 	nth, % for \nth{90}
+	physics, % better physics notation
+	siunitx, % units
+	url, % for URLs
 }
@@
 
+% better looking captions
+\usepackage[
+  format=plain,
+  labelfont={bf,it,footnotesize},
+  textfont={it,footnotesize}
+]{caption}
+
 \usepackage[acronym]{glossaries}
 \makeglossaries
+\newacronym{cca}{CCA}{canonical correlation analysis}
+\newacronym{ecmwf}{ECMWF}{European Centre for Medium-Range Weather Forecasts}
+\newacronym{enso}{ENSO}{El Ni\~{n}o-Southern Oscillation}
+\newacronym{eof}{EOF}{empirical orthogonal function}
+\newacronym{iri}{IRI}{International Research Institute for Climate and Society}
+\newacronym{hxlr}{HXLR}{heteroscedastic extended logistic regression}
+\newacronym{lprb}{LPRB}{Lower Paraguay River Basin}
+\newacronym{mjo}{MJO}{Madden-Julian Oscillation}
+\newacronym{mos}{MOS}{model output statistics}
+\newacronym{sallj}{SALLJ}{South American low-level jet}
+\newacronym{pcr}{PCR}{principal components regression}
+\newacronym{s2s}{S2S}{sub-seasonal to seasonal}
+\newacronym{sacz}{SACZ}{South Atlantic Convergence Zone}
+\newacronym{scad}{SCAD}{South Central Atlantic Dipole}
+\newacronym{sst}{SST}{sea-surface temperature}
+\newacronym{xlr}{XLR}{homoscedastic extended logistic regression}
 \newacronym{wt}{WT}{weather type}
-\newacronym{sallj}{SALLJ}{South American low-level jet}
-\newacronym{enso}{ENSO}{El Ni\~{n}o-Southern Oscillation}
-\newacronym{mjo}{MJO}{Madden-Julian Oscillation}
-\newacronym{sacz}{SACZ}{South Atlantic Convergence Zone}
-\newacronym{iri}{IRI}{International Research Institute for Climate and Society}
-\newacronym{ecmwf}{ECMWF}{European Centre for Medium-Range Weather Forecasts}
-\newacronym{s2s}{S2S}{sub-seasonal to seasonal}
-\newacronym{eof}{EOF}{empirical orthogonal function}
-\newacronym{mos}{MOS}{model output statistics}
-\newacronym{xlr}{XLR}{homoscedastic extended logistic regression}
-\newacronym{hxlr}{HXLR}{heteroscedastic extended logistic regression}
-\newacronym{pcr}{PCR}{principal components regression}
-\newacronym{cca}{CCA}{canonical correlation analysis}
-\newacronym{sst}{SST}{sea-surface temperature}
-\newacronym{scad}{SCAD}{South Central Atlantic Dipole}
-\newacronym{lprb}{LPRB}{Lower Paraguay River Basin}
 
@@
 
+\date{}
 \title{Heavy rainfall in Paraguay during the 2015-2016 austral summer: causes and sub-seasonal-to-seasonal predictive skill}
-\authors{James Doss-Gollin\correspondingauthor{James Doss-Gollin. Columbia Water Center. Dept. of Earth and Environmental Engineering, Columbia University, 500 W. 120th St., New York, NY. USA.}}
-\affiliation{Columbia Water Center.
-Dept.
-of Earth and Environmental Engineering, Columbia University, 500 W.
-120th St., New York, NY.
-USA \\ Columbia Water Center, Columbia University, 500 W.
-120th St., New York, NY.
-USA}
-\email{james.doss-gollin@columbia.edu}
-\extraauthor{\'Angel G. Mu\~noz}
-\extraaffil{Atmospheric and Oceanic Sciences (AOS). Princeton University. Princeton, NJ 08540-6649. USA  \\
-	International Research Institute for Climate and Society (IRI). The Earth Institute. Columbia University. Palisades, NY. 10964-1000. USA}
-\extraauthor{Simon J. Mason}
-\extraaffil{International Research Institute for Climate and Society (IRI). The Earth Institute. Columbia University. Palisades, NY. 10964-1000. USA}
-\extraauthor{Max Past\'{e}n}
-\extraaffil{Direcci\'{o}n de Meteorolog\'{i}a e Hidrolog\'{i}a. Paraguay \\ Facultad Polit\'ecnica. Universidad Nacional de Asunci\'{o}n. Paraguay}
+\author[1,2]{James Doss-Gollin}
+\author[3,4]{\'{A}ngel G. Mu\~{n}oz}
+\author[4]{Simon J. Mason}
+\author[5,6]{Max Past\'{e}n}
+\affil[1]{Department of Earth and Environmental Engineering, Columbia University}
+\affil[2]{Columbia Water Center}
+\affil[3]{Atmospheric and Oceanic Sciences (AOS), Princeton University}
+\affil[4]{International Research Institute for Climate and Society (IRI), The Earth Institute, Columbia University.}
+\affil[5]{Direcci\'{o}n de Meteorolog\'{i}a e Hidrolog\'{i}a, Paraguay}
+\affil[6]{Facultad Polit\'ecnica, Universidad Nacional de Asunci\'{o}n, Paraguay}
 
-%-------------------------------------------------------------------------------
-% Abstract
-% Up to 250 words in length
-%-------------------------------------------------------------------------------
+% These go last!
+\usepackage[hidelinks]{hyperref}
+\usepackage{cleveref}
 
-\abstract{
+\begin{document}
+
+This is a pre-print of the paper ``Heavy rainfall in Paraguay during the 2015-16 austral summer: causes and sub-seasonal-to-seasonal predictive skill'', by James Doss-Gollin, \'{A}ngel G.
+Mu\~{n}oz, Simon J.
+Mason, and Max Past\'{e}n, published in Journal of Climate.
+You can find the final, type-set version at \url{https://journals.ametsoc.org/doi/abs/10.1175/JCLI-D-17-0805.1} or see the full citation reference in the bibliography under \citet{DossGollin:2018bn}.
+
+Codes used to build this paper are available at \url{https://github.com/jdossgollin/2018-paraguay-floods}.
+
+\clearpage
+
+\maketitle
+
+\begin{abstract}
 	During the austral summer 2015-16, severe flooding displaced over \num{170000} people on the Paraguay River system in Paraguay, Argentina, and Southern Brazil.
@@
 	A continuous monitoring of climate drivers impacting rainfall in the region, and the use of statistically corrected heavy precipitation seasonal and sub-seasonal forecasts, may help improve flood preparedness in this and other regions.
-}
-
-\begin{document}
-
-\maketitle
+\end{abstract}
 
@@
 The aim of this paper is to diagnose the drivers of the November-February (NDJF) 2015-16 rainfall and flooding events, and to assess the skill of the associated subseasonal-to-seasonal predictions.
+
+\begin{figure}
+	\centering
+	\includegraphics[width=6.5in]{study_area.jpg}
+	\caption{
+		Topographical map of the study area.
+		Colors indicate $\log_{10}$ of elevation, in \si{\meter}, from the Global Land One-Km Base Elevation project available at \texttt{http://iridl.ldeo.columbia.edu/SOURCES/.NOAA/.NGDC/.GLOBE/.topo/}.
+		(a): all of South America.
+		The domains of the \acrlong{lprb} and the domain used for weather typing are indicated in red and blue, respectively.
+		(b): The \acrfull{lprb}.
+		As for (a), the \gls{lprb} is marked with a red box.
+		Streamflow time series shown in \cref{fig:streamflow} were taken from the fours stations indicated.
+		The Paraguay River and its tributaries, from the Natural Earth database (\texttt{www.naturalearthdata.com}), are also shown.
+		Full station names are: Bah\'ia Negra (Bne); Concepci\'on (Conc); Asunci\'on (Asu); Pilar (Pil).
+	}\label{fig:study-area}
+\end{figure}
+\begin{figure}
+	\centering
+	\includegraphics[width=6.5in]{anomalies_ndjf1516.pdf}
+	\caption{
+		Monthly composite anomalies observed during NDJF 2015-16.
+		Top row (a-d) shows streamfunction anomalies at \SI{850}{\hecto\pascal}.
+		Bottom row (e-h) shows rainfall anomalies, in units of \si{\milli\meter\per\day}.
+	}\label{fig:anomalies}
+\end{figure}
 
@@
 Computation was performed in the python environment using stable open source packages \citep{Hunter2007,McKinney2010,vanderWalt2011,Hoyer2017}.
-All codes to reproduce or modify this analysis are available at \url{https://github.com/jdossgollin/PYFloods}.
+All codes to reproduce or modify this analysis are available at the permanent link \url{https://doi.org/10.5281/zenodo.1243104}.
 
@@
 
+\begin{table}
+	\centering
+	\caption{
+		\Gls{mos} methods used to correct the \gls{ecmwf} sub-seasonal forecasts.
+		Spatial domain for predictand is always the same (\SIrange{39}{17}{\degree S}; \SIrange{66}{49}{\degree W}).
+		Two initializations are used: Nov 12th and 16th, 2015.
+	}\label{tab:mos-methods}
+	\begin{center}
+		\begin{tabular}{L{0.75in} L{1.15in} L{3.5in}}
+			\toprule
+			Model & Region (Predictor) & Final predictor(s) selected \\
+			%
+			\midrule
+			%
+			\emph{Raw} & \SIrange{39}{17}{\degree S}; \SIrange{66}{49}{\degree W} & Ensemble mean, computed using members from the  two initializations. No correction performed. \\
+			%
+			\emph{XLR} & \SIrange{39}{17}{\degree S}; \SIrange{66}{49}{\degree W} & Ensemble mean, computed using members from the  two initializations.  \\
+			%
+			\emph{HLXR} & \SIrange{39}{17}{\degree S}; \SIrange{66}{49}{\degree W} & Ensemble mean and spread, computed using  members from the two initializations.\\
+			%
+			\emph{PCR} & \SIrange{60}{0}{\degree S}; \SIrange{80}{30}{\degree W} & Linear combination of model's \glspl{eof} computed using both initializations as independent predictors (10 \glspl{eof}).\\
+			%
+			\emph{CCA} & \SIrange{60}{0}{\degree S}; \SIrange{80}{30}{\degree W} & Canonical modes computed using both initializations as independent predictors.
+(10 predictor \glspl{eof}, 4 predictand \glspl{eof}, 4 canonical modes) \\
+			%
+			\bottomrule
+			%
+		\end{tabular}
+	\end{center}
+\end{table}
+
 To evaluate model skill, we use a cross-validation approach with a 5-year window.
@@
 
+\begin{figure}
+	\centering
+	\includegraphics[width=6.5in]{Streamflow.pdf}
+	\caption{
+		River stage (height; in \si{\meter}) for the Paraguay River at four gauges along the Paraguay River.
+		The station names are shortened versions of those shown in \cref{fig:study-area}.
+		(a): Seasonality (orange) and time series of 2015-16 observations (black) at each stream gauge.
+		Seasonality was fit using local polynomial regression as implemented in the \texttt{locfit} package in the \textbf{R} statistical programming environment \citep{Loader1999}.
+		(b): Time series of daily stage measurements from 1929 to 2016 at each station.
+	}\label{fig:streamflow}
+\end{figure}
+
 Examination of \cref{fig:streamflow}b suggests multidecadal oscillation in the streamflow record.
@@
 
+\begin{figure}
+	\centering
+	\includegraphics[width=6.5in]{lagged_rain.pdf}
+	\caption{
+		Composite anomalies associated with heavy rainfall (\nth{90} percentile exceedance of area-averaged rainfall in the \acrlong{lprb}).
+		Lagged composites are shown, by column, for $t = $ \SIlist{-2;-1;0;1}{\day} relative to the date of heavy rainfall.
+		Top row (a-d) shows composite streamfunction and wind anomalies at \SIlist{850}{\hecto\pascal}.
+		Strongest 5\% of wind anomaly vectors between \SI{60}{\degree S} and \SI{10}{\degree N} (all longitudes) are also shown.
+		Bottom row (e-h) shows composite rainfall anomalies, in units of \si{\milli\meter\per\day}.
+	}\label{fig:lagged-rain}
+\end{figure}
+
 \subsection{Weather Type Analysis: Daily Circulation Patterns}\label{sec:weather-types}
@@
 This is consistent with the hypothesis that the \glspl{eof} over the study area are associated with large-scale patterns.
+
+\begin{figure}
+	\centering
+	\includegraphics[width=6.5in]{eof_loadings.pdf}
+	\caption{
+		Loadings of the four leading \glspl{eof} of daily NDJF \SI{850}{\hecto\pascal} streamfunction over the weather typing region shown in \cref{fig:study-area}.
+		Parentheses in sub-plot titles indicate the percentage of total variance explained by each \gls{eof}.
+	}\label{fig:eof-loading}
+\end{figure}
+
+\begin{figure}
+	\centering
+	\includegraphics[width=6.5in,height=9in,keepaspectratio=true]{wt_composite.pdf}
+	\caption{
+		Composite anomalies associated with each weather type.
+		Top row (a-f) shows streamfunction anomalies at \SIlist{850}{\hecto\pascal}.
+		Strongest 20\% of wind anomaly vectors over the plot area are also shown.
+		Bottom row (g-l) shows rainfall anomalies, in units of \si{\milli\meter\per\day}.
+		The relative frequency of occurrence of each weather type (in days) is presented on the top of each column.
+	}\label{fig:wt-composite}
+\end{figure}
 
@@
 This plot shows that heavy rainfall concentrated over a period spanning from mid-November 2015 through early January 2016, with shorter peaks in late January and mid-February.
+
+\begin{figure}
+	\centering
+	\includegraphics[width=6.5in]{rain_wt_201516.pdf}
+	\caption{
+		Time series of area-averaged rainfall in the \acrlong{lprb} (\cref{fig:study-area}) for each day of NDJF  2015-16.
+		Lines indicate the rainfall value, in units of \si{\milli\meter\per\day}.
+        The weather type corresponding to each day is indicated in an adjacent text label.
+		Dashed lines blue indicate (from bottom to top) the climatological 50th, \nth{90}, and 99th percentiles of NDJF area-averaged rain over the \acrlong{lprb}.
+	}\label{fig:rain-wt}
+\end{figure}
 
@@
 
+\begin{figure}
+	\centering
+	\includegraphics[width=3.5in]{seasonal_forecast.pdf}
+	\caption{
+		Seasonal model forecast for probability of exceedance of \nth{90} percentile of DJF rainfall, as issued in November 2015.
+		Color indicates the forecast probability of exceeding the \nth{90} percentile of climatological rainfall during DJF 2015-16 -- this is presented as the odds ratio as defined in \cref{eq:odds-ratio}.
+		A value greater than 1 indicates that the model forecast greater-than-average odds of rainfall exceeding the \nth{90} percentile.
+		Grid cells which observed an exceedance of the \nth{90} percentile of DJF rainfall are outlined in black.
+	}\label{fig:seas-prob-fcst}
+\end{figure}
+
 \subsection{Sub-seasonal Forecasts}
@@
 
+\begin{figure}
+	\centering
+	\includegraphics[width=6.5in]{chiclet.pdf}
+	\caption{
+		Chiclet diagram \citep[see][]{Carbin2016} of ensemble-mean precipitation anomaly forecasts over the \acrlong{lprb} (see \cref{fig:study-area}) from uncorrected \gls{ecmwf} \gls{s2s} model forecast data, as a function of the forecast target date (horizontal axis) and lead time (vertical axis).
+		Time series of CPC daily mean precipitation over the same area is plotted with $y$-axis inverted; horizontal black line denotes NDJF climatology.
+	}\label{fig:chiclet}
+\end{figure}
+
 To examine these forecasts more closely, we turn to the 14-19 day forecast of the December 1-7 2015 period.
@@
 These skill scores indicate that the model is capturing a signal, and suggest the use of \gls{mos} methods to explore the extent to which corrections in the magnitudes and spatial patterns may improve the forecast.
+
+\begin{figure}
+	\centering
+	\includegraphics[width=6.4in]{mos_forecasts.pdf}
+	\caption{
+		Raw and \gls{mos}-adjusted \gls{s2s} model forecasts and skill scores for the methods indicated in \cref{tab:mos-methods}.
+		Top row (a-e) shows the heavy rainfall forecast for 1-7 December 2015 as the odds ratio defined in \cref{eq:odds-ratio} over the target domain.
+		A value greater than 1 indicates that the model forecast greater-than-average odds of rainfall exceeding the \nth{90} percentile.
+		Second row (f-j) shows the Ignorance Score defined in \cref{eq:ignorance}, with zero indicating a perfect forecast.
+		Bottom row (k-o) shows the 2AFC skill score for each grid cell; a value greater than 50 indicates that the model outperforms climatology.
+		Columns separate different \gls{mos} models except for ``Raw'' (a,f,k), which indicates the uncorrected \gls{s2s} model output.
+		For all rows the grid cells which observed a \nth{90} percentile exceedance for 1-7 December 2015 are outlined in black.
+	}\label{fig:subs-prob-fcst}
+\end{figure}
 
@@
 
+\begin{figure}
+	\centering
+	\includegraphics[width=\textwidth]{wt_mjo_enso.pdf}
+	\caption{
+		Anomalous probability of occurrence of each weather type concurrent with observance of each \gls{mjo} phase.
+		When \gls{mjo} amplitude is less than 1, it is defined as neutral phase (0).
+		Plots are shown separately for El Ni\~no (NINO 3.4 $>1$), La Ni\~na (NINO 3.4 $< -1$), and Neutral \gls{enso} phases.
+		Only values which are significant at $\alpha=0.10$, calculated with a bootstrap of 5000 samples, are shown.
+	}\label{fig:wt-mjo-enso}
+\end{figure}
+
 By analyzing how the joint behavior physical mechanisms modulate the probability of occurrence of certain weather types, it may be possible to better understand the drivers of this and future extreme event(s).
@@
 This suggests that not only did \gls{enso}-\gls{mjo} conditions favor \gls{sallj} activity, but Atlantic-Pacific interactions specifically favored  \gls{wt} 4 occurrence, helping to explain why the most intense rainfall anomalies occurred specifically in the \gls{lprb}.
+
+\begin{figure}
+	\centering
+	\includegraphics[width=6.4in]{ChacoNoChacojet.pdf}
+	\caption{
+		Schematics of low-level jet events (red arrows) during austral summer and El Ni\~no years.
+		Most jet events are of the ``Chaco'' type, particularly when SST anomalies in the central southern Atlantic Ocean (a, see green box) are weak.
+		When a dipole SST anomaly occurs in the central southern Atlantic with the warmer pole equatorward, the meridional temperature gradient and sea-land temperature contrasts establish an anticyclonic circulation (dot-dashed line) conducive to increased occurrence of No-Chaco jet events (b).
+		Other SST anomaly configurations tend to be present outside the green box (not shown).
+		Winds in panels are typical for each case (at \SI{850}{\hecto\pascal}).
+		Reference wind vector in \si{\meter\per\second}.
+		Green box shows location of \gls{scad}.
+	}\label{fig:chaco-nochaco}
+\end{figure}
 
@@
 
-\acknowledgments
+\section*{Acknowledgments}
 The authors thank David Farnham, Upmanu Lall, Laureline Josset, and Andrew Robertson for insightful conversations and guidance.
@@
 
+\setglossarystyle{list}
+\printglossary[type=\acronymtype]
 \bibliographystyle{ametsoc2014}
@@
 
-%-------------------------------------------------------------------------------
-% TABLES
-%-------------------------------------------------------------------------------
-\begin{table}
-%
-\caption{
-	\Gls{mos} methods used to correct the \gls{ecmwf} sub-seasonal forecasts.
-	Spatial domain for predictand is always the same (\SIrange{39}{17}{\degree S}; \SIrange{66}{49}{\degree W}).
-	Two initializations are used: Nov 12th and 16th, 2015.
-}\label{tab:mos-methods}
-\begin{center}
-	\begin{tabular}{L{0.75in} L{1.15in} L{3.5in}}
-		\toprule
-		Model & Region (Predictor) & Final predictor(s) selected \\
-		%
-		\midrule
-		%
-		\emph{Raw} & \SIrange{39}{17}{\degree S}; \SIrange{66}{49}{\degree W} & Ensemble mean, computed using members from the  two initializations. No correction performed. \\
-		%
-		\emph{XLR} & \SIrange{39}{17}{\degree S}; \SIrange{66}{49}{\degree W} & Ensemble mean, computed using members from the  two initializations.  \\
-		%
-		\emph{HLXR} & \SIrange{39}{17}{\degree S}; \SIrange{66}{49}{\degree W} & Ensemble mean and spread, computed using  members from the two initializations.\\
-		%
-		\emph{PCR} & \SIrange{60}{0}{\degree S}; \SIrange{80}{30}{\degree W} & Linear combination of model's \glspl{eof} computed using both initializations as independent predictors (10 \glspl{eof}).\\
-		%
-		\emph{CCA} & \SIrange{60}{0}{\degree S}; \SIrange{80}{30}{\degree W} & Canonical modes computed using both initializations as independent predictors.
-(10 predictor \glspl{eof}, 4 predictand \glspl{eof}, 4 canonical modes) \\
-		%
-		\bottomrule
-		%
-	\end{tabular}
-\end{center}
-\end{table}
-
-%-------------------------------------------------------------------------------
-% FIGURES
-%-------------------------------------------------------------------------------
-
-\begin{figure*}
-	\noindent\includegraphics[width=6.5in]{study_area.jpg}
-	\caption{
-		Topographical map of the study area.
-		Colors indicate $\log_{10}$ of elevation, in \si{\meter}, from the Global Land One-Km Base Elevation project available at \texttt{http://iridl.ldeo.columbia.edu/SOURCES/.NOAA/.NGDC/.GLOBE/.topo/}.
-		(a): all of South America.
-		The domains of the \acrlong{lprb} and the domain used for weather typing are indicated in red and blue, respectively.
-		(b): The \acrfull{lprb}.
-		As for (a), the \gls{lprb} is marked with a red box.
-		Streamflow time series shown in \cref{fig:streamflow} were taken from the fours stations indicated.
-		The Paraguay River and its tributaries, from the Natural Earth database (\texttt{www.naturalearthdata.com}), are also shown.
-		Full station names are: Bah\'ia Negra (Bne); Concepci\'on (Conc); Asunci\'on (Asu); Pilar (Pil).
-	}\label{fig:study-area}
-\end{figure*}
-
-\begin{figure*}
-	\noindent\includegraphics[width=6.5in]{anomalies_ndjf1516.pdf}
-	\caption{
-		Monthly composite anomalies observed during NDJF 2015-16.
-		Top row (a-d) shows streamfunction anomalies at \SI{850}{\hecto\pascal}.
-		Bottom row (e-h) shows rainfall anomalies, in units of \si{\milli\meter\per\day}.
-	}\label{fig:anomalies}
-\end{figure*}
-
-\begin{figure*}
-	\noindent\includegraphics[width=6.5in]{Streamflow.pdf}
-	\caption{
-		River stage (height; in \si{\meter}) for the Paraguay River at four gauges along the Paraguay River.
-		The station names are shortened versions of those shown in \cref{fig:study-area}.
-		(a): Seasonality (orange) and time series of 2015-16 observations (black) at each stream gauge.
-		Seasonality was fit using local polynomial regression as implemented in the \texttt{locfit} package in the \textbf{R} statistical programming environment \citep{Loader1999}.
-		(b): Time series of daily stage measurements from 1929 to 2016 at each station.
-	}\label{fig:streamflow}
-\end{figure*}
-
-\begin{figure*}	\noindent\includegraphics[width=6.5in]{lagged_rain.pdf}
-	\caption{
-		Composite anomalies associated with heavy rainfall (\nth{90} percentile exceedance of area-averaged rainfall in the \acrlong{lprb}).
-		Lagged composites are shown, by column, for $t = $ \SIlist{-2;-1;0;1}{\day} relative to the date of heavy rainfall.
-		Top row (a-d) shows composite streamfunction and wind anomalies at \SIlist{850}{\hecto\pascal}.
-		Strongest 5\% of wind anomaly vectors between \SI{60}{\degree S} and \SI{10}{\degree N} (all longitudes) are also shown.
-		Bottom row (e-h) shows composite rainfall anomalies, in units of \si{\milli\meter\per\day}.
-	}\label{fig:lagged-rain}
-\end{figure*}
-
-\begin{figure*}
-	\noindent\includegraphics[width=6.5in]{eof_loadings.pdf}
-	\caption{
-		Loadings of the four leading \glspl{eof} of daily NDJF \SI{850}{\hecto\pascal} streamfunction over the weather typing region shown in \cref{fig:study-area}.
-		Parentheses in sub-plot titles indicate the percentage of total variance explained by each \gls{eof}.
-	}\label{fig:eof-loading}
-\end{figure*}
-
-\begin{figure*}
-	\noindent\includegraphics[width=6.5in,height=9in,keepaspectratio=true]{wt_composite.pdf}
-	\caption{
-		Composite anomalies associated with each weather type.
-		Top row (a-f) shows streamfunction anomalies at \SIlist{850}{\hecto\pascal}.
-		Strongest 20\% of wind anomaly vectors over the plot area are also shown.
-		Bottom row (g-l) shows rainfall anomalies, in units of \si{\milli\meter\per\day}.
-		The relative frequency of occurrence of each weather type (in days) is presented on the top of each column.
-	}\label{fig:wt-composite}
-\end{figure*}
-
-\begin{figure*}
-	\noindent\includegraphics[width=6.5in]{rain_wt_201516.pdf}
-	\caption{
-		Time series of area-averaged rainfall in the \acrlong{lprb} (\cref{fig:study-area}) for each day of NDJF  2015-16.
-		Lines indicate the rainfall value, in units of \si{\milli\meter\per\day}.
-        The weather type corresponding to each day is indicated in an adjacent text label.
-		Dashed lines blue indicate (from bottom to top) the climatological 50th, \nth{90}, and 99th percentiles of NDJF area-averaged rain over the \acrlong{lprb}.
-	}\label{fig:rain-wt}
-\end{figure*}
-
-\begin{figure*}
-	\noindent\includegraphics[width=3.5in]{seasonal_forecast.pdf}
-	\caption{
-		Seasonal model forecast for probability of exceedance of \nth{90} percentile of DJF rainfall, as issued in November 2015.
-		Color indicates the forecast probability of exceeding the \nth{90} percentile of climatological rainfall during DJF 2015-16 -- this is presented as the odds ratio as defined in \cref{eq:odds-ratio}.
-		A value greater than 1 indicates that the model forecast greater-than-average odds of rainfall exceeding the \nth{90} percentile.
-		Grid cells which observed an exceedance of the \nth{90} percentile of DJF rainfall are outlined in black.
-	}\label{fig:seas-prob-fcst}
-\end{figure*}
-
-\begin{figure*}
-	\noindent\includegraphics[width=6.5in]{chiclet.pdf}
-	\caption{
-		Chiclet diagram \citep[see][]{Carbin2016} of ensemble-mean precipitation anomaly forecasts over the \acrlong{lprb} (see \cref{fig:study-area}) from uncorrected \gls{ecmwf} \gls{s2s} model forecast data, as a function of the forecast target date (horizontal axis) and lead time (vertical axis).
-		Time series of CPC daily mean precipitation over the same area is plotted with $y$-axis inverted; horizontal black line denotes NDJF climatology.
-	}\label{fig:chiclet}
-\end{figure*}
-
-\begin{figure*}
-	\noindent\includegraphics[width=6.4in]{mos_forecasts.pdf}
-	\caption{
-		Raw and \gls{mos}-adjusted \gls{s2s} model forecasts and skill scores for the methods indicated in \cref{tab:mos-methods}.
-		Top row (a-e) shows the heavy rainfall forecast for 1-7 December 2015 as the odds ratio defined in \cref{eq:odds-ratio} over the target domain.
-		A value greater than 1 indicates that the model forecast greater-than-average odds of rainfall exceeding the \nth{90} percentile.
-		Second row (f-j) shows the Ignorance Score defined in \cref{eq:ignorance}, with zero indicating a perfect forecast.
-		Bottom row (k-o) shows the 2AFC skill score for each grid cell; a value greater than 50 indicates that the model outperforms climatology.
-		Columns separate different \gls{mos} models except for ``Raw'' (a,f,k), which indicates the uncorrected \gls{s2s} model output.
-		For all rows the grid cells which observed a \nth{90} percentile exceedance for 1-7 December 2015 are outlined in black.
-	}\label{fig:subs-prob-fcst}
-\end{figure*}
-
-\begin{figure*}
-	\includegraphics[width=\textwidth]{wt_mjo_enso.pdf}
-	\caption{
-		Anomalous probability of occurrence of each weather type concurrent with observance of each \gls{mjo} phase.
-		When \gls{mjo} amplitude is less than 1, it is defined as neutral phase (0).
-		Plots are shown separately for El Ni\~no (NINO 3.4 $>1$), La Ni\~na (NINO 3.4 $< -1$), and Neutral \gls{enso} phases.
-		Only values which are significant at $\alpha=0.10$, calculated with a bootstrap of 5000 samples, are shown.
-	}\label{fig:wt-mjo-enso}
-\end{figure*}
-
-\begin{figure*}
-	\noindent\includegraphics[width=6.4in]{ChacoNoChacojet.pdf}
-	\caption{
-		Schematics of low-level jet events (red arrows) during austral summer and El Ni\~no years.
-		Most jet events are of the ``Chaco'' type, particularly when SST anomalies in the central southern Atlantic Ocean (a, see green box) are weak.
-		When a dipole SST anomaly occurs in the central southern Atlantic with the warmer pole equatorward, the meridional temperature gradient and sea-land temperature contrasts establish an anticyclonic circulation (dot-dashed line) conducive to increased occurrence of No-Chaco jet events (b).
-		Other SST anomaly configurations tend to be present outside the green box (not shown).
-		Winds in panels are typical for each case (at \SI{850}{\hecto\pascal}).
-		Reference wind vector in \si{\meter\per\second}.
-		Green box shows location of \gls{scad}.
-	}\label{fig:chaco-nochaco}
-\end{figure*}
-
 \end{document}

### 2019-01-28 reports/preprint/preprint.tex
update preprint for OSF conventions

@@
 \usepackage[
-    margin=1in
+    margin=0.9in
 ]{geometry}
@@
 
-\usepackage[acronym]{glossaries}
+\usepackage[acronym,toc,nopostdot,nonumberlist]{glossaries}
 \makeglossaries
@@
 
+% These go last!
+\usepackage[hidelinks]{hyperref}
+\usepackage{cleveref}
+
 %-------------------------------------------------------------------------------
@@
 \title{Heavy rainfall in Paraguay during the 2015-2016 austral summer: causes and sub-seasonal-to-seasonal predictive skill}
-\author[1,2]{James Doss-Gollin}
+\author[1,2]{James Doss-Gollin\thanks{\href{mailto:james.doss-gollin@columbia.edu}{james.doss-gollin@columbia.edu}}}
 \author[3,4]{\'{A}ngel G. Mu\~{n}oz}
@@
 
-% These go last!
-\usepackage[hidelinks]{hyperref}
-\usepackage{cleveref}
-
 \begin{document}
 
-This is a pre-print of the paper ``Heavy rainfall in Paraguay during the 2015-16 austral summer: causes and sub-seasonal-to-seasonal predictive skill'', by James Doss-Gollin, \'{A}ngel G.
+This is a post-print of the paper ``Heavy rainfall in Paraguay during the 2015-16 austral summer: causes and sub-seasonal-to-seasonal predictive skill'', by James Doss-Gollin, \'{A}ngel G.
 Mu\~{n}oz, Simon J.
@@
 Codes used to build this paper are available at \url{https://github.com/jdossgollin/2018-paraguay-floods}.
+Comments to the authors may be directed to James Doss-Gollin: \href{mailto:james.doss-gollin@columbia.edu}{james.doss-gollin@columbia.edu}
 
@@
 	\centering
-	\includegraphics[width=6.5in]{study_area.jpg}
+	\includegraphics[width=5in]{study_area.jpg}
 	\caption{
